\section*{Acknowledgments}


This work was supported by Zoo Labs Foundation. We thank the Lux Network team for the FHE precompile infrastructure (\texttt{lux/precompile/fhe}), the \texttt{lux/fhe} Go TFHE library on \texttt{luxfi/lattice}, and the \texttt{luxcpp/gpu} dispatcher + \texttt{lux-private/gpu-kernels} acceleration substrate that this work consumes unchanged.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{chillotti2020tfhe}
I. Chillotti, N. Gama, M. Georgieva, and M. Izabach\`{e}ne, ``TFHE: Fast Fully Homomorphic Encryption over the Torus,'' \textit{Journal of Cryptology}, vol. 33, pp. 34--91, 2020.

\bibitem{cheon2017ckks}
J.H. Cheon, A. Kim, M. Kim, and Y. Song, ``Homomorphic Encryption for Arithmetic of Approximate Numbers,'' \textit{ASIACRYPT}, 2017.

\bibitem{lux-tfhe}
Lux Industries, ``Lux TFHE: Native Fully Homomorphic Encryption on the Lux Network,'' Technical Report, 2025.

\bibitem{lux-fhevm}
Lux Industries, ``TorusEVM: FHE-Enabled Ethereum Virtual Machine,'' Technical Report, 2025.

\bibitem{zoo-federated-wildlife}
Z. Kelling et al., ``Federated Wildlife Monitoring on the Zoo Network,'' Zoo Labs Foundation, 2025.

\bibitem{zoo-conservation-ai}
Z. Kelling et al., ``Conservation AI: Decentralized Species Monitoring and Habitat Analysis,'' Zoo Labs Foundation, 2025.

\bibitem{lux-evmgpu-benchmark}
Lux Industries, ``GPU-Accelerated EVM: Benchmarks and Architecture,'' Technical Report, 2025.

\bibitem{albrecht2018hes}
M. Albrecht et al., ``Homomorphic Encryption Standard,'' \textit{Protecting Privacy through Homomorphic Encryption}, Springer, 2021.

\bibitem{gilad2016cryptonets}
R. Gilad-Bachrach et al., ``CryptoNets: Applying Neural Networks to Encrypted Data,'' \textit{ICML}, 2016.

\bibitem{brutzkus2019low}
A. Brutzkus, R. Gilad-Bachrach, and O. Elisha, ``Low Latency Privacy Preserving Inference,'' \textit{ICML}, 2019.

\bibitem{mcmahan2017communication}
B. McMahan et al., ``Communication-Efficient Learning of Deep Networks from Decentralized Data,'' \textit{AISTATS}, 2017.

\bibitem{zhu2019deep}
L. Zhu, Z. Liu, and S. Han, ``Deep Leakage from Gradients,'' \textit{NeurIPS}, 2019.

\bibitem{lux-mchain-mpc}
Lux Industries, ``M-Chain: Threshold Multi-Party Computation on Lux Network,'' Technical Report, 2025.

\end{thebibliography}

