\section{What Is Built vs.\ What Is Designed}
\label{sec:status-summary}

A reader should not have to reach the implementation section to learn which of this
paper's claims are backed by running code and which are specifications awaiting
implementation. We therefore state the boundary here, before the construction, and
repeat the discipline at each section that makes an execution-level claim. The full
per-result ledger is Section~\ref{sec:impl-status}; the two companion
documents---the soundness companion~\cite{poisoundness} and the
emission/enforcement companion~\cite{poiemission}---carry the same boundary from the proof
side and are consistent with this table. This section is that table, surfaced.

\paragraph{The honest one-line summary.} The \emph{soundness primitive} is proven and
shipped; the end-to-end \emph{execution-proof system} it enables is now built and tested,
not a roadmap; and the residual is a named, smaller pair (a floating-point tier, and
emission coverage of architectures beyond the tested families). The matmul verifier that
catches a fabricated output with probability $\ge 1 - 2^{-122}$ (at two challenge vectors)
and never false-rejects honest integer work is implemented and unit-tested. The wire that
the earlier draft of this paper marked ``designed, not built''---a \emph{running} forward
pass committing a streamed activation-trace transcript that a challenger opens and
Freivalds-checks against the committed operands---is built: it emits from a real
quantized forward (\texttt{ProvableLinear}), composes across a whole typed graph
(\texttt{GraphTrace}, including a training step), is verified on-chain in pure Solidity
(\texttt{ComputeWitnessLib}), and gates a bond-slashing optimistic state machine
(\texttt{OptimisticEvidence}). The properties this buys---``a fabricated matmul is
provable'' and ``an honest prover is never slashable''---are theorems with passing
adversarial tests behind each, across three languages with byte-identical commitments.

\begin{table}[h]
\centering
\caption{Built-vs-designed boundary for the execution-proof scheme. ``Built'' = merged,
released under semantic versioning, and tested; the test counts are the actual passing
counts of the suites named in the last column (Rust \texttt{cargo test}, Go
\texttt{go test}, Solidity \texttt{forge test}). ``Designed'' = specified in this paper,
not yet implemented. The decisive load-bearing item---emission of the activation-trace
transcript from a live forward pass---is now in the \emph{first} block, which is why the
system-level ``cannot be faked'' claim is no longer conditional on an unbuilt wire (only on
the stated cryptographic assumptions and the determinism contract).}
\label{tab:status-summary}
\small
\begin{tabular}{@{}p{7.4cm}p{2.2cm}p{4.0cm}@{}}
\toprule
Component & Status & Evidence / where \\
\midrule
Exact-integer Freivalds matmul verifier
  & \textbf{built}, tested
  & \texttt{poi.rs} ($7$); \texttt{luxfi/crypto/poi} ($23$) \\
Activation-trace transcript (keccak Merkle) + challenger (open + Freivalds)
  & \textbf{built}, tested
  & \texttt{poi\_transcript.rs} ($8$) \\
Proof-bearing forward: returned value $=$ committed exact-integer result
  & \textbf{built}, tested
  & \texttt{poi\_forward.rs} \texttt{ProvableLinear} ($3$) \\
Full-graph composition (attention, MoE, training) via typed trace
  & \textbf{built}, tested
  & \texttt{poi\_graph.rs} \texttt{GraphTrace} ($9$) \\
Adversarial red-team battery (each forbidden attack is caught)
  & \textbf{built}, tested
  & \texttt{poi\_adversarial.rs} ($11$) \\
Soundness theorems ($1/p$ one-layer; $2^{-122}$ amplified; zero false-reject; composition)
  & \textbf{proven}
  & \S\ref{sec:freivalds}; \cite{poisoundness,poiemission} \\
Lowest-token-id argmax / canonical MoE top-$k$ (determinism D3)
  & \textbf{built}, tested
  & \texttt{sampler.rs}, \texttt{ops.rs}, \texttt{poi\_graph.rs} \\
On-chain Freivalds + Merkle witness (pure Solidity)
  & \textbf{built}, tested
  & \texttt{ComputeWitnessLib.sol} ($5$) \\
Optimistic commit/challenge/slash gate (bond paid on verified fraud only)
  & \textbf{built}, tested
  & \texttt{OptimisticEvidence.sol} ($18$) \\
Fail-closed mint gate (no compute-verifier $\Rightarrow$ no mint)
  & \textbf{built}, tested
  & \texttt{AIMiner.sol} ($12$); \texttt{ComputeProofEnforcement} ($19$) \\
Capped, halving fair-mined coin; global no-double-mint
  & \textbf{built}, tested
  & \texttt{AICoin.sol} ($11$); \texttt{AIComputeRegistry} ($6$) \\
Miner bond/slash registry
  & \textbf{built}, tested
  & \texttt{MinerStakeRegistry.sol} ($8$) \\
\Zoo{} L3 directory + endorsement (settlement registry)
  & \textbf{built}, tested
  & \texttt{L3Registry.sol} ($8$); \S\ref{sec:l3-settlement} \\
\midrule
RepOps pinned-reduction-tree fp16/bf16/fp8 mode
  & \textbf{designed}; proven
  & \S\ref{sec:repops}; \cite{poisoundness} \\
Emission coverage of architecture families beyond the tested set
  & \textbf{designed}; proven
  & \S\ref{sec:families}; \cite{poisoundness} \\
Sequential-chaining proof for SSM / diffusion latents (the \emph{chaining} obligation)
  & \textbf{designed}; proven
  & \S\ref{sec:ssm-proof}, \S\ref{sec:diffusion}; \cite{poisoundness} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Three consequences a reviewer can grep for.} A peer reviewer who reads the
repositories will find exactly these, and we state them so they are unmistakable:

\begin{enumerate}[leftmargin=1.6em]
\item \textbf{The verifier now has callers.} The earlier draft's caveat---``the verifier
has no production callers''---no longer holds. \texttt{crypto/poi}'s \texttt{VerifyMulti}
and the engine's \texttt{poi.rs} are invoked end to end: \texttt{poi\_transcript.rs} opens a
beacon-selected matmul of a real \texttt{ProvableLinear}/\texttt{GraphTrace} forward and
Freivalds-checks it, the Solidity \texttt{ComputeWitnessLib} runs the identical check
on-chain, and \texttt{OptimisticEvidence.challenge} calls \texttt{provesFraud} to slash a
bond. The \texttt{AIMiner} mint gate \emph{reverts} when no compute-verifier is wired
(\texttt{ProofGateUnavailable}), so the proofless-mint default the earlier draft flagged is
closed: out of the box a proof is mandatory, not opt-in.
\item \textbf{The ``exact, software-only'' claim is about a kernel the engine ships in the
proof path.} \texttt{ProvableLinear} computes $y=\mathrm{dequant}(\bar x\cdot W^\top)$ on a
\emph{whole-$K$ exact-$i64$ accumulator} (\texttt{poi.rs::exact\_matmul}), commits exactly
those operands, and dequantizes the committed $C$---so the returned value \emph{is} the
committed computation, with no floating-point reduction in the committed trace and zero
cross-hardware drift (the float-shadow unsoundness the earlier draft warned of is avoided by
construction). The throughput cost of the exact accumulator is paid only on proof-bearing
inferences; the fast \texttt{f8q8} path is unchanged for ordinary inference.
\item \textbf{Provable-today architecture coverage is no longer $0\%$.} The test suite emits
and verifies a real SwiGLU FFN ($\texttt{gate}\to\texttt{up}\to\texttt{down}$), the
attention spine ($QK^\top\to$ scale $\to$ deterministic integer softmax $\to\cdot V$), a
mixture-of-experts block (router $\to$ top-$k$ $\to$ gather $\to$ expert matmuls $\to$
combine), and a single SGD training step (forward $+$ backward $+$ update) as one typed
graph---so the families with a committed, verified transcript today are nonempty. What
remains a \emph{designed} ceiling rather than a deployed floor is the breadth: the
fp16/bf16/fp8 tier (RepOps) and the architecture families not yet wired to emit (the
sequential-chaining obligation for state-space and diffusion models, and the long tail of
the model zoo). ``Full modern-LLM coverage'' names the design's reach; the tested set names
its present floor, and the floor is no longer empty.
\end{enumerate}

None of this weakens the mathematics. The Freivalds soundness theorem, the binding's
anti-replay/anti-substitution lemmas, the floating-point no-go, and the whole-graph
composition theorem are correct as stated and are the Bitcoin-whitepaper-grade
contribution of the paper. What the table fixes, relative to the earlier draft, is the
\emph{tense} of the system claims: we have proven a sound primitive, shipped its core,
\emph{and built and tested} the emission and on-chain enforcement that turn that primitive
into an unfakeable settlement gate---leaving a named floating-point tier and wider
architecture emission as the honest remaining work.
