\section{Introduction}


Blockchain consensus mechanisms face a fundamental challenge: how to fairly allocate block production rights and rewards. Proof of Work (PoW) solves this by requiring computational effort, but the computation serves no purpose beyond securing the chain---an estimated 150 TWh annually for Bitcoin alone. Proof of Stake (PoS) reduces energy waste but creates plutocratic dynamics where the wealthy accumulate more wealth.

The rise of AI compute networks presents an opportunity to align consensus with useful work. AI training and inference require substantial compute resources; if this compute could simultaneously secure a blockchain, we would achieve both useful work and decentralized consensus.

\subsection{Challenges}

Designing \PoAI{} requires solving several challenges:

\begin{enumerate}
    \item \textbf{Verification}: How can validators confirm that claimed AI work was actually performed?

    \item \textbf{Quality Assessment}: Not all AI compute is equally valuable---training a useful model differs from generating noise.

    \item \textbf{Sybil Resistance}: Without proof of identity, adversaries could claim credit for others' work.

    \item \textbf{Efficiency}: Verification must be cheaper than the original computation.

    \item \textbf{Incentive Compatibility}: Rational nodes must be incentivized to perform high-quality work.
\end{enumerate}

\subsection{Contributions}

This paper makes the following contributions:

\begin{enumerate}
    \item \textbf{\PoAI{} Protocol}: A complete consensus mechanism combining \TEE{} attestation, quality scoring, and BLS aggregation (Section~\ref{sec:protocol})

    \item \textbf{Compute Verification}: Novel techniques for verifying AI workloads without re-executing them (Section~\ref{sec:verification})

    \item \textbf{Quality Scoring}: Multi-party evaluation protocol for assessing AI output value (Section~\ref{sec:quality})

    \item \textbf{Security Analysis}: Proofs of Byzantine fault tolerance and incentive compatibility (Section~\ref{sec:security})

    \item \textbf{Evaluation}: Experimental results demonstrating scalability and efficiency (Section~\ref{sec:evaluation})
\end{enumerate}

