\section{Security Analysis}

\label{sec:security}

\subsection{Byzantine Fault Tolerance}

\begin{theorem}[Safety]
If at most $f < n/3$ validators are Byzantine, no two honest validators finalize conflicting blocks.
\end{theorem}

\begin{proof}
Block finalization requires $2f+1$ valid signatures. With $n = 3f+1$ total validators and at most $f$ Byzantine, any two sets of $2f+1$ signers overlap in at least one honest validator. An honest validator signs at most one block per height.
\end{proof}

\begin{theorem}[Liveness]
In partial synchrony, if at most $f < n/3$ validators are Byzantine, honest transactions are eventually included.
\end{theorem}

\begin{proof}
After GST, message delays are bounded. With $2f+1$ honest validators, quorum can always be formed. Leader rotation ensures eventual honest leader.
\end{proof}

\subsection{Incentive Compatibility}

\begin{theorem}[Truthful Quality Reporting]
Under the multi-party evaluation protocol, reporting true quality is a dominant strategy.
\end{theorem}

\begin{proof}
The trimmed mean ignores extreme values. Lying increases variance but cannot shift the result beyond honest range. Deviating validators are identified and slashed.
\end{proof}

\begin{theorem}[No Free Riding]
Workers cannot claim rewards without performing computation.
\end{theorem}

\begin{proof}
\TEE{} attestation binds output to enclave execution, and the software-only execution proof
binds reward to genuine work with no hardware-trust assumption: a worker that claims an
output it did not compute commits an activation-trace whose matmul is caught by the
Freivalds challenge except with probability $\le 2^{-122}$ (Section~\ref{sec:freivalds};
built and tested, Section~\ref{sec:impl-status}), and the on-chain mint gate fails closed
without a valid proof (\texttt{AICoinMiner}, Section~\ref{sec:l3-settlement}). Probabilistic
challenge catches cheaters with high probability and slashing the bond exceeds the expected
gain. The execution proof is what makes this hold for an untrusted, hardware-free prover,
not only inside a trusted enclave.
\end{proof}

\subsection{Sybil Resistance}

Sybil attacks are mitigated through:
\begin{enumerate}
    \item \textbf{Stake Requirement}: Validators must stake tokens
    \item \textbf{Work Requirement}: Workers must perform verified computation
    \item \textbf{Hardware Binding}: \TEE{} attestations bound to physical devices
\end{enumerate}

