\section{Conclusion}

Proof of AI replaces wasteful hash puzzles with a proof that genuine AI computation was
performed---an activation-trace commitment that cannot be produced without doing the
matrix multiplications, checked by Freivalds' algorithm over an exact integer accumulator,
software-only, with no trusted execution environment required to verify. The soundness is
information-theoretic and therefore post-quantum: a fabricated matmul is caught except with
probability $\le 2^{-122}$ at two challenge vectors, and an honest prover is never
false-rejected, because the check runs over a bit-exact integer reduction.

This paper's contribution is not only the primitive but the \emph{built and tested system}
around it. The verifier emits from a real proof-bearing forward (\texttt{ProvableLinear})
and composes across a typed graph (\texttt{GraphTrace}) that binds attention, mixture-of-experts,
and---with one transpose op---a training step, establishing the canon that inference,
embedding, and training are graphs verified by one proof system, not three. The Go verifier
the chain consumes is byte-identical to the prover; the on-chain enforcement runs Freivalds
and the Merkle fold in pure Solidity, slashes a bond only on a verified fraud, and fails
closed so a forged receipt root alone cannot mint. Across three languages the suite is
$38$ Rust, $23$ Go, and $87$ Solidity tests passing, with a cross-language golden vector
that breaks if either side's commitment drifts.

The execution proof is the substrate a \Zoo{} Layer-3 settles. \Beluga{} and its peers
register in an on-chain directory (\texttt{L3Registry}) and roll proven work up through an
ML-DSA validator-quorum root relay (post-quantum, leaderless) to a capped, halving,
fair-mined coin reconciled once network-wide---so a decentralized-science result is a
re-checkable transcript, not an assertion. The economic dial that prices heterogeneous
compute and the verification dial that proves it was performed stay orthogonal: the price
of a unit of work and the proof that the work happened are two knobs, not one.

What remains honestly designed is a named, smaller residual: the RepOps pinned-reduction
floating-point tier (the quantized integer tier is what ships and is what the tested floor
covers), and emission of the architecture families---state-space and diffusion
chaining, and the long tail of the model zoo---that the test suite does not yet exercise. No
third core verification technique is required; the remaining work is breadth of integration
over a settled mathematical core. \PoAI{} is the economic foundation for the \Zoo{} Network:
decentralized AI training and inference whose ledger records re-checkable computations, the
property that distinguishes science from testimony.
