\section{The Software-Only Execution Proof (Design)}
\label{sec:freivalds}

This section constructs the artifact that is \emph{designed} to play, for Proof-of-AI,
the role the hash
plays for Bitcoin: a value that is a \emph{side effect} of the genuine computation,
that is \emph{cheap to verify} relative to producing it, and that \emph{cannot be
produced without performing the work}---but where the work is real AI inference and
the verification requires \emph{only software}: a model file and a CPU, no trusted
execution environment, no GPU attestation. The construction has three parts: a
commitment to the computation trace (so there is something to check against), a
probabilistic verifier for the matrix products (Freivalds' algorithm, made
bit-exact by the integer quantization path), and a determinism contract (so that
``re-derive and compare'' is well-defined across heterogeneous hardware). We then
prove soundness.

\paragraph{Status of this section.} The probabilistic verifier (the Freivalds core), the
determinism contract's two nontrivial clauses, the activation-trace commitment of
Section~\ref{sec:mmr}, and its emission from a real forward/graph pass are all \emph{built
and tested} (Section~\ref{sec:impl-status}): a \texttt{ProvableLinear}/\texttt{GraphTrace}
forward streams the trace, a challenger opens a beacon-selected matmul and Freivalds-checks
it (\texttt{poi\_transcript.rs}), and the on-chain witness re-checks it in pure Solidity
(\texttt{ComputeWitnessLib}). The soundness theorems below are therefore both correct as
mathematics \emph{and} the specification the running code implements; ``cannot be faked''
is operational for the families the suite exercises. We mark the one place (the
integer-accumulator remark) where the proof's arithmetic assumption requires the exact
whole-$K$ $i64$ accumulator (\texttt{exact\_matmul}, the proof path), which is stronger
than the $f8q8$ fast path used for ordinary, non-proof-bearing inference.

\subsection{The activation-trace commitment}
\label{sec:mmr}

A forward (or backward) pass produces a sequence of intermediate tensors
$T_1, T_2, \ldots, T_L$: the output of each matmul, each normalization, each
activation, each attention block. The prover commits to this sequence as it streams,
using a Merkle Mountain Range (MMR) over the per-tensor digests
$d_i = \mathsf{keccak}(\mathsf{enc}(T_i))$, where $\mathsf{enc}$ is the canonical
fixed-point encoding of the tensor. An MMR is appended in $O(\log i)$ work per leaf
and maintains a frontier of $O(\log L)$ peak digests; for the layer counts of
current models the frontier is a few hundred bytes (on the order of $768$ bytes for
$L$ up to $\sim 2^{24}$), so the trace commitment is streamable in constant memory
and the prover never has to retain the full trace. The committed root is the
$\mathsf{activationTraceRoot}$ of Definition~\ref{def:primitive}; the
$\mathsf{modelWeightsRoot}$ is the analogous keccak Merkle root computed once, at
load time, over the quantized weight blocks the prover holds.

We emphasize a fact about the present implementation, in the interest of stating
only what is true: the engine today emits logits without an activation-trace
commitment; the MMR of this subsection is the proposed proof-bearing addition, and
its natural integration point is the synchronous inference precompile already
present in the execution client. The commitment is cheap by design precisely so that
adding it does not change the asymptotic cost of inference.

\subsection{Freivalds verification over an exact integer accumulator}

The dominant operation in every layer is a matrix product $C = A B$ with
$A \in \mathbb{Z}^{m\times k}$, $B \in \mathbb{Z}^{k\times n}$. Re-executing it costs
$O(mnk)$. Freivalds' algorithm~\cite{freivalds1979} verifies a claimed $C$ with one
matrix--vector product on each side: sample $r \in \mathbb{F}_p^{\,n}$ and accept iff
\begin{equation}
\label{eq:freivalds}
A\,(B\,r) \;=\; C\,r \pmod p,
\end{equation}
which costs $O(mn + nk) = O(n^2)$ field operations---asymptotically cheaper than the
$O(n^3)$ product and far cheaper than re-running the layer. The random vector $r$ is
derived from the challenge, $r = \mathsf{PRG}(\mathsf{ch} \,\|\, \ell)$ for layer
index $\ell$, so by Lemma~\ref{lem:anti-replay} it is unpredictable until the beacon
block and cannot be pre-baked.

The decisive enabling fact is that the proof-bearing inference mode runs on an
\emph{exact integer} matmul. In the int8 quantization path, weights and activations
are stored as signed $8$-bit blocks and the per-block dot product accumulates in a
$32$-bit integer with \emph{no intermediate rounding}; the floating-point scale is
applied only \emph{after} the integer accumulation closes. Concretely, for a $32$-lane
block the kernel computes $\sum_j x_j y_j$ with $x_j, y_j \in [-128, 127]$, so each
product is at most $127^2 = 16129$ and a block sum is at most $516128$, comfortably
inside the $\mathbb{Z}_{2^{31}}$ range---hence the accumulation is provably exact. The
same exact-integer dot is realized in portable scalar code and in hardware
integer-dot instructions (NEON \texttt{vdotq\_s32}, AVX2 \texttt{madd}, WASM
\texttt{i32x4\_dot}), all accumulating in $32$-bit integer lanes before any
conversion to float. Because the accumulator is exact integers, the products live in
$\mathbb{Z}$, and we may run the Freivalds check~\eqref{eq:freivalds} over a prime
field $\mathbb{F}_p$ with $p = 2^{61}-1$ (a Mersenne prime, so reduction is a shift
and add) large enough that no honest matmul ever wraps. The consequence is that the
check has \emph{zero false rejections}: an honest prover passes with probability
exactly $1$, because integer equality implies equality mod $p$.

\begin{remark}[Why the proof-bearing mode is quantized, not fp16]
Floating-point matrix multiplication is \emph{not} associative and \emph{not}
reproducible across hardware: the same $A B$ computed with fp16 accumulation on two
GPUs, or on a GPU versus a CPU, can differ in the low bits due to differing reduction
orders, fused-multiply-add availability, and rounding. A Freivalds check over fp16
would therefore suffer false rejections on honest cross-hardware re-derivation, which
is fatal for a permissionless verifier. The integer path removes this \emph{provided the
entire length-$K$ dot product accumulates in i32 with no intermediate float}: under that
condition i8$\times$i8 $\to$ i32 accumulation is associative and bit-identical on every
architecture, so the proof-bearing mode is the quantized mode by necessity, not by
preference. This is the software-only analogue of fixing a single hash function: a
single, exactly reproducible arithmetic.
\end{remark}

\begin{remark}[The required whole-$K$-i32 kernel is a prerequisite, not the shipping
kernel]
\label{rem:wholek}
The exactness above is a property of a \emph{whole-$K$ i32 accumulator}, and that is a
kernel the engine does not yet ship. The quantized GEMM in the released engine (the
\texttt{f8q8} block-quantized path) accumulates each 32-lane block's i8 dot in i32 but
then \emph{converts the block partial to \texttt{f32}, scales by the block's float
$d$, and sums across blocks in floating point}---a per-block FMA chain whose reduction
order is non-associative and not bit-identical across NEON, AVX, and WASM back ends.
Exact Freivalds over that kernel would therefore \emph{false-reject honest
cross-hardware provers}, exactly the failure mode (a) of Section~\ref{sec:fp-fails}. The
$\epsilon = 0$ guarantee of this section consequently holds for a pinned whole-$K$-i32
reduction (block scales pulled out and the integer accumulation closed over the full
contraction before any float, a RepOps-style cross-block pin) that is \emph{designed and
to be built}, not for the current block-float kernel. We state this so the soundness
claim is not read as a property of code that does not yet have it.
\end{remark}

\subsection{The non-matmul remainder}

Softmax, RMS/layer normalization, GELU/SiLU activations, and the attention softmax
are not matrix products and so are not Freivalds-checked. They are handled by
\emph{commit-and-recompute}: each is committed in the trace (Section~\ref{sec:mmr}),
and a verifier challenging a step recomputes that op deterministically from the
committed inputs (which are themselves Freivalds-checked matmul outputs) and compares
to the committed result. These ops are $O(n^2)$ or cheaper and depend
element-wise on their inputs, so recomputing a challenged one is inexpensive. Under
the determinism contract below they are bit-reproducible, so the comparison has, like
Freivalds, zero false rejections.

\subsection{The determinism contract}
\label{sec:determinism}

``Re-derive and compare'' presupposes that two honest evaluations of the same model
on the same input produce \emph{byte-identical} traces. This is a property of the
runtime, not of the mathematics, and we make it an explicit, software-only contract,
digested into $\mathsf{runtimeMeasurement}$ (Section~\ref{sec:binding}) and gated by
governance. The contract has five clauses.

\begin{enumerate}[leftmargin=1.6em]
\item \textbf{Integer matmul.} Proof-bearing inference uses the int8 (or lower) path
with exact i32 accumulation. This is what makes the matmul bit-exact cross-hardware
(the remark above) and what makes Freivalds bit-exact.
\item \textbf{Temperature-0 argmax decoding.} Sampling is greedy: the next token is
$\arg\max$ over the logits, with no stochastic sampling on the proof-bearing path.
\item \textbf{A fixed, specified tie-break.} When two logits are exactly equal, the
contract fixes which token wins---\emph{the lowest token id}---so that argmax is a
total function with no implementation freedom. (We note that the current engine's
\texttt{max\_by} returns the \emph{highest} equal index; the contract therefore
specifies the rule normatively and requires the engine's tie-break to be pinned to
it. Exact ties are vanishingly rare in practice but must be specified for
reproducibility, exactly as a hash must be specified to the bit.)
\item \textbf{Pinned kernels and tokenizer.} The kernel set and tokenizer are fixed
and their identifiers are part of $\mathsf{runtimeMeasurement}$, so two provers run
literally the same code paths.
\item \textbf{A committed seed.} Any residual stochasticity (e.g.\ dropout, disabled
on the inference path) is driven by a committed seed; on the greedy path no RNG is
consulted, so the seed is inert and the decode is a deterministic function of the
prompt and the weights.
\end{enumerate}

The contract is entirely software: every clause is a configuration of the inference
runtime, checkable by reading the model file and the runtime identifiers. No clause
requires special hardware. This is the precise sense in which the execution proof is
``software-only'': both producing it and challenging it need only the model and a
CPU.

\subsection{Soundness}

We can now state and prove the central guarantee: a prover who did not perform the
genuine matmuls cannot answer a beacon-derived Freivalds challenge except with
negligible probability.

\begin{theorem}[Freivalds soundness for one layer]
\label{thm:freivalds-one}
Let $A \in \mathbb{Z}^{m \times k}$, $B \in \mathbb{Z}^{k \times n}$ be the true
operands of a layer (the committed weights and the committed input activations), and
let $\hat C$ be the prover's claimed output, committed in the activation trace. Let
$r \in \mathbb{F}_p^{\,n}$ be sampled uniformly via $\mathsf{PRG}(\mathsf{ch}\,\|\,\ell)$,
modeled as uniform and independent of $\hat C$ by Lemma~\ref{lem:anti-replay}. If
$\hat C \neq AB$ as integer matrices and $p$ exceeds the magnitude of every honest
entry of $AB$, then
\begin{equation}
\Pr_r\bigl[\,\hat C\, r \equiv A(Br) \pmod p\,\bigr] \;\le\; \frac{1}{p}.
\end{equation}
\end{theorem}

\begin{proof}
Work over $\mathbb{F}_p$; reduce all integer entries mod $p$. Let
$D = \hat C - AB \pmod p$. Because $p$ exceeds the magnitude of the honest product's
entries, $\hat C \ne AB$ over $\mathbb{Z}$ implies $D \ne 0$ over $\mathbb{F}_p$
(a genuine discrepancy cannot be an artifact of modular wraparound). The check
$\hat C r = A(Br) = (AB)r$ holds iff $D r = 0$. Since $D \ne 0$, it has some nonzero
row $D_i$, and $D r = 0$ requires $\langle D_i, r\rangle = 0$. Fix any coordinate $j$
with $D_{ij} \ne 0$. Conditioning on all coordinates of $r$ except $r_j$, the
equation $\langle D_i, r\rangle = 0$ determines $r_j$ uniquely as
$r_j = -D_{ij}^{-1}\sum_{j'\ne j} D_{ij'} r_{j'}$ (the inverse exists since
$\mathbb{F}_p$ is a field and $D_{ij}\ne 0$). As $r_j$ is uniform over the $p$ field
elements and independent of the others, it hits that single value with probability
$1/p$. Hence $\Pr[Dr=0] \le \Pr[\langle D_i, r\rangle = 0] = 1/p$. \qed
\end{proof}

Two amplifications make the soundness error cryptographically negligible. First,
using $k_{\mathrm{rep}}$ independent random vectors per layer multiplies the per-layer
error to $p^{-k_{\mathrm{rep}}}$ (equivalently, sampling $r$ from a larger space).
Second, the verifier checks a random subset of layers; a prover who cheats on even a
single layer must survive that layer's check.

\begin{theorem}[Soundness of the execution proof]
\label{thm:soundness}
Consider a forward (or backward) pass of $L$ matmul layers committed in the
activation trace, verified with $p = 2^{61}-1$, $k_{\mathrm{rep}}$ Freivalds vectors
per checked layer, and a uniformly random choice of layers to check. Suppose a
prover commits an $\mathsf{activationTraceRoot}$ and $\mathsf{outputHash}$ that are
\emph{not} the genuine result of evaluating the committed model on the committed
input under the determinism contract. Then:
\begin{enumerate}[leftmargin=1.6em]
\item \textup{(Dense cheat.)} If the prover's committed trace differs from the
genuine trace in the matmul output of \emph{every} checked layer, it passes all
checks with probability at most $p^{-k_{\mathrm{rep}}} \le 2^{-61 k_{\mathrm{rep}}}$;
for $k_{\mathrm{rep}}=2$ this is at most $2^{-122}$.
\item \textup{(Sparse cheat.)} If the prover cheats on only a $\rho$-fraction of
layers and the verifier checks $s$ layers uniformly at random, the cheat escapes all
checks only if no cheated layer is sampled, which (for cheats confined to few layers)
has probability $(1-\rho)^s$; conditioned on a cheated layer being sampled it is
caught except with probability $p^{-k_{\mathrm{rep}}}$.
\item \textup{(Determinism binding.)} A committed output hash inconsistent with the
committed (and checked) final-layer logits is detected with certainty, because the
argmax/tie-break of Section~\ref{sec:determinism} is recomputed from the
Freivalds-verified logits.
\end{enumerate}
\end{theorem}

\begin{proof}
Each part composes Theorem~\ref{thm:freivalds-one} with independence. For (1), the
$k_{\mathrm{rep}}$ vectors on a checked discrepant layer are independent, so by
Theorem~\ref{thm:freivalds-one} the layer passes with probability
$\le (1/p)^{k_{\mathrm{rep}}}$; a dense cheat is discrepant on every checked layer, so
the prover must win at least one such layer's check, bounded by
$p^{-k_{\mathrm{rep}}}$. For (2), the verifier's $s$ sampled layers miss the cheated
fraction with probability $(1-\rho)^s$ (sampling without replacement only lowers
this); if any cheated layer is sampled, part~(1)'s bound applies to it. For (3), the
final-layer pre-activation logits are a matmul output and are Freivalds-checked; the
verifier recomputes the argmax under the pinned tie-break and compares to the
committed token. Since both are deterministic functions of the verified logits, any
inconsistency is detected with probability $1$. \qed
\end{proof}

\subsection{From ``probably'' to ``certainly'': interactive bisection}

The sparse-cheat term $(1-\rho)^s$ in Theorem~\ref{thm:soundness} shrinks with more
sampled layers but does not reach zero by spot-checking alone: a cheat confined to a
single layer is missed with probability $(1-1/L)^s$. To convert this into
\emph{certainty} at logarithmic cost we add an interactive bisection in the style of
optimistic rollups~\cite{arbitrum,gensynverde}. The prover and a challenger disagree
on the final output; they recursively bisect the trace---each round the prover
commits the midpoint activation, and the parties recurse into the half where they
disagree---until they isolate a \emph{single} matmul whose inputs both agree on but
whose output they dispute. That one matmul is then either Freivalds-checked or
re-executed on-chain (it is a single layer, hence cheap), and the dispute is resolved
deterministically: exactly one party is shown wrong. The number of rounds is
$O(\log L)$, so the on-chain footprint is logarithmic in the trace length, and the
guarantee is upgraded from ``caught with overwhelming probability'' to ``caught with
certainty, after $O(\log L)$ interaction.''

\begin{corollary}[Software-only verifiability, by construction]
\label{cor:software-only}
In the construction as specified (with the trace emitted and the whole-$K$-i32 kernel of
Remark~\ref{rem:wholek} in place), producing the execution proof requires the model and
the prover's compute; checking
it---whether by spot-check (Theorem~\ref{thm:soundness}) or by bisection to
certainty---requires only the model file, a CPU to compute $\mathsf{PRG}$, the
$O(n^2)$ Freivalds matrix--vector products, and keccak for the trace. No trusted
execution environment, GPU attestation, or special hardware is needed by either
party. This is a property of the design; the two prerequisites it names (trace emission,
pinned whole-$K$ reduction) are the remaining build (Section~\ref{sec:impl-status}).
\end{corollary}

Corollary~\ref{cor:software-only} is the property that makes the scheme egalitarian: a
CPU-only validator challenges a GPU prover, and any user both generates and verifies a
proof against the same network-shared difficulty (the soundness parameters $p$,
$k_{\mathrm{rep}}$, $s$, and the bond). The emission and the off-chain challenger of
Section~\ref{sec:tiers} are built and tested (\texttt{poi\_transcript.rs}, the Go
\texttt{crypto/poi} watcher, and the on-chain \texttt{ComputeWitnessLib}), so this is a
live property for the proof-bearing path, not only an intended one---the contrast with
Bitcoin's ASIC-mediated work, drawn out in Section~\ref{sec:difficulty}.
