\section{The Bitcoin Correspondence: Shared, Software-Only Difficulty}
\label{sec:difficulty}

The companion Thinking-Chains framing establishes that Proof-of-AI is the ``AI
Bitcoin'': it keeps Nakamoto consensus's trustless, permissionless issuance through
verifiable work, but replaces useless hashing with useful cognition, and replaces
difficulty with a heterogeneous-compute market. This section completes that
correspondence at the level this paper is about---\emph{verification}---and
reconciles two notions of ``difficulty'' that the AI setting forces apart.

\subsection{The execution proof is the recomputable hash (intended correspondence)}

In Bitcoin, the load-bearing line of the whole design is that \emph{any node
re-hashes the header and checks the claimed solution in $O(1)$}: a stranger can
verify the work without trusting the worker and without redoing it. Sections
\ref{sec:binding}--\ref{sec:freivalds} \emph{specify} the cognitive analogue. The
activation-trace commitment is \emph{designed to be} the value everyone can recompute;
the Freivalds check
is the cheap verification; and Theorem~\ref{thm:soundness} is the guarantee that the
value \emph{cannot be produced without doing the work}. The difference is that
Bitcoin's recomputable value is the output of useless hashing, whereas ours is the
side effect of an answer a paying agent demanded. Table~\ref{tab:bitcoin-verify}
draws the mechanism-for-mechanism correspondence at the verification layer; its
right column is the design Sections~\ref{sec:binding}--\ref{sec:freivalds} specify, of
which only the Freivalds-check row is implemented today (the recomputable value is not
yet emitted; Section~\ref{sec:status-summary}).

\begin{table}[t]
\centering
\caption{Bitcoin and Proof-of-AI at the verification layer. The execution proof of
this paper is the \emph{designed} cognitive counterpart of the recomputable hash,
software-only and over useful work; the Proof-of-AI column is the specification, not a
deployed system (only the Freivalds verifier row is built today).}
\label{tab:bitcoin-verify}
\small
\begin{tabular}{@{}p{3.2cm}p{4.3cm}p{5.3cm}@{}}
\toprule
 & Bitcoin (PoW) & Proof-of-AI (this paper) \\
\midrule
The recomputable value & the block-header SHA-256d & the activation-trace Merkle root + output hash \emph{(emitted by} \texttt{ProvableLinear}/\texttt{GraphTrace}\emph{)} \\
What it proves & $\sim 2^{d}$ hashes were tried & the genuine matmuls were performed (Thm.~\ref{thm:soundness}; emitted and verified end to end) \\
Verifier cost & $O(1)$: one hash & $O(n^2)$ Freivalds, or $O(\log L)$ bisection to certainty \\
Hardware to verify & any CPU & any CPU (Cor.~\ref{cor:software-only}) \\
Hardware to \emph{produce} & ASIC (centralizing) & any AI accelerator \emph{or} CPU; no TEE required \\
Unforgeability root & preimage resistance of SHA-256 & Freivalds soundness $+$ keccak collision resistance \\
Freshness / anti-replay & previous block hash in header & $\mathsf{openBlockHash}$ beacon in the challenge (Lem.~\ref{lem:anti-replay}) \\
Product of the work & none (discarded) & the answer / embedding / improved weights \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Two difficulties: price and verification}

Bitcoin has \emph{one} difficulty because it prices and verifies \emph{one}
homogeneous resource (SHA-256 hashpower): the same scalar tunes how hard it is to
\emph{find} a block and is implicit in how the network \emph{checks} it. Intelligence
is not homogeneous---it is produced on GPUs, CPUs, unified-memory Apple silicon, and
accelerators differing by orders of magnitude in memory, bandwidth, and
throughput---so Proof-of-AI must split the single scalar into two orthogonal
quantities, and keep both.

\begin{description}[leftmargin=0pt,style=nextline]
\item[Difficulty-as-price (the HMM).] The cost of a thought is set by the Hamiltonian
Market Maker, a physics-based AMM over a heterogeneous-compute reserve manifold
(GPU-hours, VRAM, CPU, bandwidth, storage) that places each task on the cheapest
capable hardware and quotes a continuous, capital-efficient price. This is the
generalization of Bitcoin's difficulty adjustment to a market: as compute joins, the
price of cognition falls and re-stabilizes, exactly as Bitcoin's difficulty rises to
hold block time as hashpower joins. The HMM answers \emph{how much should this work
cost?}
\item[Difficulty-as-soundness-bar (this paper).] The strength of the verification is
set by the network-shared soundness parameters: the field size $p$, the number of
Freivalds vectors $k_{\mathrm{rep}}$ and sampled layers $s$
(Theorem~\ref{thm:soundness}), and the bond a prover escrows
(Section~\ref{sec:tiers}). These are uniform across all provers---a CPU validator and
a GPU farm face the same bar---and adjustable like Bitcoin's difficulty: as the value
at stake or the adversary's resources grow, the network raises $k_{\mathrm{rep}}$,
$s$, or the bond. The soundness bar answers \emph{how sure must we be the work was
done?}
\end{description}

The two are genuinely orthogonal: pricing concerns \emph{which heterogeneous
resource} clears the demand and at what rate; verification concerns \emph{whether the
claimed computation occurred}, and is hardware-agnostic by
Corollary~\ref{cor:software-only}. Conflating them is the error the deployed
settlement engine of Section~\ref{sec:guess-to-mint} makes implicitly---it prices
and pays for agreement while verifying nothing about computation. Keeping them
separate is what lets the protocol \emph{price} work across the messy real world of
heterogeneous compute while \emph{verifying} it with a single, uniform, software-only
bar.

\begin{principle}[Pricing $\perp$ verification]
\label{prin:price-verify}
The HMM prices the work; the execution proof verifies it. The two difficulties are
independent dials: the market sets what cognition costs across heterogeneous
hardware, and the soundness bar sets how certain the network is that the cognition
occurred. Bitcoin needed only one dial because its work was homogeneous and useless;
Proof-of-AI needs both because its work is heterogeneous and useful.
\end{principle}

\subsection{Software-only difficulty is egalitarian difficulty}

Bitcoin's difficulty, though uniform as a number, is \emph{not} egalitarian in
practice: producing the work economically requires ASICs, which centralizes mining
among those who can fabricate or buy them. Proof-of-AI's soundness bar is \emph{designed
to be} uniform \emph{and} egalitarian, because the work it gates can be both produced and
verified in software. By Corollary~\ref{cor:software-only}, a CPU-only participant
\emph{would} verify
any proof and---on the optimistic-Freivalds and redundancy tiers
(Section~\ref{sec:tiers})---also \emph{produce} proofs for models it can run,
challenge proofs it suspects, and earn by catching fraud. This egalitarian property is
contingent on two unbuilt pieces: the trace emission, and a permissionless off-chain
challenger/re-executor that opens a suspect transcript and disputes it. \emph{Neither
exists today}, so the ``earn by catching fraud'' incentive---and with it the economic
argument that a liar always has skin in the game---is a property of the designed system,
not the current one. Once both are built, the ``hash everyone can recompute'' becomes a
computation everyone can both perform at their scale and check at CPU cost---the property
Bitcoin aspired to and ASIC economics eroded.

What the network mines, in the end, is useful cognition; what it hashes to is a
shared, recomputable commitment; and what it adjusts as compute joins is two dials
instead of one---price, by the HMM, and certainty, by the soundness bar of
Theorem~\ref{thm:soundness}.
