\section{The Tier Matrix: One Gate, Migrating Backends}
\label{sec:tiers}

The execution proof of Section~\ref{sec:freivalds} is one of several proof systems a
deployment can require, trading off trust assumptions, cost, and hardware. This
section specifies the menu and the single gate that selects from it. The discipline
is the one used for post-quantum migration elsewhere in the stack: the policy ``what
proof does a task at this tier require?'' lives in \emph{exactly one place}, every
consumer reads it the same way, and the proof backends are interchangeable behind a
common binding so that the menu can change without touching the binding or the
callers.

\subsection{The proof backends}

A proof system is identified by a $\mathsf{proofType}$ and registered as a backend
that attests to a $\mathsf{reportData}$ (Section~\ref{sec:binding}). The deployed
registry recognizes four:
\begin{equation}
\mathsf{proofType} \in \{\,
  \underbrace{1}_{\text{CC-TEE}},\
  \underbrace{2}_{\text{zkML}},\
  \underbrace{3}_{\text{optimistic-Freivalds}},\
  \underbrace{4}_{M\text{-of-}N\ \text{TEE}}
\,\}.
\end{equation}
Each backend exposes the same interface---``does this witness attest to this
$\mathsf{reportData}$?''---and the verifier dispatches on $\mathsf{proofType}$,
failing \emph{closed} on any unregistered type. The optimistic-Freivalds backend
(type $3$) is the software-only path of this paper: a prover escrows a bond and
commits $(\mathsf{reportData}, \mathsf{activationTraceRoot}, \mathsf{modelWeightsRoot})$;
within a challenge window any party may dispute by supplying a Freivalds response (or,
on dispute, by driving the bisection of Section~\ref{sec:freivalds}); an
unchallenged commitment finalizes and the bond is reclaimable, while a proven fraud
slashes the bond to the challenger. The heavy check runs off-chain in the inference engine
(a watcher re-executes the committed trace and opens any discrepant matmul); the chain runs
the economic state machine the check's verdict drives, and---when the dispute is
submitted---re-runs the Freivalds check itself on-chain so the verdict is not a watcher's
say-so. Both halves are built and tested (Section~\ref{sec:impl-status}): the off-chain
re-executor and challenger (\texttt{poi\_transcript.rs} in Rust, \texttt{crypto/poi} in Go)
open a transcript and produce the dispute, and the on-chain state machine
(\texttt{OptimisticEvidence.sol}, $18/18$) records the bonded commitment, verifies the
exhibited fraud via \texttt{ComputeWitnessLib.provesFraud}, and slashes only on a verified
discrepancy. The optimistic backend's economic soundness---a liar is strictly unprofitable
once watched, an honest prover is never slashable---is therefore a live deterrent: a
fabricated matmul is provable except with probability $\le 2^{-122}$, and the bond pays the
challenger that proves it.

\subsection{The tiers}

Tiers map a task's value and trust requirement to a required proof strength.

\begin{table}[t]
\centering
\caption{Tier $\to$ proof matrix. One profile gate
(Section~\ref{sec:profile-gate}) maps a tier to its required $\mathsf{proofType}$;
where a tier is opted in, no valid proof of that type means no mint and no settlement.
The gate is built but defaults to permissive ($\mathsf{requiredProofType}=0$), and the
software-only backend (T2/T3) awaits the emission and challenger of
Section~\ref{sec:impl-status}; the matrix is thus the configured policy surface, live
enforcement following the build.}
\label{tab:tiers}
\small
\begin{tabular}{@{}p{0.7cm}p{3.3cm}p{4.1cm}p{4.0cm}@{}}
\toprule
Tier & Proof system & Trust assumption & Use \\
\midrule
T0 & zkML (type 2) & none beyond the SNARK's & tiny models; trustless, on-chain succinct verification \\
T1 & $M$-of-$N$ heterogeneous CC-TEE (type 4) & no single TEE vendor; $M$ of $N$ enclaves agree & high-value tasks; defense in depth across NVIDIA CC $+$ Intel TDX $+$ AMD SEV-SNP \\
T2 & single-vendor CC-TEE (type 1) \emph{or} optimistic-Freivalds (type 3) & one TEE vendor, \emph{or} one honest challenger $+$ a bond & the workhorse tier; software-only Freivalds is the intended egalitarian default (pending emission $+$ challenger) \\
T3 & redundancy $+$ plurality $+$ slash & honest plurality among bonded operators & liveness-first; CPU-only providers; the weakest proof, fastest path \\
\bottomrule
\end{tabular}
\end{table}

The matrix is read top-down by trust and bottom-up by accessibility. T0 (zkML)
assumes nothing beyond the soundness of the proof system but is limited to small
models by proving cost. T1 ($M$-of-$N$ heterogeneous confidential compute) tolerates
the compromise of any single TEE vendor by requiring agreement across enclaves from
\emph{different} hardware roots---NVIDIA confidential computing, Intel
TDX~\cite{inteltdx}, AMD SEV-SNP~\cite{amdsevsnp}---so that a break of one vendor's
attestation does not forge a proof. T2 is the workhorse: a single-vendor TEE for
those who have the hardware, \emph{or} the optimistic-Freivalds path of this paper for
those who do not, which assumes only that at least one honest party will challenge a
fraud within the window and that the prover's bond exceeds the gain from cheating. T3
falls back to the redundancy-and-plurality settlement of
Section~\ref{sec:guess-to-mint}, retained only where liveness dominates and the value
at stake is low enough that honest-plurality is acceptable---and, crucially,
\emph{even T3 can now require a sampled execution proof} from the plurality operators,
closing Proposition~\ref{prop:gap} at that tier too.

\subsection{The profile gate: one place}
\label{sec:profile-gate}

The mapping from tier to required proof is a single function in a single contract,
the on-chain analogue of the \texttt{RefuseUnderStrictPQ} gate used for post-quantum
policy. It exposes one read,
\begin{equation}
\mathsf{requiredProofType}(\mathsf{tier}) \;\to\; \mathsf{proofType},
\end{equation}
returning $0$ (permissive) for any tier governance has not explicitly opted in, and
otherwise the mandatory $\mathsf{proofType}$. The settlement path consults this single
oracle and enforces the invariant directly:

\begin{quote}\itshape
No valid compute proof of the required type $\Rightarrow$ no mint, no settlement.
\end{quote}

This is the constructive answer to Section~\ref{sec:guess-to-mint}: the verifier's
three-step gate---(1) the proof's $\mathsf{reportData}$ equals the task binding,
(2) the runtime is governance-accepted, (3) the registered backend attests---runs
\emph{before} any reward is paid, and fails closed at every step \emph{on any tier that
has been opted in}. Where the gate is engaged, reward is bound to
proof, not to signature and not to plurality; where it is not (the default), settlement
remains the authorship-plus-agreement rule of Section~\ref{sec:guess-to-mint}, and---until
the execution backend's emission and challenger are built---the only proof type a gate can
actually enforce is an attestation envelope, not the activation-trace execution proof.
Because the policy lives in one place,
there is exactly one way to ask ``what does this tier require?'', and changing a
tier's requirement is a single governed write rather than a scattered edit.

\subsection{Migration without rebinding}

The tiers are not static. The decisive structural property of this design is that
\emph{the binding does not depend on the proof system}: $\mathsf{reportData}$
(Section~\ref{sec:binding}) commits the task, model, input, output, and runtime, and
says nothing about how that commitment is believed. Consequently, as zkML proving cost
falls and larger models become trustlessly provable, tasks migrate from T2/T3 up to
T0 by a governance change to $\mathsf{requiredProofType}$ alone---\emph{zero change to
the binding, zero change to the receipts, zero change to the callers}. The same holds
in reverse: a deployment that gains confidential-compute hardware can move tasks from
software-only T2 to TEE-backed T1 without touching the proof artifacts that earlier
tasks emitted. The proof backends are orthogonal slots behind a stable binding; the
menu evolves, the interface does not. This is the property \emph{designed} to let a
permissionless network start egalitarian (software-only Freivalds for everyone, once
that backend's emission and challenger are built) and strengthen
task-by-task as trustless proving matures, without ever stranding the proofs it has
already minted against.
