\section{Ownership: Data, Models, and Weights as User-Owned Assets}
\label{sec:ownership}

\begin{quote}\small\itshape
Status note. The cryptographic core of this section is now built. The confidentiality
envelope---a single-shot prompt seal under \textbf{X-Wing} (the hybrid X25519 $+$
ML-KEM-768 KEM, RFC~9180 HPKE base mode), so the prompt is ciphertext to everyone but the
chosen operator---is implemented and tested (\texttt{luxfi/crypto/promptseal}, $6$ tests
passing), and the in-enclave custody binding (a TEE quote that binds the operator's
confidentiality key to an attested image) is implemented and tested
(\texttt{luxfi/crypto/attestation}, $9$ tests passing); both are detailed and proven in the
emission companion~\cite{poiemission}. The proof emission these compose with is also built
(Section~\ref{sec:impl-status}). What remains forward-looking is the \emph{product} layer
built on top---data/model/weight marketplaces, the full user-owned-asset NFT surface, and
the plaintext-blind confidential-compute tier wired into every operator path; modes (2)
and (3) below describe that intended envelope. Mode (1)---the user runs the model on their
own hardware---is non-custodial unconditionally. We keep the design framing for the product
surface and state plainly that the sealing and attestation primitives underneath it are
deployed.
\end{quote}

The execution proofs are intended to make a stronger claim possible than ``the network
computed correctly.'' They are designed to make \emph{sovereignty} auditable: a user
\emph{would} own their data, own
the inference performed on it, and own the model weights they improve, and \emph{would}
choose
\emph{per request} how much trust to place in whoever runs the compute---without the
choice ever weakening the guarantee that the work was genuine. The aim is that trust
become a UX dial, not a security compromise. With the confidentiality envelope unbuilt,
that decoupling holds today only in mode (1).

\subsection{The trust spectrum}

For any request, a user selects one of three execution modes, ordered by decreasing
sovereignty and increasing convenience. Every mode emits the same execution proofs of
Sections~\ref{sec:four-proofs}--\ref{sec:freivalds}, so every mode is auditable; the
modes differ only in \emph{who holds the data in the clear} and \emph{who the user
must trust for liveness}.

\begin{description}[leftmargin=0pt,style=nextline]
\item[(1) User-operated node --- maximal sovereignty.] The user runs the inference on
their own hardware. The data never leaves the device; the model and weights are local;
the proof is emitted for the user's own records or for third parties who later wish to
verify a result the user publishes. Trust in any operator is zero because there is no
operator. This is the floor of the spectrum and the reference against which the others
are measured.
\item[(2) DAO trustless nodes --- verify, don't trust \emph{(unbuilt)}.] The user
\emph{would} send work to
permissionless DAO operators under confidential compute (T1/T2) so the operator
\emph{would be} \emph{plaintext-blind}: processing ciphertext inside an enclave and never
seeing the data in the clear, while the execution proof (\PoI{}/\PoT{}) plus the slashing
bond would make a wrong or fabricated result detectable and punishable. \emph{As
deployed today, none of this confidentiality holds}: there is no enclave ciphertext path
and the operator receives plaintext. In the intended mode the user trusts no
particular operator's honesty---only the proof system and, for the TEE tiers, the
hardware root. This is the designed ``verify, don't trust'' mode: sovereignty over
correctness and confidentiality without running the hardware oneself.
\item[(3) Selected trusted operators --- convenience.] The user picks specific
operators by reputation (via the identity-and-reputation layer), accepting a
centralized-feeling, faster, cheaper path. Even here the proofs remain emittable, so
the user retains an audit trail: a trusted operator that misbehaves can be caught
after the fact by anyone who checks the proof, and loses reputation and future
routing. Trust is placed deliberately and remains accountable.
\end{description}

The three modes are points on a continuum, not separate products, because they share
the binding and the proof primitive. A user can move a workload from mode (3) to
mode (2) to mode (1) as their trust requirements rise, with no change to the artifacts
the workload produces.

\subsection{Data confidentiality at rest: primitives built, operator integration pending}

The cryptographic primitives are built: in modes (2) and (3) the prompt is sealed under
X-Wing (X25519 $+$ ML-KEM-768) HPKE to the chosen operator's published key
(\texttt{crypto/promptseal}, tested), committed on-chain only by hash, and---in the
confidential-compute tier---decryptable only inside an enclave whose attested image is
bound to that key (\texttt{crypto/attestation}, tested; \cite{poiemission}). What is
\emph{not} yet built is the \emph{integration} of these primitives into the production
inference operator path: the seal/open is not yet wired into every operator's request
handler, nor the per-operator key-management lifecycle, so in the currently deployed
operator path the prompt may be held in the clear and the ``never plaintext at rest''
property is guaranteed only in mode (1) and wherever the sealing path is enabled. The
soundness binding is independent of this and is real: the execution proof binds to
$\mathsf{inputHash}$ (Section~\ref{sec:binding}), so a proof certifies that the
\emph{committed} input was processed while revealing nothing about the input itself. That
binding is what makes \PoC{}'s privacy clause (Definition~\ref{def:poc}) more than a
promise: the contributor proves the gradient was computed over data consistent with the
committed policy (the training-step graph is Freivalds-checked,
Section~\ref{sec:impl-status}) and shares only the resulting delta, so the corpus stays
private while the contribution stays verifiable. The confidentiality envelope (X-Wing
prompt seal $+$ attestation) and the proof emission it relies on are both built and tested
(\texttt{crypto/promptseal}, \texttt{crypto/attestation}, the engine \texttt{poi\_*}
modules; Section~\ref{sec:impl-status}); the corpus-scale data-policy product around them
is the forward-looking part.

\subsection{Weights are a portable, user-owned asset}

A model improvement---the $\Delta W$ of \PoT{}/\PoC{}---is serialized in the gym's
canonical backend-independent BF16 format and committed by its root
(Section~\ref{sec:four-proofs}). Because the serialization is canonical and
backend-independent, a delta trained on one accelerator is byte-portable to any other
\emph{(this part is real---it is the gym's shipping format)}, and its provenance (which
base model, which data policy, what improvement margin) \emph{would be}
attested by the \PoC{} it was minted with \emph{(this part depends on the unbuilt proof
emission)}. The end state is that a user's weights are an on-chain
asset they own and can move, sell, or compose---with royalties flowing to the
contributors whose proven improvements it incorporates---rather than an artifact
locked inside a single provider's stack. The proof framework is what gives the asset
its value: a delta is worth more when it carries a verifiable proof that it was
genuinely trained and genuinely improves the model, than when it is merely asserted to.

\subsection{Trust as a dial, security as a floor}

The synthesis the design aims at is that the execution proofs decouple \emph{trust} from
\emph{security}.
A user would dial trust up the spectrum---from running everything themselves, to trusting
a proof system, to trusting a named operator---according to their convenience and
threat model, and the proof framework would hold the security floor constant underneath:
in every mode, a result that claims to be the model's output on the user's input
either is, or is caught. This is the intended difference between the deployed settlement
engine, where ``trusting the operators'' means accepting unverified guesses
(Section~\ref{sec:guess-to-mint}), and a proof-backed network, where trusting an
operator would be a choice about \emph{speed and cost}, never about \emph{correctness}.
\emph{This decoupling is not yet in force}: with neither the proof emission nor the
confidentiality envelope built, mode (1) (self-hosting) is the only mode that today
delivers either sovereignty or non-custody. The end state is that the
user owns the data, the inference, and the weights, and the protocol's job is
only to make all three auditable, at whatever point on the trust spectrum the user
selects; reaching it is the build of Section~\ref{sec:impl-status}.
