\section{Full Modern-LLM Coverage: Tiering, Not Floating-Point Freivalds}
\label{sec:fp-coverage}

The execution proof of Section~\ref{sec:freivalds} is exact because it runs over the
int8 path's integer accumulator. A natural objection is that ``real'' frontier models
run in floating point (fp16, bf16, fp8), so an exact-integer verifier covers only the
quantized deployment and not the full precision a model was trained in. This section
answers the objection rigorously, and the answer is the crux of the design: we do
\emph{not} force Freivalds onto floating point, because naive floating-point Freivalds
is simultaneously \emph{unsound} and \emph{unusable}. Instead, coverage of the full
model space is \emph{designed to be} achieved by \emph{tiering} the verification mode to
the arithmetic, so
that every model has a proof-bearing mode while no mode pays a soundness or liveness
penalty it cannot survive. The quantized, exact, software-only mode is then not a
limitation but a \emph{feature}.

\paragraph{Two status caveats, stated before the argument.} (i) The exact-integer mode
this section relies on requires a whole-$K$ integer accumulator. This is \emph{built and is
the proof path}: \texttt{poi.rs::exact\_matmul} is a whole-$K$ $i64$ reduction with no
cross-block float rescale, and \texttt{ProvableLinear} runs its matmuls there. The
distinction is that the engine's \emph{fast} kernel (\texttt{f8q8}) accumulates across
blocks in \texttt{f32}---non-associative and cross-hardware non-reproducible
(Remark~\ref{rem:wholek})---so it is used for ordinary inference, while a proof-bearing
inference pays the exact accumulator's throughput cost to get the ``$\epsilon=0$, zero
false-reject'' property. Running exact Freivalds over the \texttt{f8q8} fast path would
false-reject honest provers; running it over \texttt{exact\_matmul} does not, which is why
the proof path is the integer path. (ii) The quantized models are the arithmetic the zoo
\emph{deploys in}, and the families the test suite exercises (dense/SwiGLU, attention with
integer softmax, MoE, training) now \emph{emit a proof transcript}---so provable-today
coverage is nonempty, not $0\%$. What remains \emph{designed} is the breadth: the
floating-point (RepOps) tier and the architecture families not yet wired to emit
(Section~\ref{sec:status-summary}).

\subsection{Why naive floating-point Freivalds fails twice}
\label{sec:fp-fails}

Floating-point matrix multiplication breaks the Freivalds check in two independent
ways, one fatal to \emph{liveness} and one fatal to \emph{soundness}. Both stem from
the same root cause: an fp GEMM is not a function of its operands alone.

\paragraph{(a) Non-determinism $\Rightarrow$ false-reject (liveness break).}
Floating-point addition is not associative: $(a+b)+c \ne a+(b+c)$ in general under
IEEE-754 rounding. A GEMM accumulates $k$ products into each output entry, and the
\emph{order} of that accumulation is chosen by the kernel---tiling, the number of
threads in a reduction, whether a fused multiply--add (FMA) is emitted, the
warp/wavefront width, and the vendor's library version all permute it. Two honest
evaluations of the same $A,B$ in fp16 on two different GPUs, or on a GPU versus a CPU,
therefore yield outputs $C$ and $C'$ with $C \ne C'$ in the low mantissa bits. A
Freivalds check demands the equality $C r = A(Br)$ over the \emph{produced} $C$; with
$C \ne C'$, an honest re-deriving verifier on different hardware computes $A(B'r) \ne
C r$ and \emph{rejects honest work}. In a permissionless network where a CPU validator
must be able to challenge a GPU prover (Corollary~\ref{cor:software-only}), this
false-reject is not a tail event---it is the \emph{common} case, because heterogeneous
hardware is the premise. A verification rule that rejects honest provers is a liveness
failure regardless of its soundness.

\paragraph{(b) Tolerance bands $\Rightarrow$ a perturbation corridor (soundness break).}
The naive ``fix'' for (a) is to accept when $\lVert C r - A(Br)\rVert_\infty < \epsilon$
for some tolerance $\epsilon$ absorbing cross-hardware rounding. This trades the
liveness failure for a soundness failure. The acceptance region is no longer the point
$\{C = AB\}$ but a \emph{band} of radius $\epsilon$ around it, and a cheating prover may
roam anywhere inside that band. Concretely, an adversary who wishes to claim a
fabricated output $\hat C$ chooses any $\hat C$ with $\lVert (\hat C - AB) r \rVert_\infty
< \epsilon$; the Freivalds projection $r$ collapses the $m\times n$ discrepancy matrix
$D = \hat C - AB$ to the $m$-vector $Dr$, so the adversary needs only $Dr$ small, not
$D$ small---an under-determined constraint with a large solution space. The corridor is
\emph{exploitable}: it lets a prover skip or approximate work (e.g.\ run a smaller
model, a lower-rank approximation, or a truncated accumulation) and land inside
$\epsilon$, harvesting the fee while the band hides the cheat. Worse, the corridor must
be \emph{widened} to cover legitimate cross-hardware drift, and every widening enlarges
the cheating manifold. There is no setting of $\epsilon$ that admits all honest
cross-hardware runs (requires $\epsilon$ large) while excluding all profitable cheats
(requires $\epsilon$ small): the two requirements pull in opposite directions, exactly
as in the analogous attack on tolerance-based optimistic ML verifiers. We formalize
this dilemma in the companion proofs document (the floating-point soundness/liveness
no-go).

\begin{remark}[The integer path is not a downgrade---it removes both failures at once]
The integer accumulator (Section~\ref{sec:freivalds}) makes the matmul a function of
its operands alone: i8$\times$i8$\to$i32 accumulation is associative and bit-identical
on every architecture. This simultaneously kills (a)---two honest provers compute the
\emph{same} $C$, so $\epsilon = 0$ and there are zero false rejections---and (b)---with
$\epsilon = 0$ the acceptance region collapses back to the exact point $\{C = AB\}$ and
the corridor vanishes, restoring the clean $1/p$ soundness of
Theorem~\ref{thm:freivalds-one}. This is why the proof-bearing mode is quantized by
\emph{necessity}: exact arithmetic is the only arithmetic for which a software-only,
cross-hardware, tolerance-free equality check exists.
\end{remark}

\subsection{The key fact: floating-point values are integers}
\label{sec:fp-are-integers}

The tiering does not abandon floating-point models; it gives them an exact mode too,
via a fact that is easy to miss. An IEEE-754~\cite{ieee754} value is \emph{already an integer
encoding}: a sign bit, an exponent field, and a mantissa field, all integers. For
fp16/bf16/fp8 the significand is a small integer (an $11$-, $8$-, or $4$-bit mantissa
plus the implicit leading one), and a product of two such values, \emph{before
rounding}, is an exact integer times a power of two. The IEEE rounding that follows is
a \emph{deterministic, total function} of the exact product and the rounding mode---it
is not a source of nondeterminism. We make this precise.

\begin{proposition}[Floating-point GEMM is exact integer-mantissa arithmetic up to a
fixed reduction order]
\label{prop:fp-integer}
Fix an IEEE-754 format $F$ (e.g.\ bf16) with round-to-nearest-even, and fix a
\emph{reduction tree} $\tau$: a binary tree on the $k$ index positions specifying the
exact order in which the $k$ products of a dot product are summed and rounded. Then for
any operands $A,B$ in $F$, the GEMM output $\mathrm{GEMM}_F^\tau(A,B)$ is a
deterministic function of $(A,B)$ alone, computable by integer arithmetic on the
mantissa/exponent fields with an IEEE rounding step at each tree node, and identical on
every machine that honors $F$ and $\tau$. The \emph{only} freedom in a floating-point
GEMM is the choice of $\tau$; once $\tau$ is pinned, the result is bit-exact and
hardware-independent.
\end{proposition}

\begin{proof}[Argument]
Each product $a_i b_j$ of two $F$-values has a significand that is the integer product
of the two integer significands and an exponent that is the integer sum of the two
exponents; this is exact in $\mathbb{Z}$ (no rounding) because integer multiplication
is exact. A tree node combines two already-computed partial sums $u,v$ (each an
$F$-value, hence an integer significand $\times 2^{e}$): align exponents (an integer
shift), add significands (an integer add), then apply round-to-nearest-even (a total
function of the exact integer sum and the format's precision). Every operation is
integer arithmetic plus a deterministic rounding decision, so the node output is a
deterministic function of $(u,v)$. By induction over $\tau$, $\mathrm{GEMM}_F^\tau(A,B)$
is a deterministic function of $(A,B,\tau)$. The associativity counterexample of
Section~\ref{sec:fp-fails} is precisely a disagreement in $\tau$ between two kernels;
fixing $\tau$ removes it. \qed
\end{proof}

Proposition~\ref{prop:fp-integer} isolates the entire problem to one degree of freedom:
\emph{accumulation order}. It says nothing about precision loss---an fp16 GEMM still
rounds, and rounding still discards low bits---but it says the rounding is
\emph{reproducible}: every honest prover that uses the pinned tree $\tau$ produces the
same rounded bits. This is the hook on which an exact floating-point Freivalds hangs.

\subsection{RepOps: pinning the reduction tree makes floating-point bit-exact}
\label{sec:repops}

The deterministic-floating-point (``RepOps'', for \emph{reproducible operators})
verification mode for the fp16/bf16/fp8 tier is then: \emph{pin the reduction tree
$\tau$ as part of $\mathsf{runtimeMeasurement}$} (Section~\ref{sec:binding}), so that
the floating-point GEMM becomes bit-exact across hardware by
Proposition~\ref{prop:fp-integer}, and run Freivalds over the resulting exact values.
Because $\tau$ is fixed, $\epsilon = 0$: there is no cross-hardware drift to absorb
(failure (a) is gone) and no tolerance band to exploit (failure (b) is gone). The
Freivalds soundness of Theorem~\ref{thm:freivalds-one} applies verbatim, over the
exact (rounded, but reproducibly rounded) products---formally, one runs the check over
$\mathbb{F}_p$ on the integer significand--exponent representation of the pinned-order
output, with $p$ chosen to bound the dynamic range of the accumulated significands.

The cost is concrete and is the reason RepOps is a \emph{tier}, not the default: pinning
$\tau$ forfeits the vendor's hand-tuned GEMM kernels, whose speed comes precisely from
choosing $\tau$ freely (large tiles, many-thread reductions, FMA fusion, mixed
precision). A reproducible kernel fixes the reduction order and so runs slower than the
vendor optimum---the standard reproducible-summation trade-off~\cite{repbla}. The deployment therefore
pays for fp-precision proofs in throughput; whether that is worth it is a per-task
governance choice (Section~\ref{sec:profile-gate}), not a protocol constant.

\subsection{The coverage tiers, by arithmetic}
\label{sec:coverage-tiers}

Putting the pieces together, every model has at least one proof-bearing mode, selected
by its arithmetic and value. Table~\ref{tab:fp-coverage} states the menu. It refines,
rather than replaces, the trust-tier matrix of Section~\ref{sec:tiers}: that table
selects a backend by \emph{trust}; this one selects an \emph{exactness regime} by
arithmetic. The two compose---e.g.\ a bf16 model at high value runs RepOps-Freivalds
under an $M$-of-$N$ bond, while the same model quantized runs exact-integer Freivalds.

\begin{table}[t]
\centering
\caption{Coverage by arithmetic. Every modern-LLM arithmetic is given a proof-bearing mode
by tiering the verification to the arithmetic, not by a single floating-point Freivalds.
The quantized row is the egalitarian, software-only, hardware-free default and is
\emph{built and tested end to end}: it runs the whole-$K$ exact-integer accumulator
(\texttt{exact\_matmul}) that gives the $\epsilon=0$, zero-false-reject property, emits a
transcript (\texttt{ProvableLinear}/\texttt{GraphTrace}), and verifies on-chain
(\texttt{ComputeWitnessLib}). The floating-point (RepOps) row is \emph{designed} (proven
sound in~\cite{poisoundness}), awaiting the pinned-reduction kernel; the \TEE{} row is the
hardware-trust fallback (Section~\ref{sec:status-summary}).}
\label{tab:fp-coverage}
\small
\begin{tabular}{@{}p{2.8cm}p{4.5cm}p{4.9cm}@{}}
\toprule
Arithmetic & Verification mode & Property \\
\midrule
int8 / int4 / GGUF / AWQ / GPTQ / MLX-quant
  & \textbf{exact-integer Freivalds} (Section~\ref{sec:freivalds})
  & $\epsilon=0$, software-only, no hardware; soundness $1/p$ per vector
    ($2^{-122}$ dense); zero false-reject---\emph{for a pinned whole-$K$-i32 kernel
    (Rem.~\ref{rem:wholek}), to be built; \emph{not} the shipping block-\texttt{f32}
    kernel} \\[2pt]
fp16 / bf16 / fp8
  & \textbf{RepOps-Freivalds} (pinned reduction tree $\tau$;
    Section~\ref{sec:repops}) \emph{or} TEE/CC attestation (type 1/4)
  & RepOps: bit-exact via Prop.~\ref{prop:fp-integer}, costs vendor-kernel speed.
    TEE: hardware attests the fp run, no Freivalds, no $\tau$ pin \\[2pt]
small models (any arithmetic)
  & \textbf{zkML} (type 2; Section~\ref{sec:tiers})
  & trustless succinct verification; bounded by proving cost to small models \\
\bottomrule
\end{tabular}
\end{table}

The structure of Table~\ref{tab:fp-coverage} is the whole argument: full coverage is a
\emph{disjunction of exact modes}, not a single mode stretched past where it is sound.
A model that ships quantized (the common deployment, and the entire \Zoo{} zoo) is
\emph{designed to} get
the strongest of the three properties---exact, software-only, hardware-free,
tolerance-free---once the whole-$K$-i32 kernel and the trace emission are built. A model that must be served in full floating-point precision still has
a proof: pay the RepOps throughput tax for a software-only exact check, or attest in a
TEE for full-speed kernels at the cost of a hardware-trust assumption. No model is left
without a proof-bearing mode, and no mode is asked to bear a soundness or liveness load
it cannot carry.

\begin{principle}[Exactness is tiered to arithmetic]
\label{prin:exactness-tiered}
There is exactly one way to make a software-only, cross-hardware, tolerance-free
matmul check: exact arithmetic. Quantized models supply it natively (integer
accumulation); floating-point models supply it by pinning the reduction tree (RepOps)
at a throughput cost, or sidestep it via hardware attestation. ``Support every modern
LLM'' is therefore a routing decision over exact regimes---never a single floating-point
Freivalds, which is unsound (corridor) and unusable (false-reject) at once.
\end{principle}
