\section{The Formal Core: Receipt Soundness}
\label{sec:receipt-soundness}

The preceding sections build the execution proof (Sections~\ref{sec:freivalds},
\ref{sec:four-proofs}) and its binding (Section~\ref{sec:binding}). This
section states what \PoAI{} \emph{formally guarantees}, as a theorem over a single
universal object, and---equally important---draws the line at what it does
\emph{not} claim. The discipline is deliberate: \PoAI{} proves \emph{objective
lifecycle properties}, never semantic ones.

\subsection{What \PoAI{} proves, and what it refuses to claim}
\label{sec:rs-scope}

\PoAI{} proves exactly five properties of an AI contribution:

\begin{enumerate}[leftmargin=2em,itemsep=1pt]
  \item \textbf{Identity} --- which model/artifact lineage and which engine produced the work.
  \item \textbf{Execution} --- that the committed computation actually ran (not a cheaper substitute).
  \item \textbf{Provenance} --- the input, output, and lineage the work is bound to.
  \item \textbf{Accounting} --- that the work is metered and counted exactly once.
  \item \textbf{Settlement} --- that payment, mint, and governance are bound to the proof.
\end{enumerate}

\PoAI{} does \emph{not} prove that an answer is true, that a model is aligned, that
a dataset is morally good, or that an output is useful. Those are semantic judgments
about human intent; as the verification literature establishes, every such verifier
is a proxy for intent that degrades as the generator's capability grows
\cite{verificationhorizon}, and no fixed reward function for them stays sound.
\PoAI{} sidesteps that impossibility by anchoring economics to the one signal that
does \emph{not} decay---\emph{did the claimed work happen}---which is checkable to the
bit (Section~\ref{sec:freivalds}). The semantic remainder is pushed to an
adaptive, governed, reputation-weighted market (the quorum tier and
\texttt{AIReputation}); \PoAI{} makes no theorem about it.

\begin{principle}[Objective, not semantic]
\PoAI{} proves that claimed AI work was \emph{performed under committed artifacts},
and that payment, mint, and governance are \emph{bound to that proof}. It proves
nothing about the meaning, correctness, or value of the result.
\end{principle}

\subsection{The universal object}
\label{sec:rs-receipt}

Every AI contribution---inference, embedding, training, fine-tuning, quantization,
storage, streaming, evaluation, data---settles through one object,
\texttt{AIReceipt}. A receipt $R$ carries the job id, job type, requester,
contributor, the model/engine/runtime roots, the input and output commitments, the
witness root, the proof tier, the metering, and the license/attribution commitments.
Its canonical commitment is the two-step keccak chain of
Section~\ref{sec:binding}: a per-job-type envelope folds the type's extra
fields into $\mathit{promptHash}$, $\mathit{outputHash}$, and
$\mathit{runtimeMeasurement}$ \emph{before} the unchanged \texttt{ComputeProofLib}
chain runs, so that for \textsc{inference} the binding reduces \emph{byte-for-byte}
to the value the live miner already consumes, and every job type domain-separates so
two receipts of different types can never collide:
\[
\mathit{reportData} \;=\; \keccak\!\big(\textsf{DOMAIN}\;\|\;\mathit{challenge}\;\|\;
  \mathit{modelSpecHash}\;\|\;\mathit{promptHash}\;\|\;\mathit{outputHash}\;\|\;
  \mathit{runtimeMeasurement}\big),
\]
with $\mathit{receiptId} = \keccak(\textsf{DOMAIN\_RECEIPT}\,\|\,\mathit{reportData}\,\|\,
\langle\text{the full economic commitment}\rangle)$. The value path joins on
$\mathit{reportData}$; the audit log emits $\mathit{receiptId}$.

\subsection{The contract surface}
\label{sec:rs-contracts}

The object is realized by a small, orthogonal contract set (one concern each):

\begin{center}\small
\begin{tabular}{@{}ll@{}}
\toprule
\texttt{AIJob}     & request $+$ escrow \\
\texttt{AIReceipt} & the canonical commitment (binding library + universal struct) \\
\texttt{AIProof}   & read-only proof-route verification (the finalization gate) \\
\texttt{AISettle}  & pays / mints / slashes; the sole consequence authority \\
\texttt{AIModel}   & model lineage and roots (verifiable, non-equivocating) \\
\texttt{AIEngine}  & approved engine/runtime roots \\
\texttt{AIData}    & licenses, attribution DAG, royalties, public-goods pool \\
\texttt{AIStore}   & artifact availability / delivery proofs \\
\texttt{AICompute} & global no-double-mint accounting \\
\bottomrule
\end{tabular}
\end{center}

The dataflow is one way:
\[
\texttt{AIJob}\rightarrow\texttt{AIReceipt}\rightarrow\texttt{AIProof}
\rightarrow\texttt{AISettle}\rightarrow\texttt{AICompute},
\quad\text{with }\;
\texttt{AIData}/\texttt{AIModel}/\texttt{AIEngine}/\texttt{AIStore}\;\text{consulted at the seam.}
\]

\subsection{Proof tiers and the finalization gate}
\label{sec:rs-tiers}

Verification is tiered (Section~\ref{sec:tiers}), strongest-binding last:
\[
\textsf{L0 signed}\;<\;\textsf{L1 stake+challenge}\;<\;\textsf{L2 quorum}\;<\;
\textsf{L3 sampled execution witness (Freivalds)}\;<\;\textsf{L4 \TEE{} quote}\;<\;
\textsf{L5 zk}.
\]
Each job type carries a governance-set floor, and \texttt{AIProof} additionally binds
the proof \emph{system} to the job's \emph{work type} (a level gate alone is
insufficient---a storage receipt must not settle through the inference matmul
backend). Finalization requires, for the job type $\tau$ of receipt $R$,
\[
\textsf{ProofType}(R)\in\textsf{AllowedProofTypes}(\tau)\;\wedge\;
\textsf{ProofLevel}(R)\ge\textsf{MinProofLevel}(\tau),
\]
and the routed backend's witness must verify. No verification, no finalization; an
unwired or disabled backend \emph{fails closed}.

\subsection{Provable job classes}
\label{sec:rs-jobclasses}

Each class has its own verifier; all settle through the one receipt. The verifier
proves the \emph{objective} statement, never a semantic one:

\begin{center}\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Class} & \textbf{Proves} \\
\midrule
Inference     & model $M$ ran on input $X$ and committed output $Y$ \\
Embedding     & embedding model $E$ ran on $X$ and committed vector $V$ \\
Training      & dataset $D$, recipe $\rho$, base $B$ produced checkpoint $C$ under a committed transcript \\
Fine-tuning   & base $B$ $+$ adapter/delta $A$ $+$ data $D$ produced $M'$ \\
Quantization  & model $M$ transformed into quant artifact $Q$ under a committed quant recipe \\
Storage       & artifact $A$ was retrievable over time \\
Streaming     & pieces $P$ were delivered to receivers \\
Evaluation    & benchmark $\beta$ and scorer $S$ produced result $\Sigma$ \\
Data          & data commitment $D$ exists, is licensed, deduped, attributable, and accepted \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Invariants}
\label{sec:rs-invariants}

The following hold for every finalized receipt. Each is enforced on-chain; the
status column marks whether it is established \emph{by construction} (C), pinned by
the test suite (T), or---for the governance-consumption and storage/stream
availability wiring---staged at the integration seam (S).

\begin{center}\small
\begin{tabular}{@{}llc@{}}
\toprule
& \textbf{Invariant} & \textbf{Status} \\
\midrule
I1  & No receipt, no payment.                                   & C,T \\
I2  & No receipt, no mint.                                       & C,T \\
I3  & No receipt, no governance execution.                      & C,S \\
I4  & One job cannot settle twice.                              & C,T \\
I5  & One receipt cannot mint twice.                            & C,T \\
I6  & Receipt cannot change model after submission.            & C,T \\
I7  & Receipt cannot change input after submission.            & C,T \\
I8  & Receipt cannot change output after submission.           & C,T \\
I9  & Receipt cannot be replayed across chainId/domain.        & C,T \\
I10 & Proof backend cannot alter the receipt binding.          & C,T \\
I11 & Backend verifies the witness only.                       & C,T \\
I12 & \texttt{AISettle} cannot bypass \texttt{AIProof}.        & C,T \\
I13 & \texttt{AIGovern}/\texttt{AIExecute} consume only finalized receipts. & C,S \\
I14 & Royalties cannot exceed $100\%$.                          & C,T \\
I15 & Storage/stream rewards require an availability/delivery proof. & C,T \\
I16 & Training/fine-tune lineage binds base model, data, recipe, checkpoint. & C,T \\
\bottomrule
\end{tabular}
\end{center}

\noindent
I4--I8 follow from the commitment: $\mathit{receiptId}$ binds the job id, model,
input, output, contributor, and tier, so any change yields a different receipt that
cannot settle against the opened job. I5 and I9 are the one-shot
$\mathit{consumed}[\,\cdot\,]$ flag (keyed per job type) and the chain-independent
\texttt{AICompute} claim key. I10--I12 are the read-only verifier seam: the gate
pre-enforces $\mathit{reportData}\!=\!\mathit{expected}$ and routes only the witness
to the backend, and \texttt{AISettle}'s sole mint/pay path is gated on
\texttt{AIProof}. I14 is the $\Sigma\,\mathit{bps}\!=\!10{,}000$ split with a
slice-capped royalty reservation. I15/I16 are the proof-system gate of
Section~\ref{sec:rs-tiers} (a storage receipt admits only the retrievability
backend) and the domain-separated training fold bound to the lineage registry.

\subsection{Soundness theorems}
\label{sec:rs-theorems}

\begin{theorem}[Receipt Soundness]
\label{thm:receipt-soundness}
For any finalized \texttt{AIReceipt} $R$, assuming the selected proof backend is
sound, $R$ is a unique, non-replayable, non-malleable commitment to an AI
contribution that binds to exactly one job, exactly one model/artifact lineage,
exactly one input commitment, exactly one output commitment, exactly one proof
tier, and exactly one contributor; whose required proof verified \emph{before}
economic settlement; and from which at most one payment and at most one mint can
issue. $R$ cannot be replayed, substituted, double-paid, or double-mined.
\end{theorem}

\begin{proof}[Proof sketch]
Finalization requires \texttt{AIProof.validate} (the proof-system gate of
Section~\ref{sec:rs-tiers} plus the routed backend) and, for the pessimistic tiers,
$\mathit{finalized}(\mathit{reportData})$. Uniqueness and non-malleability:
$\mathit{receiptId}$ is the keccak commitment to all of $\{$job, model, input,
output, contributor, tier$\}$ (I6--I8), so a substitution is a distinct receipt that
fails the job-match check; collision resistance of keccak gives uniqueness. Cross-
domain non-replay: $\mathit{reportData}$ binds $\mathit{networkID}$ and
$\mathit{chainID}$ (I9). At-most-once settlement: the $\mathit{consumed}$ one-shot
and the \texttt{AICompute} claim (I4, I5) are set inside the same state snapshot as
the mint, so a revert un-sets them and a second attempt reverts. The backend cannot
forge the binding (I10, I11): the verifier pre-enforces equality and passes only the
witness; soundness of $R$ therefore reduces to soundness of the selected backend,
which is the hypothesis. \texttt{AISettle} has no mint or pay path that is not gated
on \texttt{AIProof} (I12).
\end{proof}

\begin{corollary}[Inference Soundness]
A finalized inference receipt binds model, engine, input, output, and a verified
proof; its $\mathit{reportData}$ equals \texttt{ComputeProofLib.expectedReportData}
byte-for-byte, so it is indistinguishable from a value the live miner accepts.
\end{corollary}

\begin{corollary}[Training-Lineage Soundness]
A finalized training (resp.\ fine-tune) receipt binds base model, dataset, recipe,
transcript, and checkpoint (resp.\ adapter/delta): the $\textsc{train}$ fold commits
the dataset and lineage roots, and the L3 transcript verifier re-derives a
beacon-sampled SGD step over the int8 accumulator, so a fabricated checkpoint is
caught except with the per-step Freivalds error.
\end{corollary}

\begin{corollary}[Accounting Soundness]
A finalized receipt produces at most one payment and at most one mint, across all
mechanisms and all chains: the per-type $\mathit{consumed}$ one-shot, the chain-
independent \texttt{AICompute} claim, and the shared halving cap make a second issue
revert.
\end{corollary}

\begin{corollary}[Governance Safety]
No \PoAI{}-governed action executes from an unfinalized or invalid receipt:
\texttt{AIGovern}/\texttt{AIExecute} consume only a finalized $\mathit{reportData}$
(I3, I13), which by Theorem~\ref{thm:receipt-soundness} carries a verified proof.
\end{corollary}

\subsection{The honest boundary}
\label{sec:rs-boundary}

Stated plainly, so the guarantee is defensible:

\begin{quote}\itshape
\PoAI{} does not prove an answer is true. It does not prove a model is aligned. It
does not prove a dataset is morally good. \PoAI{} proves that claimed AI work
happened under committed artifacts, and that payment, mint, and governance are bound
to that proof.
\end{quote}

\begin{remark}[The one-line truth]
\PoAI{} is not proof that AI is right. \PoAI{} is proof that AI work was
\emph{performed, bound, attributed, verified, and fairly settled}.
\end{remark}
