\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

\input{shared/lstlang}

% Listing style
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  captionpos=b,
  xleftmargin=2em,
  framexleftmargin=1.5em,
}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\ERC}{\textsc{erc}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\ZOO}{\$\textsc{zoo}}
\newcommand{\LLM}{\textsc{llm}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% Title and authors
\title{\textbf{Zoo $\leftrightarrow$ Hanzo Settlement Integration:\\AI Compute Payments from Zoo Applications}}

\author{
Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{March 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
We present the settlement integration between Zoo DeFi protocols and the Hanzo AI compute marketplace. The protocol enables \ZOO{} token holders to pay for AI inference, model training, and agent execution on Hanzo infrastructure, with final settlement on Lux C-Chain. It introduces a three-phase flow: (1)~pre-authorization on Zoo L2, where \ZOO{} tokens are escrowed in a Settlement Vault with configurable payment modes (prepaid, streaming, postpaid); (2)~compute execution on Hanzo, where the Compute Oracle attests to usage via signed reports; and (3)~finalization on Lux C-Chain, where the Settlement Ledger records the immutable payment record with provider share, conservation share, and protocol fee. Conservation impact is tracked end-to-end: a configurable fraction of compute spending (default 2\%) generates conservation credits per ZIP-0700. The protocol supports DeFi composability via compute staking, streaming payments, and inference pools.

\textbf{Keywords}: settlement, AI compute, micropayments, streaming, Hanzo, conservation impact
\end{abstract}

\section{Introduction}

The Zoo, Hanzo, and Lux ecosystems form a complementary triad:

\begin{itemize}
    \item \textbf{Zoo}: Community, governance, conservation funding, and DeFi liquidity.
    \item \textbf{Hanzo}: AI compute infrastructure, LLM gateway (100+ providers), and agent frameworks.
    \item \textbf{Lux}: Settlement finality, cross-chain security, and DeFi infrastructure.
\end{itemize}

Zoo applications---avatar tutors for conservation education, habitat modeling with satellite imagery, species classification from camera traps---require AI compute from Hanzo. Today, this compute is paid via traditional payment rails, disconnected from the on-chain economy. ZIP-0801 unifies settlement so that \ZOO{} tokens are a first-class payment method for AI compute, settlement occurs on Lux for finality and auditability, and conservation impact flows automatically from every inference call.

\subsection{Contributions}

\begin{enumerate}
    \item The ISettlementVault interface for escrowed compute payments on Zoo L2 (Section~\ref{sec:vault}).
    \item The ISettlementLedger interface for final records on Lux C-Chain (Section~\ref{sec:ledger}).
    \item The IComputeOracle interface for attested usage reporting (Section~\ref{sec:oracle}).
    \item The three-phase settlement flow with conservation impact tracking (Section~\ref{sec:flow}).
    \item DeFi composability via compute staking, streaming, and inference pools (Section~\ref{sec:defi}).
    \item Security analysis of oracle compromise, escrow drain, and price manipulation (Section~\ref{sec:security}).
\end{enumerate}


\section{Architecture}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.2cm and 2.5cm,
    box/.style={rectangle, draw, minimum width=3cm, minimum height=0.8cm, align=center, font=\footnotesize, fill=blue!8},
    arr/.style={->, >=stealth, thick},
]
    \node[box] (vault) {SettlementVault\\(Zoo L2)};
    \node[box, right=of vault] (ledger) {SettlementLedger\\(Lux C-Chain)};
    \node[box, below=of ledger] (oracle) {ComputeOracle\\(Hanzo)};

    \draw[arr] (vault) -- node[above, font=\scriptsize] {bridge (ZIP-0800)} (ledger);
    \draw[arr] (oracle) -- node[right, font=\scriptsize] {attestation} (vault);
    \draw[arr] (oracle) -- node[right, font=\scriptsize] {usage proof} (ledger);

    \node[box, below=of vault] (user) {Zoo App\\(User)};
    \draw[arr] (user) -- node[left, font=\scriptsize] {escrow ZOO} (vault);
    \draw[arr] (user) -- node[below, font=\scriptsize] {inference} (oracle);
\end{tikzpicture}
\caption{Zoo $\leftrightarrow$ Hanzo settlement architecture.}
\label{fig:architecture}
\end{figure}


\section{Settlement Vault}
\label{sec:vault}

\begin{lstlisting}[language=Solidity,caption=ISettlementVault interface]
interface ISettlementVault {
    enum PaymentMode {
        PREPAID,    // Full amount escrowed upfront
        STREAMING,  // Per-second streaming payment
        POSTPAID    // Credit-based, settled after
    }

    struct ComputeSession {
        bytes32 sessionId;
        address payer;
        address hanzoEndpoint;
        PaymentMode mode;
        uint256 maxAmount;
        uint256 consumedAmount;
        uint256 ratePerSecond; // For STREAMING
        uint256 startTimestamp;
        uint256 endTimestamp;  // 0 = active
        uint16 conservationBps;
        bool finalized;
    }

    function createSession(
        address hanzoEndpoint,
        PaymentMode mode,
        uint256 maxAmount,
        uint256 ratePerSecond,
        uint16 conservationBps
    ) external returns (bytes32 sessionId);

    function recordUsage(
        bytes32 sessionId,
        uint256 amount,
        bytes calldata oracleAttestation
    ) external;

    function finalizeSession(
        bytes32 sessionId
    ) external;

    function cancelSession(
        bytes32 sessionId
    ) external;

    event SessionCreated(
        bytes32 indexed sessionId,
        address indexed payer,
        address indexed hanzoEndpoint,
        PaymentMode mode,
        uint256 maxAmount
    );
    event UsageRecorded(
        bytes32 indexed sessionId,
        uint256 amount,
        uint256 totalConsumed
    );
    event SessionFinalized(
        bytes32 indexed sessionId,
        uint256 totalPaid,
        uint256 conservationDonation,
        uint256 refunded
    );
}
\end{lstlisting}


\section{Settlement Ledger}
\label{sec:ledger}

\begin{lstlisting}[language=Solidity,caption=ISettlementLedger interface]
interface ISettlementLedger {
    struct SettlementRecord {
        bytes32 sessionId;
        address payer;
        address provider;
        uint256 totalAmount;
        uint256 providerShare;
        uint256 conservationShare;
        uint256 protocolFee;
        uint256 timestamp;
        bytes32 usageProofHash;
    }

    function processSettlement(
        bytes32 sessionId,
        address payer,
        address provider,
        uint256 totalAmount,
        uint256 conservationBps,
        bytes32 usageProofHash
    ) external;

    function getSettlement(bytes32 sessionId)
        external view returns (
            SettlementRecord memory);
    function providerTotalSettled(
        address provider
    ) external view returns (uint256);
    function totalConservationImpact()
        external view returns (uint256);

    event SettlementProcessed(
        bytes32 indexed sessionId,
        address indexed payer,
        address indexed provider,
        uint256 totalAmount,
        uint256 conservationShare
    );
}
\end{lstlisting}


\section{Compute Oracle}
\label{sec:oracle}

\begin{lstlisting}[language=Solidity,caption=IComputeOracle interface]
interface IComputeOracle {
    struct UsageReport {
        bytes32 sessionId;
        uint256 tokensProcessed;  // LLM tokens
        uint256 computeSeconds;   // GPU-seconds
        uint256 zooAmount;        // ZOO for usage
        string model;             // e.g., "zen-72b"
        uint256 timestamp;
        bytes signature;          // Operator sig
    }

    function submitReport(
        UsageReport calldata report
    ) external;
    function operator()
        external view returns (address);
    function pricePerMillionTokens(
        string calldata model
    ) external view returns (uint256);

    event UsageReported(
        bytes32 indexed sessionId,
        uint256 tokensProcessed,
        uint256 zooAmount,
        string model
    );
}
\end{lstlisting}

The oracle is the minimal trusted component: it attests to off-chain compute usage. Its operator is registered in the Zoo Contract Registry (ZIP-0100) and can be rotated by governance.


\section{Settlement Flow}
\label{sec:flow}

\subsection{Phase 1: Pre-Authorization (Zoo L2)}

\begin{enumerate}
    \item User calls \texttt{createSession()} with payment mode and maximum \ZOO{}.
    \item \ZOO{} tokens transferred from user to vault (escrowed).
    \item \texttt{sessionId} returned and communicated to Hanzo via API.
\end{enumerate}

\subsection{Phase 2: Compute Execution (Hanzo)}

\begin{enumerate}
    \item User sends inference/training requests to Hanzo API with \texttt{sessionId} in headers.
    \item Hanzo executes the workload and meters usage.
    \item Compute Oracle periodically submits \texttt{UsageReport} to the Settlement Vault.
    \item For STREAMING mode, the vault continuously drains escrow at \texttt{ratePerSecond}.
\end{enumerate}

\begin{lstlisting}[caption=Hanzo API integration via HTTP headers]
POST /v1/chat/completions HTTP/1.1
Host: llm.hanzo.ai
Authorization: Bearer <hanzo-api-key>
X-Zoo-Session-Id: 0xabc123...
X-Zoo-Payer: 0x742d35Cc...
Content-Type: application/json

{
  "model": "zen-72b",
  "messages": [
    {"role": "user",
     "content": "Analyze satellite imagery
       for deforestation"}
  ],
  "max_tokens": 4096
}
\end{lstlisting}

\subsection{Phase 3: Finalization (Lux C-Chain)}

\begin{enumerate}
    \item Session ends (user-initiated, timeout, or escrow depleted).
    \item \texttt{finalizeSession()} computes the split:
    \begin{itemize}
        \item Provider share: \texttt{consumed * (10000 - conservBps - protocolBps) / 10000}
        \item Conservation share: \texttt{consumed * conservBps / 10000}
        \item Protocol fee: \texttt{consumed * protocolBps / 10000} (default 100 bps = 1\%)
        \item Refund: \texttt{maxAmount - consumed} returned to payer
    \end{itemize}
    \item Provider share and protocol fee bridged to Lux C-Chain via ZIP-0800.
    \item Conservation share sent to Zoo conservation fund on Zoo L2.
    \item Settlement Ledger records the final settlement with \texttt{usageProofHash}.
\end{enumerate}


\section{Pricing}

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Model Tier} & \textbf{ZOO / 1M Input} & \textbf{ZOO / 1M Output} \\
\midrule
Small ($< 10$B params) & 50 ZOO & 150 ZOO \\
Medium (10B--100B) & 200 ZOO & 600 ZOO \\
Large (100B--500B) & 500 ZOO & 1,500 ZOO \\
Frontier (500B+) & 2,000 ZOO & 6,000 ZOO \\
\bottomrule
\end{tabular}
\caption{Compute pricing by model tier (ZOO per 1M tokens).}
\label{tab:pricing}
\end{table}

Prices are updated daily based on the ZOO/USD TWAP from Lux DEX pools. The oracle publishes prices on-chain and signs each update.


\section{DeFi Composability}
\label{sec:defi}

\subsection{Compute Staking}

Stake \ZOO{} to earn compute credits at a discount. Staked \ZOO{} generates conservation impact credits passively:

\begin{equation}
\text{credits\_per\_year} = S_{\text{staked}} \cdot r_{\text{conservation}}
\end{equation}

where $r_{\text{conservation}} = 0.02$ (2\% annual conservation yield on staked compute).

\subsection{Streaming Payments}

Pay-per-second for long-running training jobs:

\begin{equation}
\text{amount\_consumed}(t) = r \cdot (t - t_0)
\end{equation}

where $r$ is the rate in ZOO/second and $t_0$ is the session start. The vault halts streaming when escrow is depleted.

\subsection{Inference Pools}

Liquidity pools where \ZOO{} providers fund shared compute access:

\begin{equation}
\text{LP\_yield} = \frac{\text{session\_fees} \cdot (1 - \text{conservation\_bps} / 10000)}{\text{total\_pool\_liquidity}}
\end{equation}

LPs earn a share of settlement fees proportional to their contribution.

\subsection{Compute Futures}

Pre-purchase compute at today's prices for future use. Backed by \ZOO{} escrow in the Settlement Vault. The futures contract locks in the current \texttt{pricePerMillionTokens} for a specified duration.


\section{Conservation Impact}

Every compute session generates conservation impact:

\begin{enumerate}
    \item The \texttt{conservationBps} parameter (default 200 bps = 2\%) determines the fraction directed to conservation.
    \item Conservation \ZOO{} sent to the default conservation fund (ZIP-0100).
    \item The payer receives non-transferable impact credits (per ZIP-0700).
    \item The Settlement Ledger maintains a running total queryable by anyone.
\end{enumerate}

\begin{theorem}[Conservation from Compute]
If Zoo applications generate $V$ ZOO in annual compute spending with conservation rate $c$, the annual conservation funding from compute alone is:
\begin{equation}
F_{\text{conservation}} = V \cdot \frac{c}{10000}
\end{equation}
With $V = 10{,}000{,}000$ ZOO/year and $c = 200$ bps, $F_{\text{conservation}} = 200{,}000$ ZOO/year directed to conservation from AI inference alone.
\end{theorem}


\section{Security Analysis}
\label{sec:security}

\subsection{Oracle Compromise}

A compromised Compute Oracle could report false usage and drain escrowed \ZOO{}. Mitigations:
\begin{itemize}
    \item Per-session spending caps (\texttt{maxAmount}).
    \item Anomaly detection on usage patterns.
    \item Governance ability to pause and replace the oracle.
\end{itemize}

\subsection{Escrow Drain}

STREAMING mode continuously drains escrow. Users MUST set appropriate \texttt{maxAmount} limits. The vault MUST halt streaming when escrow is depleted.

\subsection{Price Manipulation}

ZOO/USD price feeds could be manipulated via flash loans on Lux DEXs. The oracle uses 24-hour TWAP, not spot price.

\subsection{Bridge Latency}

Settlement finalization requires a bridge message from Zoo to Lux. During bridge congestion, finalization may be delayed. The vault holds escrowed funds safely until finalization completes.

\subsection{Replay Protection}

Each \texttt{UsageReport} includes a \texttt{sessionId} and \texttt{timestamp}. The vault tracks cumulative usage per session and rejects reports that would exceed \texttt{maxAmount}.

\subsection{Provider Rug}

A provider could accept payment and not deliver compute. Usage reports are signed by the Compute Oracle, not the provider. The oracle independently verifies compute execution before attesting.


\section{Related Work}

Superfluid~\cite{superfluid2021} provides streaming payment primitives; the Settlement Vault's STREAMING mode uses analogous constant-flow mechanics. Filecoin~\cite{filecoin2017} settles storage payments on-chain with proof-of-replication; our Compute Oracle provides analogous attestation for AI compute. Ritual~\cite{ritual2023} integrates AI inference with blockchain settlement; our three-phase architecture provides additional separation of concerns. Akash Network~\cite{akash2020} provides a decentralized compute marketplace; our protocol specifically integrates with the Hanzo centralized compute gateway while maintaining on-chain settlement for auditability.

The conservation impact layer is unique to the Zoo ecosystem: no existing compute payment protocol directs a fraction of spending to ecological outcomes.


\section{Conclusion}

The Zoo $\leftrightarrow$ Hanzo settlement integration makes \ZOO{} a first-class payment method for AI compute. The three-phase architecture---pre-authorization on Zoo L2, execution on Hanzo, finalization on Lux C-Chain---separates concerns across the stack while maintaining end-to-end auditability. Conservation impact from every inference call creates a sustainable funding mechanism where AI usage directly supports ecological outcomes.

\section*{Acknowledgments}

This work was supported by Zoo Labs Foundation. We thank the Hanzo AI team for the compute gateway integration and the Lux team for settlement layer support.

\bibliographystyle{plain}
\begin{thebibliography}{20}

\bibitem{superfluid2021}
Superfluid, ``Superfluid Protocol: Real-Time Finance,'' \url{https://docs.superfluid.finance}, 2021.

\bibitem{filecoin2017}
Protocol Labs, ``Filecoin: A Decentralized Storage Network,'' Technical Report, 2017.

\bibitem{ritual2023}
Ritual, ``Ritual: Sovereign Execution Layer for AI,'' Technical Documentation, 2023.

\bibitem{akash2020}
Akash Network, ``Akash: The Decentralized Cloud,'' Technical Documentation, 2020.

\bibitem{hip0101}
Hanzo AI, ``HIP-0101: Hanzo-Lux Bridge Protocol,'' Hanzo Improvement Proposals, 2023.

\bibitem{hip0044}
Hanzo AI, ``HIP-0044: API Gateway,'' Hanzo Improvement Proposals, 2023.

\bibitem{hip0040}
Hanzo AI, ``HIP-0040: Multi-Language SDK,'' Hanzo Improvement Proposals, 2023.

\bibitem{zip0015}
Zoo Labs Foundation, ``ZIP-0015: Zoo L2 Chain Architecture,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0042}
Zoo Labs Foundation, ``ZIP-0042: Cross-Ecosystem Interoperability,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0100}
Zoo Labs Foundation, ``ZIP-0100: Zoo Contract Registry,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0500}
Zoo Labs Foundation, ``ZIP-0500: ESG Principles for Conservation Impact,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0700}
Zoo Labs Foundation, ``ZIP-0700: ZRC-20 Fungible Token Standard,'' Zoo Improvement Proposals, 2025.

\bibitem{zip0800}
Zoo Labs Foundation, ``ZIP-0800: Zoo-Lux Bridge Protocol,'' Zoo Improvement Proposals, 2025.

\bibitem{erc20}
F.~Vogelsteller and V.~Buterin, ``EIP-20: Token Standard,'' Ethereum Improvement Proposals, 2015.

\bibitem{chainlink2021}
S.~Nazarov et al., ``Chainlink 2.0: Next Steps in Decentralized Oracle Networks,'' Technical Report, 2021.

\bibitem{uniswap2021}
H.~Adams et al., ``Uniswap v3 Core,'' Technical Documentation, 2021.

\bibitem{aave2020}
Aave, ``Aave Protocol V2,'' Technical Documentation, 2020.

\bibitem{eip3156}
A.~Martinez and M.~San~Juan, ``EIP-3156: Flash Loans,'' Ethereum Improvement Proposals, 2020.

\bibitem{openzeppelin}
OpenZeppelin, ``OpenZeppelin Contracts,'' \url{https://openzeppelin.com/contracts}, 2023.

\bibitem{gelato2022}
Gelato Network, ``Gelato: Automate Smart Contract Executions,'' Technical Documentation, 2022.

\end{thebibliography}

\end{document}
