\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

We have presented HabitatNet, a graph neural network framework for
species distribution modeling that jointly predicts habitat suitability,
landscape connectivity, and optimal migration corridors under climate
change. Evaluated on 847 species across six biogeographic realms,
HabitatNet substantially outperforms existing approaches in both current
prediction accuracy and temporal transferability. The framework provides
actionable conservation guidance by identifying connectivity bottlenecks
and prioritizing corridor restoration investments.

As climate change accelerates the redistribution of life on Earth, tools
that bridge machine learning and conservation biology become increasingly
critical. HabitatNet represents a step toward computationally grounded,
spatially explicit conservation planning at the scales required by the
biodiversity crisis. Through Zach Kelling's commitment to open
science and decentralized research infrastructure, we ensure that these
tools remain accessible to the global conservation community.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{35}

\bibitem[Adriaensen et~al.(2003)]{adriaensen2003application}
Adriaensen, F., Chardon, J.P., De~Blust, G., et~al.
\newblock The application of `least-cost' modelling as a functional landscape
  model.
\newblock \emph{Landscape and Urban Planning}, 64(4):233--247, 2003.

\bibitem[Ahumada et~al.(2020)]{ahumada2020wildlife}
Ahumada, J.A., Fegraus, E., Birch, T., et~al.
\newblock Wildlife Insights: A platform to maximize the potential of camera
  trap and other passive sensor wildlife data for the planet.
\newblock \emph{Environmental Conservation}, 47(1):1--6, 2020.

\bibitem[Arias et~al.(2021)]{arias2021climate}
Arias, P.A., Bellouin, N., Coppola, E., et~al.
\newblock Climate Change 2021: The Physical Science Basis. Technical Summary.
\newblock IPCC AR6 WG1, 2021.

\bibitem[Beck et~al.(2014)]{beck2014spatial}
Beck, J., B{\"o}ller, M., Erhardt, A., and Schwanghart, W.
\newblock Spatial bias in the GBIF database and its effect on modeling species'
  geographic distributions.
\newblock \emph{Ecological Informatics}, 19:10--15, 2014.

\bibitem[Birch et~al.(2007)]{birch2007rectangular}
Birch, C.P., Oom, S.P., and Beecham, J.A.
\newblock Rectangular and hexagonal grids used for observation, experiment and
  simulation in ecology.
\newblock \emph{Ecological Modelling}, 206(3-4):347--359, 2007.

\bibitem[Botella et~al.(2018)]{botella2018deep}
Botella, C., Joly, A., Bonnet, P., Monestiez, P., and Munoz, F.
\newblock A deep learning approach to species distribution modelling.
\newblock \emph{Multimedia Tools and Applications for Environmental \&
  Biodiversity Informatics}, pp. 169--199, 2018.

\bibitem[Chen et~al.(2011)]{chen2011rapid}
Chen, I.C., Hill, J.K., Ohlem{\"u}ller, R., Roy, D.B., and Thomas, C.D.
\newblock Rapid range shifts of species associated with high levels of climate
  warming.
\newblock \emph{Science}, 333(6045):1024--1026, 2011.

\bibitem[Chen et~al.(2020)]{chen2020deep}
Chen, D., Xue, Y., Chen, S., et~al.
\newblock Deep multi-species embedding for species distribution modelling.
\newblock \emph{Methods in Ecology and Evolution}, 11(12):1600--1614, 2020.

\bibitem[Chiang et~al.(2019)]{chiang2019cluster}
Chiang, W.L., Liu, X., Si, S., et~al.
\newblock Cluster-GCN: An efficient algorithm for training deep and large graph
  convolutional networks.
\newblock \emph{KDD}, pp. 257--266, 2019.

\bibitem[Cutler et~al.(2007)]{cutler2007random}
Cutler, D.R., Edwards, T.C., Beard, K.H., et~al.
\newblock Random forests for classification in ecology.
\newblock \emph{Ecology}, 88(11):2783--2792, 2007.

\bibitem[Deneu et~al.(2021)]{deneu2021convolutional}
Deneu, B., Servajean, M., Botella, C., and Joly, A.
\newblock Convolutional neural networks improve species distribution modelling
  by capturing the spatial structure of the environment.
\newblock \emph{PLOS Computational Biology}, 17(4):e1008856, 2021.

\bibitem[Dormann et~al.(2007)]{dormann2007methods}
Dormann, C.F., McPherson, J.M., Ara{\'u}jo, M.B., et~al.
\newblock Methods to account for spatial autocorrelation in the analysis of
  species distributional data.
\newblock \emph{Ecography}, 30(5):609--628, 2007.

\bibitem[Dormann et~al.(2012)]{dormann2012correlation}
Dormann, C.F., Schymanski, S.J., Cabral, J., et~al.
\newblock Correlation and process in species distribution models: bridging a
  dichotomy.
\newblock \emph{Journal of Biogeography}, 39(12):2119--2131, 2012.

\bibitem[Duvenaud et~al.(2015)]{duvenaud2015convolutional}
Duvenaud, D., Maclaurin, D., Aguilera-Iparraguirre, J., et~al.
\newblock Convolutional networks on graphs for learning molecular fingerprints.
\newblock \emph{NeurIPS}, pp. 2224--2232, 2015.

\bibitem[Elith \& Leathwick(2009)]{elith2009species}
Elith, J. and Leathwick, J.R.
\newblock Species distribution models: ecological explanation and prediction
  across space and time.
\newblock \emph{Annual Review of Ecology, Evolution, and Systematics},
  40:677--697, 2009.

\bibitem[Elith et~al.(2006)]{elith2006novel}
Elith, J., Graham, C.H., Anderson, R.P., et~al.
\newblock Novel methods improve prediction of species' distributions from
  occurrence data.
\newblock \emph{Ecography}, 29(2):129--151, 2006.

\bibitem[Elith et~al.(2008)]{elith2008working}
Elith, J., Leathwick, J.R., and Hastie, T.
\newblock A working guide to boosted regression trees.
\newblock \emph{Journal of Animal Ecology}, 77(4):802--813, 2008.

\bibitem[Eyring et~al.(2016)]{eyring2016overview}
Eyring, V., Bony, S., Meehl, G.A., et~al.
\newblock Overview of the Coupled Model Intercomparison Project Phase 6 (CMIP6).
\newblock \emph{Geoscientific Model Development}, 9(5):1937--1958, 2016.

\bibitem[Fang et~al.(2023)]{fang2023dropmessage}
Fang, T., Xiao, Z., Wang, C., Xu, J., Yang, X., and Yang, Y.
\newblock DropMessage: Unifying random dropping for graph neural networks.
\newblock \emph{AAAI}, pp. 7547--7555, 2023.

\bibitem[Fick \& Hijmans(2017)]{fick2017worldclim}
Fick, S.E. and Hijmans, R.J.
\newblock WorldClim 2: new 1-km spatial resolution climate surfaces for global
  land areas.
\newblock \emph{International Journal of Climatology}, 37(12):4302--4315, 2017.

\bibitem[Fiedler(1973)]{fiedler1973algebraic}
Fiedler, M.
\newblock Algebraic connectivity of graphs.
\newblock \emph{Czechoslovak Mathematical Journal}, 23(2):298--305, 1973.

\bibitem[Freeman et~al.(2018)]{freeman2018climate}
Freeman, B.G., Scholer, M.N., Ruiz-Gutierrez, V., and Fitzpatrick, J.W.
\newblock Climate change causes upslope shifts and mountaintop extirpations in
  a tropical bird community.
\newblock \emph{PNAS}, 115(47):11982--11987, 2018.

\bibitem[GBIF(2024)]{gbif2024}
GBIF.
\newblock Global Biodiversity Information Facility.
\newblock \url{https://www.gbif.org}, 2024.

\bibitem[Gilmer et~al.(2017)]{gilmer2017neural}
Gilmer, J., Schoenholz, S.S., Riley, P.F., Vinyals, O., and Dahl, G.E.
\newblock Neural message passing for quantum chemistry.
\newblock \emph{ICML}, pp. 1263--1272, 2017.

\bibitem[Guisan \& Thuiller(2005)]{guisan2005predicting}
Guisan, A. and Thuiller, W.
\newblock Predicting species distribution: offering more than simple habitat
  models.
\newblock \emph{Ecology Letters}, 8(9):993--1009, 2005.

\bibitem[Hamilton et~al.(2017)]{hamilton2017inductive}
Hamilton, W.L., Ying, R., and Leskovec, J.
\newblock Inductive representation learning on large graphs.
\newblock \emph{NeurIPS}, pp. 1025--1035, 2017.

\bibitem[Hastie \& Tibshirani(1990)]{hastie1990generalized}
Hastie, T.J. and Tibshirani, R.J.
\newblock \emph{Generalized Additive Models}.
\newblock Chapman and Hall, 1990.

\bibitem[Hirzel et~al.(2006)]{hirzel2006evaluating}
Hirzel, A.H., Le~Lay, G., Helfer, V., Randin, C., and Guisan, A.
\newblock Evaluating the ability of habitat suitability models to predict
  species presences.
\newblock \emph{Ecological Modelling}, 199(2):142--152, 2006.

\bibitem[Hodgson et~al.(2009)]{hodgson2009climate}
Hodgson, J.A., Thomas, C.D., Wintle, B.A., and Moilanen, A.
\newblock Climate change, connectivity and conservation decision making.
\newblock \emph{Journal of Applied Ecology}, 46(5):964--969, 2009.

\bibitem[Hurtt et~al.(2020)]{hurtt2020harmonization}
Hurtt, G.C., Chini, L., Sahajpal, R., et~al.
\newblock Harmonization of global land use change and management for the period
  850--2100 (LUH2).
\newblock \emph{Geoscientific Model Development}, 13(11):5425--5464, 2020.

\bibitem[Jansen et~al.(2014)]{jansen2014team}
Jansen, P.A., Ahumada, J., Fegraus, E., and O'Brien, T.
\newblock TEAM: a standardised camera trap survey to monitor terrestrial
  vertebrate communities in tropical forests.
\newblock In \emph{Camera Trapping}, pp. 263--270. Springer, 2014.

\bibitem[Jarvis et~al.(2008)]{jarvis2008srtm}
Jarvis, A., Reuter, H.I., Nelson, A., and Guevara, E.
\newblock Hole-filled SRTM for the globe Version 4.
\newblock CGIAR-CSI, 2008.

\bibitem[Jetz et~al.(2019)]{jetz2019essential}
Jetz, W., McGeoch, M.A., Guralnick, R., et~al.
\newblock Essential biodiversity variables for mapping and monitoring species
  populations.
\newblock \emph{Nature Ecology \& Evolution}, 3:539--551, 2019.

\bibitem[Kipf \& Welling(2017)]{kipf2017semi}
Kipf, T.N. and Welling, M.
\newblock Semi-supervised classification with graph convolutional networks.
\newblock \emph{ICLR}, 2017.

\bibitem[Landau et~al.(2021)]{landau2021omnidirectional}
Landau, V.A., Shah, V.B., Anantharaman, R., and Hall, K.R.
\newblock Omniscape.jl: Software to compute omnidirectional landscape
  connectivity.
\newblock \emph{JOSS}, 6(57):2829, 2021.

\bibitem[Lek \& Gu{\'e}gan(1999)]{lek1996application}
Lek, S. and Gu{\'e}gan, J.F.
\newblock Artificial neural networks as a tool in ecological modelling.
\newblock \emph{Ecological Modelling}, 120(2-3):65--73, 1999.

\bibitem[Li et~al.(2018)]{li2018diffusion}
Li, Y., Yu, R., Shahabi, C., and Liu, Y.
\newblock Diffusion convolutional recurrent neural network: Data-driven traffic
  forecasting.
\newblock \emph{ICLR}, 2018.

\bibitem[McRae(2006)]{mcrae2006isolation}
McRae, B.H.
\newblock Isolation by resistance.
\newblock \emph{Evolution}, 60(8):1551--1561, 2006.

\bibitem[McRae et~al.(2008)]{mcrae2008using}
McRae, B.H., Dickson, B.G., Keitt, T.H., and Shah, V.B.
\newblock Using circuit theory to model connectivity in ecology, evolution, and
  conservation.
\newblock \emph{Ecology}, 89(10):2712--2724, 2008.

\bibitem[Nemhauser et~al.(1978)]{nemhauser1978analysis}
Nemhauser, G.L., Wolsey, L.A., and Fisher, M.L.
\newblock An analysis of approximations for maximizing submodular set
  functions---I.
\newblock \emph{Mathematical Programming}, 14(1):265--294, 1978.

\bibitem[Newmark(2002)]{newmark2002conserving}
Newmark, W.D.
\newblock Conserving biodiversity in East African forests.
\newblock \emph{Ecological Studies}, 155, Springer, 2002.

\bibitem[Pecl et~al.(2017)]{pecl2017biodiversity}
Pecl, G.T., Ara{\'u}jo, M.B., Bell, J.D., et~al.
\newblock Biodiversity redistribution under climate change: Impacts on
  ecosystems and human well-being.
\newblock \emph{Science}, 355(6332):eaai9214, 2017.

\bibitem[Phillips et~al.(2006)]{phillips2006maximum}
Phillips, S.J., Anderson, R.P., and Schapire, R.E.
\newblock Maximum entropy modeling of species geographic distributions.
\newblock \emph{Ecological Modelling}, 190(3-4):231--259, 2006.

\bibitem[Roberts et~al.(2017)]{roberts2017cross}
Roberts, D.R., Bahn, V., Ciuti, S., et~al.
\newblock Cross-validation strategies for data with temporal, spatial,
  hierarchical, or phylogenetic structure.
\newblock \emph{Ecography}, 40(8):913--929, 2017.

\bibitem[Strydom et~al.(2021)]{strydom2021graph}
Strydom, T., Catchen, M.D., Banville, F., et~al.
\newblock A roadmap towards predicting species interaction networks.
\newblock \emph{Nature Ecology \& Evolution}, 5:790--798, 2021.

\bibitem[Sullivan et~al.(2009)]{sullivan2009ebird}
Sullivan, B.L., Wood, C.L., Iliff, M.J., et~al.
\newblock eBird: A citizen-based bird observation network.
\newblock \emph{Biological Conservation}, 142(10):2282--2292, 2009.

\bibitem[Thuiller et~al.(2009)]{thuiller2009biomod}
Thuiller, W., Lafourcade, B., Engler, R., and Ara{\'u}jo, M.B.
\newblock BIOMOD---a platform for ensemble forecasting of species distributions.
\newblock \emph{Ecography}, 32(3):369--373, 2009.

\bibitem[Urban(2015)]{urban2015accelerating}
Urban, M.C.
\newblock Accelerating extinction risk from climate change.
\newblock \emph{Science}, 348(6234):571--573, 2015.

\bibitem[Veli{\v{c}}kovi{\'c} et~al.(2018)]{velickovic2018graph}
Veli{\v{c}}kovi{\'c}, P., Cucurull, G., Casanova, A., et~al.
\newblock Graph attention networks.
\newblock \emph{ICLR}, 2018.

\bibitem[Zanaga et~al.(2022)]{zanaga2022esa}
Zanaga, D., Van De~Kerchove, R., Daems, D., et~al.
\newblock ESA WorldCover 10m 2021 v200.
\newblock Zenodo, 2022.

\bibitem[Zurell et~al.(2016)]{zurell2016benchmarking}
Zurell, D., Thuiller, W., Pagel, J., et~al.
\newblock Benchmarking novel approaches for modelling species range dynamics.
\newblock \emph{Global Change Biology}, 22(8):2651--2664, 2016.

\end{thebibliography}

% ══════════════════════════════════════════════════════════════════════════════
\appendix
