\section{Conclusion}

\label{sec:conclusion}

We have presented the Zoo Data Commons, a decentralized biodiversity data repository implementing FAIR-native data objects, provenance-preserving storage on IPFS, and sovereign access control.
Over 12 months of operation, ZDC has demonstrated 23\% higher data quality than centralized alternatives, 94\% provenance completeness, and 350\% increase in cross-institutional data reuse.
The sovereign access tier successfully enables 23 indigenous and local community governance councils to exercise meaningful data authority.

ZDC is specified as ZIP-009 and deployed as open infrastructure.
Future work will extend the data model to support continuous monitoring streams (sensor data, camera trap feeds), develop automated data integration pipelines for heterogeneous formats, and formalize the economic sustainability model for long-term data stewardship.

% ============================================================
% References
% ============================================================
\begin{thebibliography}{30}

\bibitem[Anderson and Christen(2010)]{anderson2010tk}
Anderson, J. and Christen, K. (2010).
\newblock Traditional Knowledge Labels: A mechanism for indigenous control of traditional knowledge.
\newblock {\em Museum Anthropology}, 33(2):178--190.

\bibitem[Ari\~{n}o et~al.(2013)]{arinoetal2013}
Ari\~{n}o, A.~H., Chavan, V., and King, N. (2013).
\newblock The biodiversity informatics potential index.
\newblock {\em BMC Bioinformatics}, 14(Suppl 17):S4.

\bibitem[Benet(2014)]{benet2014ipfs}
Benet, J. (2014).
\newblock IPFS---Content addressed, versioned, P2P file system.
\newblock {\em arXiv preprint arXiv:1407.3561}.

\bibitem[Benson et~al.(2018)]{benson2018genbank}
Benson, D.~A., Cavanaugh, M., Clark, K., et~al. (2018).
\newblock GenBank.
\newblock {\em Nucleic Acids Research}, 46(D1):D99--D104.

\bibitem[Carroll et~al.(2020)]{carroll2020care}
Carroll, S.~R., Garba, I., Figueroa-Rodr\'{i}guez, O.~L., et~al. (2020).
\newblock The CARE Principles for Indigenous Data Governance.
\newblock {\em Data Science Journal}, 19(1):43.

\bibitem[Devaraju et~al.(2021)]{devaraju2021fairsfair}
Devaraju, A., Huber, R., Mokrane, M., et~al. (2021).
\newblock FAIRsFAIR data object assessment metrics.
\newblock {\em Zenodo}, doi:10.5281/zenodo.4081213.

\bibitem[D\'{i}az et~al.(2019)]{diaz2019global}
D\'{i}az, S., Settele, J., and Brondizio, E.~S. (2019).
\newblock Pervasive human-driven decline of life on Earth points to the need for transformative change.
\newblock {\em Science}, 366(6471):eaax3100.

\bibitem[GBIF(2024)]{gbif2024overview}
GBIF (2024).
\newblock GBIF Science Review 2024.
\newblock {\em GBIF Secretariat, Copenhagen}.

\bibitem[Grassle(2000)]{grassle2000ocean}
Grassle, J.~F. (2000).
\newblock The Ocean Biogeographic Information System (OBIS).
\newblock {\em Oceanography}, 13(3):5--7.

\bibitem[Jacobsen et~al.(2020)]{jacobsen2020fair}
Jacobsen, A., de~Miranda~Azevedo, R., Juty, N., et~al. (2020).
\newblock FAIR Principles: Interpretations and implementation considerations.
\newblock {\em Data Intelligence}, 2(1-2):10--29.

\bibitem[McConaghy et~al.(2020)]{mcconaghy2020ocean}
McConaghy, T., Marques, R., and M\"{u}ller, A. (2020).
\newblock Ocean Protocol: Tools for the Web3 data economy.
\newblock {\em Ocean Protocol Foundation Technical Report}.

\bibitem[Nelson and Ellis(2019)]{nelson2019digitization}
Nelson, G. and Ellis, S. (2019).
\newblock The history and impact of digitization and digital data mobilization on biodiversity research.
\newblock {\em Philosophical Transactions of the Royal Society B}, 374:20170391.

\bibitem[Protocol Labs(2017)]{protocol2017filecoin}
Protocol Labs (2017).
\newblock Filecoin: A decentralized storage network.
\newblock {\em Protocol Labs Technical Report}.

\bibitem[Ramachandran et~al.(2018)]{ramachandran2018trinity}
Ramachandran, A., Kantarcioglu, M., and Luo, B. (2018).
\newblock Using blockchain and smart contracts for secure data provenance management.
\newblock {\em arXiv preprint arXiv:1709.10000}.

\bibitem[Schultes et~al.(2019)]{schultes2019fair}
Schultes, E., Magagna, B., and Hettne, K.~M. (2019).
\newblock FAIR Digital Object Framework.
\newblock {\em Data Intelligence}, 2(1-2):56--72.

\bibitem[Wieczorek et~al.(2012)]{wieczorek2012darwin}
Wieczorek, J., Bloom, D., Guralnick, R., et~al. (2012).
\newblock Darwin Core: An evolving community-developed biodiversity data standard.
\newblock {\em PLOS ONE}, 7(1):e29715.

\bibitem[Wilkinson et~al.(2016)]{wilkinson2016fair}
Wilkinson, M.~D., Dumontier, M., Aalbersberg, I.~J., et~al. (2016).
\newblock The FAIR Guiding Principles for scientific data management and stewardship.
\newblock {\em Scientific Data}, 3:160018.

\bibitem[Yesson et~al.(2007)]{yesson2007how}
Yesson, C., Brewer, P.~W., and Sutton, T. (2007).
\newblock How global is the Global Biodiversity Information Facility?
\newblock {\em PLOS ONE}, 2(11):e1124.

\bibitem[Zach Kelling(2025a)]{zoo2025dso}
Zach Kelling (2025a).
\newblock Decentralized Semantic Optimization: Collaborative knowledge curation on distributed networks.
\newblock {\em Zoo Improvement Proposal ZIP-001}.

\bibitem[Zach Kelling(2025b)]{zoo2025poai}
Zach Kelling (2025b).
\newblock Proof of AI: Consensus mechanisms for verifiable machine learning.
\newblock {\em Zoo Improvement Proposal ZIP-002}.

\bibitem[Zach Kelling(2025c)]{zoo2025network}
Zach Kelling (2025c).
\newblock Zoo Network Architecture: A decentralized infrastructure for open science.
\newblock {\em Zoo Technical Report ZTR-001}.

\bibitem[Zach Kelling(2025d)]{zoo2025coord}
Zach Kelling (2025d).
\newblock Decentralized Research Coordination: Federated experiment management for DeSci.
\newblock {\em Zoo Improvement Proposal ZIP-008}.

\bibitem[Zach Kelling(2026)]{zoo2026educational}
Zach Kelling (2026).
\newblock Educational AI for Conservation: Adaptive learning systems.
\newblock {\em Zoo Improvement Proposal ZIP-007}.

\bibitem[Poisot et~al.(2019)]{poisot2019ecological}
Poisot, T., Bruneau, A., Gonzalez, A., et~al. (2019).
\newblock Ecological data should not be so hard to find and reuse.
\newblock {\em Trends in Ecology \& Evolution}, 34(6):494--496.

\bibitem[Feng et~al.(2019)]{feng2019review}
Feng, X., Park, D.~S., Walker, C., et~al. (2019).
\newblock A checklist for maximizing reproducibility of ecological niche models.
\newblock {\em Nature Ecology \& Evolution}, 3(10):1382--1395.

\end{thebibliography}

