\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% geometry loaded by zoocover
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{listings}

% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\TChain}{\textsc{T-Chain}}
\newcommand{\KChain}{\textsc{K-Chain}}
\newcommand{\TFHE}{\textsc{TFHE}}
\newcommand{\CKKS}{\textsc{CKKS}}
\newcommand{\FHE}{\textsc{FHE}}
\newcommand{\MEV}{\textsc{MEV}}
\newcommand{\AMM}{\textsc{AMM}}
\newcommand{\DEX}{\textsc{DEX}}
\newcommand{\KEEPER}{\textsc{Keeper}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=black
}

% Title information
\title{\textbf{Confidential DeFi: Private Trading, Sealed Auctions,\\and Encrypted Order Matching on Zoo Network}\\
\large Eliminating MEV Through Fully Homomorphic Encryption}
\author{
Antje Worring, Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{May 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
Decentralized finance leaks value at every layer. Front-running extracts \$1.5B+ annually from Ethereum traders. Sandwich attacks inflate slippage. Information asymmetry between informed and uninformed participants produces a market where the fastest extractor wins, not the best price-discoverer. We present \textbf{Confidential DeFi} on \Zoo{} Network: a suite of protocols where orders are encrypted at submission, matched homomorphically without decryption, and settled via threshold decryption of only the matched pairs. The core primitive is an \FHE{}-native \DEX{} virtual machine that evaluates encrypted limit orders using \TFHE{} boolean circuits for price comparison and \CKKS{} approximate arithmetic for volume matching. Sealed-bid auctions evaluate encrypted bids without revealing amounts until the auction concludes. A private \AMM{} processes encrypted swap amounts, exposing only aggregate pool state to liquidity providers. All decryption is coordinated by \Zoo{} \TChain{} with a 67-of-100 validator quorum, ensuring no single party can observe orderflow. We benchmark the system at 340 encrypted order matches per second on GPU-accelerated validators, with end-to-end latency of 2.8 seconds from submission to settlement---practical for a blockchain \DEX{} where block times are 2 seconds. Compared to existing solutions (Flashbots Protect, MEV-Share, Penumbra, Renegade), \Zoo{} Confidential DeFi provides stronger privacy guarantees (full orderflow encryption vs.\ partial hiding) with competitive throughput.

\textbf{Keywords}: confidential DeFi, FHE, MEV, encrypted order matching, sealed-bid auction, private AMM, threshold decryption
\end{abstract}

\tableofcontents
\newpage

%----------------------------------------------------------------------

\input{sections/01-introduction}
\input{sections/02-cryptographic-primitives}
\input{sections/03-encrypted-orderflow}
\input{sections/04-fhe-order-matching}
\input{sections/05-sealed-bid-auctions}
\input{sections/06-private-amm}
\input{sections/07-threshold-settlement}
\input{sections/08-comparison-to-existing-solutions}
\input{sections/09-benchmarks}
\input{sections/10-security-analysis}
\input{sections/99-conclusion}

\end{document}
