\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

\input{shared/lstlang}

% Listing style
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  captionpos=b,
  xleftmargin=2em,
  framexleftmargin=1.5em,
}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\ERC}{\textsc{erc}}
\newcommand{\ZOO}{\$\textsc{zoo}}
\newcommand{\ZRC}{\textsc{zrc}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% Title and authors
\title{\textbf{ZRC-721: Zoo Non-Fungible Token Standard with Yield Transferability and Conservation Semantics}}

\author{
Antje Worring, Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{April 2026}

\begin{document}

\zoocoverpage

\begin{abstract}
We present \textbf{ZRC-721}, the canonical non-fungible token standard for the Zoo L2 chain (Chain~ID: 200200). ZRC-721 is fully ERC-721 compatible and extends the interface with three Zoo-specific mechanisms rooted in the project's 2021--2022 origins: (1)~\textbf{yield transferability until burn}, where NFT yield and collateral transfer to new owners upon sale and are realized (claimed) when the animal is ``freed'' (burned)---a pattern from the January 2022 \$ZOO deck that predates ERC-6551 token-bound accounts by over a year; (2)~\textbf{generation-based rarity}, preserving the original 16 generation-one base animals from the October 2021 collection (later expanded to 20); and (3)~\textbf{conservation impact attribution}, where each NFT tracks cumulative conservation contributions as non-transferable metadata. The standard also supports cross-chain bridging via Lux Warp messaging with yield state preservation, continuing the ``play on any chain'' philosophy established by the February 2022 ZOO Bridge.

\textbf{Keywords}: non-fungible token, ERC-721, yield transferability, conservation, token-bound accounts, Zoo Network
\end{abstract}

\section{Introduction}

Zoo NFTs predate the current Zoo L2 chain. The original collection launched on BSC in Q4 2021, integrated with PancakeSwap for \$ZOO token liquidity and egg purchases~\cite{zoo-oct-2021-deck}. The October 2021 deck described the collection:

\begin{quote}
``Contains 1 of 16 Generation one Base Animals.''
\end{quote}

These 16 generation-one base animals formed the initial collection, later expanded to 20 in subsequent materials. Each NFT was an ERC-721 token on BSC with a distinctive property: \textbf{yield was transferable via NFT ownership}. The January 2022 deck stated~\cite{zoo-jan-2022-deck}:

\begin{quote}
``NFT yield and collateral is transferable to new owners until the Animal is freed from ZOO.''
\end{quote}

Here, ``freeing'' an animal from Zoo corresponds to burning the NFT, which realizes all accumulated yield and collateral. This yield-transferability-until-burn pattern was a significant innovation: it meant that Zoo NFTs were not merely collectibles but yield-bearing instruments whose economic value transferred with ownership. This is precisely the semantic pattern later formalized by ERC-6551 token-bound accounts (May 2023)~\cite{erc-6551}, making Zoo's 2022 implementation a functional predecessor by more than a year.

ZRC-721 formalizes this pattern as a standard for the Zoo L2 chain, preserving backward compatibility with ERC-721 while encoding yield transferability, generation-based rarity, and conservation semantics directly in the token interface.

\subsection{Historical Context}

\begin{table}[t]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Date} & \textbf{Milestone} & \textbf{Source} \\
\midrule
Oct 2021 & 16 generation-one base animals on BSC & \cite{zoo-oct-2021-deck} \\
Oct 2021 & PancakeSwap integration for \$ZOO + eggs & \cite{zoo-oct-2021-deck} \\
Oct 2021 & ``Bridge between ETH and Binance'' roadmapped & \cite{zoo-oct-2021-deck} \\
Jan 2022 & ``NFT yield transferable to new owners until freed'' & \cite{zoo-jan-2022-deck} \\
Feb 2022 & ZOO Bridge live: NFTs bridgeable across chains & \cite{zoo-jan-2022-deck} \\
May 2023 & ERC-6551 proposed (Zoo pattern predates by 16 months) & \cite{erc-6551} \\
2025 & Migration to Zoo L2 (chain ID 200200) & This paper \\
2026 & ZRC-721 standard formalized & This paper \\
\bottomrule
\end{tabular}
\caption{Zoo NFT chronology. The yield-transferability-until-burn pattern (January 2022) predates ERC-6551 by 16 months.}
\label{tab:chronology}
\end{table}

\subsection{Contributions}

\begin{enumerate}
    \item \textbf{Yield Transferability}: Formalization of the 2022 yield-transferability-until-burn pattern as a standard interface (Section~\ref{sec:yield}).

    \item \textbf{Generation System}: Encoding the original 16 generation-one base animals and the expansion mechanism (Section~\ref{sec:generation}).

    \item \textbf{Conservation Hooks}: Per-NFT impact attribution tracking (Section~\ref{sec:conservation}).

    \item \textbf{Bridge Compatibility}: Yield-preserving cross-chain transfers via Lux Warp (Section~\ref{sec:bridge}).

    \item \textbf{Formal Specification}: Complete Solidity interface and invariant analysis (Section~\ref{sec:spec}).
\end{enumerate}

\section{Yield Transferability Until Burn}
\label{sec:yield}

\subsection{The 2022 Pattern}

The original Zoo NFT mechanism on BSC implemented the following:

\begin{enumerate}
    \item Each NFT $N$ accrues yield $Y_N(t)$ from Zoo staking pools over time $t$.
    \item On transfer of $N$ from owner $A$ to owner $B$, the accumulated yield $Y_N(t)$ transfers with the NFT. Owner $B$ inherits all rights to the accumulated yield.
    \item Collateral locked against $N$ also transfers with ownership.
    \item When the animal is ``freed'' (burned), the owner at burn time receives all accumulated yield $Y_N(t)$ and all collateral.
    \item A freed animal cannot be transferred, bridged, or yield further.
\end{enumerate}

This is precisely a token-bound account: the NFT ``owns'' yield and collateral, which transfer with the token.

\subsection{ZRC-721 Yield Interface}

\begin{lstlisting}[language=Solidity,caption=ZRC-721 Yield Interface]
interface IZRC721Yield {
    /// @notice Returns accumulated yield for token
    /// @param tokenId The NFT to query
    /// @return amount Accumulated yield in ZOO wei
    function accumulatedYield(uint256 tokenId)
        external view returns (uint256 amount);

    /// @notice Returns collateral locked against token
    /// @param tokenId The NFT to query
    /// @return amount Locked collateral in ZOO wei
    function lockedCollateral(uint256 tokenId)
        external view returns (uint256 amount);

    /// @notice Free the animal: burn NFT, realize yield + collateral
    /// @param tokenId The NFT to free
    /// @dev Transfers accumulatedYield + lockedCollateral to msg.sender
    /// @dev Burns the NFT permanently
    function free(uint256 tokenId) external;

    /// @notice Check if animal has been freed
    /// @param tokenId The NFT to query
    /// @return True if freed (burned with yield realized)
    function isFreed(uint256 tokenId)
        external view returns (bool);

    /// @notice Emitted when yield is accumulated
    event YieldAccumulated(
        uint256 indexed tokenId,
        uint256 amount,
        uint256 totalAccumulated
    );

    /// @notice Emitted when animal is freed
    event AnimalFreed(
        uint256 indexed tokenId,
        address indexed owner,
        uint256 yieldRealized,
        uint256 collateralReleased
    );
}
\end{lstlisting}

\subsection{Yield Transfer on Sale}

When an ERC-721 \texttt{transferFrom} or \texttt{safeTransferFrom} is executed, the yield state is \emph{not reset}. The new owner inherits the full accumulated yield:

\begin{definition}[Yield Transfer Invariant]
For any ZRC-721 token $N$ with accumulated yield $Y_N$ and collateral $C_N$, a transfer from owner $A$ to owner $B$ preserves:
\begin{align}
Y_N^{\text{after}} &= Y_N^{\text{before}} \\
C_N^{\text{after}} &= C_N^{\text{before}}
\end{align}
The new owner $B$ has exclusive rights to $Y_N$ and $C_N$.
\end{definition}

\begin{theorem}[Yield Conservation]
For any sequence of transfers and yield accumulations, the yield realized on burn equals the total yield accumulated over the NFT's lifetime:
\[
Y_{\text{realized}} = \sum_{i=0}^{T} \Delta Y_i
\]
where $\Delta Y_i$ is yield accumulated in epoch $i$ and $T$ is the burn epoch.
\end{theorem}

\begin{proof}
Yield is monotonically non-decreasing (only accumulation, no withdrawal before burn) and transfer-invariant (no yield is lost or gained on transfer). The burn operation (``free'') is the only yield realization event. Therefore, realized yield equals cumulative accumulation.
\end{proof}

\section{Generation System}
\label{sec:generation}

\subsection{Original 16 Base Animals}

The October 2021 collection comprised 16 generation-one base animals:

\begin{lstlisting}[language=Solidity,caption=ZRC-721 Generation Interface]
interface IZRC721Generation {
    /// @notice Returns the generation of the token
    /// @param tokenId The NFT to query
    /// @return gen Generation number (1 = original base animals)
    function generation(uint256 tokenId)
        external view returns (uint256 gen);

    /// @notice Returns the species identifier
    /// @param tokenId The NFT to query
    /// @return species Species string (e.g., "base_animal_01")
    function species(uint256 tokenId)
        external view returns (string memory species);

    /// @notice Returns total species count for a generation
    /// @param gen Generation number
    /// @return count Number of species in generation
    function speciesCount(uint256 gen)
        external view returns (uint256 count);
}
\end{lstlisting}

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Generation} & \textbf{Base Animals} & \textbf{Source} \\
\midrule
Gen 1 (Oct 2021) & 16 & Original collection \\
Gen 1 expanded & 20 & Later materials \\
Gen 2+ & Variable & Governance-approved \\
\bottomrule
\end{tabular}
\caption{Zoo NFT generations. The original 16 base animals form the canonical Gen 1 collection.}
\label{tab:generations}
\end{table}

\subsection{Egg Mechanism}

The original BSC deployment used an egg mechanism integrated with PancakeSwap:

\begin{enumerate}
    \item User purchases eggs using \$ZOO via PancakeSwap
    \item Each egg hatches into one of the 16 generation-one base animals
    \item Hatching is random with equal probability across species
    \item The hatched NFT immediately begins accruing yield
\end{enumerate}

ZRC-721 preserves the egg mechanism as an optional minter extension.

\section{Conservation Impact Attribution}
\label{sec:conservation}

Each ZRC-721 NFT tracks cumulative conservation contributions:

\begin{lstlisting}[language=Solidity,caption=ZRC-721 Conservation Interface]
interface IZRC721Conservation {
    /// @notice Returns cumulative conservation impact
    /// @param tokenId The NFT to query
    /// @return amount Total conservation contribution in ZOO wei
    function conservationImpact(uint256 tokenId)
        external view returns (uint256 amount);

    /// @notice Record a conservation contribution
    /// @param tokenId The NFT crediting the contribution
    /// @param amount Contribution amount in ZOO wei
    /// @dev Only callable by verified conservation fund
    function recordImpact(uint256 tokenId, uint256 amount)
        external;

    /// @notice Emitted on conservation contribution
    event ConservationContribution(
        uint256 indexed tokenId,
        uint256 amount,
        uint256 totalImpact
    );
}
\end{lstlisting}

Conservation impact is \textbf{non-transferable}: it remains bound to the NFT (not the owner) and survives transfers. This creates a provenance record of the NFT's conservation history.

\section{Cross-Chain Bridge Compatibility}
\label{sec:bridge}

\subsection{Historical: ZOO Bridge (2022)}

The February 2022 ZOO Bridge~\cite{zoo-jan-2022-deck} was one of the earliest cross-chain NFT bridges, described as allowing ``each users to play ZOO on any chain.'' It listed 16 supported asset types spanning EVM and non-EVM chains. ZRC-721 is designed for bridge compatibility from the start, continuing this tradition.

\subsection{Warp Bridge Interface}

ZRC-721 tokens are bridgeable via Lux Warp messaging with yield state preservation:

\begin{lstlisting}[language=Solidity,caption=ZRC-721 Bridge Interface]
interface IZRC721Bridge {
    /// @notice Encode token for bridge transfer
    /// @param tokenId The NFT to bridge
    /// @return payload Warp-compatible payload including yield state
    function encodeBridgePayload(uint256 tokenId)
        external view returns (bytes memory payload);

    /// @notice Decode and mint bridged token
    /// @param payload Warp message payload
    /// @param recipient Destination owner
    /// @return tokenId Minted wrapped token ID
    function decodeBridgeMint(
        bytes calldata payload,
        address recipient
    ) external returns (uint256 tokenId);

    /// @notice Emitted on bridge lock
    event BridgeLocked(
        uint256 indexed tokenId,
        uint256 destChainId,
        bytes payload
    );

    /// @notice Emitted on bridge mint
    event BridgeMinted(
        uint256 indexed tokenId,
        uint256 srcChainId,
        address indexed recipient
    );
}
\end{lstlisting}

The bridge payload includes the full yield state, generation, species, and conservation impact, ensuring that bridged NFTs retain all ZRC-721 semantics on the destination chain.

\subsection{Yield Preservation Across Chains}

\begin{definition}[Bridge Yield Invariant]
For a ZRC-721 token $N$ bridged from chain $A$ to chain $B$:
\begin{enumerate}
    \item $N$ is locked in escrow on chain $A$; wrapped $N'$ minted on chain $B$
    \item $Y_{N'} = Y_N$ at time of bridge (yield state preserved)
    \item $N'$ continues accruing yield on chain $B$
    \item Freeing $N'$ on chain $B$ (burn) triggers yield realization and releases $N$ on chain $A$
    \item Total yield: $Y_N + \Delta Y_{N'}$ (no yield lost)
\end{enumerate}
\end{definition}

This directly implements the 2022 deck's guarantee: yield is transferable (including across chains) until the animal is freed.

\section{Formal Specification}
\label{sec:spec}

\subsection{Complete Interface}

\begin{lstlisting}[language=Solidity,caption=ZRC-721 Complete Interface]
interface IZRC721 is
    IERC721,
    IERC721Metadata,
    IERC721Enumerable,
    IZRC721Yield,
    IZRC721Generation,
    IZRC721Conservation,
    IZRC721Bridge
{
    /// @notice Returns the ZRC-721 standard version
    function zrcVersion() external pure returns (string memory);
}
\end{lstlisting}

\subsection{Invariants}

\begin{enumerate}
    \item \textbf{Yield monotonicity}: $Y_N(t+1) \geq Y_N(t)$ for all tokens $N$ and times $t < t_{\text{burn}}$.
    \item \textbf{Transfer preservation}: $Y_N$ and $C_N$ are unchanged by \texttt{transferFrom}.
    \item \textbf{Burn finality}: After \texttt{free(tokenId)}, the token cannot be transferred, bridged, or yield further.
    \item \textbf{Conservation permanence}: Conservation impact is non-transferable and monotonically non-decreasing.
    \item \textbf{Generation immutability}: A token's generation and species cannot be changed after minting.
\end{enumerate}

\section{Relation to ERC-6551}

ERC-6551~\cite{erc-6551} (proposed May 2023) creates token-bound accounts: smart contract accounts owned by NFTs. Zoo's yield-transferability-until-burn pattern (January 2022) achieves equivalent semantics through a different mechanism:

\begin{table}[t]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Property} & \textbf{Zoo NFTs (Jan 2022)} & \textbf{ERC-6551 (May 2023)} \\
\midrule
NFT owns assets & Yield + collateral & Arbitrary tokens \\
Transfer semantics & Assets move with NFT & Account moves with NFT \\
Realization & Burn (``free the animal'') & Direct withdrawal \\
Composability & Yield-specific & General-purpose \\
Timeline & January 2022 & May 2023 \\
\bottomrule
\end{tabular}
\caption{Zoo's yield-transferability pattern vs. ERC-6551. Zoo's implementation predates ERC-6551 by 16 months.}
\label{tab:erc6551}
\end{table}

ZRC-721 acknowledges this lineage by making yield transferability a first-class interface rather than an add-on extension.

\section{Related Work}

\textbf{Zoo Oct 2021 Deck}: Original collection of 16 generation-one base animals on BSC with PancakeSwap integration~\cite{zoo-oct-2021-deck}. The Q4 roadmap included ``Bridge between ETH and Binance.''

\textbf{Zoo Jan 2022 Deck}: Defined yield transferability: ``NFT yield and collateral is transferable to new owners until the Animal is freed from ZOO''~\cite{zoo-jan-2022-deck}. Confirmed February 2022 ZOO Bridge launch.

\textbf{ERC-721}: The base non-fungible token standard~\cite{erc-721}. ZRC-721 is fully compatible.

\textbf{ERC-6551}: Token-bound accounts~\cite{erc-6551}. Zoo's yield-transferability-until-burn pattern is a functional predecessor.

\textbf{ERC-4907}: Rental NFTs with time-limited usage rights. ZRC-721's yield mechanism is orthogonal: yield accrues to the owner (not a renter), and realization requires burn.

\section{Conclusion}

ZRC-721 formalizes the non-fungible token standard for Zoo L2, encoding three mechanisms with roots in Zoo's 2021--2022 origins: yield transferability until burn (from the January 2022 \$ZOO deck, predating ERC-6551 by 16 months), generation-based rarity (preserving the original 16 generation-one base animals from October 2021), and conservation impact attribution. The standard is fully ERC-721 compatible, bridge-ready via Lux Warp messaging (continuing the ``play on any chain'' philosophy of the February 2022 ZOO Bridge), and designed to preserve the yield-transferability-until-burn pattern that made Zoo NFTs yield-bearing instruments before the concept was standardized.

\section*{Acknowledgments}

This work was supported by Zoo Labs Foundation.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{zoo-oct-2021-deck}
Zoo Labs, ``\$ZOO Token and NFT Ecosystem,'' Pitch Deck, October 2021. ``Contains 1 of 16 Generation one Base Animals.'' Q4 roadmap: ``Bridge between ETH and Binance.''

\bibitem{zoo-jan-2022-deck}
Zoo Labs, ``\$ZOO Ecosystem Update,'' Pitch Deck, January 2022. ``NFT yield and collateral is transferable to new owners until the Animal is freed from ZOO.'' ZOO Bridge: ``Our bridge allows each users to play ZOO on any chain.''

\bibitem{erc-721}
W. Entriken et al., ``EIP-721: Non-Fungible Token Standard,'' Ethereum Improvement Proposal, January 2018.

\bibitem{erc-6551}
J. Chung et al., ``ERC-6551: Non-fungible Token Bound Accounts,'' Ethereum Improvement Proposal, May 2023.

\bibitem{lux-warp-messaging}
Lux Industries, ``Warp Messaging: Verified Cross-Chain Communication on Lux Network,'' Technical Report, 2025.

\end{thebibliography}

\end{document}
