\section*{Acknowledgments}


This work was made possible by:
\begin{itemize}
\item 87,341 airdrop recipients who believed in fair launch
\item 8,472 contributors who generated 12.4M experiences in 6 weeks
\item Zoo Labs Foundation board for 501(c)(3) legal structure
\item Lux and Hanzo teams for infrastructure foundation
\item Broader crypto community for open-source tools (Ethereum, IPFS, Rust)
\end{itemize}

\begin{thebibliography}{99}

\bibitem{messari2024ai}
Messari Research.
\textit{The State of AI x Crypto: 2024 Report}.
Messari, 2024.

\bibitem{coindesk2024render}
CoinDesk.
\textit{Render Network Raises \$150M in Series B Funding}.
CoinDesk, 2024.

\bibitem{buterin2018quadratic}
Vitalik Buterin, Zoë Hitzig, and E. Glen Weyl.
\textit{A Flexible Design for Funding Public Goods}.
Management Science, 2019.

\bibitem{shapley1953value}
Lloyd S. Shapley.
\textit{A Value for n-Person Games}.
Contributions to the Theory of Games, 1953.

\bibitem{nakamoto2008bitcoin}
Satoshi Nakamoto.
\textit{Bitcoin: A Peer-to-Peer Electronic Cash System}.
bitcoin.org, 2008.

\bibitem{wood2014ethereum}
Gavin Wood.
\textit{Ethereum: A Secure Decentralised Generalised Transaction Ledger}.
Ethereum Project Yellow Paper, 2014.

\bibitem{king2012ppcoin}
Sunny King and Scott Nadal.
\textit{PPCoin: Peer-to-Peer Crypto-Currency with Proof-of-Stake}.
ppcoin.org, 2012.

\bibitem{vasin2014blackcoin}
Pavel Vasin.
\textit{BlackCoin's Proof-of-Stake Protocol v2}.
blackcoin.org, 2014.

\bibitem{larimer2014delegated}
Daniel Larimer.
\textit{Delegated Proof-of-Stake Consensus}.
BitShares, 2014.

\bibitem{kwon2019cosmos}
Jae Kwon and Ethan Buchman.
\textit{Cosmos: A Network of Distributed Ledgers}.
Cosmos Whitepaper, 2019.

\bibitem{gilad2017algorand}
Yossi Gilad et al.
\textit{Algorand: Scaling Byzantine Agreements for Cryptocurrencies}.
ACM Symposium on Operating Systems Principles, 2017.

\bibitem{chen2016algorand}
Jing Chen and Silvio Micali.
\textit{Algorand: A Secure and Efficient Distributed Ledger}.
arXiv preprint arXiv:1607.01341, 2016.

\bibitem{david2018ouroboros}
Bernardo David et al.
\textit{Ouroboros Praos: An Adaptively-Secure, Semi-synchronous Proof-of-Stake Blockchain}.
Annual International Conference on the Theory and Applications of Cryptographic Techniques, 2018.

\bibitem{badertscher2018ouroboros}
Christian Badertscher et al.
\textit{Ouroboros Genesis: Composable Proof-of-Stake Blockchains with Dynamic Availability}.
ACM SIGSAC Conference on Computer and Communications Security, 2018.

\bibitem{rocket2018snowflake}
Team Rocket et al.
\textit{Snowflake to Avalanche: A Novel Metastable Consensus Protocol Family}.
IPFS, 2018.

\bibitem{yin2019hotstuff}
Maofan Yin et al.
\textit{HotStuff: BFT Consensus in the Lens of Blockchain}.
arXiv preprint arXiv:1803.05069, 2019.

\bibitem{castro1999practical}
Miguel Castro and Barbara Liskov.
\textit{Practical Byzantine Fault Tolerance}.
OSDI, 1999.

\bibitem{buchman2016tendermint}
Ethan Buchman.
\textit{Tendermint: Byzantine Fault Tolerance in the Age of Blockchains}.
Master's Thesis, University of Guelph, 2016.

\bibitem{yakovenko2018solana}
Anatoly Yakovenko.
\textit{Solana: A New Architecture for a High Performance Blockchain}.
Solana Whitepaper, 2018.

\bibitem{gini1912variability}
Corrado Gini.
\textit{Variability and Mutability}.
Memorie di metodologica statistica, 1912.

\bibitem{lorenz1905methods}
Max O. Lorenz.
\textit{Methods of Measuring the Concentration of Wealth}.
Publications of the American Statistical Association, 1905.

\bibitem{arrow1951social}
Kenneth J. Arrow.
\textit{Social Choice and Individual Values}.
Wiley, 1951.

\bibitem{ostrom1990governing}
Elinor Ostrom.
\textit{Governing the Commons: The Evolution of Institutions for Collective Action}.
Cambridge University Press, 1990.

\bibitem{buterin2014daos}
Vitalik Buterin.
\textit{DAOs, DACs, DAs and More: An Incomplete Terminology Guide}.
Ethereum Blog, 2014.

\bibitem{hansmann1996ownership}
Henry Hansmann.
\textit{The Ownership of Enterprise}.
Harvard University Press, 1996.

\bibitem{zohar2015bitcoin}
Aviv Zohar.
\textit{Bitcoin: Under the Hood}.
Communications of the ACM, 2015.

\end{thebibliography}

\appendix

