\section*{Acknowledgments}


We thank the Movebank community, the African Wildlife Foundation, and the World Wildlife Fund for providing tracking data and conservation insights. This research was supported by the Zoo Labs Foundation.

\begin{thebibliography}{30}

\bibitem{kays2015}
R.~Kays, M.~C.~Crofoot, W.~Jetz, and M.~Wikelski, ``Terrestrial animal tracking as an eye on life and planet,'' \emph{Science}, vol.~348, no.~6240, p.~aaa2478, 2015.

\bibitem{movebank2024}
Movebank, ``Movebank: Archive, manage, share animal movement data,'' 2024. [Online]. Available: \url{https://www.movebank.org}

\bibitem{wri2023}
Wildlife Risk Intelligence, ``Tracking Data Security Incident Report 2022-047,'' WRI Technical Report, 2023.

\bibitem{hebblewhite2022}
M.~Hebblewhite, ``Public access to wildlife tracking data: balancing conservation with science,'' \emph{Animal Biotelemetry}, vol.~10, p.~14, 2022.

\bibitem{eurotrack2020}
Eurotrack Consortium, ``Animal Tracking Data Collection: A European Federated Infrastructure,'' 2020.

\bibitem{wikelski2007}
M.~Wikelski, R.~W.~Kays, N.~J.~Kasdin, et al., ``Going wild: what a global small-animal tracking system could do for experimental biologists,'' \emph{Journal of Experimental Biology}, vol.~210, no.~2, pp.~181--186, 2007.

\bibitem{gruteser2003}
M.~Gruteser and D.~Grunwald, ``Anonymous usage of location-based services through spatial and temporal cloaking,'' in \emph{MobiSys}, 2003, pp.~31--42.

\bibitem{sweeney2002}
L.~Sweeney, ``$k$-anonymity: A model for protecting privacy,'' \emph{International Journal of Uncertainty, Fuzziness and Knowledge-Based Systems}, vol.~10, no.~5, pp.~557--570, 2002.

\bibitem{dwork2006}
C.~Dwork, F.~McSherry, K.~Nissim, and A.~Smith, ``Calibrating noise to sensitivity in private data analysis,'' in \emph{TCC}, 2006, pp.~265--284.

\bibitem{andres2013}
M.~E.~Andr{\'e}s, N.~E.~Bordenabe, K.~Chatzikokolakis, and C.~Palamidessi, ``Geo-indistinguishability: differential privacy for location-based systems,'' in \emph{CCS}, 2013, pp.~901--914.

\bibitem{maxwell2021}
S.~L.~Maxwell, R.~A.~Fuller, T.~M.~Brooks, and J.~E.~M.~Watson, ``Biodiversity: the ravages of guns, nets and bulldozers,'' \emph{Nature}, vol.~536, pp.~143--145, 2016.

\bibitem{kshetri2021}
N.~Kshetri, ``Blockchain and sustainable supply chain management in developing countries,'' \emph{International Journal of Information Management}, vol.~60, p.~102376, 2021.

\bibitem{howson2019}
P.~Howson, ``Tackling climate change with blockchain,'' \emph{Nature Climate Change}, vol.~9, pp.~644--645, 2019.

\bibitem{wildchain2023}
WildChain Consortium, ``WildChain: Blockchain Framework for Conservation Data Management,'' Technical Report, 2023.

\bibitem{groth2016}
J.~Groth, ``On the Size of Pairing-Based Non-interactive Arguments,'' in \emph{EUROCRYPT}, 2016, pp.~305--326.

\bibitem{bunz2018}
B.~B{\"u}nz, J.~Bootle, D.~Boneh, A.~Poelstra, P.~Wuille, and G.~Maxwell, ``Bulletproofs: Short Proofs for Confidential Transactions and More,'' in \emph{IEEE S\&P}, 2018.

\bibitem{damgard2012}
I.~Damg{\aa}rd, V.~Pastro, N.~P.~Smart, and S.~Zakarias, ``Multiparty computation from somewhat homomorphic encryption,'' in \emph{CRYPTO}, 2012, pp.~643--662.

\bibitem{mcrae2008}
B.~H.~McRae, B.~G.~Dickson, T.~H.~Keitt, and V.~B.~Shah, ``Using circuit theory to model connectivity in ecology,'' \emph{Ecology}, vol.~89, no.~10, pp.~2712--2724, 2008.

\bibitem{worton1989}
B.~J.~Worton, ``Kernel methods for estimating the utilization distribution in home-range studies,'' \emph{Ecology}, vol.~70, no.~1, pp.~164--168, 1989.

\bibitem{johnson2008}
D.~S.~Johnson, J.~M.~London, M.-A.~Lea, and J.~W.~Durban, ``Continuous-time correlated random walk model for animal telemetry data,'' \emph{Ecology}, vol.~89, no.~5, pp.~1208--1215, 2008.

\bibitem{patterson2008}
T.~A.~Patterson, L.~Thomas, C.~Wilcox, O.~Ovaskainen, and J.~Matthiopoulos, ``State-space models of individual animal movement,'' \emph{Trends in Ecology \& Evolution}, vol.~23, no.~2, pp.~87--94, 2008.

\bibitem{nathan2008}
R.~Nathan, W.~M.~Getz, E.~Revilla, et al., ``A movement ecology paradigm for unifying organismal movement research,'' \emph{PNAS}, vol.~105, no.~49, pp.~19052--19059, 2008.

\bibitem{fleming2015}
C.~H.~Fleming, J.~M.~Calabrese, T.~Mueller, et al., ``From Fine-Scale Foraging to Home Ranges,'' \emph{The American Naturalist}, vol.~185, no.~5, pp.~E154--E167, 2015.

\bibitem{boyce2002}
M.~S.~Boyce, P.~R.~Vernier, S.~E.~Nielsen, and F.~K.~A.~Schmiegelow, ``Evaluating resource selection functions,'' \emph{Ecological Modelling}, vol.~157, no.~2-3, pp.~281--300, 2002.

\bibitem{manly2002}
B.~F.~J.~Manly, L.~L.~McDonald, D.~L.~Thomas, T.~L.~McDonald, and W.~P.~Erickson, \emph{Resource Selection by Animals}, 2nd ed., Kluwer Academic, 2002.

\bibitem{zoo2026dao}
Zach Kelling, ``Zoo DAO: Decentralized Governance for Open Science Research Coordination,'' ZOO-2026-007, 2026.

\bibitem{bonneau2015}
J.~Bonneau, A.~Miller, J.~Clark, A.~Narayanan, J.~A.~Kroll, and E.~W.~Felten, ``SoK: Research perspectives and challenges for bitcoin and cryptocurrencies,'' in \emph{IEEE S\&P}, 2015.

\bibitem{kranstauber2012}
B.~Kranstauber, R.~Kays, S.~D.~LaPoint, M.~Wikelski, and K.~Safi, ``A dynamic Brownian bridge movement model,'' \emph{Journal of Animal Ecology}, vol.~81, no.~4, pp.~738--746, 2012.

\end{thebibliography}

