\section{Conclusion}

\label{sec:conclusion}

\cevm{} achieves 2.3~Gops/sec on a single GPU---$23.8\times$ over the fastest CPU EVM---while preserving full determinism through conflict detection and sequential re-execution.
Lean~4 verification provides mathematical guarantees that parallel execution produces identical state roots to sequential execution.
Multi-GPU linear scaling to 8.9~Gops/sec on four H100s meets the throughput demands of Zoo's AI-native L2 workloads.

Zoo Network's integration of \cevm{} with DEX VM, FHE VM, and AI Mining VM establishes a template for GPU-accelerated blockchain execution that maintains EVM compatibility while unlocking throughput previously exclusive to centralized systems.

Future work includes intra-precompile parallelism for DEX and FHE VMs, predictive conflict analysis via ML \cite{zoo-tokenomics}, ZK proof co-scheduling on GPU \cite{zoo-fhe}, and cross-L2 parallel execution across \Lux{} subnets \cite{zoo-dex}.

%% ---------------------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{wood2014ethereum}
G. Wood, ``Ethereum: A secure decentralised generalised transaction ledger,'' Ethereum Project Yellow Paper, 2014.

\bibitem{zoo-network-architecture}
Z. Kelling, ``Zoo Network: Decentralized AI training and inference layer with semantic optimization,'' Zoo Labs Foundation, 2025.

\bibitem{zoo-tokenomics}
Z. Kelling, ``Zoo Network tokenomics: Fair distribution through complete airdrop,'' Zoo Labs Foundation, 2025.

\bibitem{lux-gpu-evm}
Lux Partners, ``GPU-accelerated EVM execution for high-throughput blockchain networks,'' Lux Network Whitepaper Series, 2025.

\bibitem{monad2024}
Monad Labs, ``Monad: A high-performance EVM-compatible L1 with parallel execution,'' 2024.

\bibitem{sei2024}
Sei Labs, ``Sei v2: The first parallelized EVM blockchain,'' 2024.

\bibitem{megaeth2024}
MegaETH, ``MegaETH: Real-time blockchain with sub-millisecond latency,'' 2024.

\bibitem{solana2020}
A. Yakovenko, ``Solana: A new architecture for a high performance blockchain,'' 2020.

\bibitem{kevm2018}
E. Hildenbrandt et al., ``KEVM: A complete formal semantics of the Ethereum Virtual Machine,'' IEEE CSF, 2018.

\bibitem{zoo-fhe}
Z. Kelling, ``Fully homomorphic encryption on Zoo Network,'' Zoo Labs Foundation, forthcoming.

\bibitem{zoo-dex}
Z. Kelling, ``Decentralized exchange architecture on Zoo L2,'' Zoo Labs Foundation, forthcoming.

\end{thebibliography}

