\section{The Cute \AI{} Assistant}
\label{sec:cute-ai}

\subsection{2021 Commitment}

\S21 of the founder whitepaper described an \AI{} assistant fused
with each Zoo \NFT{}: emotionally attuned, conversational, capable of
recognizing user context and personalizing responses, and ``ready to
cheer you up.'' Each species would have a distinct personality
profile.

\subsection{2025-12-15 Realization}

Ships as the on-device avatar within \textbf{Zoo Desktop}
(Tauri + React, \texttt{\textasciitilde/work/zoo/app/}). Backed by the
\Zen{} model family co-developed with \Hanzo{} \AI{}: \texttt{eco-1},
\texttt{coder-1}, \texttt{nano-1} (see \textsc{zip-0010}). Personality
is realized through:

\begin{itemize}[nosep,leftmargin=1.2em]
\item Per-species system prompt and persona file shipped with each
  \NFT{} (deterministic; on-chain hash committed under
  \textsc{zip-0204}).
\item Persistent semantic memory backed by the Experience Ledger
  (\textsc{zip-0401}, content-addressable semantic memory).
\item Real-time conversational UX (\textsc{zip-0417}).
\item Multimodal conservation knowledge (\textsc{zip-0406}).
\item Local-first inference on the user's device (Hanzo Network
  GPU/CPU), with fallback to a Hanzo cluster when local resources are
  insufficient.
\end{itemize}

\subsection{Personality Profiles (2021 Catalogue)}

The eight V1 species each have a distinct, named personality
articulated in the 2021 paper \S21.1 and instantiated in
2025-12-15:

\begin{table}[h]
\centering
\small
\begin{tabular}{p{4cm}p{10cm}}
\toprule
\textbf{Species} & \textbf{Persona (canonical, 2021)} \\
\midrule
Red Wolf & Loyal pack-mate; encourages teamwork. \\
Amur Leopard & Reserved scholar; surfaces obscure facts. \\
Nubian Giraffe & Optimistic horizon-scout; long-term planner. \\
Sumatran Elephant & Memory-keeper; reminds the user of past
  interactions. \\
Pygmy Hippo & Shy comedian; light comfort and small talk. \\
Javan Rhinoceros & Gentle giant; encourages slow deep work. \\
Siberian Tiger & Focused mentor; pushes mastery and discipline. \\
Emperor Penguin & Stoic companion; weather-aware, calm. \\
\bottomrule
\end{tabular}
\caption{Canonical V1 personalities, instantiated in 2025-12-15 as
prompt + behaviour files.}
\end{table}

\subsection{Architecture}

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{Frontend.} Zoo Desktop (Tauri 2 + React 19), with the
  avatar rendered via the in-app 3D scene; voice synthesis local-only.
\item \textbf{Inference.} \Zen{} models served by Hanzo Network on a
  per-\textsc{llm} chain (one chain per model, each model is a
  sovereign \Quasar{} appchain --- see \cite{zoo-perllm-chains}).
\item \textbf{Memory.} Experience Ledger (content-addressable;
  semantic-vector indexed); BitDelta personalization
  (\textsc{zip-0007}); DeltaSoup community merging
  (\textsc{zip-0009}).
\item \textbf{Identity.} Avatar is bound to the user's Zoo DID
  (\textsc{zip-0033}); on-device key custody.
\item \textbf{Receipts.} Every inference call emits a TEE-attested
  receipt anchored on A-Chain (\textsc{lp-134}, Hanzo \AI{} Chain).
\end{itemize}

\subsection{Pre-ChatGPT Provenance}

\textbf{Notable historical fact}: \S21 of the October 31, 2021
whitepaper described a \textsc{gui}-based conversational \AI{} avatar
\emph{thirteen months before} \textsc{chatgpt}'s public launch
(November 30, 2022). Zoo's avatar \AI{} concept --- including persistent
semantic memory and personality binding --- predates the mainstream
\textsc{llm}-chat paradigm. The Experience Ledger
(\texttt{papers/zoo-experience-ledger/}) predates mainstream vector
databases and \textsc{chatgpt}'s native conversation memory feature
(February 2024).
