\section{Augmented Reality App (Zoo Field)}
\label{sec:ar}

\subsection{2021 Commitment}

\S22 described an \AR{} app that lets the user encounter their Zoo
\NFT{} animal in real-world space, supporting in-the-field engagement
and conservation education.

\subsection{2025-12-15 Realization}

Ships as \textbf{Zoo Field}, a mobile \AR{} companion published under
\texttt{\textasciitilde/work/zoo/apps/}. Anchored to:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{Geo-attested quest completion.} On-site visits to
  partnered sanctuaries and conservation sites complete quests under
  \textsc{zip-0034} (XP \& Quests) with \texttt{ATTESTATION}
  verification.
\item \textbf{Species-ID classification.} On-device species
  classification (\textsc{zip-0510}, Species Protection Monitoring)
  recognizes real animals and rewards the user for verified
  observations.
\item \textbf{Habitat overlays.} Visualizes habitat-restoration data
  fed from \textsc{zip-0505} (Reforestation Verification) and
  \textsc{zip-0520} (Habitat Conservation).
\item \textbf{Citizen-science contribution.} Submitted observations
  flow into the Citizen Science Contribution Protocol
  (\textsc{zip-0602}) and are eligible for impact-attestation.
\end{itemize}

\subsection{Architecture}

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{Stack.} Native iOS (\textsc{arkit}) and Android
  (\textsc{arcore}) with shared Rust core for protocol bindings.
\item \textbf{Inference.} Local on-device for species classification
  and pose estimation; offloaded to Hanzo Network for harder cases.
\item \textbf{Identity.} DID-bound (\textsc{zip-0033}); offline-capable
  with deferred attestation submission.
\item \textbf{Privacy.} Geo data is selectively disclosed on Z-Chain
  (\textsc{lp-134}); raw \textsc{gps} is never written publicly.
\end{itemize}

\subsection{Field-Trip Mode}

A guided mode for sanctuary visits (Foundation events). The visitor's
DID accrues XP and unlocks limited-edition wildlife photography \NFT{}s
(\textsc{zip-0205}).
