\section{Metaverse Companion}
\label{sec:metaverse}

\subsection{2021 Commitment}

\S23 described a metaverse companion that follows the user across
surfaces, persists state, and integrates with games (mini-games,
Coin Runner, Memory Bridge --- \S24).

\subsection{2025-12-15 Realization}

The companion is the persistent avatar binding the user's Zoo DID
(\textsc{zip-0033}) to:

\begin{itemize}[nosep,leftmargin=1.2em]
\item Wallet (Lux Wallet, white-label per \textsc{lp-134}).
\item \NFT{} portfolio (Zoo \NFT{} Liquidity Protocol,
  \texttt{contracts/specs/zoo-nft-liquidity-protocol.md}).
\item Chat (Zoo Desktop avatar, \S\ref{sec:cute-ai}).
\item \AR{} (Zoo Field, \S\ref{sec:ar}).
\item Virtual Habitat Simulation (\textsc{zip-0300}).
\item Multiplayer Conservation Game (\textsc{zip-0303}).
\item Spatial Web agents (\textsc{zip-0011}, \textsc{zip-0433}).
\end{itemize}

\subsection{State Continuity}

The companion's state (XP, KEEPER, attestations, semantic memory) is
canonically held under the user's Zoo DID. Every surface reads from
A-Chain attestation; writes back through Hanzo \AI{} Chain receipts.
Cross-surface continuity is therefore a function of identity, not a
function of any single platform.

\subsection{Game-Asset Interoperability}

\textsc{zip-0304} (Game Asset Interoperability) defines how items move
across mini-games. \textsc{zip-0210} (Cross-Collection Composability)
defines how Zoo \NFT{}s compose with non-Zoo collections through
attested trait inheritance.

\subsection{Why Holographic}

The companion realizes the ``holographic'' axis named in the 2021
paper's 2025 update (\S28): every sub-DAO, per-\textsc{llm} chain, and
regional chapter inherits the same governance topology and the same
identity primitive. Whether the user is on Desktop, Field (\AR{}), the
DEX, or a sub-DAO sub-domain, the same DID drives the same companion.
That is the engineering meaning of ``holographic'' in 2025-12-15
\Zoo{}: a single template, replicated across every surface, anchored
on the same chain.

\subsection{Architecture Diagram (Logical)}

\begin{verbatim}
            ┌──────────────┐
            │   Zoo DID    │  (ZIP-0033, A-Chain attested)
            └──────┬───────┘
                   │
       ┌───────────┴───────────┐
       │                       │
  ┌────▼──────┐         ┌──────▼─────┐
  │ Companion │ <------>│ Experience │ (Ledger; ZIP-0401)
  │  State    │         │  Memory    │
  └────┬──────┘         └────────────┘
       │
  ┌────┴──────────────────────────┐
  │  Surfaces                     │
  │  - Zoo Desktop (avatar chat)  │
  │  - Zoo Field (AR)             │
  │  - Zoo Wallet                 │
  │  - Zoo NFT Marketplace        │
  │  - Sub-DAOs (per-LLM chains)  │
  └───────────────────────────────┘
\end{verbatim}
