\section{Roadmap (Surface-Specific)}
\label{sec:roadmap}

\subsection{Shipped (as of 2025-12-15)}

\begin{itemize}[nosep,leftmargin=1.2em]
\item Zoo Desktop avatar with \Zen{} models, Experience Ledger
  memory, and DID binding.
\item Per-\textsc{llm} chain provisioning for the eight V1 personality
  profiles.
\item Zoo Field \AR{} v1 (iOS / Android) with species ID and
  geo-attested quest completion.
\end{itemize}

\subsection{2025-12-25 Activation}

\begin{itemize}[nosep,leftmargin=1.2em]
\item Zoo L1 mainnet cutover: companion state migrates from staging to
  mainnet.
\item DID issuance opens for all existing 2021--2024 holders.
\item \textsc{zip-0034} Quest Registry seeded with the Founders'
  Quests.
\end{itemize}

\subsection{Post-Mainnet (after 2025-12-25)}

\begin{itemize}[nosep,leftmargin=1.2em]
\item Zoo Field \AR{} v2: visual SLAM, multi-species scene, deeper
  habitat overlays.
\item Multiplayer Conservation Game (\textsc{zip-0303}) public beta.
\item Coin Runner mini-game (canonical 2021 design, \S24.2.1).
\item Memory Bridge mini-game (canonical 2021 design, \S24.2.2).
\end{itemize}

\subsection{Cross-References}

\begin{itemize}[nosep,leftmargin=1.2em]
\item 2021 Whitepaper \S21 (Cute AI), \S22 (AR), \S23 (Metaverse),
  \S24 (Roadmap).
\item Companion paper: \emph{Zoo Per-LLM Chains}.
\item Companion paper: \emph{Zoo Experience Ledger}.
\item Companion paper: \emph{Zoo Spatial Web Agents}.
\item Lux \textsc{lp-132} (QuasarGPU), \textsc{lp-134} (chain
  topology), \textsc{lp-137} (GPU residency invariant).
\end{itemize}

\begin{thebibliography}{9}
\bibitem{zoo-perllm-chains} Zoo Labs Foundation, \emph{Zoo Per-LLM
Chains: One Sovereign Chain per Model}, 2025.
\end{thebibliography}
