\documentclass[twocolumn,11pt]{article}
\usepackage{shared/zoocover}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[numbers]{natbib}
\usepackage{abstract}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
}

\title{\textbf{ZIP-0803: Encrypted SQLite Replication for Zoo Services}}

\author{
Zach Kelling\thanks{zach@lux.network}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
We specify a sidecar-based replication protocol for Zoo Network services
that use Hanzo Base (per-org encrypted SQLite).  The system streams
Write-Ahead Log (WAL) frames through \texttt{hanzoai/replicate} to
S3-compatible object storage, encrypts at-rest with \texttt{hanzoai/age}
(X25519 + ChaCha20-Poly1305), and achieves $\leq 1$\,s Recovery Point
Objective (RPO) with seconds-scale Recovery Time Objective (RTO).
Each pod replicates to a distinct S3 prefix keyed by pod identity,
enabling StatefulSet-aware restore.  The design reuses the same
\texttt{ghcr.io/hanzoai/replicate:latest} image deployed across
Liquidity and Lux ecosystems, requiring zero Zoo-specific code.
\end{abstract}

\section{Motivation}

Zoo Network services are migrating from PostgreSQL to Hanzo Base
(per-org encrypted SQLite) following the Web5 local-first architecture
described in the Zoo research program~\cite{web5}.  SQLite provides
sub-microsecond read latency, per-tenant data isolation via
deterministic key derivation, and offline-first operation.

However, SQLite's single-writer model means each service instance
owns its database file.  Without replication:

\begin{enumerate}
\item A pod restart loses all data accumulated since the last
  PersistentVolumeClaim (PVC) snapshot.
\item Disaster recovery requires manual PVC restoration from
  cloud provider snapshots (minutes to hours).
\item Cross-region failover is impossible without an external
  replication mechanism.
\end{enumerate}

This proposal introduces continuous WAL streaming to S3 as the
replication primitive, providing:

\begin{itemize}
\item $\leq 1$\,s RPO via 1-second sync intervals
\item Seconds-scale RTO via init-container restore
\item At-rest encryption via \texttt{hanzoai/age}
\item Zero application code changes (sidecar pattern)
\end{itemize}

\section{Architecture}

\subsection{Components}

\begin{description}
\item[\texttt{hanzoai/replicate}] Litestream-based WAL streaming
  daemon.  Watches SQLite WAL, ships frames to S3.  Runs as a
  sidecar container sharing the data volume with the primary
  service container.

\item[\texttt{hanzoai/age}] X25519 + ChaCha20-Poly1305 file
  encryption.  Used for encrypting WAL snapshots and segments
  at rest in S3.  Compatible with \texttt{luxfi/age} (same
  wire format).

\item[\texttt{hanzoai/s3}] S3-compatible object storage.  Deployed
  as a ClusterIP service within the Zoo K8s cluster
  (\texttt{s3.zoo.svc.cluster.local:9000}).  Backed by
  DigitalOcean Spaces or MinIO depending on environment.

\item[ConfigMap] Shared \texttt{replicate-config} ConfigMap
  containing the Litestream YAML configuration.  All services
  reference the same ConfigMap; pod identity (\texttt{POD\_NAME})
  provides per-replica S3 path namespacing.
\end{description}

\subsection{Data Flow}

\begin{verbatim}
Service writes to SQLite
  -> WAL frames written to shared PVC
  -> Replicate sidecar reads WAL
  -> Frames uploaded to S3 (1s interval)
  -> Optional: age encryption before upload

Pod restart:
  -> Init container runs replicate restore
  -> Fetches latest snapshot + WAL segments
  -> Reconstructs SQLite database
  -> Primary container starts with warm data
\end{verbatim}

\subsection{K8s Sidecar Pattern}

Each service pod contains three components:

\begin{enumerate}
\item \textbf{Init container} (\texttt{replicate-restore}): Runs
  \texttt{replicate restore -if-db-not-exists -if-replica-exists}
  to restore from S3 on cold start.  No-op if database already
  exists on the PVC.

\item \textbf{Primary container}: The service binary (e.g.,
  \texttt{zoo-bridge}, future Base-powered services).  Reads and
  writes \texttt{/app/data/data.db}.

\item \textbf{Sidecar container} (\texttt{replicate}): Runs
  \texttt{replicate replicate} to continuously stream WAL changes
  to S3.  Shares the \texttt{/app/data} volume with the primary.
\end{enumerate}

\subsection{S3 Path Layout}

\begin{verbatim}
s3://zoo-data/
  {pod-name}/
    generations/
      {generation-id}/
        snapshots/
          {index}.snapshot.lz4
        wal/
          {index}.wal.lz4
\end{verbatim}

Each pod (e.g., \texttt{zoo-bridge-0}, \texttt{zoo-bridge-1})
replicates to a distinct prefix.  The \texttt{POD\_NAME}
environment variable is injected via the Kubernetes downward API
(\texttt{metadata.name} fieldRef).

\section{Specification}

\subsection{Replicate Configuration}

\begin{lstlisting}[language=yaml,caption={replicate-config ConfigMap}]
dbs:
  - path: /app/data/data.db
    replicas:
      - type: s3
        bucket: zoo-data
        path: ${POD_NAME}
        endpoint: ${LITESTREAM_S3_ENDPOINT}
        region: sfo3
        force-path-style: true
        sync-interval: 1s
        snapshot-interval: 1h
\end{lstlisting}

\begin{itemize}
\item \texttt{sync-interval: 1s} --- WAL frames shipped every
  second, achieving $\leq 1$\,s RPO.
\item \texttt{snapshot-interval: 1h} --- Full snapshot every hour
  bounds restore time (restore replays at most 1 hour of WAL).
\item \texttt{force-path-style: true} --- Required for
  DigitalOcean Spaces and MinIO.
\end{itemize}

\subsection{Encryption}

WAL segments and snapshots are encrypted at rest using
\texttt{hanzoai/age} with per-environment X25519 key pairs:

\begin{lstlisting}[language=yaml,caption={Encryption configuration}]
dbs:
  - path: /app/data/data.db
    replicas:
      - type: s3
        bucket: zoo-data
        path: ${POD_NAME}
        encryption:
          type: age
          key: ${AGE_SECRET_KEY}
\end{lstlisting}

Key management:
\begin{itemize}
\item Encryption keys stored in Zoo KMS (K-Chain) or
  Kubernetes secrets (bootstrapping).
\item One key pair per environment (devnet, testnet, mainnet).
\item Key rotation: generate new pair, re-snapshot with new key,
  update secret.  Old WAL segments remain readable with old key
  until next snapshot.
\end{itemize}

\subsection{Service Applicability}

\begin{table}[h]
\centering
\caption{Zoo services and replication targets}
\begin{tabular}{@{}llll@{}}
\toprule
Service & Kind & Data & Replicate \\
\midrule
zoo-bridge & Deployment & SQLite (future) & Yes \\
zoo-gateway & CRD (ZooGateway) & Stateless & No \\
zoo-explorer & Deployment & PostgreSQL & No \\
zoo-chain & StatefulSet & Node data (PVC) & No\footnotemark \\
zoo-operator & Deployment & Stateless & No \\
sql & StatefulSet & PostgreSQL & No \\
kv & StatefulSet & Valkey & No \\
ingress & Deployment & Stateless & No \\
\bottomrule
\end{tabular}
\end{table}
\footnotetext{Chain node data replication uses PVC snapshots and
  chain sync, not WAL streaming.  Replicate is for SQLite only.}

As services migrate from PostgreSQL to Base/SQLite, they gain
replicate sidecars with zero code changes --- only a K8s patch.

\subsection{Recovery Guarantees}

\begin{description}
\item[RPO $\leq 1$\,s] WAL frames synced every second.
  In the worst case, the last second of writes is lost.

\item[RTO $\approx$ seconds] Init container restores latest
  snapshot ($\leq 1$ hour old) plus WAL replay.  Typical
  restore: 2--5 seconds for databases under 1\,GB.

\item[Durability] S3 provides 11 nines of durability.
  Snapshots are LZ4-compressed and optionally age-encrypted.

\item[Consistency] SQLite WAL mode guarantees atomic commits.
  Replicate ships complete WAL frames, preserving transaction
  boundaries.
\end{description}

\section{Implementation}

\subsection{K8s Patches}

Patches are applied as standalone YAML files in
\texttt{k8s/patches/} rather than inline kustomize overlays.
This allows selective application per-service and per-environment.

Required patches:
\begin{enumerate}
\item \texttt{replicate-config.yaml} --- Shared ConfigMap
\item \texttt{replicate-bridge.yaml} --- Bridge deployment patch
  (when migrated to SQLite)
\end{enumerate}

\subsection{S3 Service}

Deploy \texttt{hanzoai/s3} as a StatefulSet with block storage:

\begin{lstlisting}[language=yaml,caption={S3 service (MinIO)}]
apiVersion: apps/v1
kind: StatefulSet
metadata:
  name: s3
spec:
  replicas: 1
  template:
    spec:
      containers:
        - name: s3
          image: ghcr.io/hanzoai/s3:latest
          ports:
            - containerPort: 9000
          env:
            - name: MINIO_ROOT_USER
              valueFrom:
                secretKeyRef:
                  name: s3-credentials
                  key: ACCESS_KEY
            - name: MINIO_ROOT_PASSWORD
              valueFrom:
                secretKeyRef:
                  name: s3-credentials
                  key: SECRET_KEY
\end{lstlisting}

\subsection{Apply Script}

A shell script (\texttt{scripts/apply-replicate.sh}) orchestrates
the full deployment:

\begin{enumerate}
\item Verifies \texttt{s3-credentials} secret exists
\item Applies the shared ConfigMap
\item Patches each target service
\item Restarts patched workloads
\item Waits for rollout completion
\end{enumerate}

\section{Comparison with Alternatives}

\begin{table}[h]
\centering
\caption{Replication approaches}
\begin{tabular}{@{}lcccc@{}}
\toprule
Approach & RPO & RTO & Complexity & Cost \\
\midrule
PVC snapshots & hours & minutes & low & low \\
PostgreSQL streaming & 0 & seconds & high & high \\
\textbf{WAL to S3} & \textbf{1\,s} & \textbf{seconds} & \textbf{low} & \textbf{low} \\
CockroachDB & 0 & 0 & very high & very high \\
\bottomrule
\end{tabular}
\end{table}

WAL-to-S3 replication provides the best tradeoff for Zoo's
requirements: near-zero RPO, fast recovery, minimal operational
complexity, and compatibility with the Base/SQLite-per-org
architecture.

\section{Cross-Ecosystem Compatibility}

The \texttt{ghcr.io/hanzoai/replicate:latest} image is shared
across all ecosystems:

\begin{itemize}
\item \textbf{Liquidity}: ATS, BD, TA, IAM, KMS (5 services)
\item \textbf{Lux}: Future Base-powered services
\item \textbf{Zoo}: Bridge and future Base-powered services
\end{itemize}

The only per-ecosystem differences are:
\begin{itemize}
\item S3 bucket name (\texttt{liquidity-data} vs \texttt{zoo-data})
\item S3 endpoint (GCS vs DigitalOcean Spaces)
\item Region (\texttt{us-central1} vs \texttt{sfo3})
\item Age encryption key pairs
\end{itemize}

All configuration is injected via ConfigMap and secrets.
Zero code divergence between ecosystems.

\section{Security Considerations}

\begin{enumerate}
\item \textbf{S3 credentials}: Stored in Kubernetes secrets,
  never in ConfigMaps or container images.  Accessed via
  \texttt{LITESTREAM\_ACCESS\_KEY\_ID} and
  \texttt{LITESTREAM\_SECRET\_ACCESS\_KEY} environment variables.

\item \textbf{At-rest encryption}: WAL segments encrypted with
  \texttt{hanzoai/age} (X25519 + ChaCha20-Poly1305).  Even if
  the S3 bucket is compromised, data is unreadable without the
  age secret key.

\item \textbf{Per-org isolation}: Base already encrypts each
  org's SQLite shard with a unique DEK derived via HKDF.
  Replicate streams the encrypted bytes --- it never sees
  plaintext org data.

\item \textbf{Network policy}: The replicate sidecar only needs
  egress to the S3 service (\texttt{s3.zoo.svc.cluster.local:9000}).
  A NetworkPolicy can restrict this.

\item \textbf{Key rotation}: Age key rotation requires a
  re-snapshot.  The apply script can trigger this via
  \texttt{replicate snapshots create}.
\end{enumerate}

\section{Post-Quantum Security}

The Zoo ecosystem has achieved comprehensive post-quantum (PQ) coverage
across all cryptographic layers, consistent with LP-102 (Lux) and
HIP-0302 (Hanzo).  All previously planned mitigations are now deployed.

\subsection{PQ Scorecard}

\begin{table}[h]
\centering
\caption{Complete post-quantum scorecard.  All layers PQ-safe except EVM chain signing.}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Layer} & \textbf{Algorithm} & \textbf{PQ Status} & \textbf{Status} \\
\midrule
Disk encryption     & AES-256 (sqlcipher)          & Safe (128-bit PQ)    & Deployed \\
HKDF derivation     & HKDF-SHA-256                 & Safe (128-bit PQ)    & Deployed \\
Field encryption    & AES-256-GCM                  & Safe (128-bit PQ)    & Deployed \\
S3 backup           & age ML-KEM-768+X25519 (\texttt{age1pq}) & Safe (FIPS 203) & Deployed \\
TLS                 & X25519MLKEM768 (first curve) & Safe (hybrid PQ)     & Deployed \\
JWT signing         & ML-DSA-65 (FIPS 204)         & Safe (Module-LWE+SIS)& Deployed \\
JWKS validation     & ML-DSA-65                    & Safe                 & Deployed \\
MPC transport       & Hybrid KEM (X25519+ML-KEM-768) & Safe              & Deployed \\
MPC key shares      & PQ KEM encryption            & Safe                 & Deployed \\
Master keys         & Cloud HSM (FIPS 140-2 L3)    & Hardware isolation   & Deployed \\
Chain signing       & secp256k1 ECDSA              & \textbf{Not PQ-safe} & EVM constraint \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Deployed PQ Components}

\begin{description}
\item[S3 backup:] All age encryption uses ML-KEM-768 + X25519 hybrid
  recipients (\texttt{age1pq} prefix).  Label enforcement prevents
  downgrade to classical-only recipients.
\item[TLS:] X25519MLKEM768 as first curve in TLS 1.3 supported groups.
\item[JWT signing:] ML-DSA-65 (FIPS~204) for IAM token issuance.
  JWKS validation uses ML-DSA-65 public keys.
\item[MPC:] Hybrid KEM for transport; PQ KEM encryption for key shares.
\item[HSM:] Cloud HSM (FIPS 140-2 L3) for master keys.  Key material
  never leaves HSM boundary.  Three keyrings per ecosystem.
\end{description}

\subsection{Harvest-Now-Decrypt-Later: Closed}

The combination of ML-KEM-768 hybrid encryption (S3), X25519MLKEM768
TLS (transit), ML-DSA-65 JWT signing (auth), PQ KEM for MPC, and
Cloud HSM for master keys closes the harvest-now-decrypt-later attack
vector completely.  An adversary who captures data today cannot decrypt
it with a future quantum computer.

\section{References}

\begin{thebibliography}{9}

\bibitem{web5}
Z.~Kelling, ``Web5: Local-First Applications with Chain-Anchored
Trust,'' Zoo Labs Foundation, April 2025.

\bibitem{litestream}
B.~Johnson, ``Litestream: Streaming Replication for SQLite,''
\url{https://litestream.io}, 2021.

\bibitem{age}
F.~Valsorda, ``age: A simple, modern and secure file encryption
tool,'' \url{https://age-encryption.org}, 2019.

\bibitem{zip014}
Zoo Labs Foundation, ``ZIP-014: Zoo KMS Integration via Lux KMS,''
2025.

\bibitem{zip015}
Zoo Labs Foundation, ``ZIP-015: Zoo L2 Chain Architecture,''
2025.

\bibitem{xwing}
M.~Connolly, ``X-Wing: The Hybrid KEM You've Been Looking For,''
Internet-Draft, IETF CFRG,
\url{https://www.ietf.org/archive/id/draft-connolly-cfrg-xwing-kem-05.html},
2024.

\bibitem{fips203}
NIST, ``Module-Lattice-Based Key-Encapsulation Mechanism Standard
(ML-KEM),'' FIPS~203, August 2024.

\end{thebibliography}

\vspace{1em}
\noindent\textbf{Status:} Draft\\
\noindent\textbf{Type:} Infrastructure\\
\noindent\textbf{Category:} Core\\
\noindent\textbf{Requires:} ZIP-014, ZIP-015\\
\noindent\textbf{References:} LP-102, HIP-0302

\vspace{0.5em}
\noindent Copyright \copyright{} 2026 Zoo Labs Foundation.
All rights reserved.

\end{document}
