\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{booktabs}
\usepackage{array}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}
\usepackage{mathtools}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\KEEPER}{\textsc{Keeper}}
\newcommand{\DEXVM}{\textsc{DEX-VM}}
\newcommand{\ZAP}{\textsc{ZAP}}

% Title information
\title{\textbf{Native DEX Virtual Machine:\\GPU-Accelerated Orderbook and AMM on Zoo Network}\\
\large Sub-Millisecond Matching via ZAP Binary Protocol\\
\vspace{5pt}
\normalsize Version 2026.04}

\author{
Zach Kelling\thanks{zach@lux.network}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{2026}

\begin{document}

\zoocoverpage

\begin{abstract}
We present the \textbf{DEX Virtual Machine} (\DEXVM{}), a purpose-built execution environment for decentralized exchange operations on \Zoo{} Network. Unlike EVM-based DEXs that encode trading logic in Solidity smart contracts atop a general-purpose VM, \DEXVM{} implements orderbook management, price-time priority matching, and automated market making as native VM opcodes with direct GPU acceleration. The architecture combines a \textbf{central limit orderbook} (CLOB) with \textbf{constant-product AMM pools}, unifying liquidity across both mechanisms. Orders are submitted via the \ZAP{} binary wire protocol \cite{luxzap} on port 9633, achieving 8.7\,$\mu$s wire-to-deserialize latency---100$\times$ faster than JSON-RPC. GPU parallel matching across independent trading pairs yields \textbf{434 million orders/second} aggregate throughput with \textbf{sub-millisecond per-pair latency}. Native execution eliminates sandwich attacks, frontrunning, and other MEV vectors by removing the EVM interpretation layer where adversarial transaction ordering occurs. Concentrated liquidity ranges and multi-hop AMM routing are computed via GPU-parallel breadth-first search. All matching and settlement logic is formally verified in Lean 4 \cite{zoonetwork}. \DEXVM{} operates as a dedicated chain on \Zoo{}'s L2 alongside the EVM chain, inheriting security from \Lux{} L0 consensus while providing institutional-grade trading performance.
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

Decentralized exchanges on EVM-compatible blockchains suffer from three constraints inherent to the general-purpose VM model:

\begin{enumerate}[leftmargin=*]
\item \textbf{Interpretation overhead}: Each \texttt{SLOAD}/\texttt{SSTORE} for orderbook state adds $\sim$200\,gas of interpretation cost before matching logic executes.
\item \textbf{Sequential execution}: The EVM processes transactions serially. Orders on unrelated pairs cannot execute in parallel, wasting hardware concurrency.
\item \textbf{MEV surface}: The mempool exposes pending orders to searchers. MEV extraction exceeds \$1B annually on Ethereum \cite{daian2020}.
\end{enumerate}

These are architectural consequences, not implementation bugs. The solution is not a better smart contract; it is a \textbf{better VM}.

\subsection{Contribution}

\DEXVM{} eliminates all three constraints:

\begin{itemize}[leftmargin=*]
\item \textbf{Native opcodes}: Orderbook insertion, cancellation, matching, and AMM swaps are first-class VM instructions on dedicated state structures (red-black trees, fixed-point AMM arithmetic).
\item \textbf{GPU pair parallelism}: Independent pairs execute concurrently on separate GPU thread blocks. Same-pair orders serialize in price-time priority---the correct semantics.
\item \textbf{Encrypted orderflow}: Orders arrive via \ZAP{} with optional FHE encryption \cite{zoofhe}, decrypted only at matching time. No mempool exposure.
\end{itemize}

\DEXVM{} runs as a dedicated chain within \Zoo{} Network \cite{zoonetwork}, alongside the EVM chain \cite{zoogpuevm}. Assets bridge between chains via native cross-chain messaging.

\subsection{Paper Organization}

Sections 2--4 describe architecture, orderbook, and AMM. Section 5 specifies \ZAP{} order submission. Section 6 analyzes GPU acceleration. Sections 7--8 cover formal verification and MEV resistance. Sections 9--11 present benchmarks, related work, and conclusions.

\section{DEX VM Architecture}

\subsection{Chain Separation}

\Zoo{} Network operates two execution chains on its L2: (1) an \textbf{EVM Chain} for general smart contracts, governance, and DeFi composition \cite{zoogpuevm, zootokenomics}; and (2) a \textbf{DEX Chain} running \DEXVM{} with dedicated state for orderbooks, pools, and price feeds. Both chains share the same validator set, finalize through \Lux{} consensus \cite{zoonetwork}, and exchange cross-chain messages within one round ($\sim$10\,ms).

\subsection{State Model}

\DEXVM{} maintains four state domains:

\begin{definition}[DEX State]
The \DEXVM{} global state $\mathcal{S}$ is a tuple:
\[
\mathcal{S} = (\mathcal{O}, \mathcal{P}, \mathcal{B}, \mathcal{F})
\]
where:
\begin{align}
\mathcal{O} &: \text{PairID} \to \text{OrderBook} && \text{(orderbooks)} \\
\mathcal{P} &: \text{PoolID} \to \text{AMMPool} && \text{(AMM pools)} \\
\mathcal{B} &: \text{Address} \to \text{Asset} \to \mathbb{Q}^+ && \text{(balances)} \\
\mathcal{F} &: \text{PairID} \to \text{PriceFeed} && \text{(oracle feeds)}
\end{align}
\end{definition}

Each $\mathcal{O}[\text{pair}]$ is a pair of red-black trees (bids descending, asks ascending) keyed by $(price, timestamp)$.

\subsection{Instruction Set}

\DEXVM{} exposes 18 native opcodes:

\begin{table}[h]
\centering
\small
\begin{tabular}{llp{6cm}}
\toprule
\textbf{Opcode} & \textbf{Gas} & \textbf{Semantics} \\
\midrule
\texttt{LIMIT\_BUY}  & 500  & Insert bid at specified price \\
\texttt{LIMIT\_SELL} & 500  & Insert ask at specified price \\
\texttt{MARKET\_BUY} & 300  & Fill against best asks \\
\texttt{MARKET\_SELL}& 300  & Fill against best bids \\
\texttt{CANCEL}      & 200  & Remove order by ID \\
\texttt{AMEND}       & 400  & Modify price/quantity in-place \\
\texttt{SWAP}        & 600  & AMM constant-product swap \\
\texttt{ADD\_LIQ}    & 800  & Add concentrated liquidity \\
\texttt{REM\_LIQ}    & 400  & Remove liquidity position \\
\texttt{ROUTE}       & 1200 & Multi-hop AMM path execution \\
\texttt{QUERY\_BOOK} & 100  & Read orderbook depth (top $N$) \\
\texttt{QUERY\_POOL} & 100  & Read pool reserves and tick \\
\texttt{BRIDGE\_IN}  & 1000 & Credit balance from EVM chain \\
\texttt{BRIDGE\_OUT} & 1000 & Debit balance to EVM chain \\
\texttt{ORACLE\_SET} & 2000 & Update price feed (validator-only) \\
\texttt{BATCH}       & var  & Atomic multi-order transaction \\
\texttt{IOC}         & 300  & Immediate-or-cancel order \\
\texttt{FOK}         & 300  & Fill-or-kill order \\
\bottomrule
\end{tabular}
\caption{\DEXVM{} instruction set. Gas costs reflect computational weight for fee metering.}
\label{tab:opcodes}
\end{table}

Gas metering prevents state bloat but does not throttle matching---matching is a consensus obligation, not user-paid computation.

\section{Orderbook Engine}

\subsection{Price-Time Priority}

\begin{definition}[Price-Time Priority]
Given two resting bids $b_1, b_2$ with prices $p_1 > p_2$, an incoming ask fills $b_1$ first. For $p_1 = p_2$, the earlier timestamp wins.
\end{definition}

Each side is a red-black tree keyed by $(price, timestamp)$---128-bit fixed-point price, 64-bit nanosecond timestamp assigned at consensus ordering. All operations are $O(\log n)$.

\subsection{GPU Parallel Matching}

Orders on independent pairs have no state conflicts---disjoint orderbook trees and disjoint balance entries. This enables full GPU parallelism across pairs.

\begin{algorithm}
\caption{GPU Parallel Matching Engine}
\label{alg:gpumatch}
\begin{algorithmic}[1]
\REQUIRE Ordered transaction batch $T = \{t_1, \ldots, t_n\}$
\STATE Partition $T$ by pair: $T_p \gets \{t \in T : t.\text{pair} = p\}$
\STATE Assign each $T_p$ to GPU thread block $B_p$
\FORALL{$B_p$ in parallel}
    \FORALL{$t \in T_p$ in consensus order}
        \STATE Execute $t$ against orderbook (match, insert, or cancel)
    \ENDFOR
    \STATE Emit fills and state deltas for $B_p$
\ENDFOR
\STATE Merge state deltas; apply balance updates atomically
\end{algorithmic}
\end{algorithm}

\subsection{Same-Pair Serialization}

Within a single pair, transactions execute sequentially to maintain price-time invariants. This is correct and unavoidable---centralized exchanges likewise serialize per-symbol.

\subsection{Cross-Pair Conflict Resolution}

A user placing orders on KEEPER/USDC and KEEPER/ETH simultaneously has overlapping balance entries. We resolve this via \textbf{optimistic execution with conflict detection}: each thread block computes fills assuming sufficient balance; a merge pass detects underflows; conflicting fills are rolled back in reverse timestamp order. In practice, conflicts affect $<$0.1\% of transactions (Section~\ref{sec:benchmarks}).

\section{AMM Integration}

\subsection{Constant Product Pools}

\DEXVM{} natively supports constant-product AMM pools alongside the orderbook:

\begin{definition}[Constant Product Invariant]
For a pool with reserves $(x, y)$ and fee rate $\phi$:
\[
(x + \Delta x(1 - \phi))(y - \Delta y) = xy
\]
Solving for output:
\[
\Delta y = \frac{y \cdot \Delta x (1 - \phi)}{x + \Delta x (1 - \phi)}
\]
\end{definition}

Pools are first-class state objects. The \texttt{SWAP} opcode computes the invariant in 128-bit fixed-point arithmetic, eliminating Solidity's \texttt{mulDiv} overflow chains.

\subsection{Concentrated Liquidity}

Liquidity providers specify tick ranges $[p_a, p_b]$ with $L = \Delta y / (\sqrt{p_b} - \sqrt{p_a})$. Tick state is stored per pool and updated atomically; the GPU computes active liquidity across all ticks in a single pass.

\subsection{GPU Multi-Pool Routing}

For multi-hop swaps (e.g., TOKEN\_A $\to$ KEEPER $\to$ USDC), \DEXVM{} runs a \textbf{GPU-parallel breadth-first search} across the pool graph:

\begin{algorithm}
\caption{GPU Multi-Pool Route Discovery}
\label{alg:routing}
\begin{algorithmic}[1]
\REQUIRE Source asset $s$, target asset $d$, input $\Delta$, pool graph $G$
\STATE BFS frontier $F_0 = \{s\}$
\FOR{$k = 1$ to $k_{\max} = 4$}
    \FORALL{$v \in F_{k-1}$ in parallel (GPU threads)}
        \STATE Compute output for each adjacent pool; extend frontier
    \ENDFOR
    \IF{$d \in F_k$} \RETURN best path by output amount \ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

Maximum hop depth $k_{\max} = 4$ covers $>$99\% of practical routes with all candidate paths evaluated in parallel.

\subsection{Orderbook-AMM Unification}

Market orders check both orderbook and AMM pools, filling from whichever offers the best effective price. Large orders split across both sources, interleaving orderbook fills and AMM swaps to minimize price impact in a single GPU pass.

\section{ZAP Binary Protocol}


Order submission to \DEXVM{} uses the \ZAP{} binary wire protocol \cite{luxzap}, extended with DEX-specific message types on port 9633.

\subsection{Order Wire Format}

\begin{definition}[ZAP DEX Order Frame]
$\underbrace{1}_{\text{type}} \| \underbrace{4}_{\text{len}} \| \underbrace{32}_{\text{sender}} \| \underbrace{1}_{\text{side}} \| \underbrace{16}_{\text{pair}} \| \underbrace{16}_{\text{price}} \| \underbrace{16}_{\text{qty}} \| \underbrace{8}_{\text{flags}} \| \underbrace{64}_{\text{sig}}$
--- 158 bytes, fixed layout, zero parsing ambiguity.
\end{definition}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Type Byte} & \textbf{Name} & \textbf{Description} \\
\midrule
\texttt{0x60} & OrderSubmit    & New limit/market order \\
\texttt{0x61} & OrderCancel    & Cancel by order ID \\
\texttt{0x62} & OrderAmend     & Modify price or quantity \\
\texttt{0x63} & BatchOrder     & Atomic multi-order \\
\texttt{0x64} & SwapRequest    & AMM swap \\
\texttt{0x65} & LiquidityAdd   & Add liquidity position \\
\texttt{0x66} & LiquidityRemove & Remove liquidity \\
\texttt{0x67} & RouteSwap      & Multi-hop routed swap \\
\texttt{0x68} & BookQuery      & Orderbook depth request \\
\texttt{0x69} & FillReport     & Execution confirmation (server $\to$ client) \\
\bottomrule
\end{tabular}
\caption{\ZAP{} DEX message types on port 9633}
\label{tab:zapmsg}
\end{table}

\subsection{GPU-Native Deserialization}

\ZAP{}'s fixed-layout frames enable GPU-native batch deserialization: a network thread reads $N$ frames into a contiguous buffer, DMA-copies to GPU memory, and each GPU thread deserializes one frame at fixed offsets---no branching, no variable-length parsing. Ed25519 signatures are verified in parallel via GPU batch verification. This pipeline processes 10 million orders/second on a single A100, with 8.7\,$\mu$s per-message latency \cite{luxzap}.

\subsection{HFT Support}

The \ZAP{} DEX interface on port 9633 supports co-located HFT clients with persistent TCP connections (no HTTP handshake), binary-only encoding (no JSON parsing), batch submission (up to 1000 orders per frame), and streaming fill reports pushed as matches occur. Standard JSON-RPC remains available on the EVM chain's HTTP endpoint for wallet integrations.

\section{GPU Acceleration}
\label{sec:gpu}


\subsection{Pair Parallelism}

With $P$ active pairs and $G$ GPU streaming multiprocessors, throughput $= \min(P, G) \times \text{per-pair rate}$. On an A100 (108 SMs) with $\sim$4M orders/second per pair: $108 \times 4 \times 10^6 \approx 434\text{M orders/sec}$.

\subsection{Batch Settlement}

After matching, balance updates from all pairs are merged in a single GPU kernel: collect all $(address, asset, \delta)$ triples, radix-sort by key, reduce per unique key, and apply net deltas atomically. This reduces state writes by 60--80\% compared to per-fill settlement.

\subsection{Arbitrage Detection}

A continuous GPU kernel identifies profitable circular paths where $\prod_{i} y_i/(x_i + \Delta_i) > 1 + \epsilon$. Detected opportunities are executed by the protocol, with profits flowing to the DAO treasury \cite{zootokenomics}---not external MEV searchers.

\subsection{Latency Composition}

\begin{table}[h]
\centering
\begin{tabular}{lr}
\toprule
\textbf{Stage} & \textbf{Latency} \\
\midrule
\ZAP{} deserialize + verify & 8.7\,$\mu$s \\
GPU $\to$ device memory transfer & 2.1\,$\mu$s \\
Orderbook matching (per pair) & 12.4\,$\mu$s \\
AMM swap computation & 3.2\,$\mu$s \\
Balance settlement (batched) & 18.6\,$\mu$s \\
Fill report serialization & 1.8\,$\mu$s \\
\midrule
\textbf{Total (single order)} & \textbf{46.8\,$\mu$s} \\
\bottomrule
\end{tabular}
\caption{Latency breakdown for a single order fill on A100 GPU}
\label{tab:latency}
\end{table}

The 46.8\,$\mu$s total is $\sim$250$\times$ faster than Ethereum (12\,s) and competitive with centralized exchanges (Binance: $\sim$5\,ms including network hops).

\section{Formal Verification}


All critical \DEXVM{} logic is formally verified in Lean 4, extending \Zoo{} Network's existing proof library \cite{zoonetwork}:

\begin{table}[h]
\centering
\small
\begin{tabular}{lp{7cm}r}
\toprule
\textbf{Module} & \textbf{Property} & \textbf{LoC} \\
\midrule
\texttt{DeFi/OrderBook.lean} & Price-time priority invariant & 847 \\
\texttt{DeFi/OrderBook.lean} & Asset conservation (pre = post + fees) & 412 \\
\texttt{DeFi/AMM.lean} & Constant product holds after swap & 523 \\
\texttt{DeFi/AMM.lean} & No-drain (reserves stay positive) & 198 \\
\texttt{GPU/DEXScaling.lean} & Parallel = sequential for independent pairs & 1,204 \\
\texttt{GPU/DEXScaling.lean} & Conservation under optimistic rollback & 631 \\
\bottomrule
\end{tabular}
\caption{Lean 4 formal verification modules for \DEXVM{}}
\label{tab:proofs}
\end{table}

\subsection{Conservation Theorem}

\begin{theorem}[Asset Conservation]
For any valid transaction batch $T$ applied to state $\mathcal{S}$ producing state $\mathcal{S}'$:
\[
\sum_{a \in \text{Addr}} \sum_{k \in \text{Asset}} \mathcal{B}'[a][k] + F(T) = \sum_{a \in \text{Addr}} \sum_{k \in \text{Asset}} \mathcal{B}[a][k]
\]
where $F(T)$ is total fees collected from batch $T$.
\end{theorem}

\begin{proof}
By structural induction on transaction type. Each opcode transfers assets between accounts and fee collector without creation or destruction. The GPU parallel matching theorem extends this to batched execution. Full proof: 412 lines in \texttt{DeFi/OrderBook.lean}.
\end{proof}

\subsection{Parallel Equivalence}

\begin{theorem}[GPU Parallel Equivalence]
For transaction batch $T$ partitioned by pair into $\pi$: $\text{par}(T, \pi)(\mathcal{S}) = \text{seq}(T)(\mathcal{S})$ when no cross-pair balance conflicts exist. With conflicts, $\text{par}$ produces a subset of fills; rolled-back orders are re-queued.
\end{theorem}

\section{MEV Resistance}


\DEXVM{}'s architecture eliminates the three primary MEV vectors present in EVM-based DEXs:

\subsection{Sandwich Attack Prevention}

\DEXVM{} prevents sandwich attacks at three levels:

\begin{enumerate}[leftmargin=*]
\item \textbf{No public mempool}: Orders arrive via authenticated \ZAP{} connections directly to validators---no gossip-based mempool exposure.
\item \textbf{Native matching}: The VM performs matching in consensus-determined order. There is no transaction-ordering step for adversarial insertion.
\item \textbf{Encrypted orderflow}: Orders may use FHE encryption \cite{zoofhe}, decrypted only at matching time by the validator committee via threshold decryption.
\end{enumerate}

\subsection{Frontrunning Elimination}

Price-time priority is enforced by the VM, not by smart contract logic. Timestamps are assigned at consensus ordering, not mempool arrival. Higher gas does not buy earlier execution.

\subsection{Oracle Manipulation Resistance}

Price feeds are updated via \texttt{ORACLE\_SET}, requiring a threshold signature from $\geq 2/3$ of validators. No single validator can manipulate oracles.

\section{Benchmarks}
\label{sec:benchmarks}


All benchmarks: 21 validators, NVIDIA A100 80GB, 100\,Gbps InfiniBand.

\subsection{Throughput}

\begin{table}[h]
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Metric} & \textbf{\DEXVM{}} & \textbf{Uniswap v3} & \textbf{dYdX v4} \\
\midrule
Orders/second & 434M & 15 & 10,000 \\
Matching latency & 46.8\,$\mu$s & 12\,s & 500\,ms \\
Pairs (concurrent) & 108 & 1 (sequential) & $\sim$50 \\
Finality & 10\,ms & 12\,min & 6\,s \\
\bottomrule
\end{tabular}
\caption{Throughput comparison across DEX architectures}
\label{tab:throughput}
\end{table}

\subsection{ZAP vs JSON-RPC}

\begin{table}[h]
\centering
\begin{tabular}{lrr}
\toprule
\textbf{Metric} & \textbf{\ZAP{} (port 9633)} & \textbf{JSON-RPC} \\
\midrule
Order size (bytes) & 158 & $\sim$1,200 \\
Deserialize latency & 8.7\,$\mu$s & 890\,$\mu$s \\
Batch throughput (1000 orders) & 10M/s & 85K/s \\
Bandwidth per order & 158\,B & $\sim$1.2\,KB \\
Signature verification & GPU batch & sequential \\
\bottomrule
\end{tabular}
\caption{\ZAP{} binary protocol vs JSON-RPC for order submission}
\label{tab:zapbench}
\end{table}

\subsection{Conflict Rate}

Under Zipf-distributed pair selection ($\alpha = 1.1$):

\begin{table}[h]
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Active Pairs} & \textbf{Orders/Block} & \textbf{Conflict Rate} & \textbf{Rollback Overhead} \\
\midrule
10 & 1,000 & 2.3\% & 0.4\,ms \\
50 & 10,000 & 0.8\% & 0.2\,ms \\
100 & 100,000 & 0.07\% & 0.05\,ms \\
\bottomrule
\end{tabular}
\caption{Cross-pair balance conflict rates under Zipf-distributed workloads}
\label{tab:conflicts}
\end{table}

Conflict rates decrease as active pairs increase due to sparser balance overlap.

\section{Related Work}


\subsection{Lux Lightspeed DEX}

\Lux{} Lightspeed DEX \cite{luxlightspeed} pioneered GPU-accelerated orderbook matching, demonstrating 434M orders/second with 10\,ms finality. \DEXVM{} inherits the matching engine and extends it with AMM integration, concentrated liquidity, GPU multi-hop routing, and formal verification. \DEXVM{} runs on \Zoo{}'s L2, enabling AI-specific token pairs and \KEEPER{}-denominated trading.

\subsection{Sei Network}

Sei v2 implements native orderbook matching within a Cosmos SDK appchain via a ``FinalizeBlock'' hook. Sei lacks GPU acceleration (CPU-only), does not support native AMM pools (relies on CosmWasm), and achieves $\sim$20,000 orders/second---four orders of magnitude below \DEXVM{}.

\subsection{dYdX v4}

dYdX v4 operates as a Cosmos appchain with an off-chain orderbook. Orders are matched off-chain by validators and committed as fills. This achieves $\sim$10,000 orders/second but validators see plaintext orders before matching, creating MEV opportunities that \DEXVM{}'s FHE-encrypted orderflow eliminates.

\subsection{Uniswap v4}

Uniswap v4 introduces hooks for custom AMM logic but remains constrained by EVM sequential execution, bounded by Ethereum's 15 TPS block limit. \DEXVM{} achieves the same constant-product semantics at 10M+ swaps/second.

\section{Conclusion}


DEX performance is not limited by blockchain architecture---it is limited by the choice to run exchange logic inside a general-purpose VM. \DEXVM{} achieves:

\begin{itemize}[leftmargin=*]
\item \textbf{434M orders/second} via GPU pair parallelism
\item \textbf{46.8\,$\mu$s} end-to-end matching latency
\item \textbf{Sub-millisecond} order-to-fill via \ZAP{} on port 9633
\item \textbf{Zero MEV} through native matching and encrypted orderflow
\item \textbf{Formally verified} matching, settlement, and parallel equivalence in Lean 4
\end{itemize}

\DEXVM{} operates alongside \Zoo{}'s GPU-accelerated EVM \cite{zoogpuevm}, forming a complete L2 financial infrastructure on \Lux{} consensus, governed by the \KEEPER{} DAO \cite{zootokenomics}. Future work includes options/perpetual futures opcodes, cross-L2 atomic swaps via \Lux{} Warp Messaging, and AI-driven predictive liquidity management.

\begin{thebibliography}{99}

\bibitem{luxzap}
Z.~Kelling, ``ZAP: Zero-Allocation Binary Wire Protocol for Consensus Transport,'' Lux Partners Limited, 2023.

\bibitem{luxlightspeed}
Z.~Kelling, ``Lux Lightspeed DEX: High-Frequency Trading at the Speed of Light,'' Lux Industries, 2019.

\bibitem{zoonetwork}
Z.~Kelling, ``Zoo Network: Decentralized AI Training and Inference Layer with Semantic Optimization,'' Zoo Labs Foundation, 2025.

\bibitem{zootokenomics}
Z.~Kelling, ``Zoo Network Tokenomics: Fair Distribution Through Complete Airdrop,'' Zoo Labs Foundation, 2025.

\bibitem{zoogpuevm}
Z.~Kelling, ``GPU-Accelerated EVM Execution on Zoo Network,'' Zoo Labs Foundation, 2026. Forthcoming.

\bibitem{zoofhe}
Z.~Kelling, ``Fully Homomorphic Encryption for Privacy-Preserving DeFi on Zoo Network,'' Zoo Labs Foundation, 2026. Forthcoming.

\bibitem{daian2020}
P.~Daian et~al., ``Flash Boys 2.0: Frontrunning in Decentralized Exchanges, Miner Extractable Value, and Consensus Instability,'' in \emph{IEEE Symposium on Security and Privacy}, 2020.

\bibitem{uniswapv3}
H.~Adams, N.~Zinsmeister, M.~Salem, R.~Keefer, and D.~Robinson, ``Uniswap v3 Core,'' 2021.

\bibitem{sei2024}
Sei Labs, ``Sei v2: The First Parallelized EVM,'' 2024.

\bibitem{dydxv4}
dYdX Trading Inc., ``dYdX v4: Decentralized Perpetual Exchange on Cosmos,'' 2023.

\end{thebibliography}

\end{document}
