\section{Native Securities}
\label{sec:native-securities}

The \Zoo{} \DEX{} lists tokenised securities as first-class assets,
not as wrapped derivatives or synthetic exposures. Each security
class settles directly on \Zoo{} D-Chain with \Quasar{} 4.0 triple-cert
finality \cite{lp-020}.

\subsection{Equities (fractional shares)}

Tokenised equity positions, fractional down to one millionth of a
share, settle on \Zoo{} D-Chain. The transfer-agent hook in the
\Liquidity{} precompile (\S\ref{sec:liquidity-integration}) updates
the official record at \Liquidity{}.io's registered transfer agent
on every trade. A holder's beneficial-owner record is updated within
the same block as the trade. Voting rights on equity tokens propagate
to the holder's wallet via the same record.

\paragraph{Fractional liquidity.}
A position of $0.0001$ shares is tradeable. The matching engine has
no minimum-quantity floor; the only floor is the gas cost per trade,
which under \Zoo{} 4.0 \GPU{}-native gas is sub-cent for typical
order sizes.

\paragraph{Corporate actions.}
Splits, dividends, mergers, and tender offers propagate via on-chain
events emitted by the issuer's \Liquidity{}-attested action contract.
Holder wallets reconcile automatically.

\subsection{Fixed income}

\paragraph{Corporate bonds.}
Tokenised corporate debt with on-chain coupon distributions. Issuer
deposits coupon principal in a custody contract; the contract
distributes pro-rata to holders on the coupon date; holders' wallets
reflect the distribution as native \texttt{\$ZOO}-denominated income
(or a configurable denomination per bond series).

\paragraph{Treasuries.}
Tokenised US Treasury exposures (T-bills, notes, bonds) settled
through the \Liquidity{} broker-dealer interface. Real-time
mark-to-market against published rates.

\paragraph{Yield strips.}
Coupon-stripping into separately tradeable principal and interest
tokens. Useful for laddered yield products served to retail.

\subsection{Real-world assets (RWA)}

\paragraph{REITs.}
Tokenised real estate trust shares with quarterly distributions.

\paragraph{Commodities.}
Tokenised gold, silver, copper, energy --- spot exposure backed by
warehouse-attested holdings. The attestations live on A-Chain
\cite{lp-134} so verification of physical backing is on-chain.

\paragraph{Private equity.}
Tokenised LP interests in private equity funds, gated by
accredited-investor status checks via the Z-Chain selective
disclosure path (\S\ref{sec:privacy}).

\paragraph{Carbon and conservation credits.}
On-chain carbon credits and conservation impact tokens (consistent
with \Zoo{} Labs Foundation's 501(c)(3) mission), tradeable like any
other asset and burnable on-chain when retired against an emissions
or impact reporting obligation.

\subsection{Tokenised derivatives}

\paragraph{Perpetual futures.}
Continuous funding-rate perpetuals on equity, commodity, and crypto
underlyings. Funding-rate calculation, mark-price oracle, and
liquidation engine all run on-chain. Cross-margined accounts permit
capital efficiency across positions.

\paragraph{Dated futures.}
Calendar futures on commodities and equity indexes, with monthly
expiry rolls and physical or cash settlement at expiry.

\paragraph{Options.}
European-style options on supported underlyings with on-chain
exercise. American-style and exotic structures gated by regulator
approval and \Liquidity{} broker-dealer review.

\subsection{Compliance-by-construction}

Every security listing requires:

\begin{enumerate}[leftmargin=1.4em]
  \item Issuer onboarding through \Liquidity{}.io (the SEC ATS,
        FINRA broker-dealer, and registered transfer agent control
        plane).
  \item A \texttt{SecurityClass} contract on \Zoo{} D-Chain that
        delegates the regulator-relevant checks (transfer
        eligibility, holder eligibility, jurisdiction) to the
        \Liquidity{} precompile.
  \item A continuous attestation feed from the issuer's transfer
        agent to A-Chain, which the \DEX{} reads before allowing
        listing changes.
  \item Z-Chain \ZKP{} circuits for any holder-side eligibility
        checks where the holder must prove a property (accreditation
        status, geography, source-of-funds).
\end{enumerate}

Trades on a security class that has lost good-standing status (e.g.,
a delisting, a corporate-action freeze, a regulatory hold) are
automatically blocked at the precompile level; the matching engine
does not even see the order. This is the compliance-by-construction
discipline: the rules are not policy in a contract; they are gates
in the precompile that the matching engine cannot bypass.

\subsection{Settlement economics}

T+0 settlement is the default. The \Quasar{} 4.0 cert provides
sub-1.1s finality \cite{lp-020}; the transfer-agent hook fires within
the same block; the official record is updated atomically with the
chain state. There is no T+1 / T+2 settlement window because there
is nothing to wait for.

\subsection{Why this is materially different from a CEX listing}

A CEX listing tokenises a hold-on-balance-sheet position. The
exchange owns the asset; the user owns a claim against the exchange.
A \Zoo{} D-Chain listing is the asset. The transfer-agent record
\emph{is} the chain record. There is no exchange-balance-sheet
intermediary. This is what ``native securities'' means in this
paper: the chain is the ledger of record, not a mirror of one.
