\section{Activation Roadmap}
\label{sec:roadmap}

\Zoo{} \DEX{} activates on April 20, 2026, alongside \Lux{},
\Hanzo{}, and Pars adopting the \Liquidity{} Protocol formally
launched on April 1, 2026 by \Liquidity{}.io
\cite{liquidity-protocol-formal,zoo-adopts-liquidity-protocol}.

\subsection{Phase 0 --- spec freeze (2026-04-01)}

This paper, the supporting ZIPs (Zoo \DEX{} V2/V3/V4 contracts,
\Liquidity{} precompile interface, retail UX, conservation
allocation), and the activation client release freeze on
2026-04-01 (the \Liquidity{} Protocol formal launch). After this
date only spec-clarification PRs are accepted; behaviour-changing
PRs slip to a post-activation patch release.

\subsection{Phase 1 --- testnet bring-up (2026-04-01 through 2026-04-13)}

\begin{itemize}[leftmargin=1.2em]
  \item \Zoo{} D-Chain testnet validators come online with
        activation-candidate binaries.
  \item AMM V2 / V3 deploy on testnet with seed liquidity for the
        initial listing set.
  \item AMM V4 precompile target dummied to a testnet \Lux{} \DEX{}
        instance for cross-chain test paths.
  \item \Liquidity{} precompile testnet image deployed; KYC, BD,
        and TA hooks exercised end-to-end against \Liquidity{}.io
        sandbox endpoints.
  \item Initial securities listings onboarded in dry-run mode:
        equity tokens, treasury tokens, REIT tokens, perpetual
        contracts on a representative basket.
  \item Wallet release-candidates exercise the trade flow on
        testnet for retail UX validation.
\end{itemize}

\subsection{Phase 2 --- canary (2026-04-13 through 2026-04-19)}

\begin{itemize}[leftmargin=1.2em]
  \item \Zoo{} D-Chain canary mainnet brought up with a subset of
        validators.
  \item Initial market-makers onboarded with rebate-tier
        agreements.
  \item Closed-beta retail user cohort tests trade flow on a small
        subset of listings.
  \item Cross-chain routing exercised via B-Chain to \Lux{} C-Chain
        canary.
  \item \Liquidity{}-precompile attestation flow exercised on
        production \Liquidity{}.io endpoints.
\end{itemize}

\subsection{Phase 3 --- activation (2026-04-20)}

\begin{itemize}[leftmargin=1.2em]
  \item \Zoo{} D-Chain mainnet open to all validators.
  \item AMM V2 / V3 / V4 production-active.
  \item Local Zoo \DEX{} CLOB precompile activated.
  \item Continuous limit book and batch auction modes available
        for retail and institutional flow.
  \item Order types matrix (limit, market, stop, IOC, FOK,
        post-only, hidden, iceberg) available across all listings.
  \item Wallet GA with \Zoo{} \DEX{} access.
  \item Initial securities listing cohort opens for trading:
        equities (tokenised fractional shares), fixed income
        (corporate bonds, treasuries), RWA (REITs, commodities,
        carbon credits), derivatives (perpetuals, dated futures,
        European options).
  \item NFT migration second cohort opens for trading on the \DEX{}
        with native fractional V3 pools.
  \item Conservation allocation begins flowing to the foundation's
        impact program from the first trade onward.
\end{itemize}

\subsection{Phase 4 --- post-activation roll-forward}

\Zoo{} \DEX{} continues to ship features incrementally. Roadmap
items targeted for post-activation patch releases:

\begin{itemize}[leftmargin=1.2em]
  \item Phased pair listings: weekly cohorts of new tokenised
        securities listings as issuers complete \Liquidity{}.io
        onboarding.
  \item Market-maker programme expansion: tiered rebate schedules,
        bespoke risk-allocation lines, per-pair maker grants from
        the protocol treasury.
  \item Institutional bridges: prime-broker integrations for
        tokenised-securities flow, custody-vault integrations for
        large holders, on-ramp/off-ramp via existing exchanges via
        the \Liquidity{} Protocol's BD interface.
  \item International listings: cohort-by-cohort onboarding of
        non-US issuers under the \Liquidity{} Protocol's
        jurisdiction-aware compliance perimeter.
\end{itemize}

\subsection{Cross-paper relationships}

\Zoo{} \DEX{} sits at the intersection of several \Zoo{} papers:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{Upstream:} \Zoo{} 4.0 \cite{zoo-4-0} (February 14,
        2026 L1 graduation). \DEX{} runs on the GPU-native L1
        substrate; the AMM curve math, order-book matching, and
        cert generation all execute on the GPU.
  \item \textbf{Upstream:} \Zoo{} 3.0 \cite{zoo-3-0} (October 31,
        2025 full PQ stack). The \DEX{}'s settlement Q-day
        resilience, NFT-loan PQ signing, and validator key
        management all inherit the 3.0 PQ posture.
  \item \textbf{Lateral:} \Liquidity{} Protocol formal proof
        \cite{liquidity-protocol-formal} (April 1, 2026
        \Liquidity{}.io launch). Compliance perimeter and audit
        interface are inherited.
  \item \textbf{Lateral:} Zoo Adopts the \Liquidity{} Protocol
        \cite{zoo-adopts-liquidity-protocol} (April 20, 2026
        adoption note). The proof artifact for Zoo's specific
        adoption.
  \item \textbf{Lateral:} Zoo 2025 securities and \DAO{}
        \cite{zoo-2025-securities-and-dao} (\$113T market thesis).
        The market-sizing and governance frame for the \DEX{}'s
        securities listings.
  \item \textbf{Lateral:} Per-\LLM{} chains \cite{zoo-per-llm-chains}.
        Per-\LLM{} token markets list on the \DEX{} alongside
        traditional securities.
  \item \textbf{Origin:} \Zoo{} 2021 original whitepaper
        \cite{zoo-2021}. The \DEX{} is the realisation of the 2021
        ``NFT Liquidity Protocol'' coinage.
\end{itemize}

\subsection{Out-of-scope for this launch}

\begin{itemize}[leftmargin=1.2em]
  \item Synthetic exposures and structured products beyond the
        listed perpetual / dated future / European option matrix.
  \item Cross-jurisdictional retail accounts beyond the initial
        US-centric \Liquidity{}.io perimeter; international
        retail onboarding ships post-activation.
  \item Margin lending against tokenised-equity collateral; ships
        in a follow-up release with dedicated risk parameters.
\end{itemize}

\subsection{One-line summary}

\Zoo{} \DEX{} is what happens when you take \Zoo{} 4.0's
\GPU{}-native sovereign L1, settle native securities on a
\Lux{}-D-Chain white-label, integrate the \Liquidity{} Protocol's
compliance perimeter, and ship a Robinhood-style retail UX with
zero commission, fractional shares, 24/7 markets, and sub-1.1s
T+0 settlement. April 20, 2026.
