\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

\input{shared/lstlang}

% Listing style
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  captionpos=b,
  xleftmargin=2em,
  framexleftmargin=1.5em,
}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\ERC}{\textsc{erc}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\ZOO}{\$\textsc{zoo}}
\newcommand{\TBA}{\textsc{tba}}
\newcommand{\MPC}{\textsc{mpc}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% Title and authors
\title{\textbf{Zoo $\leftrightarrow$ Lux Bridge: Multi-Chain Asset Transfer via Warp Messaging}}

\author{
Antje Worring, Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{February 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
We present the cross-chain bridge protocol between Zoo L2 (Chain~ID: 200200) and Lux Network (C-Chain, X-Chain, B-Chain). The bridge is built on the Lux Teleport architecture (LP-6332) and Warp messaging (ZIP-0022), and secured by MPC threshold signatures with a $\lceil 2n/3 \rceil$ supermajority requirement using the GG20 threshold ECDSA scheme. It supports bridging of ZRC-20 fungible tokens, ZRC-721 wildlife NFTs with full metadata preservation, ZRC-1155 multi-tokens, and native ZOO/LUX gas tokens. The protocol ensures that conservation metadata, provenance chains, and token-bound account states are faithfully preserved across chains. We describe the bridge gateway contracts, MPC signer set, metadata serialization, TBA freeze/mirror mechanism, fee structure, and liveness/safety guarantees.

\textbf{Keywords}: cross-chain bridge, Warp messaging, MPC, threshold ECDSA, Teleport, metadata preservation
\end{abstract}

\section{Introduction}

The Zoo ecosystem exists on Zoo L2 (Chain~ID: 200200), a rollup anchored to the Lux primary network. Zoo tokens, wildlife NFTs, and conservation assets must be portable to Lux C-Chain for DeFi liquidity, to Lux X-Chain for exchange listing, and to external chains for broader ecosystem access.

Two bridge architectures serve different trust models. The Lux Warp messaging bridge~\cite{warp2023} is the native, trustless path for Zoo L2 $\leftrightarrow$ Lux C-Chain transfers: Warp messages are validated by the Lux primary network validators, requiring no additional trust assumptions. For transfers to external EVM chains, a lock-and-mint bridge secured by an MPC threshold signer set provides Byzantine fault tolerance.

The bridge's distinguishing feature is conservation metadata preservation. When Tiger \#42's ZRC-721 NFT bridges from Zoo to Lux, its species metadata (IUCN status, GPS coordinates, taxonomy), provenance chain (field observations, veterinary records), and token-bound account state (donation balance, conservation history) must arrive intact on the destination chain.

\subsection{Contributions}

\begin{enumerate}
    \item The IZooBridgeGateway interface supporting four asset types (Section~\ref{sec:gateway}).
    \item MPC threshold signature verification using GG20 (Section~\ref{sec:mpc}).
    \item Conservation metadata serialization for cross-chain transfer (Section~\ref{sec:metadata}).
    \item TBA freeze/mirror mechanism for NFT bridge transfers (Section~\ref{sec:tba}).
    \item Liveness and safety guarantees with emergency withdrawal fallback (Section~\ref{sec:liveness}).
    \item Security analysis of MPC key compromise, replay, and censorship attacks (Section~\ref{sec:security}).
\end{enumerate}


\section{Architecture}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.2cm and 3.5cm,
    box/.style={rectangle, draw, minimum width=3cm, minimum height=0.7cm, align=center, font=\footnotesize, fill=blue!8},
    mpc/.style={rectangle, draw, minimum width=3cm, minimum height=0.7cm, align=center, font=\footnotesize, fill=green!8},
    arr/.style={<->, >=stealth, thick},
]
    \node[box] (zoo) {ZooBridgeGateway\\(Zoo L2)};
    \node[mpc, right=of zoo] (mpc) {MPC Signer Set\\($n \geq 7$, $t = \lceil 2n/3 \rceil$)};
    \node[box, right=of mpc] (lux) {LuxBridgeGateway\\(Lux C-Chain)};

    \draw[arr] (zoo) -- (mpc);
    \draw[arr] (mpc) -- (lux);

    \node[box, below=0.8cm of zoo] (locker) {Token Locker\\+ Metadata Serializer\\+ TBA Freezer};
    \node[box, below=0.8cm of lux] (minter) {Token Minter\\+ Metadata Deserializer\\+ TBA Mirror};

    \draw[->, >=stealth] (zoo) -- (locker);
    \draw[->, >=stealth] (lux) -- (minter);
\end{tikzpicture}
\caption{Zoo $\leftrightarrow$ Lux bridge architecture.}
\label{fig:architecture}
\end{figure}


\section{Bridge Gateway}
\label{sec:gateway}

\begin{lstlisting}[language=Solidity,caption=IZooBridgeGateway interface]
interface IZooBridgeGateway {
    enum AssetType {
        NATIVE,   // ZOO or LUX gas token
        ZRC20,    // Fungible token
        ZRC721,   // Non-fungible token
        ZRC1155   // Multi-token
    }

    struct BridgeMessage {
        bytes32 messageId;
        uint256 sourceChainId;
        uint256 destinationChainId;
        address sender;
        address recipient;
        AssetType assetType;
        address tokenContract;
        uint256 tokenId;   // 0 for fungibles
        uint256 amount;    // 1 for NFTs
        bytes metadata;    // Serialized conservation
        uint256 nonce;     // Replay protection
        uint256 deadline;  // Message expiry
    }

    function bridgeOut(
        uint256 destinationChainId,
        address recipient,
        AssetType assetType,
        address tokenContract,
        uint256 tokenId,
        uint256 amount
    ) external payable returns (bytes32 messageId);

    function bridgeIn(
        BridgeMessage calldata message,
        bytes calldata mpcSignature
    ) external;

    function mpcSignerSet()
        external view returns (address);
    function processedMessages(bytes32 messageId)
        external view returns (bool);

    event BridgeOutInitiated(
        bytes32 indexed messageId,
        address indexed sender,
        uint256 destinationChainId,
        AssetType assetType,
        address tokenContract,
        uint256 tokenId,
        uint256 amount
    );
    event BridgeInCompleted(
        bytes32 indexed messageId,
        address indexed recipient,
        uint256 sourceChainId,
        AssetType assetType,
        address tokenContract,
        uint256 tokenId,
        uint256 amount
    );
}
\end{lstlisting}

\subsection{Warp Messaging Path}

For Zoo L2 $\leftrightarrow$ Lux C-Chain transfers, the bridge uses native Warp messaging~\cite{warp2023}:

\begin{lstlisting}[language=Solidity,caption=Warp bridge integration (from ZIP-0022)]
IWarpMessenger constant WARP =
    IWarpMessenger(
        0x0200000000000000000000000000000000000005
    );

function bridgeOutViaWarp(
    uint256 amount, address recipient
) external {
    IERC20(zooToken).transferFrom(
        msg.sender, address(this), amount);
    bytes memory payload = abi.encode(
        recipient, amount);
    WARP.sendWarpMessage(payload);
}

function bridgeInViaWarp(uint32 index) external {
    (IWarpMessenger.WarpMessage memory msg,
        bool valid) =
        WARP.getVerifiedWarpMessage(index);
    require(valid, "invalid warp message");
    require(msg.sourceChainID == peerChainId,
        "wrong source");
    (address recipient, uint256 amount) =
        abi.decode(msg.payload,
            (address, uint256));
    IERC20(zooToken).transfer(recipient, amount);
}
\end{lstlisting}

Warp messages are validated by the Lux primary network validators. No MPC signer set is needed for this path.


\section{MPC Threshold Signatures}
\label{sec:mpc}

For external chain bridges, the protocol uses GG20~\cite{gg20} threshold ECDSA:

\begin{table}[t]
\centering
\begin{tabular}{lc}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Minimum signers ($n$) & 7 \\
Threshold ($t$) & $\lceil 2n/3 \rceil$ \\
Key generation & Distributed key generation (DKG) \\
Signature scheme & Threshold ECDSA (secp256k1) \\
Rotation period & 30 days \\
Slashing & 10,000 LUX stake; slashed for equivocation \\
\bottomrule
\end{tabular}
\caption{MPC signer set parameters.}
\label{tab:mpc}
\end{table}

\begin{lstlisting}[language=Solidity,caption=IMPCSignerSet interface]
interface IMPCSignerSet {
    function verifyThresholdSignature(
        bytes32 messageHash,
        bytes calldata signature
    ) external view returns (bool valid);

    function signerSetInfo()
        external view returns (
            uint256 totalSigners,
            uint256 threshold,
            uint256 rotationEpoch,
            bytes32 publicKeyHash
        );

    function proposeRotation(
        address[] calldata newSigners,
        uint256 newThreshold
    ) external;

    event SignerSetRotated(
        uint256 epoch,
        uint256 totalSigners,
        uint256 threshold,
        bytes32 publicKeyHash
    );
}
\end{lstlisting}

MPC threshold ECDSA produces a single on-chain signature regardless of signer set size. This costs constant gas for verification, unlike multisig which scales linearly with signer count.

\begin{theorem}[Byzantine Fault Tolerance]
With $n = 7$ signers and threshold $t = 5$, the bridge tolerates $f = 2$ Byzantine (compromised) signers. An attacker must compromise $t = 5$ signers to forge a bridge message. The economic cost of compromise is $5 \times 10{,}000 = 50{,}000$ LUX in slashable stake.
\end{theorem}


\section{Asset Bridging Mechanics}

\subsection{Native Token (ZOO/LUX)}

\begin{itemize}
    \item \textbf{Zoo to Lux}: ZOO locked in ZooBridgeGateway. Wrapped ZOO (wZOO) minted on Lux as LRC-20 (LP-3800).
    \item \textbf{Lux to Zoo}: wZOO burned on Lux. ZOO unlocked on Zoo L2.
    \item LUX follows the reverse path (locked on Lux, wrapped on Zoo).
\end{itemize}

\subsection{ZRC-20 Tokens}

Tokens locked on the source chain. A bridged representation is minted on the destination per LP-3800. Conservation extensions (donation rate, impact credits) are restored from on-chain state on return.

\subsection{ZRC-721 NFTs}

\begin{enumerate}
    \item NFT locked on Zoo L2.
    \item Species metadata serialized via ABI encoding and included in \texttt{BridgeMessage.metadata}.
    \item Provenance chain snapshot uploaded to IPFS; CID included in metadata.
    \item TBA frozen (ZIP-0703).
    \item Mirror LRC-721 minted on Lux with full metadata.
    \item On return: mirror burned, original unlocked, TBA unfrozen, ``bridged\_back'' provenance record added.
\end{enumerate}

\subsection{ZRC-1155 Multi-Tokens}

Multiple token IDs and amounts can be bridged in a single message. Soulbound tokens MUST NOT be bridged; the gateway rejects any \texttt{bridgeOut} for non-transferable token IDs.


\section{Metadata Serialization}
\label{sec:metadata}

\begin{lstlisting}[language=Solidity,caption=ZRC-721 metadata serialization]
bytes memory metadata = abi.encode(
    speciesMetadata,        // SpeciesMetadata
    provenanceSnapshotCID,  // IPFS CID
    conservationFundAddress,
    royaltySplit,           // (conservBps, creatorBps)
    tbaAddress,             // TBA on source chain
    tbaBalanceSnapshot      // ZOO balance in TBA
);
\end{lstlisting}

ABI encoding ensures deterministic serialization. The destination chain deserializes the metadata and reconstructs the on-chain state. The provenance CID provides a full history reference without including the entire provenance chain in the bridge message.


\section{TBA Freeze/Mirror}
\label{sec:tba}

\begin{algorithm}[t]
\caption{TBA Freeze on Bridge-Out}
\label{alg:freeze}
\begin{algorithmic}[1]
\STATE \textbf{Input}: NFT $N$, destination chain $D$
\STATE Lock $N$ in bridge gateway
\STATE $\text{TBA.state()} \leftarrow \text{TBA.state()} + 1$ \COMMENT{Invalidate pending ops}
\STATE $\text{TBA.frozen} \leftarrow \text{true}$
\STATE Serialize TBA balance snapshot
\STATE Include in bridge message metadata
\STATE Relay to destination chain $D$
\STATE Create mirror TBA on $D$ with same salt
\end{algorithmic}
\end{algorithm}

The freeze and NFT lock are atomic within a single \texttt{bridgeOut} transaction. This prevents a race condition where the TBA is accessed between the lock and the freeze.

\begin{theorem}[Cross-Chain Asset Safety]
A TBA's assets cannot be double-spent across chains. The freeze mechanism ensures exactly one chain has an active TBA at any time. Asset access on the inactive chain requires a bridge transfer, which is mediated by the MPC signer set.
\end{theorem}


\section{Fee Structure}

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Operation} & \textbf{Fee} & \textbf{Recipient} \\
\midrule
Native token bridge & 0.1\% & MPC signers \\
ZRC-20 bridge & 0.05\% & 50\% signers, 50\% conservation \\
ZRC-721 bridge & Flat 10 ZOO & Conservation fund \\
ZRC-1155 batch & 5 ZOO per type & Conservation fund \\
\bottomrule
\end{tabular}
\caption{Bridge fee schedule.}
\label{tab:fees}
\end{table}


\section{Liveness and Safety}
\label{sec:liveness}

\textbf{Liveness}: If the MPC signer set is unavailable for $> 72$ hours, a governance-controlled emergency withdrawal path activates. Users can reclaim locked assets after a 7-day timelock.

\textbf{Safety}: Messages expire after \texttt{deadline} (default 24 hours). Expired messages are rejected. Each \texttt{messageId} is processed at most once.

\textbf{Finality}: Bridge messages are relayed only after source chain finality. Zoo L2 inherits sub-second finality from the Lux primary network. Lux C-Chain requires 2~seconds.

\textbf{Rate limiting}: Per ZIP-0022, the bridge enforces per-transaction, hourly, and daily volume limits:

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Adjustable} \\
\midrule
Per-transaction limit & 1,000,000 ZOO & Governance \\
Hourly volume limit & 10,000,000 ZOO & Governance \\
Daily volume limit & 50,000,000 ZOO & Governance \\
Emergency pause & 25\% of locked supply in 24h & Fixed \\
\bottomrule
\end{tabular}
\caption{Bridge rate limiting parameters.}
\label{tab:ratelimit}
\end{table}

\begin{theorem}[Supply Invariant]
The bridge enforces:
\begin{equation}
S_{\text{zoo\_l2}} + S_{\text{locked}} = S_{\text{initial}} + S_{\text{emissions}}
\end{equation}
An on-chain audit function verifies this invariant and pauses the bridge if violated.
\end{theorem}


\section{Security Analysis}
\label{sec:security}

\subsection{MPC Key Compromise}

If $t$ signers are compromised, the bridge is vulnerable to fraudulent minting. Mitigations: 30-day key rotation, geographic distribution of signers, HSMs, and slashing for equivocation.

\subsection{Replay Attacks}

Each \texttt{messageId} is tracked in \texttt{processedMessages}. The nonce and deadline fields provide additional replay protection.

\subsection{Metadata Tampering}

The \texttt{messageHash} covers all fields including \texttt{metadata}. MPC signature verification ensures integrity.

\subsection{Censorship}

If MPC relayers censor specific messages, the 72-hour liveness fallback allows governance to force-process messages.

\subsection{TBA Freeze Race Condition}

The TBA must be frozen before the bridge message is relayed. The \texttt{bridgeOut} function atomically locks the NFT and freezes the TBA in a single transaction.

\subsection{Signer Set Rotation}

During rotation, both old and new signer sets are valid for a 24-hour overlap period. Messages signed by either set are accepted during this window.


\section{Related Work}

Lux Teleport (LP-6332)~\cite{teleport2022} provides the transport layer; ZIP-0800 adds the application layer for conservation metadata. LayerZero~\cite{layerzero2022} provides generic cross-chain messaging but lacks conservation metadata support. Wormhole~\cite{wormhole2022} uses a guardian set similar to our MPC signer set. Axelar~\cite{axelar2022} provides general message passing with threshold signatures. Chainlink CCIP~\cite{ccip2023} offers cross-chain interoperability with oracle-based verification. The Cosmos IBC protocol~\cite{ibc2019} provides trustless interchain communication for Tendermint-based chains; Lux Warp messaging provides analogous guarantees within the Lux ecosystem.


\section{Conclusion}

The Zoo $\leftrightarrow$ Lux bridge combines trustless Warp messaging for intra-ecosystem transfers with MPC threshold signatures for external chain access. Conservation metadata preservation ensures that wildlife NFTs, impact credits, and token-bound account states survive cross-chain transfers intact. Rate limiting and the supply invariant provide defense-in-depth against the bridge exploits that have plagued cross-chain DeFi.

\section*{Acknowledgments}

This work was supported by Zoo Labs Foundation. We thank the Lux team for the Teleport architecture and Warp messaging protocol.

\bibliographystyle{plain}
\begin{thebibliography}{20}

\bibitem{warp2023}
Lux Partners, ``Lux Warp Messaging,'' \url{https://docs.lux.network/warp}, 2023.

\bibitem{teleport2022}
Lux Partners, ``LP-6332: Teleport Bridge Architecture,'' Lux Protocol Standards, 2022.

\bibitem{gg20}
R.~Gennaro and S.~Goldfeder, ``One Round Threshold ECDSA with Identifiable Abort,'' \textit{IACR ePrint} 2020/540, 2020.

\bibitem{lp3800}
Lux Partners, ``LP-3800: Bridged Asset Standard,'' Lux Protocol Standards, 2022.

\bibitem{lp6000}
Lux Partners, ``LP-6000: B-Chain Bridge Specification,'' Lux Protocol Standards, 2022.

\bibitem{layerzero2022}
R.~Zarick et al., ``LayerZero: An Omnichain Interoperability Protocol,'' Technical Report, 2022.

\bibitem{wormhole2022}
Wormhole, ``Wormhole: Generic Message Passing Protocol,'' Technical Documentation, 2022.

\bibitem{axelar2022}
S.~Haouari et al., ``Axelar: Connecting All Blockchains,'' Technical Documentation, 2022.

\bibitem{ccip2023}
Chainlink, ``CCIP: Cross-Chain Interoperability Protocol,'' Technical Documentation, 2023.

\bibitem{ibc2019}
C.~Goes et al., ``The Interblockchain Communication Protocol,'' Cosmos Network, 2019.

\bibitem{zip0022}
Zoo Labs Foundation, ``ZIP-0022: Multi-Chain Bridge Standard,'' Zoo Improvement Proposals, 2025.

\bibitem{zip0042}
Zoo Labs Foundation, ``ZIP-0042: Cross-Ecosystem Interoperability,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0100}
Zoo Labs Foundation, ``ZIP-0100: Zoo Contract Registry,'' Zoo Improvement Proposals, 2024.

\bibitem{zip0700}
Zoo Labs Foundation, ``ZIP-0700: ZRC-20 Fungible Token Standard,'' Zoo Improvement Proposals, 2025.

\bibitem{zip0701}
Zoo Labs Foundation, ``ZIP-0701: ZRC-721 NFT Standard,'' Zoo Improvement Proposals, 2025.

\bibitem{zip0703}
Zoo Labs Foundation, ``ZIP-0703: Token-Bound Accounts for Wildlife,'' Zoo Improvement Proposals, 2025.

\bibitem{hip0101}
Hanzo AI, ``HIP-0101: Hanzo-Lux Bridge Protocol,'' Hanzo Improvement Proposals, 2023.

\bibitem{ronin2022}
S.~Gudgeon et al., ``Ronin Bridge Hack: A Post-Mortem,'' Technical Analysis, 2022.

\bibitem{nomad2022}
Nomad, ``Nomad Bridge Exploit: Root Cause Analysis,'' Technical Report, 2022.

\bibitem{eip7683}
Across Protocol, ``EIP-7683: Cross-Chain Intents,'' Ethereum Improvement Proposals, 2023.

\end{thebibliography}

\end{document}
