\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{shapes,arrows,positioning,calc,decorations.pathreplacing}
\usepackage{listings}
% Solidity language definition (listings does not include it by default)
\lstdefinelanguage{Solidity}{
    morekeywords={contract,function,returns,return,if,else,for,while,do,
        mapping,address,uint256,uint,int,bool,string,bytes,bytes32,
        public,private,internal,external,view,pure,payable,memory,storage,
        calldata,event,emit,modifier,require,struct,enum,interface,
        abstract,override,virtual,is,new,delete,import,pragma,solidity,
        library,using,constructor,msg,sender,block,timestamp,this,super},
    sensitive=true,
    comment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]",
}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{longtable}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Zen}{\textsc{Zen}}
\newcommand{\Quasar}{\textsc{Quasar}}
\newcommand{\PoAI}{\textsc{PoAI}}
\newcommand{\TEE}{\textsc{tee}}
\newcommand{\LLM}{\textsc{llm}}
\newcommand{\GPU}{\textsc{gpu}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\LoRA}{\textsc{LoRA}}
\newcommand{\BlockSTM}{\textsc{Block-STM}}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray},
    tabsize=2,
    captionpos=b,
}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\title{\textbf{Per-LLM Chains}\\
\large One Sovereign Chain per Model: Unified Training, Research, and Pooled GPU Hosting on \Quasar{} 3.0}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{December 15, 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
We introduce the \emph{Per-LLM Chain} architecture: a sovereign \Quasar{}-certified appchain dedicated to the lifetime of a single large language model. Each chain binds together training-data provenance, compute receipts, research artifact registries, contributor attribution graphs, and inference revenue distribution into a single consensus-ordered ledger. Per-LLM chains settle on \Lux{} \Quasar{} 3.0 (\textsc{lp-020}) via the \Quasar{}-Native App Stack (\textsc{lp-133}), use the \Quasar{}\GPU{} adapter (\textsc{lp-132}) for deterministic GPU scheduling, anchor TEE-attested compute receipts to the \Lux{} A-Chain (\textsc{lp-134}), and federate cloud capacity through \Lux{} Cloud (\textsc{lp-136}). The canonical deployment hosts the \Zen{} model family---eighteen open-weight frontier models from \texttt{zen4-nano} (0.8B) through \texttt{zen4-ultra-max} (1T+)---each with its own chain. The \Zoo{} per-LLM stack complements the \Hanzo{} AI Chain (which operates the inference and commerce marketplace) by providing the upstream training, research, and capacity layer. Spec freeze: 2025-12-15. Activation alongside \Quasar{} 3.0: 2025-12-25.

\textbf{Keywords}: per-LLM chains, decentralized training, GPU pooling, model provenance, \Quasar{} 3.0, appchain, attribution graphs, \Zen{} models
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-introduction}
\input{sections/02-architecture}
\input{sections/03-unified-training}
\input{sections/04-unified-research}
\input{sections/05-pooled-gpu-hosting}
\input{sections/06-zen-models}
\input{sections/07-cross-chain}
\input{sections/08-token-economics}
\input{sections/09-privacy-provenance}
\input{sections/10-activation}
\input{sections/11-securities-and-dao}
\input{sections/12-zoo-3-0-substrate}

\begin{thebibliography}{99}

\bibitem{lp-020-quasar-3}
Lux Foundation.
\newblock LP-020: Quasar Consensus 3.0.
\newblock Lux Improvement Proposal, December 2025.
\newblock \url{https://github.com/luxfi/lps/lp-020}.

\bibitem{lp-010-blockstm}
Lux Foundation.
\newblock LP-010: Quasar Block-STM 3 Parallel Execution.
\newblock Lux Improvement Proposal, December 2025.
\newblock \url{https://github.com/luxfi/lps/lp-010}.

\bibitem{lp-105-quasar-consensus}
Lux Foundation.
\newblock LP-105: Quasar Consensus Family Specification.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-132-quasargpu}
Lux Foundation.
\newblock LP-132: QuasarGPU Adapter for Deterministic GPU Scheduling.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-137-residency}
Lux Foundation.
\newblock LP-137: GPU-Residency Invariant.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{zoo-3-0-launch}
Zoo Labs Foundation.
\newblock Zoo 3.0 Launch: GPU-Native L1.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{lp-133-app-stack}
Lux Foundation.
\newblock LP-133: Quasar-Native App Stack for Sovereign Appchains.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-134-topology}
Lux Foundation.
\newblock LP-134: Lux Chain Topology and A-Chain/B-Chain Specification.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-136-cloud}
Lux Foundation.
\newblock LP-136: Lux Cloud Capacity Federation Layer.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-013-fchain}
Lux Foundation.
\newblock LP-013: F-Chain FHE-Protected Inference.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{hanzo-ai-chain}
Hanzo AI.
\newblock The Hanzo AI Chain: Inference Marketplace and Commerce Layer.
\newblock Hanzo Whitepaper, December 2025.

\bibitem{zen-models}
Zen LM.
\newblock The Zen Model Family: Open-Weight Frontier Models.
\newblock \url{https://huggingface.co/zenlm}, December 2025.

\bibitem{zoo-poai}
Zoo Labs Foundation.
\newblock PoAI Consensus and Decentralized Semantic Optimization.
\newblock Zoo Whitepaper, 2025.

\bibitem{zoo-gym}
Zoo Labs Foundation.
\newblock Zoo Gym: Unified Training Platform.
\newblock Zoo Whitepaper, 2025.

\bibitem{zoo-2025-securities-and-dao}
Zoo Labs Foundation.
\newblock Zoo 2025: Digital Securities, Quantum-Secure Settlement, and the Holographic Consensus DAO.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{block-stm}
Gelashvili, R., et al.
\newblock Block-STM: Scaling Blockchain Execution by Turning Ordering Curse to a Performance Blessing.
\newblock arXiv:2203.06871, 2022.

\bibitem{tee-attestation}
Costan, V., Devadas, S.
\newblock Intel SGX Explained.
\newblock IACR Cryptology ePrint Archive, 2016.

\bibitem{nvidia-cc}
NVIDIA.
\newblock H100 Confidential Computing.
\newblock NVIDIA Technical Brief, 2024.

\bibitem{dp-sgd}
Abadi, M., et al.
\newblock Deep Learning with Differential Privacy.
\newblock CCS, 2016.

\end{thebibliography}

\end{document}
