\section{\Zoo{} \DEX{}}
\label{sec:zoo-dex}

\Zoo{} ships a native \AMM{} \DEX{}. The design is intentionally
Uniswap-shaped at the surface so existing tooling, audits, and pool
analytics translate directly. Underneath, the curve evaluation runs on
the \GPU{} per \S\ref{sec:gpu-native}.

\subsection{V2 --- constant product}

V2 implements the canonical $x \cdot y = k$ \AMM{}. Pools are pairs of
ZRC-20 tokens; LP shares are minted as ZRC-20 receipt tokens. Fees default
to 30 bps with per-pool override. A V2 swap on \Zoo{} 3.0 costs roughly
the same gas as a Uniswap V2 swap on a CPU \textsc{evm}, but block-level
throughput is materially higher because hundreds of swaps execute in
parallel on the device.

\subsection{V3 --- concentrated liquidity}

V3 implements concentrated liquidity in the Uniswap V3 style
\cite{uniswap-v3}: liquidity providers position capital in price ranges
defined by tick boundaries. Tick crossings, fee accumulation, and
range-position accounting are device-side. Multi-fee-tier pools
(5 / 30 / 100 bps) coexist for the same pair.

\subsection{V4 --- \Lux{} \DEX{} precompile}

V4 is not a separate \AMM{}. V4 is a precompile interface that lets
\Zoo{} contracts route swaps to the \Lux{} \DEX{} \cite{lux-dex},
which lives at \texttt{\textasciitilde/work/lx/dex}. The \Lux{} \DEX{}
is the cross-chain liquidity router and the canonical venue for
\Lux{}-ecosystem CLOB-style matching; \Zoo{} V4 plugs into it via
B-Chain (\textsc{lp-134}) routed precompile calls. Use V4 when:

\begin{itemize}[leftmargin=1.2em]
  \item Routing across chains where the deepest book lives on \Lux{}.
  \item Matching against external order flow that the \Lux{} \DEX{}
        already aggregates.
  \item Serving institutional flow that expects CLOB semantics.
\end{itemize}

Use V2/V3 when:

\begin{itemize}[leftmargin=1.2em]
  \item Trading \Zoo{}-native assets (per-\LLM{} tokens, wildlife \NFT{}
        fractional pools, \DAO{} governance pools).
  \item Latency budget is sub-block; staying chain-local avoids the
        B-Chain hop.
  \item LP yield must be programmable per pool by Zoo \DAO{} policy.
\end{itemize}

\subsection{Optional local Zoo \DEX{} precompile}

For \Zoo{}-internal use cases that want CLOB semantics without the
B-Chain hop, \Zoo{} ships an optional in-chain \DEX{} precompile that
implements the same matching primitives as the \Lux{} \DEX{} but
against \Zoo{}-local order books. The precompile shares its kernel
implementation with \Lux{} \DEX{} so behaviour is identical; only
the venue and gas accounting differ.

\subsection{Per-\LLM{} token markets}

The per-\LLM{} chain framework \cite{zoo-per-llm-chains} mints a
distinct gas/utility token per \Zen{} model. These tokens trade against
\$ZOO and against each other on \Zoo{} \DEX{} V2/V3. Inference revenue
flowing through the per-\LLM{} chain attribution graph settles into
the relevant V3 pool, providing a price discovery mechanism for model
quality.

\subsection{\NFT{} fractional pools}

\NFT{}s migrated to \Zoo{} L1 (\S\ref{sec:nft-strategy}) can be wrapped
into ERC-4626 vault shares and traded as concentrated-liquidity V3
positions. Combined with F-Chain \FHE{} (\S\ref{sec:privacy}), the
fractional positions can trade with confidential balances --- material
for high-value collections.

\subsection{Versioning policy}

\Zoo{} \DEX{} V2 and V3 are stable. V4 is the precompile interface and
will track the \Lux{} \DEX{} interface contract; breaking changes to
\Lux{} \DEX{} require a corresponding \Zoo{} \DEX{} V4 revision. There
is no separate ``Zoo V4 \AMM{}''; V4 is the precompile, full stop.
