\section{\NFT{} Strategy}
\label{sec:nft-strategy}

\Zoo{} 3.0 launches with an explicit go-to-market for \NFT{} migration.
\Zoo{} intends to make competitive offers to the leading \NFT{}
collections in the wildlife/conservation, AI/agent, and metaverse
categories for migration to \Zoo{} L1. The goal is for \Zoo{} to host a
top-2 \NFT{} project portfolio inside its first activation cycle.

\subsection{Why a collection should migrate}

\paragraph{Native \FHE{} privacy.}
High-value collectibles benefit from confidential trading: the floor
price reveals less to predators, holders can build positions without
front-running, and treasury balances stay private until disclosure.
F-Chain (\textsc{lp-013}) makes this a chain-level primitive, not a
per-collection bolt-on.

\paragraph{\GPU{}-fast minting and transfer.}
Drop mechanics that bottleneck on signature verification clear faster
on a \GPU{}-native chain. A 10k-supply drop that exceeds CPU \textsc{evm}
block budgets clears inside a normal block on \Zoo{} 3.0.

\paragraph{Native AI provenance.}
A-Chain (AIVM, \textsc{lp-134}) anchors TEE-attested inference
receipts. For collections whose value derives from generative outputs
(AI-art collections, agent \NFT{}s, on-chain creatures with AI
behaviour), provenance is a chain-level guarantee, not a marketing
claim.

\paragraph{Per-\NFT{} token-bound accounts.}
ZRC-721 \NFT{}s on \Zoo{} L1 have native ERC-6551-style token-bound
accounts. The account can hold \$ZOO, per-\LLM{} tokens, other
\NFT{}s, and accrue yield from staking. Combined with the \Zoo{} \DEX{}
(\S\ref{sec:zoo-dex}) and bridge (\S\ref{sec:zoo-bridge}), the \NFT{}
becomes a programmable agent rather than a static JPEG.

\paragraph{\DAO{}-aligned governance.}
Migrated collections plug into Zoo \DAO{} voting weight. Holders
participate in conservation grants, per-\LLM{} chain governance, and
treasury allocation.

\subsection{Migration mechanics}

\begin{enumerate}[leftmargin=1.4em]
  \item \textbf{Snapshot.} \Zoo{} Bridge (\S\ref{sec:zoo-bridge}) takes a
        snapshot of holders on the source chain.
  \item \textbf{Lock-and-mint.} The source-chain contract is escrowed;
        the \Zoo{} L1 native ZRC-721 is minted to the same wallet
        address.
  \item \textbf{Metadata mirror.} On-chain metadata moves over with
        provenance hash chain intact; off-chain assets re-pin to
        \Zoo{}'s decentralised storage layer.
  \item \textbf{Treasury onboarding.} The collection's treasury is
        onboarded to a \Zoo{} \DAO{}-governed multisig with optional
        F-Chain \FHE{} balance privacy.
  \item \textbf{Royalty redirection.} Future royalties accrue to the
        L1 contract; the bridge keeps the source-chain contract
        readable for legacy explorers.
\end{enumerate}

\subsection{What \Zoo{} offers a migrating collection}

The migration package includes:

\begin{itemize}[leftmargin=1.2em]
  \item Validator-tier infrastructure for the collection's contracts.
  \item Native integration with the \Zoo{} \DEX{} for floor and
        fractional liquidity.
  \item Marketing co-promotion across the \Zoo{} foundation channels.
  \item Treasury-grant matching for conservation- or AI-aligned
        collections from the Zoo Fund.
  \item Optional F-Chain confidential balances and Z-Chain selective
        disclosure for holder identity.
  \item Free \Zoo{} AI Desktop integration for collections that ship
        agent companions.
\end{itemize}

\subsection{Categories}

\Zoo{} prioritises three collection categories:

\begin{enumerate}[leftmargin=1.4em]
  \item \textbf{Wildlife and conservation.} Aligned with the Zoo Labs
        Foundation 501(c)(3) mission. Collections whose ongoing
        impact accrues to species preservation.
  \item \textbf{AI agents and generative art.} Collections that
        benefit from A-Chain provenance and Zen model integration.
  \item \textbf{Metaverse and gaming.} Collections that map to the
        Zoo metaverse companion app and gym progression.
\end{enumerate}

\subsection{Note on naming}

This paper does not name specific target collections; competitive
offers are confidential until accepted. The architecture above is
designed to be a credible offer for any collection in the listed
categories, on terms no other chain can match.
