\section{Rollback Plan}


If mainnet exhibits critical failures (consensus divergence, bridge exploit,
data corruption), the following rollback procedure activates.

\subsection{Chain Halt Conditions}

A chain halt is triggered if \textbf{any} of the following occur:

\begin{enumerate}[nosep]
  \item Consensus divergence: two or more validators on different forks for $>$60s.
  \item Bridge exploit: unauthorized mint or release detected.
  \item Data corruption: state root mismatch across $>$1 validator.
  \item Critical vulnerability: zero-day with active exploitation.
\end{enumerate}

\subsection{Halt Procedure}

\begin{tabularx}{\textwidth}{@{}l X l@{}}
\toprule
\textbf{\#} & \textbf{Action} & \textbf{Owner} \\
\midrule
1 & CTO authorizes chain halt                                          & CTO    \\
2 & Disable bridge (freeze lock/mint contracts on both sides)          & Bridge \\
3 & Coordinate validator shutdown via authenticated channel             & DevOps \\
4 & All 5 founding validators stop block production                    & DevOps \\
5 & Public status page updated: \texttt{status.zoo.ngo}                & Comms  \\
6 & Post-mortem investigation begins (target: root cause within 4h)    & Core   \\
\bottomrule
\end{tabularx}

\subsection{Genesis Reset (Last Resort)}

If the chain state is unrecoverable:

\begin{enumerate}[nosep]
  \item Export last-known-good state snapshot from all validators.
  \item Generate new genesis from snapshot (preserving balances and contract state).
  \item Validators re-key via K-Chain key ceremony.
  \item Re-execute bootstrap sequence (Section~3.1).
  \item Bridge re-activation with fresh MPC keys via T-Chain.
  \item Independent audit of new genesis state before restart.
\end{enumerate}

\noindent This procedure destroys transaction history after the snapshot.
It is used only when all other recovery options are exhausted.

\subsection{Communication Plan}

\begin{tabularx}{\textwidth}{@{}l X l@{}}
\toprule
\textbf{\#} & \textbf{Action} & \textbf{Owner} \\
\midrule
1 & Status page updated within 5 minutes of halt decision         & Comms  \\
2 & Community notification via Discord, Twitter, blog              & Comms  \\
3 & Hourly updates until resolution                                & Comms  \\
4 & Post-mortem published within 72 hours                          & Core   \\
\bottomrule
\end{tabularx}

%% ---------------------------------------------------------------------------
