\section{Architecture}


\subsection{Components}

\begin{description}
\item[\texttt{hanzoai/replicate}] Litestream-based WAL streaming
  daemon.  Watches SQLite WAL, ships frames to S3.  Runs as a
  sidecar container sharing the data volume with the primary
  service container.

\item[\texttt{hanzoai/age}] X25519 + ChaCha20-Poly1305 file
  encryption.  Used for encrypting WAL snapshots and segments
  at rest in S3.  Compatible with \texttt{luxfi/age} (same
  wire format).

\item[\texttt{hanzoai/s3}] S3-compatible object storage.  Deployed
  as a ClusterIP service within the Zoo K8s cluster
  (\texttt{s3.zoo.svc.cluster.local:9000}).  Backed by
  DigitalOcean Spaces or MinIO depending on environment.

\item[ConfigMap] Shared \texttt{replicate-config} ConfigMap
  containing the Litestream YAML configuration.  All services
  reference the same ConfigMap; pod identity (\texttt{POD\_NAME})
  provides per-replica S3 path namespacing.
\end{description}

\subsection{Data Flow}

\begin{verbatim}
Service writes to SQLite
  -> WAL frames written to shared PVC
  -> Replicate sidecar reads WAL
  -> Frames uploaded to S3 (1s interval)
  -> Optional: age encryption before upload

Pod restart:
  -> Init container runs replicate restore
  -> Fetches latest snapshot + WAL segments
  -> Reconstructs SQLite database
  -> Primary container starts with warm data
\end{verbatim}

\subsection{K8s Sidecar Pattern}

Each service pod contains three components:

\begin{enumerate}
\item \textbf{Init container} (\texttt{replicate-restore}): Runs
  \texttt{replicate restore -if-db-not-exists -if-replica-exists}
  to restore from S3 on cold start.  No-op if database already
  exists on the PVC.

\item \textbf{Primary container}: The service binary (e.g.,
  \texttt{zoo-bridge}, future Base-powered services).  Reads and
  writes \texttt{/app/data/data.db}.

\item \textbf{Sidecar container} (\texttt{replicate}): Runs
  \texttt{replicate replicate} to continuously stream WAL changes
  to S3.  Shares the \texttt{/app/data} volume with the primary.
\end{enumerate}

\subsection{S3 Path Layout}

\begin{verbatim}
s3://zoo-data/
  {pod-name}/
    generations/
      {generation-id}/
        snapshots/
          {index}.snapshot.lz4
        wal/
          {index}.wal.lz4
\end{verbatim}

Each pod (e.g., \texttt{zoo-bridge-0}, \texttt{zoo-bridge-1})
replicates to a distinct prefix.  The \texttt{POD\_NAME}
environment variable is injected via the Kubernetes downward API
(\texttt{metadata.name} fieldRef).

