\section{Specification}


\subsection{Replicate Configuration}

\begin{lstlisting}[language=yaml,caption={replicate-config ConfigMap}]
dbs:
  - path: /app/data/data.db
    replicas:
      - type: s3
        bucket: zoo-data
        path: ${POD_NAME}
        endpoint: ${LITESTREAM_S3_ENDPOINT}
        region: sfo3
        force-path-style: true
        sync-interval: 1s
        snapshot-interval: 1h
\end{lstlisting}

\begin{itemize}
\item \texttt{sync-interval: 1s} --- WAL frames shipped every
  second, achieving $\leq 1$\,s RPO.
\item \texttt{snapshot-interval: 1h} --- Full snapshot every hour
  bounds restore time (restore replays at most 1 hour of WAL).
\item \texttt{force-path-style: true} --- Required for
  DigitalOcean Spaces and MinIO.
\end{itemize}

\subsection{Encryption}

WAL segments and snapshots are encrypted at rest using
\texttt{hanzoai/age} with per-environment X25519 key pairs:

\begin{lstlisting}[language=yaml,caption={Encryption configuration}]
dbs:
  - path: /app/data/data.db
    replicas:
      - type: s3
        bucket: zoo-data
        path: ${POD_NAME}
        encryption:
          type: age
          key: ${AGE_SECRET_KEY}
\end{lstlisting}

Key management:
\begin{itemize}
\item Encryption keys stored in Zoo KMS (K-Chain) or
  Kubernetes secrets (bootstrapping).
\item One key pair per environment (devnet, testnet, mainnet).
\item Key rotation: generate new pair, re-snapshot with new key,
  update secret.  Old WAL segments remain readable with old key
  until next snapshot.
\end{itemize}

\subsection{Service Applicability}

\begin{table}[h]
\centering
\caption{Zoo services and replication targets}
\begin{tabular}{@{}llll@{}}
\toprule
Service & Kind & Data & Replicate \\
\midrule
zoo-bridge & Deployment & SQLite (future) & Yes \\
zoo-gateway & CRD (ZooGateway) & Stateless & No \\
zoo-explorer & Deployment & PostgreSQL & No \\
zoo-chain & StatefulSet & Node data (PVC) & No\footnotemark \\
zoo-operator & Deployment & Stateless & No \\
sql & StatefulSet & PostgreSQL & No \\
kv & StatefulSet & Valkey & No \\
ingress & Deployment & Stateless & No \\
\bottomrule
\end{tabular}
\end{table}
\footnotetext{Chain node data replication uses PVC snapshots and
  chain sync, not WAL streaming.  Replicate is for SQLite only.}

As services migrate from PostgreSQL to Base/SQLite, they gain
replicate sidecars with zero code changes --- only a K8s patch.

\subsection{Recovery Guarantees}

\begin{description}
\item[RPO $\leq 1$\,s] WAL frames synced every second.
  In the worst case, the last second of writes is lost.

\item[RTO $\approx$ seconds] Init container restores latest
  snapshot ($\leq 1$ hour old) plus WAL replay.  Typical
  restore: 2--5 seconds for databases under 1\,GB.

\item[Durability] S3 provides 11 nines of durability.
  Snapshots are LZ4-compressed and optionally age-encrypted.

\item[Consistency] SQLite WAL mode guarantees atomic commits.
  Replicate ships complete WAL frames, preserving transaction
  boundaries.
\end{description}

