\section{Post-Quantum Security}


The Zoo ecosystem has achieved comprehensive post-quantum (PQ) coverage
across all cryptographic layers, consistent with LP-102 (Lux) and
HIP-0302 (Hanzo).  All previously planned mitigations are now deployed.

\subsection{PQ Scorecard}

\begin{table}[h]
\centering
\caption{Complete post-quantum scorecard.  All layers PQ-safe except EVM chain signing.}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Layer} & \textbf{Algorithm} & \textbf{PQ Status} & \textbf{Status} \\
\midrule
Disk encryption     & AES-256 (sqlcipher)          & Safe (128-bit PQ)    & Deployed \\
HKDF derivation     & HKDF-SHA-256                 & Safe (128-bit PQ)    & Deployed \\
Field encryption    & AES-256-GCM                  & Safe (128-bit PQ)    & Deployed \\
S3 backup           & age ML-KEM-768+X25519 (\texttt{age1pq}) & Safe (FIPS 203) & Deployed \\
TLS                 & X25519MLKEM768 (first curve) & Safe (hybrid PQ)     & Deployed \\
JWT signing         & ML-DSA-65 (FIPS 204)         & Safe (Module-LWE+SIS)& Deployed \\
JWKS validation     & ML-DSA-65                    & Safe                 & Deployed \\
MPC transport       & Hybrid KEM (X25519+ML-KEM-768) & Safe              & Deployed \\
MPC key shares      & PQ KEM encryption            & Safe                 & Deployed \\
Master keys         & Cloud HSM (FIPS 140-2 L3)    & Hardware isolation   & Deployed \\
Chain signing       & secp256k1 ECDSA              & \textbf{Not PQ-safe} & EVM constraint \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Deployed PQ Components}

\begin{description}
\item[S3 backup:] All age encryption uses ML-KEM-768 + X25519 hybrid
  recipients (\texttt{age1pq} prefix).  Label enforcement prevents
  downgrade to classical-only recipients.
\item[TLS:] X25519MLKEM768 as first curve in TLS 1.3 supported groups.
\item[JWT signing:] ML-DSA-65 (FIPS~204) for IAM token issuance.
  JWKS validation uses ML-DSA-65 public keys.
\item[MPC:] Hybrid KEM for transport; PQ KEM encryption for key shares.
\item[HSM:] Cloud HSM (FIPS 140-2 L3) for master keys.  Key material
  never leaves HSM boundary.  Three keyrings per ecosystem.
\end{description}

\subsection{Harvest-Now-Decrypt-Later: Closed}

The combination of ML-KEM-768 hybrid encryption (S3), X25519MLKEM768
TLS (transit), ML-DSA-65 JWT signing (auth), PQ KEM for MPC, and
Cloud HSM for master keys closes the harvest-now-decrypt-later attack
vector completely.  An adversary who captures data today cannot decrypt
it with a future quantum computer.

