\section{Conclusion}
\label{sec:conclusion}
%% ============================================================

\Beluga{} demonstrates that AI model economics benefit from application-specific blockchain infrastructure. By graduating from smart contracts on a general-purpose L2 to a sovereign L3 with custom precompiles, \Beluga{} achieves three properties simultaneously:

\begin{enumerate}
    \item \textbf{Performance}: Native precompiles reduce inference pricing from $\sim$500K gas to 5K gas, enable real-time pricing updates, and support 2,400 TPS for inference transactions.
    \item \textbf{Specialization}: Every transaction type (model registration, inference request, training bounty, data contribution, evaluation) has dedicated on-chain support with appropriate fee structures.
    \item \textbf{Composability}: Full EVM compatibility and Teleport bridging maintain interoperability with the Zoo DeFi ecosystem, enabling model-backed lending, inference derivatives, and cross-chain data markets.
\end{enumerate}

The migration path from Phase 1 (Zoo contracts) to Phase 3 (sovereign L3) provides a practical roadmap that validates each component before committing to independent block production. The fixed 100M \BLG{} supply, allocated across DAO treasury, inference rewards, training bounties, and DEX liquidity, creates a sustainable economy where value flows to participants who train models, curate data, host inference, evaluate outputs, and govern the network.

\Beluga{} is not an isolated experiment. It is one component of the Zoo ecosystem---alongside Proof of AI consensus, the Experience Ledger, Gym training infrastructure, and Agent NFTs---that collectively builds an open, community-owned AI economy on the \Lux{} Network.

%% ============================================================
%% References
%% ============================================================

\begin{thebibliography}{99}

% Zoo papers
\bibitem{zoo-poai}
Z.~Kelling, ``Proof of AI: A Consensus Mechanism for Verifiable Compute Contribution,'' Zoo Labs Foundation, 2024.

\bibitem{zoo-dso}
Z.~Kelling, ``Experience Ledger: Decentralized Semantic Optimization for Large Language Models,'' Zoo Labs Foundation, 2025.

\bibitem{zoo-experience-ledger}
Z.~Kelling, ``Experience Ledger: Content-Addressable Semantic Memory for AI Systems,'' Zoo Labs Foundation, 2025.

\bibitem{zoo-gym}
Z.~Kelling, ``Gym: Democratizing AI Model Training Through Comprehensive Open-Source Infrastructure,'' Zoo Labs Foundation, 2025.

\bibitem{zoo-agent-nft}
Z.~Kelling, ``Agent NFTs: A Token Standard for Autonomous AI Agents with Economic Agency,'' Zoo Labs Foundation, 2021.

\bibitem{zoo-tokenomics}
Z.~Kelling, ``Zoo Network Tokenomics: Fair Distribution Through Complete Airdrop,'' Zoo Labs Foundation, 2025.

% Lux papers
\bibitem{lux-quasar}
Lux Industries, ``Quasar: Post-HotStuff BFT Consensus with Sub-Second Finality,'' Lux Network Technical Report, 2024.

\bibitem{lux-teleport}
Lux Industries, ``Teleport: Trustless Cross-Chain Token Transfers via BLS Aggregate Signatures,'' Lux Network Technical Report, 2024.

\bibitem{lux-precompiles}
Lux Industries, ``Lux Precompile Addressing Standard (LP-9000),'' Lux Improvement Proposal, 2024.

\bibitem{lux-standard}
Lux Industries, ``@luxfi/standard: Smart Contract Library for Lux Network,'' \url{https://github.com/luxfi/standard}, 2024.

% Training methods
\bibitem{hu2021lora}
E.~J.~Hu et al., ``LoRA: Low-Rank Adaptation of Large Language Models,'' \textit{arXiv:2106.09685}, 2021.

\bibitem{dettmers2023qlora}
T.~Dettmers et al., ``QLoRA: Efficient Finetuning of Quantized Language Models,'' \textit{NeurIPS}, 2023.

\bibitem{shao2024deepseekmath}
Z.~Shao et al., ``DeepSeekMath: Pushing the Limits of Mathematical Reasoning in Open Language Models,'' \textit{arXiv:2402.03300}, 2024.

% AI compute networks
\bibitem{bittensor2023}
Opentensor Foundation, ``Bittensor: A Peer-to-Peer Intelligence Market,'' Whitepaper, 2023.

\bibitem{render2023}
Render Network, ``The Render Network Whitepaper,'' 2023.

\bibitem{akash2020}
Akash Network, ``Akash: The Decentralized Cloud,'' Whitepaper, 2020.

\bibitem{ocean2019}
Ocean Protocol Foundation, ``Ocean Protocol: Tools for the Web3 Data Economy,'' Whitepaper, 2019.

\bibitem{singularitynet2019}
SingularityNET, ``SingularityNET: A Decentralized, Open Market and Inter-Network for AIs,'' Whitepaper, 2019.

% App-chain precedent
\bibitem{dydx2023}
dYdX Foundation, ``dYdX v4: An Open-Source Protocol for Decentralized Perpetual Trading,'' 2023.

% Cryptography
\bibitem{boneh2001bls}
D.~Boneh et al., ``Short Signatures from the Weil Pairing,'' \textit{ASIACRYPT}, 2001.

\bibitem{lux-fhe}
Lux Industries, ``Fully Homomorphic Encryption on Lux K-Chain,'' Lux Network Technical Report, 2025.

\bibitem{lux-mpc}
Lux Industries, ``Threshold Multi-Party Computation for Lux Validator Sets,'' Lux Network Technical Report, 2025.

% Per-LLM chains and Quasar 3.0 stack
\bibitem{zoo-per-llm-chains}
Zoo Labs Foundation, ``Per-LLM Chains: Unified Training, Research, and Pooled GPU Hosting on Quasar 3.0,'' Zoo Whitepaper, December 2025.

\bibitem{lp-020-quasar-3}
Lux Foundation, ``LP-020: Quasar Consensus 3.0,'' Lux Improvement Proposal, December 2025.

\bibitem{lp-132-quasargpu}
Lux Foundation, ``LP-132: QuasarGPU Adapter for Deterministic GPU Scheduling,'' Lux Improvement Proposal, December 2025.

\bibitem{lp-133-app-stack}
Lux Foundation, ``LP-133: Quasar-Native App Stack,'' Lux Improvement Proposal, December 2025.

\bibitem{lp-134-topology}
Lux Foundation, ``LP-134: Lux Chain Topology and A-Chain/B-Chain Specification,'' Lux Improvement Proposal, December 2025.

\bibitem{lp-136-cloud}
Lux Foundation, ``LP-136: Lux Cloud Capacity Federation,'' Lux Improvement Proposal, December 2025.

\bibitem{hanzo-ai-chain}
Hanzo AI, ``The Hanzo AI Chain: Inference Marketplace and Commerce Layer,'' Hanzo Whitepaper, December 2025.

\end{thebibliography}

