\section{Related Work}

\label{sec:related}

\subsection{Reproducibility Infrastructure}

Computational reproducibility has received substantial infrastructure investment.
Workflow management systems such as Nextflow \citep{ditommaso2017nextflow}, Snakemake \citep{molder2021sustainable}, and Galaxy \citep{afgan2018galaxy} capture analysis pipelines as executable specifications.
Container technologies (Docker, Singularity) preserve software environments.
\citet{stodden2018empirical} found that even with these tools, only 26\% of computational results in published papers could be independently reproduced, primarily due to missing data, undocumented parameters, and environment-specific dependencies.

\subsection{Pre-Registration and Open Science}

Pre-registration requires researchers to specify hypotheses and analysis plans before data collection \citep{nosek2018preregistration}.
Registered Reports extend pre-registration by committing journals to publish results regardless of outcome \citep{chambers2013registered}.
While effective at reducing publication bias, these mechanisms rely on voluntary compliance and centralized platforms susceptible to single points of failure.

\subsection{Blockchain for Science}

Blockchain applications in science have focused on several areas: timestamping research claims \citep{koepsell2020blockchain}, managing intellectual property \citep{hasan2019blockchain}, decentralized peer review \citep{tenorio2019open}, and funding through decentralized autonomous organizations \citep{wright2021desci}.
\citet{bartling2020blockchain} surveyed the landscape and identified experiment management and reproducibility verification as underexplored application domains.

\subsection{Provenance Systems}

Data provenance systems track the lineage of computational artifacts.
The W3C PROV specification \citep{moreau2013prov} provides a standard data model for provenance.
\citet{mcphillips2015yesworkflow} introduced YesWorkflow for extracting provenance from scripts.
ProvDB \citep{miao2018provdb} stores provenance alongside data in relational databases.
These systems operate within single organizations and lack mechanisms for cross-institutional verification or economic incentive alignment.

% ============================================================
% 3. Protocol Design
% ============================================================
