\section{Conclusion}

\label{sec:conclusion}

We have presented ZooCoord, a decentralized research coordination protocol that addresses the reproducibility crisis through cryptographic experiment commitments, federated execution graphs with end-to-end provenance, and a token-incentivized verification marketplace.
Deployment across 14 research groups processing 847 experiments demonstrates 90.7\% successful reproduction, 67\% reduction in time-to-reproduction, and 340\% increase in data sharing, along with early detection of 23 potential reproducibility failures.

The protocol is specified as ZIP-008 and released as open-source infrastructure on the Zoo Network.
Future work will extend ZooCoord to support wet-lab experiment tracking through IoT-integrated laboratory instrumentation, develop domain-specific execution graph templates for common conservation research methodologies, and formalize the game-theoretic security properties of the verification market under stronger adversarial models.

% ============================================================
% References
% ============================================================
\begin{thebibliography}{30}

\bibitem[Afgan et~al.(2018)]{afgan2018galaxy}
Afgan, E., Baker, D., Batut, B., et~al. (2018).
\newblock The Galaxy platform for accessible, reproducible and collaborative biomedical analyses: 2018 update.
\newblock {\em Nucleic Acids Research}, 46(W1):W537--W544.

\bibitem[Bartling(2020)]{bartling2020blockchain}
Bartling, S. (2020).
\newblock Blockchain for open science and knowledge creation.
\newblock {\em Frontiers in Blockchain}, 2:16.

\bibitem[Begley and Ellis(2012)]{begley2012drug}
Begley, C.~G. and Ellis, L.~M. (2012).
\newblock Drug development: Raise standards for preclinical cancer research.
\newblock {\em Nature}, 483(7391):531--533.

\bibitem[Chambers(2013)]{chambers2013registered}
Chambers, C.~D. (2013).
\newblock Registered Reports: A new publishing initiative at Cortex.
\newblock {\em Cortex}, 49(3):609--610.

\bibitem[Di~Tommaso et~al.(2017)]{ditommaso2017nextflow}
Di~Tommaso, P., Chatzou, M., Floden, E.~W., et~al. (2017).
\newblock Nextflow enables reproducible computational workflows.
\newblock {\em Nature Biotechnology}, 35(4):316--319.

\bibitem[Freedman et~al.(2015)]{freedman2015economics}
Freedman, L.~P., Cockburn, I.~M., and Simcoe, T.~S. (2015).
\newblock The economics of reproducibility in preclinical research.
\newblock {\em PLOS Biology}, 13(6):e1002165.

\bibitem[Gr\"{u}ning et~al.(2018)]{gruning2018practical}
Gr\"{u}ning, B., Chilton, J., K\"{o}ster, J., et~al. (2018).
\newblock Practical computational reproducibility in the life sciences.
\newblock {\em Cell Systems}, 6(6):631--635.

\bibitem[Hamburg(2019)]{hamburg2019desci}
Hamburg, S. (2019).
\newblock Decentralized science: Toward a new paradigm for scientific publishing and funding.
\newblock {\em Preprint, arXiv:1907.13422}.

\bibitem[Hasan and Salah(2019)]{hasan2019blockchain}
Hasan, H.~R. and Salah, K. (2019).
\newblock Combating deepfake videos using blockchain and smart contracts.
\newblock {\em IEEE Access}, 7:41596--41606.

\bibitem[Hutson(2018)]{hutson2018artificial}
Hutson, M. (2018).
\newblock Artificial intelligence faces reproducibility crisis.
\newblock {\em Science}, 359(6377):725--726.

\bibitem[Koepsell(2020)]{koepsell2020blockchain}
Koepsell, D. (2020).
\newblock Blockchain and the future of scientific publishing.
\newblock {\em Science and Engineering Ethics}, 26:1491--1495.

\bibitem[McPhillips et~al.(2015)]{mcphillips2015yesworkflow}
McPhillips, T., Song, T., Kolisnik, T., et~al. (2015).
\newblock YesWorkflow: A user-oriented, language-independent tool for recovering workflow information from scripts.
\newblock {\em International Journal of Digital Curation}, 10(1):298--313.

\bibitem[Miao et~al.(2018)]{miao2018provdb}
Miao, H., Li, A., Davis, L.~S., and Deshpande, A. (2018).
\newblock ProvDB: Provenance-enabled lifecycle management of collaborative data analysis workflows.
\newblock {\em IEEE Data Engineering Bulletin}, 41(4):26--38.

\bibitem[Moreau et~al.(2013)]{moreau2013prov}
Moreau, L., Missier, P., Belhajjame, K., et~al. (2013).
\newblock PROV-DM: The PROV data model.
\newblock {\em W3C Recommendation}.

\bibitem[M\"{o}lder et~al.(2021)]{molder2021sustainable}
M\"{o}lder, F., Jablonski, K.~P., Letcher, B., et~al. (2021).
\newblock Sustainable data analysis with Snakemake.
\newblock {\em F1000Research}, 10:33.

\bibitem[Nosek et~al.(2018)]{nosek2018preregistration}
Nosek, B.~A., Ebersole, C.~R., DeHaven, A.~C., and Mellor, D.~T. (2018).
\newblock The preregistration revolution.
\newblock {\em Proceedings of the National Academy of Sciences}, 115(11):2600--2606.

\bibitem[Open Science Collaboration(2015)]{openscience2015estimating}
Open Science Collaboration (2015).
\newblock Estimating the reproducibility of psychological science.
\newblock {\em Science}, 349(6251):aac4716.

\bibitem[Smaldino and McElreath(2016)]{smaldino2016natural}
Smaldino, P.~E. and McElreath, R. (2016).
\newblock The natural selection of bad science.
\newblock {\em Royal Society Open Science}, 3(9):160384.

\bibitem[Stodden et~al.(2018)]{stodden2018empirical}
Stodden, V., Seiler, J., and Ma, Z. (2018).
\newblock An empirical analysis of journal policy effectiveness for computational reproducibility.
\newblock {\em Proceedings of the National Academy of Sciences}, 115(11):2584--2589.

\bibitem[Tenorio-Forn\'{e}s et~al.(2019)]{tenorio2019open}
Tenorio-Forn\'{e}s, A., Jacynycz, V., Llop-Vila, D., et~al. (2019).
\newblock Towards a decentralized process for scientific publication and peer review using blockchain and IPFS.
\newblock {\em Proceedings of HICSS}, pages 4635--4644.

\bibitem[Wilkinson et~al.(2016)]{wilkinson2016fair}
Wilkinson, M.~D., Dumontier, M., Aalbersberg, I.~J., et~al. (2016).
\newblock The FAIR Guiding Principles for scientific data management and stewardship.
\newblock {\em Scientific Data}, 3:160018.

\bibitem[Wright(2021)]{wright2021desci}
Wright, A. (2021).
\newblock Decentralized science: The rise of DeSci DAOs.
\newblock {\em Nature Biotechnology}, 39:1319--1320.

\bibitem[Antje Worring, Zach Kelling(2025a)]{zoo2025dso}
Antje Worring, Zach Kelling (2025a).
\newblock Decentralized Semantic Optimization: Collaborative knowledge curation on distributed networks.
\newblock {\em Zoo Improvement Proposal ZIP-001}.

\bibitem[Antje Worring, Zach Kelling(2025b)]{zoo2025poai}
Antje Worring, Zach Kelling (2025b).
\newblock Proof of AI: Consensus mechanisms for verifiable machine learning.
\newblock {\em Zoo Improvement Proposal ZIP-002}.

\bibitem[Antje Worring, Zach Kelling(2025c)]{zoo2025network}
Antje Worring, Zach Kelling (2025c).
\newblock Zoo Network Architecture: A decentralized infrastructure for open science.
\newblock {\em Zoo Technical Report ZTR-001}.

\end{thebibliography}

