\section{Sustainability}
\label{sec:sustainability}

Blockchain technology has the potential to be a game-changer when it comes to sustainability and the environment. Unlike traditional centralized banking systems, blockchain is built on a decentralized and distributed network that has the power to reduce carbon emissions and improve energy efficiency. According to a study conducted by the Cambridge Center for Alternative Finance, the annual electricity consumption of Bitcoin mining is only a fraction of the energy used by the traditional banking system.

The impact of human consumption on the environment cannot be disregarded, and companies must take responsibility for their environmental footprint. At Zoo Labs, we understand the importance of sustainability, and we have made it a priority in our approach to creating products and content. We believe that blockchain technology can have a positive impact on the environment and can be more sustainable than centralized banking.

Moreover, blockchain can be used to track and verify supply chains, which can help reduce waste and ensure that goods are produced sustainably. This is particularly important for industries such as agriculture and fashion, which have been known to contribute to deforestation, pollution, and other environmental problems.

Zoo is not just a game; it is also a sustainable investment opportunity for users. With NFTs, users can own in-game assets, which can be scarce, interoperable, and immutable. As the world becomes increasingly digitized, Zoo and its community will happily capitalize on saving the environment and drive NFT-game adoption to the general public.

As more companies and industries adopt blockchain technology, we can expect to see a shift towards more sustainable and eco-friendly practices. By leveraging the power of blockchain, we can create a more transparent and accountable economy that benefits both people and the planet for generations to come.

\section{Supporting Non-Profits}
\label{sec:nonprofits}

Zoo uses its digital platform to advocate for endangered species, managing resources via its DAO, expanding collaborations, and inspiring tangible conservation actions within its Game-Fi multiverse.

Zoo's dedication to the preservation of endangered species extends beyond the digital realm. We understand the crucial role that non-profit organizations play in safeguarding vulnerable animal species and are committed to forming meaningful partnerships with these organizations. By collaborating with non-profits that share our values and focus on the specific animals represented by our NFTs, we aim to generate awareness and support for their protection.

Through these partnerships, we strive to foster community collaboration and create compelling content that raises awareness and encourages action. Our platform serves as a catalyst for engagement and education, enabling us to amplify the conservation message and drive meaningful impact.

The Zoo DAO, as the governing body, takes responsibility for managing the treasury and ensuring that donations are allocated to impactful initiatives. The DAO's decision-making process prioritizes directing funds towards non-profit partnerships that align with our mission and have a proven track record in conservation efforts.

As Zoo continues to grow and evolve, we will forge new collaborations with non-profit organizations that share our passion for animal preservation. By expanding our network of partners, we can broaden our reach and actively contribute to initiatives aimed at protecting endangered species.

The metaverse-ready animals within the Zoo ecosystem serve as powerful ambassadors for conservation. By leveraging the creative capabilities and engagement of our community, we can make a lasting impact on the preservation of endangered species. The DAO remains steadfast in its commitment to the core objective of saving these animals, ensuring that the projects and initiatives we support align with this mission.

\section{Zoo Foundation: Conservation \texorpdfstring{$\times$}{x} Education}
\label{sec:foundation}

The Zoo Foundation was established to harness blockchain technology and the global community it connects to help preserve our world's most critically endangered animal species. We want to end the extinction of animals and we want to help the world see how important they are to all of us. We also want to inspire people to change animals' lives for the better. We are deeply committed to creating digital twins of endangered species to raise awareness, enact change, and create an animal sanctuary for all Zoo NFT holders.

Through the Zoo Foundation, Zoo Labs will provide support to our many partners in wildlife conservation. This support will come through monetary contribution and by providing opportunities for people of all ages, anywhere on Earth, to learn about the beautiful creatures the foundation works towards preserving.

The Zoo Foundation will also host fundraising events with our partners in wildlife conservation and other strategic partners to promote donation and awareness worldwide. Through financial and human support we can furthermore raise awareness and mitigate the devastating effects of the illegal wildlife trade and how people can get involved in stopping it.

It is our driving purpose to deepen the connection humans have with animals by creating sustainable animal sanctuaries for endangered species. We will share the Zoo Sanctuary with local visitors and educators, host fundraisers and events, and get the Zoo Community as involved as possible in saving endangered animals all over the world. The Zoo protocol allows users to directly save a life through a monthly or one-time contribution.
