\section{2025 Update --- Securities, Quantum-Secure Settlement, Holographic DAO, and Open Information}
\label{sec:2025-update}

\textit{This section is an editorial addendum dated 2025-12-15, four
years after the original whitepaper. It exists to surface five planks
that descend from the founding vision but were not specified in the
2021 text. The original document above is unchanged; this addendum is
the canonical bridge from the founder vision to the activation state
on 2025-12-25.}

\subsection{Updated team and mission (2025)}

\textbf{Founder}: Antje Worring (author of this October~31, 2021
whitepaper). \textbf{Chief Scientist}: Zach Kelling. \textbf{Mission} (as
of 2025-12-15): \emph{An open AI protocol for research aligned with
environment-saving, endangered AI, and the intersection of AI,
blockchain, metaverse, and gaming.} Full team page is the canonical
\texttt{team.md} in the Zoo \textsc{zip} repository. Allied
organizations: Lux Network (chain infrastructure), Hanzo AI (Techstars
'17, AI compute and Zen LM co-development), and a regulated securities
partner (exclusive partner for retail-scale digital-securities trading).

\subsection{\$113 trillion digital-securities market}

The total addressable market for digital securities globally is
estimated at approximately \textbf{\$113 trillion} (BIS / WFE 2025
estimates of global capitalization across equities, fixed income,
derivatives, and tokenized real-world assets). Zoo positions itself as
the open AI / wildlife / metaverse layer in this stack. Lux provides
the regulated chain infrastructure. \textbf{A regulated securities partner} is Zoo's
exclusive partner for retail-scale securities trading, carrying the
regulated ATS / broker-dealer / transfer-agent perimeter.

\paragraph{Tagline.} \emph{Robinhood-style capabilities for everyone ---
actually make money like the 1\% finally.}

\paragraph{Concrete capabilities.} Zero-commission trading, fractional
shares to the cent, instant settlement on Quasar 3.0, tokenized
real-world securities (stocks, bonds, commodities, REITs, private
equity), 24/7 markets, and a public verifiable order book. Compliance
(KYC, AML, suitability, best execution) lives at the regulated securities partner's
perimeter; the chain is the audit-eligible settlement layer.

\paragraph{Honest framing.} \$113T is a market-size figure, not a
Zoo-specific TVL. Zoo Labs Foundation is a 501(c)(3) and does not
operate a regulated market center; the regulated securities partner does.

\subsection{Quantum-secure settlement (Quasar 3.0)}

Every securities trade, every Zoo block, every NFT mint, and every DAO
vote settles under \textbf{Quasar 3.0} (Lux LP-020). Quasar 3.0
composes three independent cryptographic hardness assumptions:

\begin{itemize}[nosep,leftmargin=1.2em]
\item BLS12-381 aggregate signature (classical), Lux LP-075.
\item Corona Ring-LWE threshold signature (post-quantum lattice),
   Lux LP-073.
\item ML-DSA-65 identity proof via Z-Chain Groth16 rollup
   (post-quantum lattice with succinct proof), Lux LP-070.
\end{itemize}

An adversary must break \emph{all three} to forge a Lux/Zoo block. A
BLS compromise on Q-day does not break finality: the Corona +
ML-DSA pair stays secure. M-Chain (Lux LP-019) holds threshold custody
of governance and settlement keys; F-Chain (Lux LP-013) provides FHE
substrate. The retail user can therefore truthfully say: trade
post-quantum-safe, custody post-quantum-safe, audit-trail
post-quantum-safe.

\subsection{Homomorphic / Holographic Consensus DAO}

The Zoo DAO described in §19 of the 2021 text now formalizes as a
\emph{homomorphic} (votes tallied while encrypted via F-Chain FHE,
Lux LP-013) and \emph{holographic} (every sub-DAO is structurally a
copy of the meta-DAO; per-LLM chains, research verticals, and regional
chapters all inherit the same governance topology) consensus DAO.

Voting weight is

\[
\text{weight}(p) \;=\; \alpha \cdot a_p \;+\; \beta \cdot i_p \;+\; \gamma \cdot c_p \;+\; \delta \cdot s_p
\]

\noindent with $\alpha + \beta + \gamma + \delta = 1$, where the four
axes are advocacy ($a_p$), involvement ($i_p$), contribution ($c_p$),
and token stake ($s_p$). Default initial weights at 2025-12-25:
$(0.20, 0.30, 0.40, 0.10)$. Token stake is bounded at 10\% to prevent
plutocracy. Anti-Sybil: contribution and involvement scores require
TEE attestation through A-Chain (Lux LP-134); advocacy must be on
attestation-rooted social platforms. Threshold custody of governance
keys lives on M-Chain MPC (Lux LP-019).

\subsection{Democratized end-to-end access to information}

The Open AI Protocol named in §\ref{sec:open-ai-protocol} now operates
as an end-to-end-verifiable pipeline:

\begin{enumerate}[nosep,leftmargin=1.4em]
\item User --- proves eligibility via zero-knowledge selective
   disclosure (Z-Chain Groth16 attestations).
\item Model --- weights, architecture, and training-data lineage all
   committed on-chain via Zoo per-LLM chains.
\item Data --- every training shard's hash is on the receipts tree;
   contributors are attributable.
\item Output --- every inference call produces a TEE-attested receipt
   anchored on A-Chain via the Hanzo AI Chain.
\item Reward --- royalties flow back along the contribution graph
   automatically.
\end{enumerate}

Closed labs (OpenAI, Anthropic) treat weights, data, and architecture
as competitive secrets. Zoo treats them as public goods whose
provenance is the protocol. This is the durable distinction between
the open AI protocol named in the 2021 founding vision and the
proprietary AI services that came after.

\subsection{Cross-references}

\begin{itemize}[nosep,leftmargin=1.2em]
\item Companion paper: \emph{Zoo 2025: Digital Securities,
   Quantum-Secure Settlement, and the Holographic Consensus DAO}
   (full-length treatment of the five planks).
\item Companion paper: \emph{Zoo Per-LLM Chains: One Sovereign Chain
   per Model}.
\item Lux LP-020 (Quasar 3.0), LP-013 (F-Chain), LP-019 (M-Chain),
   LP-070 (ML-DSA), LP-073 (Corona), LP-075 (BLS), LP-105 (Quasar
   3.0 paper), LP-134 (A/B/M/F topology).
\item Hanzo AI Chain whitepaper.
\item Zoo \textsc{zip-0017} (DAO governance), \textsc{zip-0026}
   (reputation), \textsc{zip-0570} (impact thesis).
\end{itemize}

\medskip
\noindent\textit{Editorial addendum closes 2025-12-15.}
