\section{Conclusion}

\label{sec:conclusion}

We have presented ZooEd, a decentralized adaptive learning system for conservation education that combines Bayesian knowledge tracing, ecologically grounded spaced repetition, and multi-modal multi-language assessment.
A 14-month deployment across 23 sites in 11 countries demonstrates substantial improvements in knowledge retention (34\%), field task transfer (28\%), and learner engagement (41\%) compared to traditional curricula.

Key technical contributions include the Conservation Competency Ontology spanning 2,847 learning objectives, the SREC algorithm that leverages IoT sensor data for contextual reinforcement, and a calibrated LLM-based assessment pipeline supporting 47 languages.

All components are released under the Zoo Improvement Proposal framework as ZIP-007, with the aspiration that adaptive, accessible conservation education becomes a global public good.
Future work will extend the system to support collaborative learning scenarios, integrate augmented reality for field-based instruction, and develop specialized modules for climate change adaptation education.

% ============================================================
% References
% ============================================================
\begin{thebibliography}{30}

\bibitem[Ardoin et~al.(2020)]{ardoin2020environmental}
Ardoin, N.~M., Bowers, A.~W., and Gaillard, E. (2020).
\newblock Environmental education outcomes for conservation: A systematic review.
\newblock {\em Biological Conservation}, 241:108224.

\bibitem[Artetxe et~al.(2020)]{artetxe2020cross}
Artetxe, M., Ruder, S., and Yogatama, D. (2020).
\newblock On the cross-lingual transferability of monolingual representations.
\newblock {\em Proceedings of ACL}, pages 4623--4637.

\bibitem[Brewer(2020)]{brewer2020conservation}
Brewer, C. (2020).
\newblock Conservation education partnerships in practice.
\newblock {\em Conservation Biology}, 34(3):701--710.

\bibitem[Corbett and Anderson(1994)]{corbett1994knowledge}
Corbett, A.~T. and Anderson, J.~R. (1994).
\newblock Knowledge tracing: Modeling the acquisition of procedural knowledge.
\newblock {\em User Modeling and User-Adapted Interaction}, 4(4):253--278.

\bibitem[Couldry and Mejias(2019)]{couldry2019data}
Couldry, N. and Mejias, U.~A. (2019).
\newblock {\em The Costs of Connection: How Data Is Colonizing Human Life and Appropriating It for Capitalism}.
\newblock Stanford University Press.

\bibitem[D\'{i}az et~al.(2019)]{diaz2019global}
D\'{i}az, S., Settele, J., and Brondizio, E.~S. (2019).
\newblock Pervasive human-driven decline of life on Earth points to the need for transformative change.
\newblock {\em Science}, 366(6471):eaax3100.

\bibitem[Fan et~al.(2021)]{fan2021beyond}
Fan, A., Bhosale, S., and Schwenk, H. (2021).
\newblock Beyond English-centric multilingual machine translation.
\newblock {\em Journal of Machine Learning Research}, 22(107):1--48.

\bibitem[Fisher et~al.(2022)]{fisher2022growth}
Fisher, R., O'Leary, R.~A., and Low-Choy, S. (2022).
\newblock The growth and structure of conservation science.
\newblock {\em Conservation Science and Practice}, 4(1):e12594.

\bibitem[Ghosh et~al.(2020)]{ghosh2020context}
Ghosh, A., Heffernan, N., and Lan, A.~S. (2020).
\newblock Context-aware attentive knowledge tracing.
\newblock {\em Proceedings of KDD}, pages 2330--2339.

\bibitem[Grech and Camilleri(2017)]{grech2017blockchain}
Grech, A. and Camilleri, A.~F. (2017).
\newblock Blockchain in education.
\newblock {\em JRC Science for Policy Report}.

\bibitem[Horn et~al.(2018)]{horn2018inaturalist}
Horn, G.~V., Aodha, O.~M., and Song, Y. (2018).
\newblock The iNaturalist species classification and detection dataset.
\newblock {\em Proceedings of CVPR}, pages 8769--8778.

\bibitem[Johnson et~al.(2018)]{johnson2018wildlife}
Johnson, M.~F., Hannah, C., and Acton, L. (2018).
\newblock Wildlife identification apps and conservation education.
\newblock {\em People and Nature}, 1(2):123--135.

\bibitem[Kulik and Fletcher(2016)]{kulik2016effectiveness}
Kulik, J.~A. and Fletcher, J.~D. (2016).
\newblock Effectiveness of intelligent tutoring systems: A meta-analytic review.
\newblock {\em Review of Educational Research}, 86(1):42--78.

\bibitem[Maffi(2005)]{maffi2005linguistic}
Maffi, L. (2005).
\newblock Linguistic, cultural, and biological diversity.
\newblock {\em Annual Review of Anthropology}, 34:599--617.

\bibitem[Piech et~al.(2015)]{piech2015deep}
Piech, C., Spencer, J., and Huang, J. (2015).
\newblock Deep knowledge tracing.
\newblock {\em Advances in Neural Information Processing Systems}, 28.

\bibitem[Rousell and Cater-Steel(2019)]{rousell2019gamification}
Rousell, D. and Cater-Steel, A. (2019).
\newblock Gamification in environmental education: A systematic review.
\newblock {\em Environmental Education Research}, 25(7):1021--1037.

\bibitem[Settles and Meeder(2016)]{settles2016trainable}
Settles, B. and Meeder, B. (2016).
\newblock A trainable spaced repetition model for language learning.
\newblock {\em Proceedings of ACL}, pages 1848--1858.

\bibitem[Sharples and Domingue(2016)]{sharples2016blockchain}
Sharples, M. and Domingue, J. (2016).
\newblock The blockchain and kudos: A distributed system for educational record, reputation and reward.
\newblock {\em Proceedings of EC-TEL}, pages 490--496.

\bibitem[Shin et~al.(2021)]{shin2021saint}
Shin, D., Shim, Y., and Yu, H. (2021).
\newblock SAINT+: Integrating temporal features for EdNet correctness prediction.
\newblock {\em LAK21 Conference Proceedings}, pages 490--496.

\bibitem[Stern et~al.(2014)]{stern2014value}
Stern, M.~J., Powell, R.~B., and Hill, D. (2014).
\newblock Environmental education program evaluation in the new millennium.
\newblock {\em Environmental Education Research}, 20(5):581--611.

\bibitem[VanLehn(2011)]{vanlehn2011relative}
VanLehn, K. (2011).
\newblock The relative effectiveness of human tutoring, intelligent tutoring systems, and other tutoring systems.
\newblock {\em Educational Psychologist}, 46(4):197--221.

\bibitem[Zhang et~al.(2017)]{zhang2017dynamic}
Zhang, J., Shi, X., and King, I. (2017).
\newblock Dynamic key-value memory networks for knowledge tracing.
\newblock {\em Proceedings of WWW}, pages 765--774.

\bibitem[Antje Worring, Zach Kelling(2025a)]{zoo2025dso}
Antje Worring, Zach Kelling (2025a).
\newblock Decentralized Semantic Optimization: Collaborative knowledge curation on distributed networks.
\newblock {\em Zoo Improvement Proposal ZIP-001}.

\bibitem[Antje Worring, Zach Kelling(2025b)]{zoo2025poai}
Antje Worring, Zach Kelling (2025b).
\newblock Proof of AI: Consensus mechanisms for verifiable machine learning.
\newblock {\em Zoo Improvement Proposal ZIP-002}.

\bibitem[Antje Worring, Zach Kelling(2025c)]{zoo2025network}
Antje Worring, Zach Kelling (2025c).
\newblock Zoo Network Architecture: A decentralized infrastructure for open science.
\newblock {\em Zoo Technical Report ZTR-001}.

\end{thebibliography}

