\section*{Acknowledgments}


We thank the Snapshot Serengeti team, Wildlife Insights, and the LILA BC initiative for providing benchmark datasets. This work was supported by the Antje Worring, Zach Kelling and the Hanzo AI research division.

\begin{thebibliography}{40}

\bibitem{wwf2022}
WWF, ``Living Planet Report 2022: Building a Nature-Positive Society,'' World Wildlife Fund, 2022.

\bibitem{iucn2023}
IUCN, ``The IUCN Red List of Threatened Species,'' Version 2023-1, 2023.

\bibitem{steenweg2017}
R.~Steenweg, M.~Hebblewhite, R.~Kays, et al., ``Scaling-up camera traps: monitoring the planet's biodiversity with networks of remote sensors,'' \emph{Frontiers in Ecology and the Environment}, vol.~15, no.~1, pp.~26--34, 2017.

\bibitem{swanson2015}
A.~Swanson, M.~Kosmala, C.~Lintott, et al., ``Snapshot Serengeti, high-frequency annotated camera trap images of 40 mammalian species in an African savanna,'' \emph{Scientific Data}, vol.~2, p.~150026, 2015.

\bibitem{tabak2019}
M.~A.~Tabak, M.~S.~Norouzzadeh, D.~W.~Wolfson, et al., ``Machine learning to classify animal species in camera trap images: Applications in ecology,'' \emph{Methods in Ecology and Evolution}, vol.~10, no.~4, pp.~585--590, 2019.

\bibitem{beery2019}
S.~Beery, G.~Van Horn, and P.~Perona, ``Recognition in Terra Incognita,'' in \emph{ECCV}, 2018, pp.~456--473.

\bibitem{beery2018}
S.~Beery, G.~Van Horn, and P.~Perona, ``Recognition in Terra Incognita,'' in \emph{Proceedings of the European Conference on Computer Vision}, 2018.

\bibitem{norouzzadeh2018}
M.~S.~Norouzzadeh, A.~Nguyen, M.~Kosmala, et al., ``Automatically identifying, counting, and describing wild animals in camera-trap images with deep learning,'' \emph{PNAS}, vol.~115, no.~25, pp.~E5716--E5725, 2018.

\bibitem{beery2019mega}
S.~Beery, D.~Morris, and S.~Yang, ``Efficient Pipeline for Camera Trap Image Review,'' arXiv:1907.06772, 2019.

\bibitem{willi2019}
M.~Willi, R.~T.~Pitman, A.~W.~Cardoso, et al., ``Identifying animal species in camera trap images using deep learning and citizen science,'' \emph{Methods in Ecology and Evolution}, vol.~10, no.~1, pp.~80--91, 2019.

\bibitem{dosovitskiy2021}
A.~Dosovitskiy, L.~Beyer, A.~Kolesnikov, et al., ``An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale,'' in \emph{ICLR}, 2021.

\bibitem{pantazis2021}
O.~Pantazis, G.~Brostow, K.~E.~Jones, and O.~Mac~Aodha, ``Focus on the positives: self-supervised learning for biodiversity monitoring,'' in \emph{ICCV}, 2021.

\bibitem{weinstein2022}
B.~G.~Weinstein, S.~Marconi, S.~Bohlman, A.~Zare, and E.~P.~White, ``A general deep learning model for bird detection in high-resolution airborne imagery,'' \emph{Ecological Applications}, vol.~32, no.~8, p.~e2694, 2022.

\bibitem{jolly1965}
G.~M.~Jolly, ``Explicit estimates from capture-recapture data with both death and immigration-stochastic model,'' \emph{Biometrika}, vol.~52, no.~1/2, pp.~225--247, 1965.

\bibitem{seber1965}
G.~A.~F.~Seber, ``A note on the multiple-recapture census,'' \emph{Biometrika}, vol.~52, no.~1/2, pp.~249--259, 1965.

\bibitem{buckland2001}
S.~T.~Buckland, D.~R.~Anderson, K.~P.~Burnham, J.~L.~Laake, D.~L.~Borchers, and L.~Thomas, \emph{Introduction to Distance Sampling}, Oxford University Press, 2001.

\bibitem{mackenzie2002}
D.~I.~MacKenzie, J.~D.~Nichols, G.~B.~Lachman, S.~Droege, J.~A.~Royle, and C.~A.~Langtimm, ``Estimating site occupancy rates when detection probabilities are less than one,'' \emph{Ecology}, vol.~83, no.~8, pp.~2248--2255, 2002.

\bibitem{karanth1998}
K.~U.~Karanth, ``Estimating tiger \emph{Panthera tigris} populations from camera-trap data using capture-recapture models,'' \emph{Biological Conservation}, vol.~71, no.~3, pp.~333--338, 1998.

\bibitem{schneider2020}
S.~Schneider, G.~W.~Taylor, and S.~Kremer, ``Similarity learning networks for animal individual re-identification,'' in \emph{WACV}, 2020.

\bibitem{chandler2011}
R.~B.~Chandler and J.~A.~Royle, ``Spatially explicit models for inference about density in unmarked or partially marked populations,'' \emph{The Annals of Applied Statistics}, vol.~5, no.~2A, pp.~1351--1376, 2011.

\bibitem{tucker1979}
C.~J.~Tucker, ``Red and photographic infrared linear combinations for monitoring vegetation,'' \emph{Remote Sensing of Environment}, vol.~8, no.~2, pp.~127--150, 1979.

\bibitem{huete2002}
A.~Huete, K.~Didan, T.~Miura, et al., ``Overview of the radiometric and biophysical performance of the MODIS vegetation indices,'' \emph{Remote Sensing of Environment}, vol.~83, no.~1-2, pp.~195--213, 2002.

\bibitem{mcgarigal2012}
K.~McGarigal, S.~A.~Cushman, and E.~Ene, ``FRAGSTATS v4: Spatial Pattern Analysis Program for Categorical Maps,'' University of Massachusetts, Amherst, 2012.

\bibitem{helber2019}
P.~Helber, B.~Bischke, A.~Dengel, and D.~Borth, ``EuroSAT: A Novel Dataset and Deep Learning Benchmark for Land Use and Land Cover Classification,'' \emph{IEEE JSTARS}, vol.~12, no.~7, pp.~2217--2226, 2019.

\bibitem{ortega2021}
S.~Ortega~Adarme, M.~Queiroz~Feitosa, C.~Aparecido~De~Almeida, and G.~A.~O.~P.~Costa, ``Evaluation of Deep Learning Techniques for Deforestation Detection in the Brazilian Amazon and Cerrado Biomes,'' \emph{Remote Sensing}, vol.~12, no.~6, p.~910, 2020.

\bibitem{mcrae2008}
B.~H.~McRae, B.~G.~Dickson, T.~H.~Keitt, and V.~B.~Shah, ``Using circuit theory to model connectivity in ecology, evolution, and conservation,'' \emph{Ecology}, vol.~89, no.~10, pp.~2712--2724, 2008.

\bibitem{dalal2005}
N.~Dalal and B.~Triggs, ``Histograms of oriented gradients for human detection,'' in \emph{CVPR}, 2005, pp.~886--893.

\bibitem{lowe2004}
D.~G.~Lowe, ``Distinctive image features from scale-invariant keypoints,'' \emph{International Journal of Computer Vision}, vol.~60, no.~2, pp.~91--110, 2004.

\bibitem{tan2020}
M.~Tan, R.~Pang, and Q.~V.~Le, ``EfficientDet: Scalable and Efficient Object Detection,'' in \emph{CVPR}, 2020.

\bibitem{tan2019}
M.~Tan and Q.~V.~Le, ``EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks,'' in \emph{ICML}, 2019.

\bibitem{guo2017}
C.~Guo, G.~Pleiss, Y.~Sun, and K.~Q.~Weinberger, ``On Calibration of Modern Neural Networks,'' in \emph{ICML}, 2017.

\bibitem{efford2004}
M.~Efford, ``Density estimation in live-trapping studies,'' \emph{Oikos}, vol.~106, no.~3, pp.~598--610, 2004.

\bibitem{zoo2026commons}
Antje Worring, Zach Kelling, ``Zoo Data Commons: A Decentralized Repository for Open Biodiversity Datasets,'' Antje Worring, Zach Kelling Paper ZOO-2026-011, 2026.

\end{thebibliography}

