\section{V1 Species Catalogue and Conservation Partners}
\label{sec:catalogue}

The 2021 whitepaper enumerated eight V1 species (\S18) with associated
non-profit partners. The 2025-12-15 catalogue retains those species as
foundational and adds Generation~1 candidates through the \DAO{} (\textsc{zip-0030},
On-Chain Species Registry).

\subsection{V1 Catalogue (2021 Genesis Cohort)}

\begin{table}[h]
\centering
\small
\begin{tabular}{p{4cm}p{4cm}p{6cm}}
\toprule
\textbf{Species (Common)} & \textbf{Scientific} & \textbf{2025-12-15 Partner Set} \\
\midrule
Red Wolf & \emph{Canis rufus} & Wolf Conservation Center, Defenders of Wildlife. \\
Amur Leopard & \emph{Panthera pardus orientalis} & WWF Russia, ALTA (Amur Leopard \& Tiger Alliance). \\
Nubian Giraffe & \emph{Giraffa camelopardalis camelopardalis} & Giraffe Conservation Foundation. \\
Sumatran Elephant & \emph{Elephas maximus sumatranus} & WWF Indonesia, Sumatran Elephant Conservation Initiative. \\
Pygmy Hippopotamus & \emph{Choeropsis liberiensis} & ZSL West Africa Programme, IUCN SSC Pygmy Hippo Specialist Group. \\
Javan Rhinoceros & \emph{Rhinoceros sondaicus} & International Rhino Foundation, Save the Rhino. \\
Siberian (Amur) Tiger & \emph{Panthera tigris altaica} & WCS Russia, Panthera. \\
Emperor Penguin & \emph{Aptenodytes forsteri} & British Antarctic Survey programmes; Pew Antarctic Project. \\
\bottomrule
\end{tabular}
\caption{V1 \NFT{} catalogue with current partner set. Partner status
is reviewed quarterly by the \DAO{} under \textsc{zip-0530}
(Community Partnerships, \textsc{fpic}).}
\end{table}

Every catalogue species has a corresponding empathetic AI companion
(\textsc{zip-0402}) and a Habitat \NFT{} for fractional conservation
funding (\textsc{zip-0203}). Generation~1 candidates --- additional
endangered species nominated by the community --- are admitted through
the On-Chain Species Registry (\textsc{zip-0030}) and require a Tier 2
non-profit partnership commitment.

\subsection{Selection Criteria}

\begin{enumerate}[nosep,leftmargin=1.4em]
\item IUCN Red List category Vulnerable, Endangered, or Critically
  Endangered.
\item At least one named non-profit field operation partner with
  established track record.
\item No legal blockers to grant disbursement in partner jurisdiction.
\item Available genomic / population data for the Conservation AI
  Pillar (\textsc{zip-0500} series).
\end{enumerate}

\subsection{Multilateral Alignment}

The catalogue and grant cycles align with multilateral conservation
frameworks documented in \textsc{zip-0550} (Conservation Standards
Alignment): IUCN Red List, CITES Appendices, and the Global Biodiversity
Framework targets. Compliance with each framework is on-chain
attested through A-Chain (\textsc{lp-134}).
