\section{Treasury and Grant Cycles}
\label{sec:treasury}

\subsection{Custody}

The Zoo Foundation Treasury is held under M-Chain MPC threshold custody
(Lux \textsc{lp-019}) with a 2-of-3 split between board signatories,
the Foundation operations multisig, and a community-elected
\textsc{dao}{} representative. Every disbursement is co-signed and
emits an on-chain receipt anchored on Q-Chain (\Quasar{} 3.0,
\textsc{lp-020}).

\subsection{Funding Sources}

Treasury inflows come from four channels:

\begin{enumerate}[nosep,leftmargin=1.4em]
\item \textbf{Sustainability tax} --- 1--3\% on every protocol
  transaction (see \texttt{contracts/specs/zoo-sustainability-tax.md}).
\item \textbf{Direct micro-donations} --- streaming via
  \textsc{zip-0112}.
\item \textbf{Conservation bonds} --- yield-bearing tokenized donations
  (\textsc{zip-0101}).
\item \textbf{Validator-rewards conservation split} --- a percentage of
  validator rewards (\textsc{zip-0019}).
\end{enumerate}

\subsection{Grant Cycle (Quarterly)}

Each cycle follows the framework in \textsc{zip-0023} (Community Grant
Program) and \textsc{zip-0104} (Research Funding \textsc{dao} Treasury):

\begin{enumerate}[nosep,leftmargin=1.4em]
\item \textbf{Open call.} Proposals submitted on-chain; gated by DID
  Tier (\textsc{zip-0033}) and minimum reporting commitment.
\item \textbf{Review.} Domain reviewers (catalogue partners, in-situ
  ecologists, citizen-science contributors) score proposals.
  Reviewers' DIDs are published; reputation drift is bounded by
  \textsc{zip-0026}.
\item \textbf{Vote.} Weighted vote per the 2021 paper \S28 formula:
  $\text{weight}(p) = \alpha \cdot a_p + \beta \cdot i_p + \gamma \cdot c_p + \delta \cdot s_p$
  with stake capped at 10\% of weight.
\item \textbf{Disbursement.} M-Chain co-signs payment; on-chain
  receipt anchors evidence to the Evidence Locker
  (\textsc{zip-0560}).
\item \textbf{Reporting.} Recipients commit to the Reproducibility
  Attestation framework (\textsc{zip-0606}) and a 90-day field-results
  report.
\end{enumerate}

\subsection{Allocation Limits}

\begin{itemize}[nosep,leftmargin=1.2em]
\item Operations expense capped at 15\% of treasury per cycle (the
  remainder must be programmatic or in-situ).
\item Per-grant cap of 10\% of treasury (avoids single-project
  concentration).
\item Reserve floor of 25\% retained against multi-year obligations.
\end{itemize}

The \textsc{dao}{} can override these limits only by supermajority vote
under \textsc{zip-0017} (Treasury Spend $\geq 60\%$, $\geq 10\%$
quorum, 72-hour timelock).
