\section{Impact Measurement}
\label{sec:impact}

\subsection{Framework}

Every grant cycle's outcome is measured under the Conservation Impact
Measurement framework (\textsc{zip-0501}) and indexed in the Evidence
Locker (\textsc{zip-0560}). Impact data is published as open
biodiversity data (\textsc{zip-0601}, Open Biodiversity Database
Standard).

\subsection{Verifiable Metrics}

\begin{table}[h]
\centering
\small
\begin{tabular}{p{5cm}p{4cm}p{5.5cm}}
\toprule
\textbf{Metric} & \textbf{Source} & \textbf{Verification} \\
\midrule
Population census delta & Field partner & Multi-observer attestation; satellite cross-check (\textsc{zip-0425}). \\
Habitat hectares restored & Reforestation log & Satellite verification (\textsc{zip-0505}). \\
Anti-poaching incidents prevented & Anti-poaching alert network & On-chain alert log (\textsc{zip-0503}). \\
Educational reach & Avatar-tutor sessions (\textsc{zip-0008}) & Session attestations on A-Chain (\textsc{lp-134}). \\
Citizen-science contributions & Citizen-science protocol (\textsc{zip-0602}) & Reproducibility attestation (\textsc{zip-0606}). \\
\bottomrule
\end{tabular}
\caption{Verifiable impact metrics. Every metric has a defined source,
a defined verifier, and a defined chain anchor.}
\end{table}

\subsection{Annual Impact Report}

The Foundation publishes an annual impact report each calendar year
end. The 2025 edition is the first operating under \Quasar{} 3.0 and
the first end-to-end-verifiable report in the Foundation's history.

\subsection{Open Data}

All non-sensitive impact data is published under \textsc{zip-0601}
(Open Biodiversity Database Standard), licensed CC-BY for research
reuse. Sensitive location data (e.g., poaching-vulnerable species
coordinates) is published behind selective-disclosure attestations
(Z-Chain, \textsc{lp-134}) gated by researcher DID tier.

\subsection{Cross-References}

\begin{itemize}[nosep,leftmargin=1.2em]
\item 2021 Whitepaper \S5--7 (Sustainability, Non-Profits, Foundation).
\item Companion paper: \emph{Zoo Foundation Mission} (Pillars I/II/III).
\item Companion paper: \emph{Zoo Conservation AI}.
\item \textsc{zip-0500} series (Conservation).
\item \textsc{zip-0570} (Zoo Labs Impact Thesis).
\item \texttt{contracts/specs/zoo-sustainability-tax.md}.
\item Lux \textsc{lp-019} (M-Chain), \textsc{lp-020} (\Quasar{} 3.0),
  \textsc{lp-134} (chain topology).
\end{itemize}

\begin{thebibliography}{9}
\bibitem{zoo-2021} A.~Worring, \emph{Zoo: A Game-Fi Multiverse Ecosystem
for the Preservation of Endangered Species}, Zoo Labs Foundation,
October 31, 2021. Reconstructed 2025-12-15.
\end{thebibliography}
