\documentclass[11pt,a4paper]{article}
\usepackage{shared/zoocover}

% ── Packages ──────────────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
% geometry loaded by zoocover
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric}
\usepackage{float}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{fancyhdr}
\usepackage{abstract}
\usepackage{multirow}

% ── Page Style ────────────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Zach Kelling}
\fancyhead[R]{\small Verifiable Carbon Credits}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ── Metadata ──────────────────────────────────────────────────────────────────
\title{%
  \textbf{Verifiable Carbon Credits: MRV Infrastructure\\
  for Nature-Based Solutions}\\[0.5em]
  \large Zach Kelling Paper ZOO-MRV-2026
}

\author{
  Zach Kelling\\
  \texttt{research@zoo.ngo}\\[0.5em]
  \small With contributions from the Zoo Open Science Network
}

\date{June 2025}

% ── Macros ────────────────────────────────────────────────────────────────────
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Loss}{\mathcal{L}}
\newcommand{\tCO}{tCO$_2$e}

\begin{document}

\zoocoverpage

% ── Abstract ──────────────────────────────────────────────────────────────────
\begin{abstract}
\noindent
Nature-based carbon credits suffer a crisis of credibility. Recent
analyses reveal that 70--90\% of certified forest carbon offsets may
represent phantom emissions reductions, undermining market integrity and
climate goals. The root cause is inadequate Measurement, Reporting, and
Verification (MRV) infrastructure that relies on infrequent manual
audits, self-reported baselines, and opaque methodologies. We present
\textbf{ZooMRV}, an open-source digital MRV platform that combines
multi-spectral satellite imagery, LiDAR-derived biomass estimation,
eddy covariance flux networks, and on-chain attestation to produce
continuously verified carbon credit certificates. ZooMRV processes
Sentinel-2, Landsat-9, and GEDI LiDAR data through a deep learning
pipeline (CarbonNet) that estimates above-ground biomass with a root
mean square error of 18.4~Mg/ha---a 34\% improvement over the current
operational standard. Credits are issued as non-fungible tokens on the
Zoo Network with immutable provenance chains linking each credit to
specific monitoring data, spatial boundaries, and verification reports.
We demonstrate the system on 47 nature-based solution (NBS) projects
across tropical forests, mangroves, peatlands, and grasslands, totaling
2.3 million hectares. ZooMRV detects 23\% more reversals (biomass loss
events) than traditional audit cycles and reduces the mean time between
event and detection from 14 months to 12 days. The platform is governed
through Zoo DAO processes and released under open licenses.

\medskip
\noindent\textbf{Keywords:} carbon credits, MRV, satellite monitoring,
biomass estimation, nature-based solutions, blockchain verification,
remote sensing
\end{abstract}

\newpage
\tableofcontents
\newpage

% ══════════════════════════════════════════════════════════════════════════════

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-carbonnet-biomass-estimation-model}
\input{sections/04-dynamic-baselining}
\input{sections/05-continuous-monitoring-pipeline}
\input{sections/06-on-chain-credit-registry}
\input{sections/07-evaluation-on-nbs-projects}
\input{sections/08-discussion}
\input{sections/09-implementation}
\input{sections/10-future-work}
\input{sections/99-conclusion}

\end{document}
