\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{caption}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={Zoo 2025: Digital Securities, Quantum-Secure Settlement, and the Holographic Consensus DAO},
    pdfauthor={Antje Worring, Zach Kelling},
    pdfsubject={Zoo Labs Foundation - Securities, Governance, and Quantum-Secure Settlement},
    pdfkeywords={digital securities, holographic consensus, FHE voting, Quasar 3.0, ML-DSA, Corona, BLS}
}

\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Quasar}{\textsc{Quasar}}
% Brand-neutral descriptor for the regulated securities operator partner.
\newcommand{\Partner}{the regulated securities partner}
\newcommand{\PartnerCap}{The regulated securities partner}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\FHE}{\textsc{fhe}}
\newcommand{\MPC}{\textsc{mpc}}
\newcommand{\TEE}{\textsc{tee}}
\newcommand{\BLS}{\textsc{bls}}
\newcommand{\MLDSA}{\textsc{ml-dsa}}
\newcommand{\PQ}{\textsc{pq}}

\title{\textbf{Zoo 2025}\\
\large Digital Securities, Quantum-Secure Settlement,\\and the Holographic Consensus DAO}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{December 15, 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
\noindent We articulate five intertwined planks that together describe \Zoo{}
Labs Foundation as it stands at the \Quasar{} 3.0 activation date
(2025-12-25): (i) the canonical team and mission that descend from the
October 31, 2021 founding whitepaper; (ii) \Zoo{}'s positioning against the
$\sim$\$113 trillion global digital-securities market through its exclusive
partnership with \Partner{} and the \Lux{} chain stack; (iii) the
quantum-secure settlement guarantee provided by \Quasar{} 3.0
(\Lux{} \textsc{lp-020}), which composes three independent cryptographic
hardness assumptions ($\BLS{}_{12\text{-}381}$, Corona Ring-LWE, and
$\MLDSA{}\text{-}65$ via Z-Chain Groth16); (iv) a Homomorphic / Holographic
Consensus \DAO{} with weighted voting over \emph{advocacy}, \emph{involvement},
\emph{contribution}, and \emph{token stake}; and (v) democratized end-to-end
information access powered by per-\textsc{llm} chains, public verifiable
inference receipts, and zero-knowledge selective disclosure. This paper is
the canonical cross-reference for these five planks. Spec freeze:
2025-12-15. Activation alongside \Quasar{} 3.0: 2025-12-25.

\textbf{Keywords}: digital securities, holographic consensus, \FHE{} voting,
\Quasar{} 3.0, post-quantum settlement, weighted governance, open AI protocol
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-team}
\input{sections/02-securities-113t}
\input{sections/03-quantum-secure}
\input{sections/04-holographic-dao}
\input{sections/05-democratized-access}
\input{sections/06-conclusion}

\begin{thebibliography}{99}

\bibitem{lp-020-quasar-3}
Lux Foundation.
\newblock LP-020: Quasar Consensus 3.0.
\newblock Lux Improvement Proposal, December 2025.
\newblock \url{https://github.com/luxfi/lps/lp-020}.

\bibitem{lp-105-quasar}
Lux Foundation.
\newblock LP-105: Quasar Consensus Family Specification (Quasar 3.0 paper).
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-013-fchain}
Lux Foundation.
\newblock LP-013: F-Chain FHE-Protected Substrate.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-019-mchain}
Lux Foundation.
\newblock LP-019: M-Chain MPC and Threshold Custody.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-070-mldsa}
Lux Foundation.
\newblock LP-070: ML-DSA-65 Post-Quantum Signature Profile.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-073-corona}
Lux Foundation.
\newblock LP-073: Corona Ring-LWE Threshold Signatures.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-075-bls}
Lux Foundation.
\newblock LP-075: BLS12-381 Aggregate Signatures.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-134-topology}
Lux Foundation.
\newblock LP-134: Lux Chain Topology - A/B/M/F Chains.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{hanzo-ai-chain}
Hanzo AI.
\newblock The Hanzo AI Chain: Inference Marketplace and Commerce Layer.
\newblock Hanzo Whitepaper, December 2025.

\bibitem{zoo-per-llm-chains}
Zoo Labs Foundation.
\newblock Per-LLM Chains: One Sovereign Chain per Model.
\newblock Zoo Whitepaper, December 2025.

\bibitem{zoo-foundation-mission}
Zoo Labs Foundation.
\newblock Zoo Labs Foundation: AI and Blockchain Research for Conservation, Education, and Frontier Science.
\newblock Zoo Whitepaper, October 2021.

\bibitem{zoo-2021-original}
Worring, A.
\newblock Zoo: A Game-Fi Multiverse Ecosystem for the Preservation of Endangered Species.
\newblock Zoo Labs Foundation, October 31, 2021.

\bibitem{liquidity-perimeter}
Regulated securities operator.
\newblock Regulated ATS / Broker-Dealer / Transfer-Agent Perimeter.
\newblock Public regulatory filings (SEC, FINRA), 2025.

\bibitem{bis-tokenization}
Bank for International Settlements.
\newblock Tokenisation: A Primer for the Securities Markets.
\newblock BIS Quarterly Review, 2025.

\bibitem{wfe-market-cap}
World Federation of Exchanges.
\newblock Market Capitalisation Statistics.
\newblock WFE Statistical Database, 2025.

\bibitem{groth16}
Groth, J.
\newblock On the Size of Pairing-Based Non-Interactive Arguments.
\newblock EUROCRYPT, 2016.

\bibitem{nist-mldsa}
NIST.
\newblock FIPS 204: Module-Lattice-Based Digital Signature Standard.
\newblock NIST, 2024.

\bibitem{ring-lwe}
Lyubashevsky, V., Peikert, C., Regev, O.
\newblock On Ideal Lattices and Learning With Errors over Rings.
\newblock EUROCRYPT, 2010.

\end{thebibliography}

\end{document}
