\documentclass[11pt,letterpaper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{titling}
\usepackage{fancyhdr}
\usepackage{setspace}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=black,
    pdftitle={Zoo Whitepaper - Founder Vision (October 31, 2021)},
    pdfauthor={Antje Worring},
    pdfsubject={Zoo Labs Foundation - Game-Fi Multiverse for Endangered Species},
    pdfkeywords={NFT, DeFi, Game-Fi, AI, Metaverse, Conservation, DAO}
}

\graphicspath{{figures/}}

% Title information
\title{\textbf{Zoo}\\\large A Game-Fi Multiverse Ecosystem for the Preservation of Endangered Species}
\author{Antje Worring\\\textit{Founder, Zoo Labs Foundation}\\\texttt{antje@zoo.ngo}}
\date{October 31, 2021}

% Page style
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Zoo Whitepaper}}
\fancyhead[R]{\small\textit{October 31, 2021}}
\fancyfoot[L]{\scriptsize Originally published October 31, 2021. Reconstructed from \texttt{zoolabs.gitbook.io/whitepaper/} on 2025-12-15 for archival purposes.}
\fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\begin{document}

% --- Cover page ---
\begin{titlepage}
  \thispagestyle{empty}
  \centering
  \vspace*{2cm}
  {\Huge\bfseries Zoo\par}
  \vspace{0.6cm}
  {\Large A Game-Fi Multiverse Ecosystem for the\\Preservation of Endangered Species\par}
  \vspace{2cm}
  {\large \textbf{Antje Worring}\par}
  \vspace{0.2cm}
  {\itshape Founder, Zoo Labs Foundation\par}
  \vspace{0.1cm}
  {\small \texttt{antje@zoo.ngo}\par}
  \vspace{2cm}
  {\large \textbf{October 31, 2021}\par}
  \vfill
  {\itshape A founder-vision whitepaper at the intersection of\\Open AI Protocol, Blockchain, Metaverse, and Gaming.\par}
  \vspace{1cm}
  {\small Zoo Labs Foundation \quad $\cdot$ \quad \texttt{zoo.ngo}\par}
  \vspace{0.4cm}
  {\scriptsize Originally published October 31, 2021. Reconstructed from \texttt{zoolabs.gitbook.io/whitepaper/}\\on 2025-12-15 for archival purposes. All graphics, prose, and structure derive from the\\canonical Zoo Labs Foundation whitepaper.\par}
  \vspace{0.3cm}
  {\scriptsize\itshape Forward note: Zoo 3.0 (2025-12-25) brings the original 2021 vision to a\\GPU-native sovereign L1 --- see \emph{Zoo 3.0 Launch} (zoo-3-0-launch).\par}
\end{titlepage}

% Abstract pseudo-page (front matter)
\pagenumbering{roman}
\setcounter{page}{1}

\begin{center}
{\Large \textbf{Foreword from the Founder}}
\end{center}
\medskip

\noindent When I founded Zoo Labs Foundation, I wanted to build something that married three forces I believed would define the next decade: open AI, blockchain, and the metaverse. The thread connecting them, for me, was always the animals. I grew up loving wildlife, and I could not stand by while species after species disappeared. Zoo is my answer: a Game-Fi ecosystem where every digital animal is at once a financial primitive, an emotionally intelligent AI companion, and a conservation token whose economics route real value to the non-profits saving their real-world counterparts.

This document is the canonical record of that founding vision, dated to the night I first wrote it. It captures \emph{why} Zoo exists, the protocol-level mechanics that make it work, the V1 drop catalogue, and the roadmap that took us from a Tamagotchi-like pet game to a full multiverse companion. Decentralization, scarcity, on-chain provenance, and an open AI protocol are non-negotiable pillars. Conservation is the alignment objective.

To everyone who joins this journey: thank you. Every animal you raise, every NFT you mint, every conversation you have with one of our intelligent avatars contributes to the survival of a real species in the real world.

\vspace{0.4cm}
\noindent --- Antje Worring, Founder, Zoo Labs Foundation \\
\noindent \textit{October 31, 2021}

\vspace{1cm}

\tableofcontents

\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}

\input{sections/01-abstract-mission}
\input{sections/02-utility-market}
\input{sections/03-sustainability-conservation}
\input{sections/04-differentiators}
\input{sections/05-token-collateral}
\input{sections/06-marketplace-gameplay}
\input{sections/07-animals-feeding}
\input{sections/08-ai-ar-metaverse}
\input{sections/09-roadmap-extras}
\input{sections/10-2025-update}

\clearpage
\section*{Colophon}
\addcontentsline{toc}{section}{Colophon}

This document is a faithful reconstruction of the Zoo Labs Foundation whitepaper, originally published online at \url{https://zoolabs.gitbook.io/whitepaper/}. The reconstruction was performed on 2025-12-15 to preserve the founder vision in archival LaTeX form. Every section corresponds to a page in the source whitepaper. All images were retrieved from the live gitbook CDN and are reproduced here under the Zoo Labs Foundation's own publication.

The original publication date of October 31, 2021 reflects when Antje Worring first articulated this vision as the canonical founder document for the Zoo Labs Foundation.

For the most current version of the Zoo ecosystem, including post-2021 evolutions of the AI protocol and DAO governance, see the companion papers in the Zoo Labs research series at \url{https://zoo.ngo/papers}.

\end{document}
