\section{Protocol Model}


\subsection{System Model}

We consider a set of $n$ validators $\mathcal{V} = \{v_1, \ldots, v_n\}$ with stake weights $\{w_1, \ldots, w_n\}$ where $\sum_{i=1}^{n} w_i = W$. The protocol tolerates up to $f$ Byzantine validators subject to:

\begin{equation}
\sum_{i \in \mathcal{F}} w_i < \frac{W}{3}
\end{equation}

where $\mathcal{F} \subseteq \mathcal{V}$ is the set of Byzantine validators.

\subsection{Dual-Path Finality}

\begin{definition}[Quasar Finality]
A block $B$ at height $h$ achieves \Quasar{} finality when both conditions hold:
\begin{align}
\textsc{BLS-Finality}(B) &: \sum_{i \in \mathcal{S}_{\text{BLS}}} w_i \geq \frac{2W}{3} \label{eq:bls-fin} \\
\textsc{Ringtail-Finality}(B) &: \textsc{Verify}_{\Ringtail{}}(\sigma_{\text{thresh}}, B) = \top \label{eq:rt-fin}
\end{align}
where $\mathcal{S}_{\text{BLS}}$ is the set of validators that signed $B$ with BLS, and $\sigma_{\text{thresh}}$ is the \Ringtail{} threshold signature on $B$.
\end{definition}

Both paths execute concurrently. The block proposer collects BLS signatures and \Ringtail{} signature shares in parallel. Finality is declared when both thresholds are met.

\subsection{Threat Model}

\Quasar{} provides security under two adversary classes:

\begin{itemize}
\item \textbf{Classical adversary}: Controls $< n/3$ validators and has polynomial-time computational resources. BLS path alone suffices.
\item \textbf{Quantum adversary}: Controls $< n/3$ validators and has access to a cryptographically relevant quantum computer. The \Ringtail{} path provides security; BLS path may be compromised but cannot override \Ringtail{} finality without controlling $\geq n/3$ stake.
\end{itemize}

The dual requirement ensures that an adversary must break \textbf{both} BLS and \Ringtail{} simultaneously to violate safety.

