\section{Governance Architecture}

\label{sec:architecture}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Token Structure}

Zoo DAO uses a dual-token system:

\begin{enumerate}
  \item \textbf{\ZOO{} token} (fungible, transferable): Governance token
    for voting credits, staking, and economic coordination. Maximum
    supply: 1 billion.

  \item \textbf{ZooRep} (non-transferable, soulbound): Reputation tokens
    earned through verified contributions. Determines eligibility for
    governance roles and review assignments.
\end{enumerate}

\paragraph{ZooRep earning mechanisms.}
\begin{table}[H]
\centering
\caption{ZooRep earning rates by contribution type.}
\label{tab:rep}
\begin{tabular}{lcc}
\toprule
\textbf{Contribution} & \textbf{ZooRep earned} & \textbf{Decay rate} \\
\midrule
Published research (open access) & 100--500 & 10\%/year \\
Peer review completion & 20--50 & 15\%/year \\
Data contribution (verified) & 10--100 & 10\%/year \\
Successful grant project & 200--1000 & 5\%/year \\
Citizen science observations & 1--5 & 20\%/year \\
Code contribution (merged) & 10--50 & 15\%/year \\
\bottomrule
\end{tabular}
\end{table}

Decay ensures reputation reflects recent activity.

\subsection{Governance Bodies}

\begin{enumerate}
  \item \textbf{General Assembly}: All \ZOO{} holders. Votes on
    constitutional changes and treasury allocation ceilings. 10\% quorum.

  \item \textbf{Research Council}: Members with $\geq$500 ZooRep.
    Reviews grants and sets research priorities. 7-member elected body,
    staggered 2-year terms.

  \item \textbf{Review Board}: Members with $\geq$200 ZooRep in
    relevant domains. Conducts peer review.

  \item \textbf{Treasury Committee}: 5 multi-sig holders (3-of-5)
    managing DAO treasury. Elected annually.

  \item \textbf{Domain Working Groups}: Self-organizing groups focused
    on specific research areas. Minimum 5 members with $\geq$100 ZooRep.
\end{enumerate}

\subsection{Proposal Types}

\begin{table}[H]
\centering
\caption{Proposal types and governance requirements.}
\label{tab:proposals}
\begin{tabular}{p{3cm}p{2.5cm}p{2cm}p{2.5cm}p{2cm}}
\toprule
\textbf{Type} & \textbf{Proposer req.} & \textbf{Quorum} &
  \textbf{Voting method} & \textbf{Duration} \\
\midrule
Research grant & 50 ZooRep & 5\% & Quadratic & 14 days \\
Parameter change & 200 ZooRep & 10\% & Token-weighted & 7 days \\
Constitutional & 500 ZooRep & 20\% & Supermajority & 21 days \\
Emergency & Research Council & 33\% & Simple majority & 48 hours \\
\bottomrule
\end{tabular}
\end{table}


% ══════════════════════════════════════════════════════════════════════════════
