\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

Zoo DAO demonstrates that decentralized governance can produce faster,
fairer, and more impactful research funding. Quadratic voting allocates
funds more efficiently by capturing preference intensity while resisting
plutocracy. Incentivized peer review produces higher-quality, faster
reviews. Retroactive funding rewards genuine impact over proposal writing
skill.

After 18 months, the DAO has funded 143 projects across 34 countries with
demonstrably better allocation quality than simulated alternatives. The
Zoo DAO model provides a viable blueprint for open, community-governed
science. Through Zoo Improvement Proposals at \texttt{zips.zoo.ngo}, the
governance framework itself evolves through the democratic processes it
enables.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{28}

\bibitem[Alberts et~al.(2014)]{alberts2014rescuing}
Alberts, B., Kirschner, M.W., Tilghman, S., and Varmus, H.
\newblock Rescuing US biomedical research from its systemic flaws.
\newblock \emph{PNAS}, 111(16):5773--5777, 2014.

\bibitem[Azoulay et~al.(2019)]{azoulay2019funding}
Azoulay, P., Graff~Zivin, J.S., Li, D., and Sampat, B.N.
\newblock Public R\&D investments and private-sector patenting.
\newblock \emph{NBER Working Paper}, 2019.

\bibitem[Barbereau et~al.(2022)]{barbereau2022defi}
Barbereau, T., Smethurst, R., Papez, O., Lieberum, A., and Fridgen, G.
\newblock DeFi, not so decentralized: The measured distribution of voting rights.
\newblock \emph{HICSS}, 2022.

\bibitem[Benkler(2006)]{benkler2006wealth}
Benkler, Y.
\newblock \emph{The Wealth of Networks}.
\newblock Yale University Press, 2006.

\bibitem[Boudreau et~al.(2016)]{boudreau2016looking}
Boudreau, K.J., Guinan, E.C., Lakhani, K.R., and Riedl, C.
\newblock Looking across and looking beyond the knowledge frontier.
\newblock \emph{Management Science}, 62(10):2765--2783, 2016.

\bibitem[Bush(1945)]{bush1945science}
Bush, V.
\newblock Science, the Endless Frontier.
\newblock US Government Printing Office, 1945.

\bibitem[Buterin et~al.(2019)]{buterin2019flexible}
Buterin, V., Hitzig, Z., and Weyl, E.G.
\newblock A flexible design for funding public goods.
\newblock \emph{Management Science}, 65(11):5171--5187, 2019.

\bibitem[Buterin(2014)]{buterin2014ethereum}
Buterin, V.
\newblock Ethereum: A next-generation smart contract platform.
\newblock Ethereum Whitepaper, 2014.

\bibitem[Compound(2020)]{compound2020governance}
Compound.
\newblock Compound Governance.
\newblock \url{https://compound.finance/governance}, 2020.

\bibitem[Coordinape(2021)]{coordinape2021}
Coordinape.
\newblock A framework for decentralized coordination.
\newblock Technical report, 2021.

\bibitem[Franzoni \& Sauermann(2014)]{franzoni2011recombinant}
Franzoni, C. and Sauermann, H.
\newblock Crowd science: Organization of research in open collaborative projects.
\newblock \emph{Research Policy}, 43(1):1--20, 2014.

\bibitem[Frey \& Gallus(2017)]{frey2017awards}
Frey, B.S. and Gallus, J.
\newblock Towards an economics of awards.
\newblock \emph{Journal of Economic Surveys}, 31(1):190--200, 2017.

\bibitem[Gitcoin(2023)]{gitcoin2023}
Gitcoin.
\newblock Gitcoin Grants: Quadratic funding for public goods.
\newblock \url{https://gitcoin.co/grants}, 2023.

\bibitem[Hamburg(2019)]{hamburg2019desci}
Hamburg, S.
\newblock Decentralized science: Toward a new model.
\newblock \emph{Science}, 365(6450):220--221, 2019.

\bibitem[Heller \& Eisenberg(1998)]{heller1998can}
Heller, M.A. and Eisenberg, R.S.
\newblock Can patents deter innovation?
\newblock \emph{Science}, 280(5364):698--701, 1998.

\bibitem[Ioannidis(2005)]{ioannidis2005most}
Ioannidis, J.P.A.
\newblock Why most published research findings are false.
\newblock \emph{PLOS Medicine}, 2(8):e124, 2005.

\bibitem[Lalley \& Weyl(2018)]{lalley2018quadratic}
Lalley, S.P. and Weyl, E.G.
\newblock Quadratic voting: How mechanism design can radicalize democracy.
\newblock \emph{AEA Papers and Proceedings}, 108:33--37, 2018.

\bibitem[Lee et~al.(2013)]{lee2013bias}
Lee, C.J., Sugimoto, C.R., Zhang, G., and Cronin, B.
\newblock Bias in peer review.
\newblock \emph{JASIST}, 64(1):2--17, 2013.

\bibitem[MakerDAO(2017)]{mcdao2017}
MakerDAO.
\newblock The Maker Protocol: Multi-collateral Dai.
\newblock Technical report, 2017.

\bibitem[Murray et~al.(2016)]{murray2016grand}
Murray, F., Stern, S., Campbell, G., and MacCormack, A.
\newblock Grand innovation prize: Incentivizing solutions.
\newblock \emph{Science}, 352(6282):151--152, 2016.

\bibitem[Optimism(2023)]{optimism2023retropgf}
Optimism Collective.
\newblock Retroactive Public Goods Funding Round 3.
\newblock \url{https://optimism.io/retropgf}, 2023.

\bibitem[Ostrom(1990)]{ostrom1990governing}
Ostrom, E.
\newblock \emph{Governing the Commons}.
\newblock Cambridge University Press, 1990.

\bibitem[Posner \& Weyl(2018)]{posner2018radical}
Posner, E.A. and Weyl, E.G.
\newblock \emph{Radical Markets}.
\newblock Princeton University Press, 2018.

\bibitem[SourceCred(2020)]{sourcecred2020}
SourceCred.
\newblock A reputation protocol for open communities.
\newblock \url{https://sourcecred.io}, 2020.

\bibitem[Stephan(2012)]{stephan2012economics}
Stephan, P.
\newblock \emph{How Economics Shapes Science}.
\newblock Harvard University Press, 2012.

\bibitem[UNESCO(2021)]{unesco2021science}
UNESCO.
\newblock UNESCO Science Report 2021.
\newblock UNESCO Publishing, 2021.

\bibitem[Weyl et~al.(2022)]{weyl2022decentralized}
Weyl, E.G., Ohlhaver, P., and Buterin, V.
\newblock Decentralized society: Finding web3's soul.
\newblock \emph{SSRN}, 2022.

\bibitem[Wright(2021)]{wright2021daos}
Wright, A.
\newblock The rise of decentralized autonomous organizations.
\newblock \emph{Stanford J.\ Blockchain Law \& Policy}, 4(2):152--237, 2021.

\end{thebibliography}

