\section{Zoo Animal Utility}
\label{sec:animal-utility}

Each Zoo Animal NFT provides a comprehensive set of functions for its owner. The following outlines all utility available from the first NFT drop.

\subsection{Offers}
NFT owners can switch on the offers button to allow any user in the Zoo ecosystem to find their NFT and make an offer on the animal. Owners have a dedicated offers page where they can accept, decline, or counter-offer. If an offer is accepted, the full ownership of the NFT is transferred to the purchaser. This means the NFT, its collateral, and its earned value to date are transferred to the new owner. The current boost level (as defined by the collateral) remains the same. Despite the change of ownership, the NFT sustains its original utility.

\subsection{Buy/Sell}
Owners can list their NFT on the Zoo Market for a set price, or put it on auction with or without a floor price. If the selling price is met with a purchase, the NFT is sold and ownership is transferred on-chain. The NFT, its collateral, and its earned value to date are transferred. The current boost level remains the same, and the NFT sustains its original utility.

\subsection{Collateralize}
NFT owners can infinitely collateralize their animals for further rewards (Boosts) by using almost any currency, fiat or crypto. The collateral is locked into the NFT and can only be released if the NFT is burned; however, owners can also sell the NFT with its collateral value.

\subsection{Boosts}
Boosts are determined by collateral added. They provide an additional reward on top of the daily reward that comes with the animal:
\begin{equation}
\text{Total Reward} = \text{Daily Reward} + (\text{Collateral} \times \text{Boost} \times \text{Days}/365)
\end{equation}

\noindent The boost tiers are as follows:
\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Collateral Range} & \textbf{Boost Multiplier} \\
\midrule
\$10--\$99   & 1.1$\times$ \\
\$100--\$999  & 1.2$\times$ \\
\$1{,}000--\$9{,}999 & 1.3$\times$ \\
\$10{,}000+  & 1.4$\times$ \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Earns}
Each NFT Animal has a set daily reward. Even if the owner never ``feeds'' the animal, it will earn value on a daily basis. If collateral is added, the daily reward is amplified:
\begin{equation}
\text{Earning} = (\text{Principal} + \text{Collateral}) \times \text{Boost} + (\text{Daily Amount} \times \text{Days})
\end{equation}
\noindent Estimated earning potential over time can be calculated with the Zoo Reward Calculator.

\subsection{Create Pool}
A pool is a set price determined by the owner for two or more NFTs to be sold as a collection. Once the price is set, anyone can purchase the pool. The price is displayed in the preferred currency (set by the pool owner), and the buyer can pay in any currency, crypto or fiat. After payment is sent, it is converted to \$ZOO and paid out to the seller.

\subsection{Breeding}
If a user is able to breed two of the same adult species, they will mint an NFT Egg of that species' status, whether Endangered, Rare, or Sublime. There are only 7 generations of animals allowed, therefore only 6 breedable generations.

\subsection{Hatch Eggs}
Hatching an egg unveils what animal is inside. The animal is the same status as the parents (Endangered, Rare, or Sublime). Once hatched, the Egg NFT is burned forever and a new Baby Animal NFT is born, which can be fed more crypto or fiat to mint the teen version and eventually the adult version.

\subsection{Burns}
Burning a Zoo NFT returns the rewards earned so far and the collateral locked into it. A Zoo sustainability tax is applied upon burning.

\subsection{Metaverse}
Future versions of Zoo will include metaverse land where animals can live. There will also be an app that allows placement of animals into any physical setting with VR. Furthermore, owners can utilize the NFT file formats (FBX and WebGL) to upload their animals into any type of metaverse landscape.
