\section{Zoo Foundation: Conservation $\times$ Education}

The Zoo Foundation was established to harness blockchain technology and the global community it connects to help preserve our world's most critically endangered animal species. We want to end the extinction of animals and we want to help the world see how important they are to all of us. We also want to inspire people to change animals' lives for the better. We are committed to creating digital twins of endangered species to raise awareness, enact change, and create an animal sanctuary for all Zoo NFT holders.

Through the Zoo Foundation, Zoo Labs will provide support to our many partners in wildlife conservation. This support will come through monetary contribution and by providing opportunities for people of all ages, anywhere on Earth, to learn about the beautiful creatures the foundation works towards preserving.

The Zoo Foundation will also host fundraising events with our partners in wildlife conservation and other strategic partners to promote donation and awareness worldwide.

Through financial and human support we can furthermore raise awareness and mitigate the devastating effects of the illegal wildlife trade and how people can get involved in stopping it.

It is our driving purpose to deepen the connection humans have with animals by creating sustainable animal sanctuaries for endangered species. We aim to share the Zoo Sanctuary with local visitors and educators, host fundraisers and events, and get the Zoo Community as involved as possible in saving endangered animals all over the world. The Zoo protocol allows users to directly save a life through a monthly or one-time contribution.
