\section{Gameplay Functions}

\subsection{Pre-Drop Whitelist}

The Zoo experience begins with NFT Eggs and an NFT Egg Drop. Each Zoo NFT Egg Drop features a series of animals from egg, to juvenile, adolescent, and mature forms. These drops may be themed based upon something related to the natural world, a cultural celebration, current events, partnerships and licensing agreements, or any number of other potential sources of inspiration. In the time before each NFT Egg Drop, new or existing users have the opportunity to purchase one random NFT Egg, of potentially any available rarity tier, for the price of the most common Egg. When the NFT Eggs are distributed to the presale purchasers, their egg type and its related value will be revealed. Some users will end up with the rarest and most valuable NFT Egg for the price of the lowest cost and most common NFT Egg.

\subsection{Original Egg Price}

The original price of the three varieties of Eggs determines the cost to mint the future animals. The original prices are as follows:

\begin{center}
\begin{tabular}{lrl}
\toprule
\textbf{Egg Type} & \textbf{Price} & \textbf{Distribution} \\
\midrule
Sublime Egg  & 1 ETH   & 5\% (most rare)   \\
Rare Egg     & 0.5 ETH & 15\% (second most rare) \\
Endangered Egg & 0.1 ETH & 85\% (most common) \\
\bottomrule
\end{tabular}
\end{center}

Once a user has purchased their NFT Egg and it resides in their wallet, they may choose to hatch their egg into a baby NFT Animal according to each drop and its rarity tiers. To mint the teen version, the user must ``feed'' the NFT $1\times$ the original price of its Egg. When doing this, the user is not spending those tokens---they are staking them into the NFT.

After the NFT Animal has hatched and the user has experienced it in its teen form, the user can choose to ``feed'' the NFT's original base price again at $2\times$ to make their NFT Animal grow into its adult form, thereby minting another NFT. The user will have collateral in the adult NFT worth $2\times$ the original price of its Egg. Now the user can find another adult of the same species and breed them, resulting in a new egg of the same rarity, by paying the breeding fee which is $3\times$ the price of the original egg.

From this point forward, the user can continue to feed their NFT Animal crypto, both \$ZOO and virtually any other cryptocurrency, to stake the tokens into the NFT Animal's liquidity pool which results in the Boosts (see Section~\ref{sec:collateral-backed}).
