\section{NFT Marketplace}

The Zoo proprietary, integrated marketplace is the home for exclusive Zoo Labs NFT Animals and NFT Eggs.

The Zoo Marketplace offers users the chance to purchase items using either cryptocurrency, like the \$ZOO Token, or through the use of credit/debit cards directly in the marketplace. Designed with usability in mind, the Zoo team built the most seamless checkout flow with either crypto or fiat to eliminate the most confusing terminology for newcomers in DeFi. Once a purchase is made, the seller is paid out in whatever currency they listed their NFT for, regardless of the payment method used.

Users can also use their credit/debit cards to purchase \$ZOO directly, removing one of the most prohibitive elements of the transaction process and opening up investment to many new individuals.

The Marketplace also facilitates ``Asset Transfers,'' allowing one user to transfer their NFT plus its stored value to another through a mutually agreed upon transaction. If a user sells or trades an NFT with another user, they are also transferring the full collateral amount of the NFT to the user they are transacting with.

\subsection{Leaderboard}

Users can compare their Zoo statistics with friends, family, and with every NFT holder globally. The leaderboard also allows users to make an offer bid on any NFT regardless of whether it is listed on the marketplace or not. Items are shown based on the NFT's total value or the user's cumulative NFT collection value.
