\section{Gen~0 NFT Drop}

In the Gen~0 NFT drop, an enhanced reward system creates an engaging and rewarding experience for the Zoo community. Each Gen~0 animal NFT comes with its own unique set of daily rewards, designed to incentivize active participation and offer ongoing benefits to collectors. Whether additional \$ZOO tokens, exclusive NFTs, or other enticing in-game rewards, the daily rewards system ensures that there is always something to look forward to.

Boosts have a direct impact on the daily rewards. By adding collateral to the animals they own (known as ``Feeding''), players can boost their daily rewards. These boosts amplify the potential rewards earned, providing an additional incentive for collectors to enhance the value of their animals over time.

The initial rewards are carved out during the Gen~0 NFT drop, with 50\% of the entire token supply allocated towards rewards. This allocation sets the foundation for a robust and sustainable reward system. Any further rewards, reward modifications, or the duration of the reward period are subject to the decisions made by the Zoo \DAO{}.

The daily rewards for each Gen~0 animal have a finite duration: daily rewards continue for a period of two years after the animal is minted. This time frame allows for a meaningful and engaging reward experience while maintaining a healthy ecosystem.

\subsection{Breeding and Hatching Guide}

\paragraph{Breeding Basics.}
Breeding requires two adult NFT animals of the same species but can be of any generation. The cost of breeding is determined by the original egg price (OEP), which varies based on rarity:

\begin{center}
\begin{tabular}{lr}
\toprule
\textbf{Egg Type} & \textbf{OEP} \\
\midrule
Sublime    & 1 ETH \\
Rare       & 0.5 ETH \\
Endangered & 0.05 ETH \\
\bottomrule
\end{tabular}
\end{center}

Upon successful breeding, an egg is minted and can be hatched for free or sold on the marketplace.

\paragraph{Growth Stages.}
\begin{itemize}[nosep]
\item Baby $\to$ Teen: lock in collateral equal to $1\times$ OEP
\item Teen $\to$ Adult: lock in collateral equal to $2\times$ OEP
\item Breeding fee: $3\times$ OEP (directed to the \DAO{}, not staked as collateral)
\end{itemize}

\paragraph{Generational Limits.}
Each egg generation has a limit of one baby, one teen, and one adult that can be minted. Adults with breeding privileges can breed with another adult of the same species to mint a subsequent egg generation. The first generation can breed 7 times, the second 6, and so on, until the seventh generation which has no breeding privileges.

\subsection{Daily Reward Schedule}

The rarity category of the animal determines its daily allowance based on age:

\begin{center}
\begin{tabular}{lccc}
\toprule
\textbf{Rarity} & \textbf{Baby} & \textbf{Teen} & \textbf{Adult} \\
\midrule
Sublime    & 1 \$ZOO    & 2 \$ZOO    & 3 \$ZOO \\
Rare       & 0.25 \$ZOO & 0.5 \$ZOO  & 1 \$ZOO \\
Endangered & 0.05 \$ZOO & 0.1 \$ZOO  & 0.2 \$ZOO \\
\bottomrule
\end{tabular}
\end{center}

\noindent Breeding privileges are based on the generation of the animal, while the rarity category determines the daily allowance based on the age of the animal.
