\section{Roadmap}

At Zoo Labs, we are set on a journey towards creating a more vibrant, immersive, and interactive NFT universe. The roadmap for future upgrades includes features that will redefine the user experience and extend the scope of engagement with NFT animals.

\subsection{Planned Features}

\begin{itemize}[nosep]
\item \textbf{Emotionally Intelligent Avatars:} NFT animals will evolve beyond static digital assets into dynamic, emotionally intelligent beings that users can interact with in the metaverse. With planned integration of AI, these animals will understand and reciprocate emotions, creating rich emotional connections.

\item \textbf{Mini-Games and Metaverse Lands:} Enhanced gaming experience through introduction of mini-games and expanded metaverse lands. Special powers and play-to-earn games offer exciting new ways to engage and reap rewards.

\item \textbf{Gold and Silver Coin Collection:} Through partnership with the Lux blockchain, players will have the opportunity to collect and redeem gold and silver coins, adding another layer of engagement and incentive.

\item \textbf{New Animal Species:} Introduction of new animals, including extinct species and mythical creatures, further broadening the scope of the NFT animal universe.

\item \textbf{Celebrity Collaborations:} Collaborations with celebrities and influencers for animal collections, creating cross-over between entertainment and digital collectibles.

\item \textbf{Unique Animal Traits:} Introduction of unique animal traits such as horns, double heads, and wings, alongside a variety of skins and powers. Each NFT animal is as unique and individual as its owner.

\item \textbf{Interactive VR/AR Pet App:} An interactive VR/AR pet app that responds emotionally and serves as a personalized AI-powered emotional support companion.

\item \textbf{\DAO{} Governance:} Formation of a \DAO{} that will oversee all game functions, manage the treasury, forge partnerships, and allocate rewards for each animal.
\end{itemize}

\subsection{Future Game Interface}

The Zoo Labs gaming landscape consists of floating islands in the vast cosmos, each distinct and teeming with life. Each island mirrors the unique attributes of the NFT animals inhabiting them, creating diverse ecosystems.

Users are given the opportunity to embody their NFT animals, experiencing the world through their eyes. They engage in play-to-earn games, navigate through captivating environments, and participate in quests and treasure hunts. The Zoo Labs universe is a thriving social space where users can connect with friends and other players from around the globe.

Intelligent avatar animals serve as AI-powered companions that bring a sense of social engagement. These avatars, capable of learning, adapting, and reacting, interact with users organically.

\subsection{Play-and-Earn Games}

\paragraph{Coin Runner.}
An endless-runner style quest for valuable gold and silver coins through the Zoo Labs temple. Players choose one of their Zoo Animals to play with, each with unique abilities and skills. They dash through winding paths, collecting gold and silver nuggets of various denominations, from small nuggets to the coveted bullion bar. Collected coins can be used to unlock power-ups, upgrades, and new animals. All rewards are redeemable for physical assets or can be transferred on Zoo Market or a secondary marketplace.

\paragraph{Memory Bridge.}
A memory-based game set in the Zoo Labs Metaverse. Players guide their Zoo Animal across a treacherous bridge by memorizing the pattern of bricks. The bridge becomes longer as the game progresses, increasing the challenge and potential rewards. Players have 4~seconds per level to step on the correct brick. Successfully completing each level earns valuable rewards, including rare items, tokens, and special powers. Every reward collected is redeemable for real-world assets.
