\section{GPU Acceleration}

\label{sec:gpu}
%----------------------------------------------------------------------

\subsection{NTT Butterfly Parallelism}

The computational bottleneck in both \TFHE{} and \CKKS{} is polynomial multiplication in $R_q$, implemented via the Number Theoretic Transform (\NTT{}). The \NTT{} butterfly network of depth $\log_2 N$ maps naturally to GPU thread blocks:

\begin{itemize}
  \item Each butterfly stage is a parallel map over $N/2$ independent pairs.
  \item Stages synchronize via \texttt{\_\_syncthreads()} within a block or global barrier across blocks.
  \item For $N = 2^{14} = 16384$ (Zoo's default polynomial degree), a single SM processes 8192 butterflies per stage.
\end{itemize}

\subsection{TFHE Bootstrap on GPU}

A single \TFHE{} bootstrap consists of:

\begin{enumerate}
  \item \textbf{Blind rotation}: $n = 630$ external products, each requiring two \NTT{} forward/inverse transforms on degree-$N$ polynomials.
  \item \textbf{Key switching}: one matrix--vector product in $\mathbb{Z}_q^{n' \times n}$.
  \item \textbf{Modulus switching}: rounding from $q$ to $q'$.
\end{enumerate}

On CPU (single-threaded, AVX-512), one bootstrap takes $\sim$12\,ms. On an NVIDIA H100 GPU:

\begin{itemize}
  \item Blind rotation: 630 external products execute as a batched \NTT{} kernel, 0.22\,ms.
  \item Key switching: cuBLAS GEMV, 0.05\,ms.
  \item Modulus switching: element-wise kernel, 0.03\,ms.
  \item \textbf{Total: 0.3\,ms} (40$\times$ speedup).
\end{itemize}

\subsection{Batch Amortization}

When multiple independent bootstraps are required (e.g., evaluating a 256-bit comparison circuit), \Zoo{} batches them into a single GPU dispatch:

\begin{equation}
  T_{\mathrm{batch}}(k) = T_{\mathrm{single}} + \alpha \cdot k, \quad \alpha \approx 0.015\,\text{ms per additional bootstrap}
\end{equation}

\noindent for $k \leq 4096$ bootstraps. Beyond this, L2 cache pressure causes throughput degradation. At $k = 256$ (a typical 8-bit integer comparison), total latency is $\sim$4.1\,ms.

\subsection{Memory Bounds}

FHE ciphertexts are large. A single \CKKS{} ciphertext at level $\ell = 15$ with $N = 16384$ occupies:

\begin{equation}
  |\mathsf{ct}| = 2 \cdot N \cdot \ell \cdot 8 = 2 \cdot 16384 \cdot 15 \cdot 8 \approx 3.75\,\text{MB}
\end{equation}

An H100 with 80\,GB HBM3 can hold $\sim$21{,}000 such ciphertexts simultaneously. \Zoo{} validators must provision at least 40\,GB GPU memory for FHE workloads. The VM enforces per-contract ciphertext storage limits via gas metering.

%----------------------------------------------------------------------
