\section{Related Work}

\label{sec:related}
%----------------------------------------------------------------------

\textbf{Lux TFHE.} The \Lux{} \texttt{lux-tfhe} paper~\cite{lux-tfhe} introduced GPU-accelerated TFHE for the Lux L0 base layer. \Zoo{} extends this work with threshold decryption, \CKKS{} support, and deeper VM integration via FHE-specific opcodes on \TChain{}.

\textbf{Zama fhEVM.} Zama's fhEVM~\cite{zama-fhevm} pioneered FHE smart contracts on EVM-compatible chains. \Zoo{} differs in three respects: (1)~\CKKS{} support for approximate arithmetic, enabling ML inference; (2)~threshold decryption rather than single-operator key management; (3)~formal verification of noise bounds and GPU scaling in Lean~4.

\textbf{Sunscreen FHE.} The Sunscreen compiler~\cite{sunscreen2023} provides a Rust-based FHE toolkit targeting BFV and CKKS. \Zoo{} builds on similar compiler infrastructure but integrates it directly into the blockchain VM with gas metering and threshold key management.

\textbf{Concrete/TFHE-rs.} Zama's \texttt{tfhe-rs} library~\cite{zama-tfhers} provides a Rust implementation of TFHE with GPU support. \Zoo{}'s GPU kernels build on this foundation, extending it with batched bootstrapping optimizations and \TChain{} integration.

\textbf{Threshold FHE in MPC.} Boneh et al.~\cite{boneh2018threshold} formalized threshold FHE for multi-party computation. \Zoo{} adapts this framework to the blockchain setting, replacing interactive MPC rounds with asynchronous partial-decryption collection anchored by BFT consensus.

%----------------------------------------------------------------------
