\section*{Acknowledgments}


The authors thank the \Zoo{} Labs Foundation engineering team for GPU kernel development, the \Lux{} cryptography team for the \texttt{lux-tfhe} foundation, and the \Hanzo{} infrastructure team for validator provisioning.

\begin{thebibliography}{99}

\bibitem{gentry2009fhe}
Craig Gentry.
\textit{Fully Homomorphic Encryption Using Ideal Lattices}.
STOC, 2009.

\bibitem{chillotti2020tfhe}
Ilaria Chillotti, Nicolas Gama, Mariya Georgieva, and Malika Izabach\`{e}ne.
\textit{TFHE: Fast Fully Homomorphic Encryption over the Torus}.
Journal of Cryptology, 33(1):34--91, 2020.

\bibitem{cheon2017ckks}
Jung Hee Cheon, Andrey Kim, Miran Kim, and Yongsoo Song.
\textit{Homomorphic Encryption for Arithmetic of Approximate Numbers}.
ASIACRYPT, 2017.

\bibitem{albrecht2015lwe}
Martin R. Albrecht, Rachel Player, and Sam Scott.
\textit{On the Concrete Hardness of Learning with Errors}.
Journal of Mathematical Cryptology, 9(3):169--203, 2015.

\bibitem{boneh2018threshold}
Dan Boneh, Rosario Gennaro, Steven Goldfeder, Aayush Jain, Sam Kim, Peter M.R. Rasmussen, and Amit Sahai.
\textit{Threshold Cryptosystems from Threshold Fully Homomorphic Encryption}.
CRYPTO, 2018.

\bibitem{zama-fhevm}
Zama.
\textit{fhEVM: Confidential Smart Contracts on the EVM Using Fully Homomorphic Encryption}.
Zama Whitepaper, 2023.

\bibitem{zama-tfhers}
Zama.
\textit{TFHE-rs: A Pure Rust Implementation of the TFHE Scheme}.
\url{https://github.com/zama-ai/tfhe-rs}, 2023.

\bibitem{sunscreen2023}
Sunscreen.
\textit{Sunscreen: A Compiler for FHE Programs}.
Sunscreen Tech Report, 2023.

\bibitem{lux-tfhe}
Lux Industries.
\textit{GPU-Accelerated TFHE for Lux Network}.
Lux Research, 2025.

\bibitem{zoo-tokenomics}
Zach Kelling.
\textit{Zoo Network Tokenomics: Fair Distribution Through Complete Airdrop}.
Zoo Labs Foundation, 2025.

\bibitem{zoo-gpu-evm}
Zach Kelling.
\textit{Zoo EVM: A Post-Quantum Safe Layer 2 Blockchain with AI-Native Precompiles}.
Zoo Labs Foundation, 2026.

\bibitem{zoo-network-architecture}
Zach Kelling.
\textit{Zoo Network: Decentralized AI Training and Inference Layer with Semantic Optimization}.
Zoo Labs Foundation, 2025.

\bibitem{zoo-threshold-signatures}
Zach Kelling.
\textit{Threshold Signatures on Zoo T-Chain} (forthcoming).
Zoo Labs Foundation, 2026.

\end{thebibliography}

