\section{Decentralized Semantic Optimization Algorithm}

\label{sec:dso}

\subsection{Operations}

The \DSO{} algorithm supports three operations:

\begin{definition}[Experience Operations]
\begin{enumerate}
    \item \textbf{Add}($e$): Insert new experience $e$ if non-redundant
    \item \textbf{Modify}($e_{\text{old}}, e_{\text{new}}$): Replace $e_{\text{old}}$ with refined $e_{\text{new}}$
    \item \textbf{Delete}($e$): Remove obsolete or incorrect experience $e$
\end{enumerate}
\end{definition}

\subsection{Extraction Pipeline}

\begin{algorithm}[t]
\caption{Semantic Advantage Extraction}
\label{alg:extraction}
\begin{algorithmic}[1]
\STATE \textbf{Input}: Query $q$, rollouts $\{o_1, \ldots, o_G\}$, ground truth $y$, library $\mathcal{E}$
\STATE \textbf{Output}: Operations $\Omega$
\STATE \COMMENT{Stage 1: Summarize trajectories}
\FOR{$i = 1$ \textbf{to} $G$}
    \STATE $s_i \leftarrow \text{LLM.summarize}(q, o_i, r_i, y)$
\ENDFOR
\STATE \COMMENT{Stage 2: Extract group advantages}
\STATE $\Omega_{\text{raw}} \leftarrow \text{LLM.extract}(q, \{s_1, \ldots, s_G\}, \mathcal{E}, y)$
\STATE \COMMENT{Stage 3: Consolidate}
\STATE $\Omega \leftarrow \text{LLM.consolidate}(\Omega_{\text{raw}}, \mathcal{E})$
\RETURN $\Omega$
\end{algorithmic}
\end{algorithm}

\textbf{Stage 1 Prompt} (Trajectory Summarization):
\begin{quote}
\textit{An agent was provided with the following experiences and produced this trajectory. Summarize step-by-step: (1) what actions were taken, (2) which experiences guided decisions, (3) identify errors or detours given the correct answer.}
\end{quote}

\textbf{Stage 2 Prompt} (Group Critique):
\begin{quote}
\textit{Review these problem-solving attempts and extract generalizable experiences. Identify patterns in successful vs. failed trajectories. Suggest updates: Add (new insights), Modify (refine existing), Delete (incorrect). Max 3 operations per group. Each experience must begin with context ("When...") and be ≤32 words.}
\end{quote}

\textbf{Stage 3 Prompt} (Batch Consolidation):
\begin{quote}
\textit{Consolidate suggested updates into final revisions. Merge similar experiences, eliminate redundancy, ensure each is clear and generalizable (≤32 words). Return JSON operations: Add/Modify/Merge/Delete.}
\end{quote}

\subsection{Conflict Resolution}

When multiple operations target the same experience $e$:

\begin{enumerate}
    \item \textbf{Delete + Modify}: Prioritize Delete (indicates fundamental incorrectness)
    \item \textbf{Modify + Modify}: Merge into single Modify via LLM synthesis
    \item \textbf{Add + Add} (similar): Check embedding similarity; if $> 0.9$, merge
\end{enumerate}

\subsection{Convergence Analysis}

\begin{definition}[Experience Library Convergence]
A library $\mathcal{E}$ converges if:
\begin{equation}
\lim_{t \to \infty} \|\mathcal{E}_{t+1} - \mathcal{E}_t\| = 0
\end{equation}
where $\|\cdot\|$ measures embedding space distance.
\end{definition}

\begin{theorem}[Convergence Under \DSO{}]
Given bounded query distribution $\mathcal{Q}$ and $G \geq 5$, \DSO{} converges to a stable library in $O(|\mathcal{Q}|)$ training steps.
\end{theorem}

\begin{proof}[Proof Sketch]
Each epoch extracts at most $k$ new experiences (where $k = 3 \times$ batch size). As coverage increases (Theorem 1), the probability of extracting novel insights decreases exponentially: $P(\text{novel} | \mathcal{E}_t) \approx e^{-|\mathcal{E}_t|/d}$. Once coverage exceeds 95\%, subsequent epochs yield mostly Modify/Delete operations with diminishing magnitude. By Lyapunov stability, $\mathcal{E}_t$ converges to a local optimum.
\end{proof}

\subsection{Hyperparameters}

\begin{table}[t]
\centering
\caption{\DSO{} Hyperparameters}
\label{tab:hyperparams}
\begin{tabular}{lcc}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\
\midrule
Group size $G$ & 5--8 & Rollouts per query \\
Sampling temp & 0.7 & Diversity in generation \\
Top-$k$ retrieval & 5 & Experiences per context \\
Sim threshold $\tau$ & 0.6 & Relevance cutoff \\
Max exp length & 32 words & Conciseness constraint \\
Epochs & 3 & Training iterations \\
LLM (extraction) & GPT-4 / DeepSeek-V3 & Introspection model \\
\bottomrule
\end{tabular}
\end{table}

