\section*{Acknowledgments}


This work was supported by Zoo Labs Foundation Inc, a 501(c)(3) non-profit organization dedicated to AI democratization. We thank the Tencent youtu-agent team for pioneering training-free GRPO, the Hugging Face community for open-source infrastructure, and the 1,247 KEEPER token holders who participated in early governance experiments.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{brown2020language}
T. Brown et al., ``Language models are few-shot learners,'' \textit{NeurIPS}, 2020.

\bibitem{christiano2017deep}
P. Christiano et al., ``Deep reinforcement learning from human preferences,'' \textit{NeurIPS}, 2017.

\bibitem{harris2019decentralized}
J. Harris and B. Waggoner, ``Decentralized and collaborative AI on blockchain,'' \textit{IEEE BigData}, 2019.

\bibitem{johnson2019billion}
J. Johnson et al., ``Billion-scale similarity search with GPUs,'' \textit{IEEE Trans. Big Data}, 2019.

\bibitem{kairouz2021advances}
P. Kairouz et al., ``Advances and open problems in federated learning,'' \textit{Found. Trends ML}, 2021.

\bibitem{kurtulmus2018trustless}
A. Kurtulmus and K. Daniel, ``Trustless machine learning contracts via blockchain,'' \textit{arXiv:1802.10185}, 2018.

\bibitem{lewis2020retrieval}
P. Lewis et al., ``Retrieval-augmented generation for knowledge-intensive NLP tasks,'' \textit{NeurIPS}, 2020.

\bibitem{mcmahan2017communication}
B. McMahan et al., ``Communication-efficient learning of deep networks from decentralized data,'' \textit{AISTATS}, 2017.

\bibitem{ouyang2022training}
L. Ouyang et al., ``Training language models to follow instructions with human feedback,'' \textit{NeurIPS}, 2022.

\bibitem{rafailov2023direct}
R. Rafailov et al., ``Direct preference optimization: Your language model is secretly a reward model,'' \textit{NeurIPS}, 2023.

\bibitem{schulman2017proximal}
J. Schulman et al., ``Proximal policy optimization algorithms,'' \textit{arXiv:1707.06347}, 2017.

\bibitem{shao2024deepseekmath}
Z. Shao et al., ``DeepSeekMath: Pushing the limits of mathematical reasoning in open language models,'' \textit{arXiv:2402.03300}, 2024.

\bibitem{tencentyoutu2025grpo}
Tencent youtu-agent team, ``Training-free group relative policy optimization,'' \textit{arXiv:2510.08191}, 2025.

\bibitem{wang2022self}
X. Wang et al., ``Self-consistency improves chain of thought reasoning in language models,'' \textit{ICLR}, 2022.

\bibitem{wei2022chain}
J. Wei et al., ``Chain-of-thought prompting elicits reasoning in large language models,'' \textit{NeurIPS}, 2022.

\bibitem{yao2023tree}
S. Yao et al., ``Tree of thoughts: Deliberate problem solving with large language models,'' \textit{NeurIPS}, 2023.

\bibitem{zoo2025zen}
Zoo Labs Foundation, ``Zen-Reranker: Multilingual dense retrieval at scale,'' \textit{Technical Report}, 2025.

\end{thebibliography}

\appendix

