\section{Conclusion}

\label{sec:conclusion}

We have presented ZooSat, a deep learning pipeline for continuous satellite-based ecological monitoring.
The Temporal Siamese Network achieves 94.2\% change detection accuracy and 87.8\% change type classification by leveraging temporal sequence modeling to suppress phenological and atmospheric confounders.
The Habitat Fragmentation Index provides a real-time, field-validated measure of habitat connectivity degradation ($r = 0.87$ with species richness change).
The Conservation Alert System delivers priority-ranked alerts 11.2 days faster than existing global monitoring systems.

An 18-month operational deployment across 47 protected areas in 14 countries generated 12,847 alerts (89.3\% true positive rate), supported 47 enforcement actions, and contributed to an estimated 4,200 hectares of prevented forest loss.
ZooSat demonstrates that deep learning applied to freely available Sentinel-2 imagery can provide conservation-grade ecological monitoring at a cost of \$0.003 per km$^2$ per month.

The system is deployed as open infrastructure on the Zoo Network and specified as Zoo Improvement Proposal ZIP-010.
Future work will extend the pipeline to incorporate radar imagery (Sentinel-1) for all-weather detection, integrate species distribution models for species-specific fragmentation assessment, and develop predictive models that forecast future deforestation fronts based on spatial and temporal patterns.

% ============================================================
% References
% ============================================================
\begin{thebibliography}{30}

\bibitem[Bandara and Patel(2022)]{bandara2022transformer}
Bandara, W.~G.~C. and Patel, V.~M. (2022).
\newblock A transformer-based Siamese network for change detection.
\newblock {\em Proceedings of IGARSS}, pages 207--210.

\bibitem[Brancalion et~al.(2020)]{brancalion2020global}
Brancalion, P.~H.~S., Niamir, A., Broadbent, E., et~al. (2020).
\newblock Global restoration opportunities in tropical rainforest landscapes.
\newblock {\em Science Advances}, 5(7):eaav3223.

\bibitem[Chen et~al.(2022)]{chen2022remote}
Chen, H., Qi, Z., and Shi, Z. (2022).
\newblock Remote sensing image change detection with transformers.
\newblock {\em IEEE Transactions on Geoscience and Remote Sensing}, 60:1--14.

\bibitem[Chen et~al.(2020)]{chen2020spatial}
Chen, H., Wu, C., Du, B., et~al. (2020).
\newblock Spatial-temporal attention-based method and a new dataset for remote sensing image change detection.
\newblock {\em Remote Sensing}, 12(10):1662.

\bibitem[Daudt et~al.(2018)]{daudt2018fully}
Daudt, R.~C., Le~Saux, B., and Boulch, A. (2018).
\newblock Fully convolutional Siamese networks for change detection.
\newblock {\em Proceedings of ICIP}, pages 4063--4067.

\bibitem[Drusch et~al.(2012)]{drusch2012sentinel}
Drusch, M., Del~Bello, U., Carlier, S., et~al. (2012).
\newblock Sentinel-2: ESA's optical high-resolution mission for GMES operational services.
\newblock {\em Remote Sensing of Environment}, 120:25--36.

\bibitem[Fahrig(2017)]{fahrig2017ecological}
Fahrig, L. (2017).
\newblock Ecological responses to habitat fragmentation per se.
\newblock {\em Annual Review of Ecology, Evolution, and Systematics}, 48:1--23.

\bibitem[FAO(2020)]{fao2020global}
FAO (2020).
\newblock {\em Global Forest Resources Assessment 2020: Main Report}.
\newblock Food and Agriculture Organization of the United Nations, Rome.

\bibitem[Finer et~al.(2018)]{finer2018combating}
Finer, M., Novoa, S., Weisse, M.~J., et~al. (2018).
\newblock Combating deforestation: From satellite to intervention.
\newblock {\em Science}, 360(6395):1303--1305.

\bibitem[Giglio et~al.(2016)]{giglio2016active}
Giglio, L., Schroeder, W., and Justice, C.~O. (2016).
\newblock The collection 6 MODIS active fire detection algorithm and fire products.
\newblock {\em Remote Sensing of Environment}, 178:31--41.

\bibitem[Haddad et~al.(2015)]{haddad2015habitat}
Haddad, N.~M., Brudvig, L.~A., Clobert, J., et~al. (2015).
\newblock Habitat fragmentation and its lasting impact on Earth's ecosystems.
\newblock {\em Science Advances}, 1(2):e1500052.

\bibitem[Hansen et~al.(2013)]{hansen2013high}
Hansen, M.~C., Potapov, P.~V., Moore, R., et~al. (2013).
\newblock High-resolution global maps of 21st-century forest cover change.
\newblock {\em Science}, 342(6160):850--853.

\bibitem[Hansen et~al.(2016)]{hansen2016humid}
Hansen, M.~C., Krylov, A., Tyukavina, A., et~al. (2016).
\newblock Humid tropical forest disturbance alerts using Landsat data.
\newblock {\em Environmental Research Letters}, 11(3):034008.

\bibitem[Maretto et~al.(2020)]{maretto2020spatio}
Maretto, R.~V., Fonseca, L.~M.~G., and Jacobs, N. (2020).
\newblock Spatio-temporal deep learning approach for land use change detection.
\newblock {\em IEEE Journal of Selected Topics in Applied Earth Observations}, 14:7297--7309.

\bibitem[Maxwell et~al.(2016)]{maxwell2016biodiversity}
Maxwell, S.~L., Fuller, R.~A., Brooks, T.~M., and Watson, J.~E.~M. (2016).
\newblock Biodiversity: The ravages of guns, nets and bulldozers.
\newblock {\em Nature}, 536(7615):143--145.

\bibitem[McGarigal(2012)]{mcgarigal2012fragstats}
McGarigal, K. (2012).
\newblock FRAGSTATS v4: Spatial pattern analysis program for categorical and continuous maps.
\newblock {\em University of Massachusetts, Amherst}.

\bibitem[Ortega et~al.(2021)]{ortega2021near}
Ortega~Adarme, M., Queiroz~Feitosa, R., and Achanccaray, P. (2021).
\newblock Near real-time deforestation detection using SAR.
\newblock {\em Remote Sensing}, 13(13):2587.

\bibitem[Planet(2020)]{planet2020nicfi}
Planet (2020).
\newblock NICFI Satellite Data Program: Planet basemaps for tropical forest monitoring.
\newblock {\em Norway's International Climate and Forests Initiative}.

\bibitem[Redford(2003)]{redford2003empty}
Redford, K.~H. (2003).
\newblock The empty forest.
\newblock {\em BioScience}, 42(6):412--422.

\bibitem[Saura and Pascual-Hortal(2007)]{saura2007new}
Saura, S. and Pascual-Hortal, L. (2007).
\newblock A new habitat availability index to integrate connectivity in landscape conservation planning.
\newblock {\em Landscape Ecology}, 22:1215--1229.

\bibitem[Singh(1989)]{singh1989review}
Singh, A. (1989).
\newblock Review article: Digital change detection techniques using remotely-sensed data.
\newblock {\em International Journal of Remote Sensing}, 10(6):989--1003.

\bibitem[Vogt et~al.(2007)]{vogt2007mapping}
Vogt, P., Riitters, K.~H., Estreguil, C., et~al. (2007).
\newblock Mapping spatial patterns with morphological image processing.
\newblock {\em Landscape Ecology}, 22:171--177.

\bibitem[Watson et~al.(2018)]{watson2018decline}
Watson, J.~E.~M., Evans, T., Venter, O., et~al. (2018).
\newblock The exceptional value of intact forest ecosystems.
\newblock {\em Nature Ecology \& Evolution}, 2:599--610.

\bibitem[Wulder et~al.(2019)]{wulder2019current}
Wulder, M.~A., Loveland, T.~R., and Roy, D.~P. (2019).
\newblock Current status of Landsat program, science, and applications.
\newblock {\em Remote Sensing of Environment}, 225:127--147.

\bibitem[Yuan et~al.(2021)]{yuan2021deep}
Yuan, X., Shi, J., and Gu, L. (2021).
\newblock A review of deep learning methods for semantic segmentation of remote sensing imagery.
\newblock {\em Expert Systems with Applications}, 169:114417.

\bibitem[Zheng et~al.(2021)]{zheng2021change}
Zheng, Z., Ma, A., Zhang, L., and Zhong, Y. (2021).
\newblock Change is everywhere: Single-temporal supervised object change detection in remote sensing imagery.
\newblock {\em Proceedings of ICCV}, pages 15193--15202.

\bibitem[Zhu(2017)]{zhu2017change}
Zhu, Z. (2017).
\newblock Change detection using Landsat time series: A review of frequencies, preprocessing, algorithms, and applications.
\newblock {\em ISPRS Journal of Photogrammetry and Remote Sensing}, 130:370--384.

\bibitem[Antje Worring, Zach Kelling(2025)]{zoo2025network}
Antje Worring, Zach Kelling (2025).
\newblock Zoo Network Architecture: A decentralized infrastructure for open science.
\newblock {\em Zoo Technical Report ZTR-001}.

\bibitem[Antje Worring, Zach Kelling(2026a)]{zoo2026datacommons}
Antje Worring, Zach Kelling (2026a).
\newblock Zoo Data Commons: A decentralized repository for open biodiversity datasets.
\newblock {\em Zoo Improvement Proposal ZIP-009}.

\bibitem[Antje Worring, Zach Kelling(2026b)]{zoo2026coord}
Antje Worring, Zach Kelling (2026b).
\newblock Decentralized Research Coordination: Federated experiment management for DeSci.
\newblock {\em Zoo Improvement Proposal ZIP-008}.

\end{thebibliography}

