\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

Verifiable carbon credits require verifiable infrastructure. ZooMRV
demonstrates that continuous satellite-based monitoring, algorithmic
baselining, and on-chain attestation can produce carbon credits with
substantially higher integrity than traditional certification. The
30.7\% over-crediting correction identified in our evaluation portfolio
underscores the urgency of MRV modernization. By reducing verification
costs by orders of magnitude while increasing accuracy and transparency,
digital MRV can restore confidence in nature-based carbon markets and
channel genuine finance toward climate mitigation.

Through Zach Kelling's commitment to open science, ZooMRV's
models, data pipelines, and smart contracts are freely available. The
platform is governed through Zoo DAO processes, ensuring that
methodology updates, buffer pool parameters, and credit issuance
standards evolve through community consensus rather than proprietary
decision-making.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{28}

\bibitem[Abadie et~al.(2010)]{abadie2010synthetic}
Abadie, A., Diamond, A., and Hainmueller, J.
\newblock Synthetic control methods for comparative case studies.
\newblock \emph{Journal of the American Statistical Association},
  105(490):493--505, 2010.

\bibitem[Badgley et~al.(2022)]{badgley2022}
Badgley, G., Freeman, J., Hamber, J.J., et~al.
\newblock Systematic over-crediting in California's forest carbon offsets
  program.
\newblock \emph{Global Change Biology}, 28(4):1433--1445, 2022.

\bibitem[Chave et~al.(2014)]{chave2014improved}
Chave, J., R{\'e}jou-M{\'e}chain, M., B{\'u}rquez, A., et~al.
\newblock Improved allometric models to estimate the aboveground biomass of
  tropical trees.
\newblock \emph{Global Change Biology}, 20(10):3177--3190, 2014.

\bibitem[Drusch et~al.(2012)]{drusch2012sentinel}
Drusch, M., Del~Bello, U., Carlier, S., et~al.
\newblock Sentinel-2: ESA's optical high-resolution mission for GMES.
\newblock \emph{Remote Sensing of Environment}, 120:25--36, 2012.

\bibitem[Dubayah et~al.(2020)]{dubayah2020global}
Dubayah, R., Blair, J.B., Goetz, S., et~al.
\newblock The Global Ecosystem Dynamics Investigation: High-resolution laser
  ranging of the Earth's forests and topography.
\newblock \emph{Science of Remote Sensing}, 1:100002, 2020.

\bibitem[Dubayah et~al.(2022)]{dubayah2022gedi}
Dubayah, R., Armston, J., Kellner, J.R., et~al.
\newblock GEDI launches a new era of biomass inference from space.
\newblock \emph{Environmental Research Letters}, 17:095001, 2022.

\bibitem[Forest Trends(2022)]{foresttrends2022}
Forest Trends' Ecosystem Marketplace.
\newblock The Art of Integrity: State of the Voluntary Carbon Markets 2022.
\newblock Forest Trends, 2022.

\bibitem[Franke et~al.(2020)]{franke2020blockchain}
Franke, L., Schletz, M., and Salomo, S.
\newblock Designing a blockchain model for the Paris Agreement's carbon market
  mechanism.
\newblock \emph{Sustainability}, 12(3):1068, 2020.

\bibitem[Fuss et~al.(2020)]{fuss2020moving}
Fuss, S., Canadell, J.G., Ciais, P., et~al.
\newblock Moving toward net-zero emissions requires new alliances for carbon
  dioxide removal.
\newblock \emph{One Earth}, 3(2):145--149, 2020.

\bibitem[Guizar-Couti{\~n}o et~al.(2022)]{guizar2022carbon}
Guizar-Couti{\~n}o, A., Jones, J.P., Sherren, K., et~al.
\newblock A global evaluation of the effectiveness of voluntary REDD+ projects.
\newblock \emph{Conservation Letters}, 15(6):e12903, 2022.

\bibitem[Hancock et~al.(2019)]{hancock2019gradient}
Hancock, S., Armston, J., Hofton, M., et~al.
\newblock The GEDI simulator: A large-footprint waveform lidar simulator for
  calibration and validation.
\newblock \emph{Earth and Space Science}, 6(2):294--310, 2019.

\bibitem[Howson(2019)]{howson2019crypto}
Howson, P.
\newblock Tackling climate change with blockchain.
\newblock \emph{Nature Climate Change}, 9(9):644--645, 2019.

\bibitem[Lang et~al.(2023)]{lang2023high}
Lang, N., Jetz, W., Schindler, K., and Wegner, J.D.
\newblock A high-resolution canopy height model of the Earth.
\newblock \emph{Nature Ecology \& Evolution}, 7:1778--1789, 2023.

\bibitem[Rodriguez-Veiga et~al.(2021)]{rodriguez2021predicting}
Rodriguez-Veiga, P., Quegan, S., Carreiras, J., et~al.
\newblock Forest biomass retrieval approaches from Earth observation.
\newblock \emph{Remote Sensing of Environment}, 256:112340, 2021.

\bibitem[Schepaschenko et~al.(2019)]{schepaschenko2019forest}
Schepaschenko, D., Shvidenko, A., Usoltsev, V., et~al.
\newblock A dataset of forest biomass structure for Eurasia.
\newblock \emph{Scientific Data}, 6:198, 2019.

\bibitem[Schletz et~al.(2020)]{schletz2020blockchain}
Schletz, M., Franke, L.A., and Salomo, S.
\newblock Blockchain application for Paris Agreement carbon market mechanism.
\newblock \emph{Sustainability}, 12(12):5069, 2020.

\bibitem[Schneider \& La~Hoz~Theuer(2019)]{schneider2019double}
Schneider, L. and La~Hoz~Theuer, S.
\newblock Environmental integrity of international carbon market mechanisms.
\newblock \emph{Annual Review of Resource Economics}, 11:203--227, 2019.

\bibitem[West et~al.(2023)]{west2023action}
West, T.A., B{\"o}rner, J., Sills, E.O., and Kontoleon, A.
\newblock Action needed to make carbon offsets from tropical forest
  conservation work for climate change mitigation.
\newblock \emph{Science}, 381(6660):873--877, 2023.

\bibitem[Xu et~al.(2021)]{xu2021changes}
Xu, L., Saatchi, S.S., Yang, Y., et~al.
\newblock Changes in global terrestrial live biomass over the 21st century.
\newblock \emph{Science Advances}, 7(27):eabe9829, 2021.

\bibitem[IPCC(2019)]{ipcc2019refinement}
IPCC.
\newblock 2019 Refinement to the 2006 IPCC Guidelines for National Greenhouse
  Gas Inventories.
\newblock IPCC, 2019.

\bibitem[Harris et~al.(2021)]{harris2021global}
Harris, N.L., Gibbs, D.A., Baccini, A., et~al.
\newblock Global maps of twenty-first century forest carbon fluxes.
\newblock \emph{Nature Climate Change}, 11:234--240, 2021.

\bibitem[Griscom et~al.(2017)]{griscom2017natural}
Griscom, B.W., Adams, J., Ellis, P.W., et~al.
\newblock Natural climate solutions.
\newblock \emph{PNAS}, 114(44):11645--11650, 2017.

\bibitem[Ciais et~al.(2014)]{ciais2014carbon}
Ciais, P., Sabine, C., Bala, G., et~al.
\newblock Carbon and other biogeochemical cycles.
\newblock In \emph{Climate Change 2013: The Physical Science Basis},
  Cambridge University Press, 2014.

\bibitem[Gorelick et~al.(2017)]{gorelick2017google}
Gorelick, N., Hancher, M., Dixon, M., et~al.
\newblock Google Earth Engine: Planetary-scale geospatial analysis.
\newblock \emph{Remote Sensing of Environment}, 202:18--27, 2017.

\bibitem[Hansen et~al.(2013)]{hansen2013high}
Hansen, M.C., Potapov, P.V., Moore, R., et~al.
\newblock High-resolution global maps of 21st-century forest cover change.
\newblock \emph{Science}, 342(6160):850--853, 2013.

\bibitem[Saatchi et~al.(2011)]{saatchi2011benchmark}
Saatchi, S.S., Harris, N.L., Brown, S., et~al.
\newblock Benchmark map of forest carbon stocks in tropical regions.
\newblock \emph{PNAS}, 108(24):9899--9904, 2011.

\bibitem[Baccini et~al.(2012)]{baccini2012estimated}
Baccini, A., Goetz, S.J., Walker, W.S., et~al.
\newblock Estimated carbon dioxide emissions from tropical deforestation.
\newblock \emph{Nature Climate Change}, 2:182--185, 2012.

\bibitem[Sexton et~al.(2016)]{sexton2016conservation}
Sexton, J.O., Noojipady, P., Song, X.P., et~al.
\newblock Conservation policy and the measurement of forests.
\newblock \emph{Nature Climate Change}, 6:192--196, 2016.

\end{thebibliography}

