\section{Zoo Labs Foundation: Mission \& Structure}


\subsection{Legal Structure \& Governance}

Zoo Labs Foundation Inc operates as a 501(c)(3) tax-exempt non-profit organization registered in the United States. This legal structure ensures:

\begin{enumerate}
    \item \textbf{Tax-Deductible Donations}: Contributors receive tax benefits, incentivizing conservation funding.
    \item \textbf{Non-Profit Mandate}: All revenue reinvests into research, open-source development, and conservation projects—no shareholder extraction.
    \item \textbf{Transparent Governance}: Board oversight, annual financial disclosures (Form 990), and public accountability.
    \item \textbf{Charitable Mission Priority}: Legal obligation to serve public good over commercial interests.
\end{enumerate}

Our governance model integrates traditional non-profit structures with decentralized autonomous organization (DAO) principles. Major research directions, funding allocations, and model governance decisions undergo community voting via KEEPER tokens on the Zoo Network blockchain \citep{zoo2024whitepaper}, ensuring transparent, democratic control.

\subsection{Funding Model}

Zoo Labs Foundation pursues diversified, mission-aligned funding:

\begin{itemize}
    \item \textbf{Individual Donations}: Tax-deductible contributions from researchers, conservationists, and AI enthusiasts.
    \item \textbf{Foundation Grants}: Partnerships with biodiversity-focused foundations (e.g., Wildlife Conservation Society, Rainforest Trust).
    \item \textbf{Research Contracts}: Government agencies (NSF, USAID) and NGOs contracting conservation genomics analysis.
    \item \textbf{Compute Donations}: Cloud providers donating GPU credits for training conservation-focused models.
    \item \textbf{Crowdfunding}: Zoo.fund platform enabling community-driven project financing.
    \item \textbf{Inference Revenue}: OpenAI-compatible API serving models, with proceeds funding conservation (voluntary pricing).
\end{itemize}

Critically, we reject revenue models requiring data sale, privacy violation, or exploitative labor. Our financial sustainability depends on demonstrating irreplaceable public value—that a world with Zoo Labs Foundation is measurably better for biodiversity, scientific progress, and equitable AI access.

\subsection{Mission Statement}

\begin{quote}
\textit{``To preserve, protect, and digitize Earth's biodiversity through cutting-edge AI and blockchain technology—from sequencing genomes of endangered species to bringing back extinct animals—while democratizing AI education and research globally. We envision a future where every species' genetic code lives immortally in decentralized ledgers, where any researcher anywhere can train frontier models, and where artificial intelligence serves life itself.''}
\end{quote}

This mission operationalizes through three research pillars, each reinforcing the others.

