\section{Research Pillar I: Conservation AI}


\subsection{Genomic Sequencing \& Analysis}

Our Conservation AI initiative digitally preserves endangered species through whole-genome sequencing, computational analysis, and immutable blockchain storage. Key projects include:

\subsubsection{African Elephant Genomics (2024-2025)}
Partnering with African Wildlife Foundation, we sequenced 47 forest elephant (\textit{Loxodonta cyclotis}) genomes from Central African populations. Using custom-trained transformer models (fine-tuned Qwen3-7B via Gym), we:
\begin{itemize}
    \item Identified 18 novel conserved genomic regions under positive selection.
    \item Mapped population structure revealing three genetically distinct subpopulations.
    \item Detected 127 single nucleotide polymorphisms (SNPs) associated with ivory morphology—enabling forensic ivory trade tracking.
    \item Stored complete genome assemblies (3.2 billion base pairs each) on Arweave with IPFS mirrors, ensuring permanent accessibility.
\end{itemize}

\subsubsection{Coral Metagenomics (Ongoing)}
Climate-driven coral bleaching threatens 75\% of reefs by 2050 \citep{hughes2018spatial}. We analyze coral holobiont metagenomes (coral + symbiotic algae + microbiome) to identify thermal tolerance mechanisms. Our AI pipeline:
\begin{enumerate}
    \item \textbf{Metagenomic Assembly}: Deep learning-based contig assembly using MetaGLM (fine-tuned via Gym).
    \item \textbf{Functional Annotation}: Transformer models predicting gene function from sequence alone, bypassing costly lab validation.
    \item \textbf{Symbiont Optimization}: Identifying thermally resistant \textit{Symbiodinium} strains for assisted evolution interventions.
\end{enumerate}

Results stored on-chain include 14 terabases of metagenomic data, 8,400 heat-shock protein variants, and AI-predicted functional annotations for 420,000 coral genes.

\subsection{De-Extinction Research}

Beyond preservation, we explore \textit{de-extinction}—resurrecting extinct species via genetic engineering. Our approach combines:

\begin{itemize}
    \item \textbf{Ancient DNA Recovery}: Extracting degraded DNA from museum specimens, permafrost remains, and sub-fossils.
    \item \textbf{Genome Reconstruction}: AI-powered gap-filling using phylogenetic context (e.g., reconstructing woolly mammoth genomes by comparing Asian elephant references).
    \item \textbf{CRISPR Design}: Machine learning models predicting optimal guide RNA sequences for genome editing.
    \item \textbf{Ethical Frameworks}: Blockchain-recorded community governance votes on de-extinction candidates.
\end{itemize}

\textbf{Case Study: Passenger Pigeon Genome} \\
In collaboration with Revive \& Restore, we reconstructed 92\% of the passenger pigeon (\textit{Ectopistes migratorius}) genome from 1870s museum specimens. Our custom DNA-BERT model filled 340 million base pair gaps by learning evolutionary constraints from extant Columbidae genomes. The complete genome (1.27 gigabases) resides permanently on Arweave, accessible via \texttt{ar://passengerpigeonv1}.

\subsection{Computational Biodiversity Monitoring}

AI enables real-time, planetary-scale biodiversity monitoring:

\begin{itemize}
    \item \textbf{Bioacoustic Surveillance}: Deploying autonomous recording units in rainforests, analyzing 50,000 hours of audio via fine-tuned Whisper models to detect species from vocalizations.
    \item \textbf{Camera Trap Analysis}: Training YOLOv9 models (via Gym's multimodal pipelines) identifying 4,300 species from camera trap images—reducing human labeling from months to hours.
    \item \textbf{eDNA Metabarcoding}: Using transformers to detect species from environmental DNA in water/soil samples.
\end{itemize}

\subsection{Blockchain for Conservation Integrity}

All genomic data, species observations, and conservation decisions record on Zoo Network blockchain:

\begin{itemize}
    \item \textbf{Immutable Provenance}: Every genome assembly includes cryptographic hash, sequencing metadata, and contributor attributions.
    \item \textbf{Access Control}: Smart contracts governing data access—public for research, restricted for commercially sensitive data (e.g., anti-poaching intelligence).
    \item \textbf{Incentive Alignment}: Contributors earning KEEPER tokens redeemable for compute credits or governance rights.
    \item \textbf{Verification}: Independent scientists verifying sequencing quality, with verification recorded on-chain.
\end{itemize}

