\section{Why AI + Blockchain}


\subsection{AI's Limitations Without Blockchain}

\begin{enumerate}
    \item \textbf{Centralized Control}: Models trained by corporations serve corporate interests—data extraction, engagement maximization, profit optimization.
    \item \textbf{Opacity}: Weight updates are uninterpretable—no one knows \textit{why} model behavior changes.
    \item \textbf{Fragility}: Models degrade unpredictably, catastrophic forgetting erases learned behaviors.
    \item \textbf{Inaccessibility}: Training costs and infrastructure requirements exclude 99\% of researchers.
\end{enumerate}

\subsection{Blockchain's Limitations Without AI}

\begin{enumerate}
    \item \textbf{Speculative Focus}: Blockchain dominated by cryptocurrency speculation, not societal utility.
    \item \textbf{Complexity}: Smart contract development requires specialized expertise, limiting adoption.
    \item \textbf{Storage Costs}: On-chain storage expensive—genomic data requires off-chain solutions (IPFS/Arweave).
    \item \textbf{User Experience}: Wallets, gas fees, and transaction confirmation confuse non-technical users.
\end{enumerate}

\subsection{Combined Benefits}

\begin{table*}[t]
\centering
\caption{AI + Blockchain Synergies}
\label{tab:synergies}
\begin{tabular}{@{}p{0.25\linewidth}p{0.35\linewidth}p{0.35\linewidth}@{}}
\toprule
\textbf{Challenge} & \textbf{AI Contribution} & \textbf{Blockchain Contribution} \\ \midrule
\textbf{Conservation Data} & Analyze genomes, predict extinction risk & Immutable storage, provenance tracking \\
\textbf{Model Training} & Fine-tune specialized models & Decentralized compute marketplace \\
\textbf{Governance} & NLP-based proposal analysis & Transparent voting, tamper-proof records \\
\textbf{Attribution} & Code/genome authorship detection & Cryptographic proof of contribution \\
\textbf{Reproducibility} & Model checkpoints & Content-addressed versioning \\
\textbf{Incentives} & Quality assessment (reward models) & Token-based contributor rewards \\
\bottomrule
\end{tabular}
\end{table*}

\textbf{Concrete Example: Coral Genome Project} \\
AI analyzes 14TB metagenomic data, predicting 420,000 gene functions. Blockchain records:
\begin{itemize}
    \item Raw sequencing data → IPFS CID \texttt{Qm...}
    \item Analysis pipeline → Git commit hash + Docker image
    \item Gene annotations → Smart contract with Merkle tree
    \item Contributor identities → Ethereum addresses
    \item Governance votes → On-chain transaction history
\end{itemize}

Result: Complete scientific reproducibility, permanent accessibility (even if Zoo Labs Foundation dissolves), contributor attribution for funding/recognition, and democratic control over data usage policies.

