\section{Call to Action: Join the Mission}


\subsection{How Researchers Can Contribute}

\begin{enumerate}
    \item \textbf{Use Gym}: Train conservation models, publish findings, cite our work.
    \item \textbf{Contribute Code}: GitHub \texttt{github.com/zooai/gym}—PRs welcome for new models, optimizations.
    \item \textbf{Share Data}: Deposit genomic data into Immortal Genome Library (with proper consent/ethics approvals).
    \item \textbf{Collaborate}: Propose joint projects (email: research@zoo.ngo).
\end{enumerate}

\subsection{How Conservationists Can Contribute}

\begin{enumerate}
    \item \textbf{Propose Species}: Nominate endangered species for genomic sequencing.
    \item \textbf{Field Deployments}: Install AI monitoring systems (we provide hardware/training).
    \item \textbf{Data Annotation}: Label camera trap images, bioacoustic recordings—earn KEEPER tokens.
    \item \textbf{Policy Advocacy}: Use Zoo research to inform conservation policy (we provide reports/testimony).
\end{enumerate}

\subsection{How Developers Can Contribute}

\begin{enumerate}
    \item \textbf{Build Applications}: Create tools leveraging Zoo APIs (species identification apps, genomic visualizers).
    \item \textbf{Blockchain Development}: Improve smart contracts, optimize IPFS integration.
    \item \textbf{Infrastructure}**: Donate GPU compute, run Zoo Network validator nodes.
    \item \textbf{Documentation}: Write tutorials, translate docs into local languages.
\end{enumerate}

\subsection{How Philanthropists Can Contribute}

\begin{enumerate}
    \item \textbf{Tax-Deductible Donations}: \texttt{zoo.ngo/donate}—501(c)(3) status ensures tax benefits.
    \item \textbf{Species Sponsorships}: Fund complete genome sequencing for specific species (\$50K-200K).
    \item \textbf{Endowments}: Establish permanent funds supporting specific research areas.
    \item \textbf{Compute Grants}: Donate cloud credits (AWS, GCP, Azure) for training runs.
\end{enumerate}

\subsection{How Institutions Can Contribute}

\begin{enumerate}
    \item \textbf{Academic Partnerships}: Joint faculty appointments, student internships.
    \item \textbf{Data Sharing Agreements}**: Deposit institutional genomic/ecological data.
    \item \textbf{Curriculum Integration}**: Adopt Gym in AI/genomics courses.
    \item \textbf{Conference Support}**: Host Zoo Labs workshops, symposia.
\end{enumerate}

