\section*{Acknowledgments}


This work stands on the shoulders of countless open-source contributors: PyTorch, Hugging Face Transformers, DeepSpeed, PEFT library developers; conservation biologists collecting field data under harsh conditions; indigenous communities sharing traditional ecological knowledge; blockchain pioneers building censorship-resistant infrastructure; and students worldwide choosing to apply AI for planetary benefit rather than profit maximization. To all who contributed code, data, funding, expertise, or moral support—you are co-authors of this future. Thank you.

\bibliographystyle{unsrtnat}
\begin{thebibliography}{99}

\bibitem{ceballos2015accelerated}
Ceballos, G., Ehrlich, P. R., Barnosky, A. D., García, A., Pringle, R. M., \& Palmer, T. M. (2015).
\textit{Accelerated modern human–induced species losses: Entering the sixth mass extinction}.
Science Advances, 1(5), e1400253.

\bibitem{ipbes2019global}
IPBES. (2019).
\textit{Global Assessment Report on Biodiversity and Ecosystem Services}.
Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services.

\bibitem{wwf2020living}
WWF. (2020).
\textit{Living Planet Report 2020: Bending the Curve of Biodiversity Loss}.
World Wildlife Fund.

\bibitem{svalbard2023report}
Svalbard Global Seed Vault. (2023).
\textit{Annual Report 2023}.
Norwegian Ministry of Agriculture and Food.

\bibitem{openai2024gpt4}
OpenAI. (2024).
\textit{GPT-4 Technical Report}.
arXiv:2303.08774.

\bibitem{anthropic2024claude}
Anthropic. (2024).
\textit{Claude 3 Model Card}.
Anthropic AI Safety Research.

\bibitem{alibaba2024qwen}
Alibaba Cloud. (2024).
\textit{Qwen2.5: Large Language Models for Diverse Applications}.
arXiv:2409.12186.

\bibitem{alibaba2024qwen3}
Alibaba Cloud. (2025).
\textit{Qwen3: Next-Generation Language Understanding}.
Technical Report.

\bibitem{meta2024llama3}
Meta AI. (2024).
\textit{Introducing Meta LLaMA 3: The Next Generation of Open Foundation Models}.
Meta Research Blog.

\bibitem{deepseek2024v3}
DeepSeek. (2024).
\textit{DeepSeek-V3 Technical Report}.
arXiv:2412.19437.

\bibitem{hughes2018spatial}
Hughes, T. P., et al. (2018).
\textit{Spatial and temporal patterns of mass bleaching of corals in the Anthropocene}.
Science, 359(6371), 80-83.

\bibitem{hu2021lora}
Hu, E. J., et al. (2021).
\textit{LoRA: Low-Rank Adaptation of Large Language Models}.
arXiv:2106.09685.

\bibitem{dettmers2023qlora}
Dettmers, T., et al. (2023).
\textit{QLoRA: Efficient Finetuning of Quantized LLMs}.
arXiv:2305.14314.

\bibitem{ouyang2022training}
Ouyang, L., et al. (2022).
\textit{Training Language Models to Follow Instructions with Human Feedback}.
NeurIPS 2022.

\bibitem{rafailov2023dpo}
Rafailov, R., et al. (2023).
\textit{Direct Preference Optimization: Your Language Model is Secretly a Reward Model}.
arXiv:2305.18290.

\bibitem{dao2023flashattention2}
Dao, T., et al. (2023).
\textit{FlashAttention-2: Faster Attention with Better Parallelism and Work Partitioning}.
arXiv:2307.08691.

\bibitem{unsloth2024}
Unsloth AI. (2024).
\textit{Unsloth: 2x Faster Language Model Training}.
GitHub: unslothai/unsloth.

\bibitem{youtu2024trainingfree}
Tencent Youtu Lab. (2024).
\textit{Training-Free Guidance for Discrete Diffusion Models and Beyond}.
arXiv:2410.08191.

\bibitem{zoo2024whitepaper}
Zoo Labs Foundation. (2024).
\textit{Zoo Network: Decentralized AI Infrastructure for Conservation}.
Technical Whitepaper, zoo.ngo/whitepaper.

\bibitem{gym2025}
Zoo Labs Foundation. (2025).
\textit{Gym: Democratizing AI Model Training}.
GitHub: github.com/zooai/gym.

\bibitem{hllm2025}
Zoo Labs Foundation. (2025).
\textit{Hamiltonian Large Language Models: Training-Free Semantic Optimization}.
arXiv:2501.00001, 2025.

\bibitem{zenreranker2025}
Zoo Labs Foundation. (2025).
\textit{ZIP-002: Zen-Reranker for Efficient Retrieval}.
Zoo Improvement Proposal.

\end{thebibliography}

