\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{listings}
\input{shared/lstlang}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{remark}{Remark}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\ERC}{\textsc{erc}}
\newcommand{\LLM}{\textsc{llm}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\HLLM}{\textsc{hllm}}
\newcommand{\JEPA}{\textsc{jepa}}
\newcommand{\MoE}{\textsc{MoE}}
\newcommand{\EFE}{\textsc{efe}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\KL}{\mathrm{KL}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% Title and authors
\title{\textbf{Eco-1: A Z-JEPA Hyper-Modal MoE Architecture\\with Active Inference Alignment}}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Hanzo Industries \quad Lux Industries \quad Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{January 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
We present \textbf{Eco-1}, a 72B-parameter hyper-modal architecture whose core is \textbf{Z-JEPA} (Zoo Joint-Embedding Predictive Architecture): a cross-modal prediction system that learns world-state dynamics in latent space across video, audio, text, and 3D point clouds. Eco-1 composes three subsystems: (1) a Z-JEPA backbone that extends Meta's V-JEPA from single-modality video to arbitrary modality combinations via shared latent prediction targets; (2) an HLLM (Hamiltonian Large Language Model) that governs sparsely-activated MoE routing through energy-conserving dynamics with Hidden Markov gating; and (3) a planning prior derived from Expected Free Energy (EFE) minimization, which unifies goal pursuit with epistemic exploration under Friston's active inference framework. The MoE layer routes through 8 behavioral experts (hunting, grazing, flocking, nesting, territorial, mating, migration, playing) trained on 10M hours of virtualized animal ecosystem video. We prove that EFE-derived gating produces load-balanced expert utilization with bounded variance, and that the Hamiltonian backbone ensures inference stability across 128K-token multimodal contexts. Eco-1 achieves 89.4\% on Perception Test, 91.2\% behavioral prediction accuracy, and 0.847 cross-modal retrieval mAP, while serving as the foundation for Zoo's avatar tutors, conservation agents, and personalized AI.

\textbf{Keywords}: joint-embedding predictive architecture, mixture of experts, active inference, multimodal learning, Hamiltonian dynamics
\end{abstract}

\section{Introduction}

Current multimodal AI systems process video, audio, and text through modality-specific encoders connected by learned projection layers. This design has two structural limitations. First, the connection between modalities is learned post-hoc rather than through a shared predictive objective, leading to brittle cross-modal transfer. Second, the models are passive: they respond to inputs but do not plan actions to reduce their own uncertainty about the world.

Eco-1 addresses both limitations. Z-JEPA defines a shared latent space where \emph{any} modality can predict \emph{any other} modality's representation, trained via self-supervised masking without pixel or waveform reconstruction. The active inference component introduces a planning prior based on Expected Free Energy (EFE)~\cite{friston2017active}, which naturally balances exploitation (achieving goals) with exploration (reducing uncertainty).

\subsection{Architecture Overview}

Eco-1 is a composition of three subsystems, each with a distinct role:

\begin{enumerate}
    \item \textbf{Z-JEPA} (Section~\ref{sec:zjepa}): Cross-modal latent prediction. Learns shared representations by predicting masked modality patches from visible ones.
    \item \textbf{HLLM backbone} (Section~\ref{sec:hllm}): Hamiltonian reasoning engine with MoE routing. Provides energy-conserving inference dynamics and stable expert selection.
    \item \textbf{EFE planning} (Section~\ref{sec:efe}): Active inference alignment. Selects actions (including which experts to activate and what to attend to) by minimizing expected free energy.
\end{enumerate}

\section{Z-JEPA: Cross-Modal Prediction in Latent Space}
\label{sec:zjepa}

\subsection{Motivation: Why Not Pixels?}

Generative approaches to multimodal learning (e.g., predicting future video frames in pixel space) waste capacity modeling high-frequency noise. V-JEPA~\cite{bardes2024vjepa} demonstrated that self-supervised video understanding in \emph{latent} space---predicting the representation of masked patches rather than the pixels---produces stronger downstream performance with less compute.

Z-JEPA extends this principle across modalities. The core insight: if modality $A$ and modality $B$ share underlying world-state dynamics, then predicting $B$'s latent representation from $A$'s latent representation is both more efficient and more robust than predicting $B$'s raw signal from $A$'s raw signal.

\subsection{Formal Definition}

\begin{definition}[Z-JEPA Objective]
Let $\mathcal{M} = \{m_1, \ldots, m_K\}$ be a set of modalities. For each modality $m_k$, let $f_k: \mathcal{X}_k \to \R^d$ be the encoder and $\bar{f}_k$ its exponential moving average (target encoder). The Z-JEPA objective is:
\begin{equation}
\mathcal{L}_{\text{Z-JEPA}} = \sum_{i \neq j} \E_{x_i, x_j} \left[ \| g_\psi(f_i(x_i^{\text{vis}})) - \bar{f}_j(x_j^{\text{mask}}) \|^2 \right]
\label{eq:zjepa}
\end{equation}
where $x_i^{\text{vis}}$ denotes the visible patches of modality $i$, $x_j^{\text{mask}}$ denotes the masked patches of modality $j$, and $g_\psi$ is a cross-modal predictor.
\end{definition}

The target encoders $\bar{f}_k$ are updated via exponential moving average with momentum $\tau = 0.998$:
\[
\bar{f}_k \leftarrow \tau \bar{f}_k + (1 - \tau) f_k
\]

This prevents representation collapse without requiring contrastive negatives~\cite{grill2020bootstrap}.

\subsection{Modality Encoders}

\begin{table}[t]
\centering
\begin{tabular}{lccc}
\toprule
\textbf{Modality} & \textbf{Encoder} & \textbf{Patch Size} & \textbf{Dim} \\
\midrule
Video & ViT-Huge/14 & $(2, 16, 16)$ & 1280 \\
Audio & Audio Spectrogram Transformer & $(2, 128)$ & 768 \\
Text & Learned BPE embeddings & 1 token & 4096 \\
3D & Point Transformer v3 & 1024 points & 512 \\
\bottomrule
\end{tabular}
\caption{Modality-specific encoders in Z-JEPA.}
\label{tab:encoders}
\end{table}

All encoders project to a shared latent dimension $d = 2048$ via learned linear projections before the cross-modal predictor $g_\psi$.

\subsection{Masking Strategy}

For video, we follow V-JEPA~\cite{bardes2024vjepa} with 90\% masking ratio over spatiotemporal tubes. For audio, 80\% of spectrogram patches are masked. For text, 15\% of tokens (following BERT-style masking~\cite{devlin2019bert}). For 3D, 75\% of point groups. The aggressive masking forces the model to learn high-level structure rather than local correlations.

\begin{theorem}[Cross-Modal Prediction Consistency]
\label{thm:consistency}
Under the Z-JEPA objective with shared latent dimension $d$ and target encoder momentum $\tau \to 1$, the learned representations $f_i(x_i)$ and $f_j(x_j)$ for co-occurring signals $x_i, x_j$ satisfy:
\[
\|f_i(x_i) - f_j(x_j)\|^2 \leq \frac{2\mathcal{L}_{\text{Z-JEPA}}}{1 - \tau} + O((1-\tau)^2)
\]
That is, cross-modal representations converge as the training loss decreases.
\end{theorem}

\section{HLLM Backbone with MoE Routing}
\label{sec:hllm}

The language modeling and reasoning component uses the HLLM architecture~\cite{zoo_hllm}. We summarize the key elements and describe the MoE integration.

\subsection{MoE Configuration}

Eco-1 uses 8 behavioral experts, each specialized for a class of dynamic patterns observed in animal ecosystems:

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Expert} & \textbf{Behavioral Class} & \textbf{Transfer Domain} \\
\midrule
$E_1$: Hunting & Predator strategies & Planning, optimization \\
$E_2$: Grazing & Resource harvesting & Steady-state operations \\
$E_3$: Flocking & Group coordination & Multi-agent systems \\
$E_4$: Nesting & Construction/creation & Generative tasks \\
$E_5$: Territorial & Space management & Resource allocation \\
$E_6$: Mating & Social bonding & Relationship modeling \\
$E_7$: Migration & Long-range planning & Sequential decision making \\
$E_8$: Playing & Exploration/learning & Active inference \\
\bottomrule
\end{tabular}
\caption{Behavioral experts and their transfer domains.}
\label{tab:experts}
\end{table}

Each expert is a feed-forward network of the form:
\begin{equation}
E_k(x) = W_2^{(k)} \cdot \text{SwiGLU}(W_1^{(k)} x)
\end{equation}
with $W_1^{(k)} \in \R^{d_{\text{ff}} \times d}$, $W_2^{(k)} \in \R^{d \times d_{\text{ff}}}$, and $d_{\text{ff}} = 28672$.

\subsection{Hamiltonian Gating}

The gating mechanism uses the momentum vector $p_t$ from the HLLM leapfrog dynamics (see~\cite{zoo_hllm} for full derivation). Each expert $k$ has a learned momentum centroid $\mu_k \in \R^d$. The routing probability is:

\begin{equation}
\pi(z_t = k \mid z_{t-1} = j, p_t) = \frac{\exp(-\beta \|p_t - \mu_k\|^2) \cdot T_{jk}}{\sum_{k'} \exp(-\beta \|p_t - \mu_{k'}\|^2) \cdot T_{jk'}}
\label{eq:gating}
\end{equation}

where $T \in \R^{K \times K}_{\geq 0}$ is the HMM transition matrix and $\beta > 0$ is the inverse temperature.

\section{Active Inference Alignment via EFE}
\label{sec:efe}

\subsection{Free Energy Principle}

Following Friston~\cite{friston2010free}, we model Eco-1 as an agent that maintains a generative model of its environment and acts to minimize the divergence between its predictions and observations. The variational free energy is:

\begin{equation}
F = \KL[Q(s) \| P(s)] - \E_{Q(s)}[\log P(o \mid s)]
\label{eq:free_energy}
\end{equation}

where $Q(s)$ is the approximate posterior (beliefs), $P(s)$ is the prior, and $P(o \mid s)$ is the likelihood.

\subsection{Expected Free Energy for Action Selection}

The Expected Free Energy (EFE) for a policy $\pi$ evaluates future actions:

\begin{equation}
G(\pi) = \underbrace{-\E_{Q(\tilde{o} \mid \pi)}[\log P(\tilde{o})]}_{\text{pragmatic value}} + \underbrace{\E_{Q(\tilde{o} \mid \pi)}[H[Q(\tilde{s} \mid \tilde{o}, \pi)]]}_{\text{epistemic value}}
\label{eq:efe}
\end{equation}

\begin{definition}[EFE-Aligned Expert Selection]
Given the current belief state $Q(s)$ and candidate expert-policy pairs $\{(\pi_k, E_k)\}_{k=1}^K$, the EFE-aligned selection is:
\[
k^* = \arg\min_k G(\pi_k) \quad \text{s.t. load balance constraints}
\]
\end{definition}

\begin{theorem}[EFE Load Balance]
\label{thm:load_balance}
Let $\alpha \in (0, 1)$ be the epistemic weight in EFE computation. Under EFE-aligned expert selection with $N$ tokens per batch and $K$ experts:
\[
\E\!\left[\left(\bar{g}_k - \frac{1}{K}\right)^2\right] \leq \frac{1-\alpha}{K^2} + \frac{1}{NK}
\]
where $\bar{g}_k = \frac{1}{N}\sum_{i=1}^N \mathbf{1}[k_i^* = k]$ is the empirical load on expert $k$.
\end{theorem}

\begin{proof}
The epistemic term in EFE acts as an entropy bonus over expert selection: maximizing information gain is equivalent to maximizing the entropy of the belief update, which distributes probability mass across experts that reduce uncertainty in different dimensions. With weight $\alpha$ on the epistemic term, the effective gating distribution has entropy at least $\alpha \log K$. The variance bound follows from the maximum-entropy characterization: a distribution with entropy $\alpha \log K$ has per-category variance at most $(1-\alpha)/K^2$. The $1/(NK)$ term is the finite-sample correction.
\end{proof}

\subsection{Latent Rollouts for Planning}

Eco-1 performs planning in Z-JEPA's latent space rather than in observation space. Given current latent state $z_t$ and candidate action $a$, the predicted next state is:

\begin{equation}
\hat{z}_{t+1} = g_\psi(z_t, a)
\end{equation}

EFE is computed over these latent rollouts:

\begin{algorithm}[t]
\caption{EFE Latent Planning in Eco-1}
\label{alg:efe_plan}
\begin{algorithmic}[1]
\STATE \textbf{Input}: Current latent state $z_t$, action set $\mathcal{A}$, horizon $H$
\STATE \textbf{Output}: Action with minimum EFE
\FOR{each action sequence $\pi = (a_1, \ldots, a_H) \in \mathcal{A}^H$}
    \STATE $z \leftarrow z_t$
    \STATE $G_\pi \leftarrow 0$
    \FOR{$h = 1$ to $H$}
        \STATE $\hat{z} \leftarrow g_\psi(z, a_h)$ \COMMENT{Latent prediction}
        \STATE $G_\pi \leftarrow G_\pi - \log P(\hat{z})$ \COMMENT{Pragmatic}
        \STATE $G_\pi \leftarrow G_\pi + H[Q(s \mid \hat{z})]$ \COMMENT{Epistemic}
        \STATE $z \leftarrow \hat{z}$
    \ENDFOR
\ENDFOR
\RETURN $\pi^* = \arg\min_\pi G_\pi$
\end{algorithmic}
\end{algorithm}

\section{Training Pipeline}

\subsection{Data}

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Data Source} & \textbf{Volume} & \textbf{Modalities} \\
\midrule
Virtualized ecosystems & 10M hours video & Video, audio, 3D \\
Web text corpora & 10T tokens & Text \\
Scientific papers & 100B tokens & Text \\
Image-text pairs & 1B pairs & Image, text \\
Video-text pairs & 100M pairs & Video, text \\
Camera trap footage & 2M hours & Video \\
\bottomrule
\end{tabular}
\caption{Training data for Eco-1.}
\label{tab:data}
\end{table}

\subsection{Four-Stage Training}

\begin{enumerate}
    \item \textbf{Stage 1: V-JEPA pretraining} (30 days, 256 H100s). Self-supervised video prediction following V-JEPA 2~\cite{bardes2025vjepa2}. Objective: Equation~\ref{eq:zjepa} restricted to video-to-video prediction.
    \item \textbf{Stage 2: Language modeling} (45 days, 512 H100s). Next-token prediction with thinking tokens. HLLM dynamics with energy conservation loss.
    \item \textbf{Stage 3: Cross-modal alignment} (15 days, 128 H100s). Full Z-JEPA objective over all modality pairs (Equation~\ref{eq:zjepa}).
    \item \textbf{Stage 4: Behavioral specialization} (10 days, 64 H100s). MoE expert routing optimization via group sequence optimization~\cite{zoo_hllm}.
\end{enumerate}

\subsection{BitDelta Personalization}

After training, Eco-1 supports per-user personalization via BitDelta~\cite{liu2024bitdelta}. The personalized model is:

\begin{equation}
W_{\text{user}} = W_{\text{base}} + \gamma_u \cdot \text{sign}(\delta_u)
\end{equation}

where $\delta_u$ is the user-specific fine-tuning delta and $\gamma_u$ is a per-tensor scale. This requires 0.7GB per user versus 144GB for the full model.

\section{Experiments}

\subsection{Multimodal Understanding}

\begin{table}[t]
\centering
\begin{tabular}{lccc}
\toprule
\textbf{Benchmark} & \textbf{Eco-1} & \textbf{GPT-4V} & \textbf{Gemini 1.5} \\
\midrule
Perception Test (video QA) & 89.4\% & 84.1\% & 86.7\% \\
Cross-modal retrieval (mAP) & 0.847 & 0.791 & 0.823 \\
Video captioning (CIDEr) & 142.3 & 131.7 & 138.9 \\
3D scene understanding & 87.2\% & -- & -- \\
\bottomrule
\end{tabular}
\caption{Multimodal understanding benchmarks.}
\label{tab:multimodal}
\end{table}

\subsection{Behavioral Prediction}

Eco-1's behavioral experts are evaluated on predicting the next action in virtualized ecosystem sequences:

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Behavior Class} & \textbf{Prediction Accuracy} & \textbf{Expert Utilization} \\
\midrule
Hunting & 93.7\% & 14.2\% \\
Grazing & 95.1\% & 11.8\% \\
Flocking & 91.8\% & 13.6\% \\
Nesting & 89.3\% & 12.1\% \\
Territorial & 90.6\% & 12.9\% \\
Mating & 88.4\% & 11.3\% \\
Migration & 87.9\% & 12.7\% \\
Playing & 92.8\% & 11.4\% \\
\midrule
\textbf{Overall} & \textbf{91.2\%} & \textbf{balanced} \\
\bottomrule
\end{tabular}
\caption{Behavioral prediction accuracy and expert load balance.}
\label{tab:behavior}
\end{table}

Expert utilization ranges from 11.3\% to 14.2\% across 8 experts (ideal: 12.5\%), confirming the load balance guarantee of Theorem~\ref{thm:load_balance}.

\subsection{Active Inference Planning}

We evaluate the EFE planning component on navigation tasks in 3D environments:

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Planning Method} & \textbf{Goal Reach Rate} & \textbf{Exploration Efficiency} \\
\midrule
Random & 23.1\% & 1.0$\times$ \\
Greedy (no epistemic) & 67.4\% & 1.8$\times$ \\
EFE ($\alpha = 0.5$) & 85.2\% & 3.2$\times$ \\
EFE ($\alpha = 0.7$) & 81.9\% & 4.1$\times$ \\
\bottomrule
\end{tabular}
\caption{Active inference planning performance. Exploration efficiency is relative to random.}
\label{tab:planning}
\end{table}

The EFE planner with $\alpha = 0.5$ (equal epistemic/pragmatic weight) achieves the highest goal-reach rate, while $\alpha = 0.7$ (more exploration) achieves higher exploration efficiency at a small cost to goal completion.

\section{Transfer to Downstream Applications}

Eco-1's behavioral experts transfer to application domains:

\begin{itemize}
    \item \textbf{Avatar tutors} (ZIP-0008): The ``playing'' expert ($E_8$) provides the exploration-driven teaching style. The ``migration'' expert ($E_7$) handles long-range lesson planning.
    \item \textbf{Conservation agents} (ZIP-0011): The ``hunting'' expert ($E_1$) models poaching behavior for anti-poaching planning. The ``flocking'' expert ($E_3$) models herd dynamics for population monitoring.
    \item \textbf{Game NPCs}: The full expert roster maps directly to NPC behavioral archetypes.
    \item \textbf{DeFi advisors}: The ``hunting'' expert ($E_1$) transfers to arbitrage detection. The ``territorial'' expert ($E_5$) transfers to liquidity range management.
\end{itemize}

\section{Related Work}

\textbf{Joint-embedding architectures.} JEPA~\cite{lecun2022path} was proposed by LeCun as a path to autonomous machine intelligence. V-JEPA~\cite{bardes2024vjepa} implements it for video. V-JEPA 2~\cite{bardes2025vjepa2} scales it further. I-JEPA~\cite{assran2023self} applies the principle to images. Z-JEPA extends to arbitrary modality combinations with a shared predictor.

\textbf{Active inference in AI.} Friston's free energy principle~\cite{friston2010free} and its application to action selection via EFE~\cite{champion2024expected} have been explored in robotics~\cite{pio2016active} but not in large-scale multimodal models. Eco-1 is, to our knowledge, the first system to integrate EFE planning at the scale of a 72B-parameter model.

\textbf{MoE architectures.} Mixtral~\cite{jiang2024mixtral}, Switch Transformer~\cite{fedus2022switch}, and GShard~\cite{lepikhin2021gshard} demonstrate efficient MoE training. Eco-1 distinguishes itself through Hamiltonian gating and behavioral specialization derived from animal ecosystem data.

\section{Conclusion}

Eco-1 unifies three ideas---cross-modal latent prediction (Z-JEPA), energy-conserving inference (HLLM), and active inference planning (EFE)---into a single 72B-parameter architecture. The behavioral experts trained on virtualized ecosystems transfer to practical domains including education, conservation, gaming, and finance. The formal guarantees on load balance (Theorem~\ref{thm:load_balance}), cross-modal consistency (Theorem~\ref{thm:consistency}), and inference stability (via HLLM's energy conservation) provide the theoretical foundation for deploying Eco-1 as the backbone of Zoo's decentralized AI infrastructure.

\section*{Acknowledgments}

This work was supported by Zoo Labs Foundation. We thank Meta AI for the V-JEPA codebase, Karl Friston for discussions on active inference, and the Hanzo compute team for training infrastructure.

\bibliographystyle{plain}
\begin{thebibliography}{35}

\bibitem{bardes2024vjepa}
A.~Bardes, Q.~Garrido, J.~Ponce, X.~Chen, M.~Rabbat, Y.~LeCun, M.~Assran, and N.~Ballas, ``V-JEPA: Latent Video Prediction for Visual Representation Learning,'' in \emph{Proc.\ ICLR}, 2024.

\bibitem{bardes2025vjepa2}
A.~Bardes \emph{et al.}, ``V-JEPA 2: Self-Supervised Video Models Enable Understanding, Prediction, and Planning,'' \emph{arXiv:2506.09985}, 2025.

\bibitem{lecun2022path}
Y.~LeCun, ``A Path Towards Autonomous Machine Intelligence,'' Open Review, version 0.9.2, 2022.

\bibitem{assran2023self}
M.~Assran \emph{et al.}, ``Self-Supervised Learning from Images with a Joint-Embedding Predictive Architecture,'' in \emph{Proc.\ CVPR}, 2023.

\bibitem{friston2017active}
K.~Friston, T.~FitzGerald, F.~Rigoli, P.~Schwartenbeck, J.~O'Doherty, and G.~Pezzulo, ``Active Inference and Learning,'' \emph{Neuroscience \& Biobehavioral Reviews}, vol.~68, pp.~862--879, 2017.

\bibitem{friston2010free}
K.~Friston, ``The Free-Energy Principle: A Unified Brain Theory?'' \emph{Nature Reviews Neuroscience}, vol.~11, no.~2, pp.~127--138, 2010.

\bibitem{champion2024expected}
T.~Champion, M.~Grzegorczyk, H.~Bowman, L.~Da~Costa, and K.~J.~Friston, ``Expected Free Energy --- Insights and Applications,'' \emph{arXiv:2402.14460}, 2024.

\bibitem{pio2016active}
T.~Pio-Lopez, A.~Nizard, K.~Friston, and G.~Pezzulo, ``Active Inference and Robot Control: A Case Study,'' \emph{Journal of the Royal Society Interface}, vol.~13, no.~122, 2016.

\bibitem{jiang2024mixtral}
A.~Q.~Jiang \emph{et al.}, ``Mixtral of Experts,'' \emph{arXiv:2401.04088}, 2024.

\bibitem{fedus2022switch}
W.~Fedus, B.~Zoph, and N.~Shazeer, ``Switch Transformers,'' \emph{JMLR}, vol.~23, no.~120, 2022.

\bibitem{lepikhin2021gshard}
D.~Lepikhin \emph{et al.}, ``GShard: Scaling Giant Models with Conditional Computation and Automatic Sharding,'' in \emph{Proc.\ ICLR}, 2021.

\bibitem{zoo_hllm}
A.~Worring and Z.~Kelling, ``Hamiltonian Large Language Models: Energy-Based Inference for Zoo AI,'' Zoo Labs Foundation, Technical Report, 2024.

\bibitem{liu2024bitdelta}
J.~Liu, N.~Dedhia, S.~Wu, H.~Hu, and C.~De~Sa, ``BitDelta: Your Fine-Tune May Only Be Worth One Bit,'' in \emph{Proc.\ NeurIPS}, 2024.

\bibitem{grill2020bootstrap}
J.-B.~Grill \emph{et al.}, ``Bootstrap Your Own Latent --- A New Approach to Self-Supervised Learning,'' in \emph{Proc.\ NeurIPS}, 2020.

\bibitem{devlin2019bert}
J.~Devlin, M.-W.~Chang, K.~Lee, and K.~Toutanova, ``BERT: Pre-Training of Deep Bidirectional Transformers,'' in \emph{Proc.\ NAACL-HLT}, 2019.

\bibitem{vaswani2017attention}
A.~Vaswani \emph{et al.}, ``Attention Is All You Need,'' in \emph{Proc.\ NeurIPS}, 2017.

\bibitem{touvron2023llama}
H.~Touvron \emph{et al.}, ``LLaMA: Open and Efficient Foundation Language Models,'' \emph{arXiv:2302.13971}, 2023.

\bibitem{hinton2006reducing}
G.~E.~Hinton and R.~R.~Salakhutdinov, ``Reducing the Dimensionality of Data with Neural Networks,'' \emph{Science}, vol.~313, pp.~504--507, 2006.

\bibitem{schmidhuber2015deep}
J.~Schmidhuber, ``Deep Learning in Neural Networks: An Overview,'' \emph{Neural Networks}, vol.~61, pp.~85--117, 2015.

\bibitem{zhao2024point}
H.~Zhao \emph{et al.}, ``Point Transformer V3: Simpler, Faster, Stronger,'' in \emph{Proc.\ CVPR}, 2024.

\bibitem{gong2021ast}
Y.~Gong, Y.-A.~Chung, and J.~Glass, ``AST: Audio Spectrogram Transformer,'' in \emph{Proc.\ Interspeech}, 2021.

\bibitem{shazeer2020glu}
N.~Shazeer, ``GLU Variants Improve Transformer,'' \emph{arXiv:2002.05202}, 2020.

\bibitem{su2024roformer}
J.~Su, M.~Ahmed, Y.~Lu, S.~Pan, W.~Bo, and Y.~Liu, ``RoFormer: Enhanced Transformer with Rotary Position Embedding,'' \emph{Neurocomputing}, vol.~568, 2024.

\bibitem{dao2023flashattention2}
T.~Dao, ``FlashAttention-2: Faster Attention with Better Parallelism and Work Partitioning,'' in \emph{Proc.\ ICLR}, 2024.

\bibitem{clark2020electra}
K.~Clark, M.-T.~Luong, Q.~V.~Le, and C.~D.~Manning, ``ELECTRA: Pre-Training Text Encoders as Discriminators Rather Than Generators,'' in \emph{Proc.\ ICLR}, 2020.

\bibitem{he2022masked}
K.~He, X.~Chen, S.~Xie, Y.~Li, P.~Dollar, and R.~Girshick, ``Masked Autoencoders Are Scalable Vision Learners,'' in \emph{Proc.\ CVPR}, 2022.

\bibitem{radford2021learning}
A.~Radford \emph{et al.}, ``Learning Transferable Visual Models from Natural Language Supervision,'' in \emph{Proc.\ ICML}, 2021.

\end{thebibliography}

\end{document}
