\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

ZooPlatform demonstrates that adaptive gamification, AI-assisted
identification, Bayesian quality modeling, and token-based incentives
can produce citizen science data that approaches professional survey
quality at a fraction of the cost. By directing observer effort toward
scientifically valuable observations through RL-powered challenges,
the platform reduces the spatial, temporal, and taxonomic biases that
limit the utility of traditional citizen science data. The integration
with Zoo DAO governance ensures that active contributors have voice in
how their data is used, creating a self-governing scientific community.

After 14 months, ZooPlatform has enrolled 23{,}847 contributors across
41 countries, generating 4.2 million observations covering 12{,}394
species. These data are freely available under CC-BY-4.0 through the Zoo
Open Science Network, contributing to the Foundation's mission of open,
decentralized biodiversity science.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{30}

\bibitem[Agrawal \& Goyal(2013)]{agrawal2013thompson}
Agrawal, S. and Goyal, N.
\newblock Thompson sampling for contextual bandits with linear payoffs.
\newblock \emph{ICML}, pp. 127--135, 2013.

\bibitem[Boakes et~al.(2010)]{boakes2010distorted}
Boakes, E.H., McGowan, P.J.K., Fuller, R.A., et~al.
\newblock Distorted views of biodiversity: Spatial and temporal bias in species
  occurrence data.
\newblock \emph{PLOS Biology}, 8(6):e1000385, 2010.

\bibitem[Bowser et~al.(2013)]{bowser2013gamifying}
Bowser, A., Hansen, D., He, Y., et~al.
\newblock Using gamification to inspire new citizen science volunteers.
\newblock \emph{Gamification}, pp. 18--25, 2013.

\bibitem[Courter et~al.(2013)]{courter2013weekend}
Courter, J.R., Johnson, R.J., Stuyck, C.M., Lang, B.A., and Kaiser, E.W.
\newblock Weekend bias in citizen science data reporting.
\newblock \emph{International Journal of Biometeorology}, 57(5):715--720, 2013.

\bibitem[Crowston et~al.(2020)]{crowston2020citizen}
Crowston, K. and Prestopnik, N.R.
\newblock Motivation and data quality in a citizen science game.
\newblock \emph{HICSS}, pp. 450--459, 2020.

\bibitem[Deci et~al.(1999)]{deci1999meta}
Deci, E.L., Koestner, R., and Ryan, R.M.
\newblock A meta-analytic review of experiments examining the effects of
  extrinsic rewards on intrinsic motivation.
\newblock \emph{Psychological Bulletin}, 125(6):627--668, 1999.

\bibitem[Eveleigh et~al.(2014)]{eveleigh2014designing}
Eveleigh, A., Jennett, C., Blandford, A., Brohan, P., and Cox, A.L.
\newblock Designing for dabblers and deterring drop-outs in citizen science.
\newblock \emph{CHI}, pp. 2985--2994, 2014.

\bibitem[GBIF(2024)]{gbif2024}
GBIF.
\newblock Global Biodiversity Information Facility.
\newblock \url{https://www.gbif.org}, 2024.

\bibitem[Go{\"e}au et~al.(2022)]{goeau2022overview}
Go{\"e}au, H., Bonnet, P., and Joly, A.
\newblock Overview of PlantCLEF 2022.
\newblock \emph{CLEF}, 2022.

\bibitem[iNaturalist(2024)]{inaturalist2024}
iNaturalist.
\newblock iNaturalist.
\newblock \url{https://www.inaturalist.org}, 2024.

\bibitem[Isaac et~al.(2014)]{isaac2014statistics}
Isaac, N.J., van~Strien, A.J., de~Zeeuw, M.P., et~al.
\newblock Statistics for citizen science: Extracting signals of change from
  noisy ecological data.
\newblock \emph{Methods in Ecology and Evolution}, 5(10):1052--1060, 2014.

\bibitem[Johnston et~al.(2020)]{johnston2020analytical}
Johnston, A., Hochachka, W.M., Strimas-Mackey, M.E., et~al.
\newblock Analytical guidelines to increase the value of community science data.
\newblock \emph{Biological Conservation}, 244:108410, 2020.

\bibitem[Kahl et~al.(2021)]{kahl2021birdnet}
Kahl, S., Wood, C.M., Eibl, M., and Klinck, H.
\newblock BirdNET: Deep learning for avian diversity monitoring.
\newblock \emph{Ecological Informatics}, 61:101236, 2021.

\bibitem[Kelling et~al.(2015)]{kelling2015can}
Kelling, S., Johnston, A., Hochachka, W.M., et~al.
\newblock Can observation skills of citizen scientists be estimated using species
  accumulation curves?
\newblock \emph{PLOS ONE}, 10(10):e0139600, 2015.

\bibitem[Preece(2016)]{preece2019citizen}
Preece, J.
\newblock Citizen science: New research challenges for human-computer
  interaction.
\newblock \emph{International Journal of Human-Computer Interaction},
  32(8):585--612, 2016.

\bibitem[Sauermann \& Franzoni(2015)]{sauermann2015crowd}
Sauermann, H. and Franzoni, C.
\newblock Crowd science user contribution patterns and their implications.
\newblock \emph{PNAS}, 112(3):679--684, 2015.

\bibitem[Simpson et~al.(2014)]{simpson2014zooniverse}
Simpson, R., Page, K.R., and De~Roure, D.
\newblock Zooniverse: Observing the world's largest citizen science platform.
\newblock \emph{WWW}, pp. 1049--1054, 2014.

\bibitem[Sullivan et~al.(2014)]{sullivan2014ebird}
Sullivan, B.L., Aycrigg, J.L., Barry, J.H., et~al.
\newblock The eBird enterprise: An integrated approach to development and
  application of citizen science.
\newblock \emph{Biological Conservation}, 169:31--40, 2014.

\bibitem[Terry et~al.(2020)]{terry2020think}
Terry, J.C.D., Roy, H.E., and August, T.A.
\newblock Thinking like a naturalist: Enhancing computer vision of citizen
  science images.
\newblock \emph{Methods in Ecology and Evolution}, 11(2):303--315, 2020.

\bibitem[Theobald et~al.(2015)]{theobald2015global}
Theobald, E.J., Ettinger, A.K., Burgess, H.K., et~al.
\newblock Global change and local solutions: Tapping the unrealized potential
  of citizen science for biodiversity research.
\newblock \emph{Biological Conservation}, 181:236--244, 2015.

\bibitem[Troudet et~al.(2017)]{troudet2017taxonomic}
Troudet, J., Grandcolas, P., Blin, A., Vignes-Lebbe, R., and Legendre, F.
\newblock Taxonomic bias in biodiversity data and societal preferences.
\newblock \emph{Scientific Reports}, 7:9132, 2017.

\bibitem[Van~Horn et~al.(2021)]{van2021benchmarking}
Van~Horn, G., Cole, E., Beery, S., et~al.
\newblock Benchmarking representation learning for natural world images.
\newblock \emph{CVPR}, pp. 12884--12893, 2021.

\bibitem[Bonney et~al.(2009)]{bonney2009citizen}
Bonney, R., Cooper, C.B., Dickinson, J., et~al.
\newblock Citizen science: A developing tool for expanding science knowledge
  and scientific literacy.
\newblock \emph{BioScience}, 59(11):977--984, 2009.

\bibitem[Dickinson et~al.(2012)]{dickinson2012current}
Dickinson, J.L., Shirk, J., Bonter, D., et~al.
\newblock The current state of citizen science as a tool for ecological
  research and public engagement.
\newblock \emph{Frontiers in Ecology and the Environment}, 10(6):291--297, 2012.

\bibitem[Kosmala et~al.(2016)]{kosmala2016assessing}
Kosmala, M., Wiggins, A., Swanson, A., and Simmons, B.
\newblock Assessing data quality in citizen science.
\newblock \emph{Frontiers in Ecology and the Environment}, 14(10):551--560, 2016.

\bibitem[Pocock et~al.(2015)]{pocock2015developing}
Pocock, M.J., Chapman, D.S., Sheppard, L.J., and Roy, H.E.
\newblock Choosing and using citizen science.
\newblock Centre for Ecology \& Hydrology, 2015.

\bibitem[Silvertown et~al.(2015)]{silvertown2015new}
Silvertown, J., Harvey, M., Greenwood, R., et~al.
\newblock Crowdsourcing the identification of organisms.
\newblock \emph{Trends in Ecology \& Evolution}, 30(3):149--150, 2015.

\bibitem[Wiggins \& Crowston(2011)]{wiggins2011citizen}
Wiggins, A. and Crowston, K.
\newblock From conservation to crowdsourcing: A typology of citizen science.
\newblock \emph{HICSS}, pp. 1--10, 2011.

\bibitem[Haklay(2013)]{haklay2013citizen}
Haklay, M.
\newblock Citizen science and volunteered geographic information.
\newblock In \emph{Crowdsourcing Geographic Knowledge}, pp. 105--122.
  Springer, 2013.

\bibitem[Chandler et~al.(2017)]{chandler2017contribution}
Chandler, M., See, L., Copas, K., et~al.
\newblock Contribution of citizen science towards international biodiversity
  monitoring.
\newblock \emph{Biological Conservation}, 213:280--294, 2017.

\end{thebibliography}

