\section{Why Sovereign AI Memory Matters}

\label{sec:why}
%----------------------------------------------------------------------

\subsection{Persistence Across Sessions}

Long-running agent tasks---multi-day research projects, ongoing code refactoring, iterative design work---require memory that outlives individual inference calls. Current systems approximate persistence through conversation logs, but these degrade rapidly: a 200k-token conversation history is neither searchable nor structured, and exceeding the context window forces lossy summarization.

A sovereign memory system treats session boundaries as irrelevant. The capsule persists on-chain; the agent attaches to it at session start and detaches at session end. No information is lost.

\subsection{Privacy From Providers}

When an agent processes medical records, legal documents, or proprietary code, the memory it forms is as sensitive as the source material. Provider-controlled memory creates an untenable situation: the user must trust the provider not to inspect, sell, or leak derived knowledge.

Sovereign memory eliminates this trust requirement. The capsule is encrypted with keys the provider never sees. The provider runs inference; the agent manages its own state.

\subsection{Portability Across Models}

Model capability evolves rapidly. Users should be free to migrate from one model to another without losing accumulated context. Sovereign memory, stored in a model-agnostic format and encrypted with user-controlled keys, makes migration a key-management operation rather than a data-migration project.

\subsection{Agent Autonomy}

Autonomous agents---those operating with delegated authority over wallets, APIs, and infrastructure---need memory they control. An Agent NFT \cite{zoo-agent-nft} that cannot remember its own transaction history, learned preferences, or accumulated expertise is not autonomous in any meaningful sense. Memory Capsules provide the state layer that Agent NFTs require.

%----------------------------------------------------------------------
