\section{Memory Economics}

\label{sec:economics}
%----------------------------------------------------------------------

\subsection{Storage Costs}

Memory Capsules consume storage on \Zoo{} Network. Costs are denominated in $\KEEPER$ tokens:

\begin{table}[h]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Operation} & \textbf{Cost} & \textbf{Unit} \\
\midrule
Capsule creation & 1,000 & $\KEEPER$ \\
Entry append & 10 & $\KEEPER$ per KB \\
CRDT sync (per merge) & 50 & $\KEEPER$ \\
\IChain{} anchor update & 100 & $\KEEPER$ \\
FHE encrypted query & 500 & $\KEEPER$ \\
Threshold reveal & 200 & $\KEEPER$ per reveal \\
\bottomrule
\end{tabular}
\caption{Memory operation costs}
\label{tab:costs}
\end{table}

\subsection{Garbage Collection}

Unbounded memory growth is economically unsustainable. The garbage collection policy:

\begin{enumerate}
    \item \textbf{TTL Expiry.} Entries past their TTL are marked for collection.
    \item \textbf{Salience Threshold.} Episodic entries below $\theta_{\textit{gc}} = 0.1$ salience are eligible.
    \item \textbf{Confidence Threshold.} Semantic entries below $\gamma = 0.05$ are eligible.
    \item \textbf{Storage Budget.} Each capsule has a configurable storage budget. When exceeded, the lowest-priority entries are collected first.
    \item \textbf{Compaction.} Tombstoned entries in the operation log are removed during periodic compaction (default: weekly).
\end{enumerate}

\begin{proposition}
Under the garbage collection policy with parameters $(\theta_{\textit{gc}} = 0.1, \gamma = 0.05, \textit{TTL}_{\textit{default}} = 90\text{ days})$, a capsule receiving 1,000 episodic entries per day converges to a steady-state size of approximately 45,000 live entries (180 MB encrypted).
\end{proposition}

\subsection{Archival Policies}

Collected entries are not destroyed---they are moved to cold storage at reduced cost:

\begin{itemize}
    \item \textbf{Hot Storage.} Active capsule on \Zoo{} validators. Full read/write, CRDT sync. Cost: 10 $\KEEPER$/KB/month.
    \item \textbf{Cold Storage.} Archived entries on decentralized storage (IPFS/Filecoin). Read-only, high latency. Cost: 0.1 $\KEEPER$/KB/month.
    \item \textbf{Glacier.} Long-term archival. Retrieval requires 24-hour notice. Cost: 0.01 $\KEEPER$/KB/month.
\end{itemize}

\subsection{Memory as an Asset}

Memory Capsules are transferable on-chain assets. An agent's accumulated knowledge has economic value independent of the agent itself. Use cases:

\begin{itemize}
    \item An agent retiring can sell its capsule to a successor agent.
    \item A training institution can sell pre-built capsules (``expert medical knowledge base'').
    \item DAO treasuries can fund public-good capsules (``complete Rust documentation memory'').
\end{itemize}

Transfer is implemented as a \DID{} ownership change on \IChain{}, with re-encryption of the vault key under the new owner's \KChain{} key.

%----------------------------------------------------------------------
