\section{Semantic Context Manager}


The Semantic Context Manager maintains experience embeddings and performs retrieval.

\subsection{Embedding Generation}

Each experience $e$ is converted to a dense vector $\mathbf{v}_e \in \mathbb{R}^d$ (typically $d = 768$ or $d = 1536$) via a pretrained encoder:
\[
\mathbf{v}_e = \text{Encoder}(e.\text{text})
\]

Common encoders:
\begin{itemize}
\item \textbf{OpenAI \texttt{text-embedding-ada-002}}: 1536-dim, high quality, proprietary.
\item \textbf{BGE-large-en-v1.5} \cite{xiao2023bge}: 1024-dim, open-source, competitive.
\item \textbf{Sentence-BERT} \cite{reimers2019sentencebert}: 768-dim, efficient.
\end{itemize}

Embeddings are stored alongside experiences in the IPFS library.

\subsection{Similarity Search}

Given query embedding $\mathbf{v}_q$, compute cosine similarity with all experience embeddings:
\[
\text{sim}(\mathbf{v}_q, \mathbf{v}_e) = \frac{\mathbf{v}_q \cdot \mathbf{v}_e}{\|\mathbf{v}_q\| \|\mathbf{v}_e\|}
\]

For large libraries ($|E| > 10^6$), use approximate nearest neighbor (ANN) search:
\begin{itemize}
\item \textbf{FAISS} \cite{johnson2019faiss}: Facebook's library for billion-scale vector search.
\item \textbf{HNSW} \cite{malkov2018hnsw}: Hierarchical Navigable Small World graphs.
\item \textbf{ScaNN} \cite{guo2020scann}: Google's optimized ANN.
\end{itemize}

\subsection{Caching and Optimization}

\begin{itemize}
\item \textbf{Precomputed Embeddings}: All experiences have embeddings computed offline; no real-time encoding during inference.
\item \textbf{Experience Deduplication}: Before adding new experiences, check for $\text{sim}(\mathbf{v}_{\text{new}}, \mathbf{v}_e) > 0.9$. If found, merge instead of adding duplicate.
\item \textbf{LRU Cache}: Frequently retrieved experiences cached in GPU memory for zero-latency access.
\end{itemize}

\subsection{Quality Scoring}

Each experience has a quality score:
\[
Q(e) = \alpha \cdot \frac{\text{upvotes}}{\text{upvotes} + \text{downvotes}} + \beta \cdot \frac{\text{usage\_count}}{\max(\text{usage\_counts})} + \gamma \cdot e.\text{confidence}
\]
where $\alpha, \beta, \gamma$ are tunable weights. Low-quality experiences are pruned when the library exceeds maximum size.

