\section{Inference Router}


The Inference Router is a Solidity smart contract coordinating query execution.

\subsection{Contract Interface}

\begin{verbatim}
contract InferenceRouter {
    struct Request {
        bytes32 queryHash;
        address user;
        uint256 timestamp;
        string ipfsCID;  // Experience library version
    }

    mapping(bytes32 => Request) public requests;
    mapping(address => uint256) public nodeStakes;

    function submitQuery(
        string memory query,
        string memory ipfsCID
    ) public payable returns (bytes32 requestId);

    function fulfillQuery(
        bytes32 requestId,
        string memory response,
        bytes memory proof
    ) public;

    function slashNode(address node) public;
}
\end{verbatim}

\subsection{Query Lifecycle}

\begin{enumerate}
\item \textbf{Submission}: User calls \texttt{submitQuery(query, ipfsCID)}, pays fee in native token (ZOO).
\item \textbf{Node Selection}: Router selects GPU node via:
   \begin{itemize}
   \item \textit{Round-robin}: Fair distribution.
   \item \textit{Stake-weighted}: Nodes with higher stakes prioritized.
   \item \textit{Performance-based}: Nodes with lower latency prioritized.
   \end{itemize}
\item \textbf{Execution}: Selected node:
   \begin{enumerate}
   \item Fetches base model (frozen weights) from IPFS.
   \item Fetches experience library at \texttt{ipfsCID}.
   \item Retrieves top-$k$ relevant experiences.
   \item Constructs prompt: \texttt{experiences + query}.
   \item Runs inference.
   \item Optionally generates proof (zkSNARK or optimistic).
   \end{enumerate}
\item \textbf{Fulfillment}: Node calls \texttt{fulfillQuery(requestId, response, proof)}. Contract verifies proof (if provided) and releases payment.
\item \textbf{Dispute}: If proof fails or response is incorrect, user can challenge. Validators re-execute query; if node is malicious, stake is slashed.
\end{enumerate}

\subsection{Proof Systems}

\subsubsection{Optimistic Proofs}

Default mode: no cryptographic proof. Users can challenge within dispute window (e.g., 1 hour). Validators re-run inference; if outputs differ, slash node.

\begin{itemize}
\item \textbf{Pros}: Zero overhead, fast finality.
\item \textbf{Cons}: Requires dispute mechanism, delayed finality.
\end{itemize}

\subsubsection{zkSNARK Proofs}

Node generates a succinct proof that inference was executed correctly:
\[
\pi \leftarrow \text{Prove}(\text{circuit}, \text{model}, \text{input}, \text{output})
\]
Contract verifies:
\[
\text{Verify}(\pi, \text{output}) = \text{True}
\]

\begin{itemize}
\item \textbf{Pros}: Instant finality, no disputes.
\item \textbf{Cons}: Proving overhead ($\sim$10x inference time), circuit complexity.
\end{itemize}

Current implementations: \texttt{zkML} \cite{zkml2023}, \texttt{Modulus Labs EZKL} \cite{ezkl2023}.

