\section{Economic Model}


\subsection{Token Utility: KEEPER}

Zoo's native token, \textbf{KEEPER}, serves three functions:

\begin{enumerate}
\item \textbf{Governance}: 1 KEEPER = 1 vote on experience proposals, treasury allocations, protocol upgrades.
\item \textbf{Staking}: GPU nodes must stake KEEPER as collateral. Malicious behavior results in slashing.
\item \textbf{Fee Payment}: Users pay for inference queries in KEEPER (or wrapped native tokens).
\end{enumerate}

\subsection{Contribution-Based Rewards}

Contributors earn KEEPER via:

\subsubsection{Data Contribution}

Users who submit datasets for Training-Free GRPO receive:
\[
\text{KEEPER}_{\text{data}} = \alpha \cdot |D| \cdot Q(D)
\]
where $|D|$ is dataset size and $Q(D)$ is quality score (based on validation accuracy improvement).

\subsubsection{Experience Submission}

Users who submit experiences (post-DAO approval) receive:
\[
\text{KEEPER}_{\text{exp}} = \beta \cdot \left(\frac{\text{upvotes}}{\text{upvotes} + \text{downvotes}}\right)^2 \cdot \log(1 + \text{usage\_count})
\]
This rewards both initial quality (voting) and long-term utility (usage).

\subsubsection{Inference Provision}

GPU nodes earn:
\[
\text{KEEPER}_{\text{inference}} = \gamma \cdot \text{num\_queries} \cdot \left(1 + \frac{\text{stake}}{\text{total\_stake}}\right)
\]
Higher stake = priority in query routing = more earnings.

\subsection{Fee Structure}

\begin{center}
\begin{tabular}{lrr}
\toprule
\textbf{Model Size} & \textbf{Fee (KEEPER)} & \textbf{USD Equiv.} \\
\midrule
Qwen3-4B & 0.01 & \$0.005 \\
Qwen3-7B & 0.02 & \$0.010 \\
Qwen3-32B & 0.05 & \$0.025 \\
Qwen3-72B & 0.10 & \$0.050 \\
DeepSeek-V3 & 0.50 & \$0.250 \\
\bottomrule
\end{tabular}
\end{center}

10\% of fees burned (deflationary), 40\% to GPU nodes, 40\% to experience contributors, 10\% to treasury.

\subsection{Anti-Sybil Mechanisms}

To prevent fake contributions:
\begin{itemize}
\item \textbf{Proof of Personhood}: Integration with WorldCoin \cite{worldcoin2023whitepaper} or BrightID \cite{brightid2021whitepaper}.
\item \textbf{Stake Requirements}: Minimum 100 KEEPER stake to submit experiences.
\item \textbf{Reputation System}: Contributors build reputation over time; low-reputation users face higher scrutiny.
\end{itemize}

