\section{Security and Byzantine Robustness}


\subsection{Threat Model}

\subsubsection{Malicious GPU Nodes}

Nodes may:
\begin{itemize}
\item Return incorrect outputs (to save computation).
\item Inject backdoor experiences (to manipulate model behavior).
\item Collude to control query routing.
\end{itemize}

\subsubsection{Malicious Contributors}

Contributors may:
\begin{itemize}
\item Submit toxic/biased experiences.
\item Sybil attack to upvote low-quality experiences.
\item Poison training data.
\end{itemize}

\subsubsection{Network-Level Attacks}

\begin{itemize}
\item \textbf{DDoS}: Flood Inference Router with junk queries.
\item \textbf{Eclipse}: Isolate honest nodes from network.
\item \textbf{Censorship}: Prevent specific experiences from propagating.
\end{itemize}

\subsection{Defenses}

\subsubsection{Proof Systems}

\begin{itemize}
\item \textbf{Optimistic}: Fast but requires dispute mechanism.
\item \textbf{zkSNARK}: Slow but provides instant finality.
\item \textbf{Hybrid}: Optimistic by default, zkSNARK for high-value queries.
\end{itemize}

\subsubsection{DAO Governance}

All experience updates require 66\% DAO approval. This prevents single-actor manipulation.

\subsubsection{Slashing}

Nodes caught returning incorrect outputs lose staked KEEPER. Slashing conditions:
\begin{enumerate}
\item Proof verification fails.
\item Output deviates from majority (if $\geq 3$ nodes execute same query).
\item Node is offline during assigned query.
\end{enumerate}

\subsubsection{Rate Limiting}

Inference Router enforces per-user rate limits (e.g., 100 queries/hour for free tier). This mitigates DDoS.

\subsubsection{Redundancy}

Critical queries can be executed by $k \geq 3$ nodes. Outputs are compared via majority voting. If nodes disagree, all are slashed except the majority.

\subsection{Cryptographic Primitives}

\begin{itemize}
\item \textbf{Merkle Trees}: Verify experience inclusion without revealing full library.
\item \textbf{Digital Signatures}: All on-chain transactions signed with ECDSA/EdDSA.
\item \textbf{Post-Quantum}: Lux layer provides CRYSTALS-Dilithium resistance.
\end{itemize}

