\section{Future Work}


\subsection{Technical Enhancements}

\begin{enumerate}
\item \textbf{zk-SNARK Integration}: Deploy efficient zkML circuits for instant proof verification.
\item \textbf{Multi-Agent Coordination}: Enable agents to share experiences and collaborate on complex tasks.
\item \textbf{Automated Hyperparameter Tuning}: Optimize group size, rollout temperature, and retrieval parameters via meta-learning.
\item \textbf{Cross-Chain Bridges}: Enable Zoo experiences to be used on other chains (Ethereum, Solana, Cosmos).
\end{enumerate}

\subsection{Research Directions}

\begin{enumerate}
\item \textbf{Meta-Learning Experiences}: Can experiences guide the generation of new experiences? (Self-improving curation)
\item \textbf{Private Experiences}: Use homomorphic encryption to enable private inference with encrypted experiences.
\item \textbf{Experience Markets}: Trade high-value domain-specific experiences (e.g., medical, legal) with royalties.
\item \textbf{Adversarial Robustness}: Study how malicious experiences can manipulate model behavior and design defenses.
\end{enumerate}

\subsection{Ecosystem Development}

\begin{enumerate}
\item \textbf{Developer Tools}: SDKs for Python, TypeScript, Rust to interact with Zoo API.
\item \textbf{Model Zoo}: Curated collection of fine-tuned models with public experience libraries.
\item \textbf{Bounties}: Incentivize contributions to underrepresented domains (low-resource languages, specialized sciences).
\item \textbf{Partnerships}: Integrate with existing AI platforms (HuggingFace, LangChain, LlamaIndex).
\end{enumerate}

