\section*{Acknowledgments}


Zoo Network is developed by Zoo Labs Foundation Inc, a 501(c)(3) non-profit organization. We thank the Hanzo Network team for providing the base compute infrastructure, the Lux Blockchain team for consensus primitives, and the open-source community for tools and libraries. Special thanks to Tencent for releasing the Training-Free GRPO paper and reference implementation.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{akash2021whitepaper}
Akash Network.
\newblock Akash: Decentralized Cloud Compute Marketplace.
\newblock Technical report, 2021.

\bibitem{arweave2018whitepaper}
Sam Williams and Viktor Tron.
\newblock Arweave: A Protocol for Economically Sustainable Information Permanence.
\newblock Technical report, 2018.

\bibitem{bender2021dangers}
Emily M. Bender, Timnit Gebru, Angelina McMillan-Major, and Shmargaret Shmitchell.
\newblock On the Dangers of Stochastic Parrots: Can Language Models Be Too Big?
\newblock In \textit{FAccT}, 2021.

\bibitem{benet2014ipfs}
Juan Benet.
\newblock IPFS - Content Addressed, Versioned, P2P File System.
\newblock \textit{arXiv:1407.3561}, 2014.

\bibitem{bi2024deepseek}
DeepSeek-AI.
\newblock DeepSeek-V3: Technical Report.
\newblock Technical report, 2024.

\bibitem{bitansky2013snarks}
Nir Bitansky, Alessandro Chiesa, Yuval Ishai, Rafail Ostrovsky, and Omer Paneth.
\newblock Succinct Non-Interactive Arguments via Linear Interactive Proofs.
\newblock In \textit{TCC}, 2013.

\bibitem{bittensor2021whitepaper}
Bittensor Foundation.
\newblock Bittensor: A Peer-to-Peer Intelligence Market.
\newblock Technical report, 2021.

\bibitem{brightid2021whitepaper}
BrightID.
\newblock BrightID: A Proof of Uniqueness Protocol.
\newblock Technical report, 2021.

\bibitem{chu2024qwen2audio}
Yunfei Chu et al.
\newblock Qwen2-Audio: Technical Report.
\newblock \textit{arXiv:2407.10759}, 2024.

\bibitem{dettmers2023qlora}
Tim Dettmers, Artidoro Pagnoni, Ari Holtzman, and Luke Zettlemoyer.
\newblock QLoRA: Efficient Finetuning of Quantized LLMs.
\newblock In \textit{NeurIPS}, 2023.

\bibitem{dwork2014algorithmic}
Cynthia Dwork and Aaron Roth.
\newblock The Algorithmic Foundations of Differential Privacy.
\newblock \textit{Foundations and Trends in Theoretical Computer Science}, 9(3-4):211-407, 2014.

\bibitem{ezkl2023}
Modulus Labs.
\newblock EZKL: Easy Zero-Knowledge Machine Learning.
\newblock \url{https://github.com/zkonduit/ezkl}, 2023.

\bibitem{guo2020scann}
Ruiqi Guo et al.
\newblock Accelerating Large-Scale Inference with Anisotropic Vector Quantization.
\newblock In \textit{ICML}, 2020.

\bibitem{hanzo2025paper}
Hanzo Network.
\newblock Hanzo: Decentralized Compute Infrastructure for AI.
\newblock Technical report, 2025.

\bibitem{hendrycks2021math}
Dan Hendrycks et al.
\newblock Measuring Mathematical Problem Solving With the MATH Dataset.
\newblock In \textit{NeurIPS}, 2021.

\bibitem{hu2021lora}
Edward J. Hu et al.
\newblock LoRA: Low-Rank Adaptation of Large Language Models.
\newblock In \textit{ICLR}, 2022.

\bibitem{huggingface2023tgi}
HuggingFace.
\newblock Text Generation Inference.
\newblock \url{https://github.com/huggingface/text-generation-inference}, 2023.

\bibitem{jiang2023mistral}
Albert Q. Jiang et al.
\newblock Mistral 7B.
\newblock \textit{arXiv:2310.06825}, 2023.

\bibitem{johnson2019faiss}
Jeff Johnson, Matthijs Douze, and Hervé Jégou.
\newblock Billion-scale similarity search with GPUs.
\newblock \textit{IEEE Transactions on Big Data}, 7(3):535-547, 2019.

\bibitem{kim2019blockchained}
Hyesung Kim, Jihong Park, Mehdi Bennis, and Seong-Lyun Kim.
\newblock Blockchained On-Device Federated Learning.
\newblock \textit{IEEE Communications Letters}, 24(6):1279-1283, 2019.

\bibitem{kwon2023vllm}
Woosuk Kwon et al.
\newblock Efficient Memory Management for Large Language Model Serving with PagedAttention.
\newblock In \textit{SOSP}, 2023.

\bibitem{lux2025whitepaper}
Lux Network.
\newblock Lux: Multi-Consensus Blockchain with Post-Quantum Cryptography.
\newblock Technical report, 2025.

\bibitem{malkov2018hnsw}
Yury A. Malkov and Dmitry A. Yashunin.
\newblock Efficient and robust approximate nearest neighbor search using Hierarchical Navigable Small World graphs.
\newblock \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, 42(4):824-836, 2018.

\bibitem{mcmahan2017federated}
H. Brendan McMahan et al.
\newblock Communication-Efficient Learning of Deep Networks from Decentralized Data.
\newblock In \textit{AISTATS}, 2017.

\bibitem{ocean2021whitepaper}
Ocean Protocol Foundation.
\newblock Ocean Protocol: A Decentralized Data Exchange Protocol.
\newblock Technical report, 2021.

\bibitem{ouyang2022training}
Long Ouyang et al.
\newblock Training language models to follow instructions with human feedback.
\newblock In \textit{NeurIPS}, 2022.

\bibitem{qwen2024qwen3}
Qwen Team.
\newblock Qwen3 Technical Report.
\newblock Technical report, 2024.

\bibitem{rafailov2023dpo}
Rafael Rafailov et al.
\newblock Direct Preference Optimization: Your Language Model is Secretly a Reward Model.
\newblock In \textit{NeurIPS}, 2023.

\bibitem{rasley2020deepspeed}
Jeff Rasley et al.
\newblock DeepSpeed: System Optimizations Enable Training Deep Learning Models with Over 100 Billion Parameters.
\newblock In \textit{KDD}, 2020.

\bibitem{reimers2019sentencebert}
Nils Reimers and Iryna Gurevych.
\newblock Sentence-BERT: Sentence Embeddings using Siamese BERT-Networks.
\newblock In \textit{EMNLP-IJCNLP}, 2019.

\bibitem{rocket2020snowman}
Team Rocket et al.
\newblock Scalable and Probabilistic Leaderless BFT Consensus through Metastability.
\newblock \textit{arXiv:1906.08936}, 2020.

\bibitem{shao2024deepseekmath}
Zhihong Shao et al.
\newblock DeepSeekMath: Pushing the Limits of Mathematical Reasoning in Open Language Models.
\newblock \textit{arXiv:2402.03300}, 2024.

\bibitem{shoeybi2019megatron}
Mohammad Shoeybi et al.
\newblock Megatron-LM: Training Multi-Billion Parameter Language Models Using Model Parallelism.
\newblock \textit{arXiv:1909.08053}, 2019.

\bibitem{singularitynet2021whitepaper}
SingularityNET Foundation.
\newblock SingularityNET: A Decentralized, Open Market and Network for AIs.
\newblock Technical report, 2021.

\bibitem{tencent2025grpo}
Tencent youtu-agent Team.
\newblock Training-Free Group Relative Policy Optimization.
\newblock \textit{arXiv:2510.08191}, 2025.

\bibitem{touvron2023llama}
Hugo Touvron et al.
\newblock LLaMA: Open and Efficient Foundation Language Models.
\newblock \textit{arXiv:2302.13971}, 2023.

\bibitem{worldcoin2023whitepaper}
Worldcoin Foundation.
\newblock Worldcoin: A Privacy-Preserving Proof-of-Personhood Protocol.
\newblock Technical report, 2023.

\bibitem{xiao2023bge}
Shitao Xiao et al.
\newblock C-Pack: Packaged Resources To Advance General Chinese Embedding.
\newblock \textit{arXiv:2309.07597}, 2023.

\bibitem{zkml2023}
zkML Community.
\newblock Zero-Knowledge Machine Learning.
\newblock \url{https://github.com/zkml-community}, 2023.

\bibitem{zoo2025hllm}
Zoo Labs Foundation.
\newblock Hamiltonian Large Language Models: Training-Free Semantic Optimization.
\newblock Technical report, 2025.

\end{thebibliography}

\appendix

