\section{Conclusion}


\KChain{} demonstrates that distributed post-quantum key management is practical at blockchain consensus timescales. By replacing centralized KMS with validator-distributed threshold shares and NIST-standardized post-quantum primitives, Zoo Network achieves:

\begin{itemize}
\item \textbf{No single point of compromise}: Key material never exists in complete form on any node.
\item \textbf{Quantum resistance}: ML-KEM and ML-DSA provide security against both classical and quantum adversaries.
\item \textbf{Performance}: GPU NTT acceleration makes post-quantum operations practical for real-time consensus.
\item \textbf{Formal guarantees}: Lean 4 proofs establish correctness of all cryptographic primitives and their threshold adaptations.
\item \textbf{Secure implementation}: Constant-time operations, locked memory, and cryptographic clearing prevent side-channel attacks.
\end{itemize}

Future work includes integration of hybrid post-quantum TLS for validator communication, homomorphic encryption for computation on encrypted key shares, and hardware security module (HSM) integration for validators requiring additional physical security guarantees.

\begin{thebibliography}{99}

\bibitem{fips203}
NIST, ``Module-Lattice-Based Key-Encapsulation Mechanism Standard (FIPS 203),'' 2024.

\bibitem{fips204}
NIST, ``Module-Lattice-Based Digital Signature Standard (FIPS 204),'' 2024.

\bibitem{fips205}
NIST, ``Stateless Hash-Based Digital Signature Standard (FIPS 205),'' 2024.

\bibitem{shamir1979share}
A. Shamir, ``How to Share a Secret,'' \textit{Communications of the ACM}, vol. 22, no. 11, 1979.

\bibitem{feldman1987practical}
P. Feldman, ``A Practical Scheme for Non-interactive Verifiable Secret Sharing,'' in \textit{FOCS}, 1987.

\bibitem{zoo-threshold-signatures}
Z. Kelling, ``Threshold Signatures on Zoo Network,'' Zoo Labs Foundation, 2026.

\bibitem{zoo-identity-chain}
Z. Kelling, ``Decentralized Identity on Zoo Network: Chain-Native DIDs and Verifiable Credentials via I-Chain,'' Zoo Labs Foundation, 2026.

\bibitem{zoo-pq-crypto}
Z. Kelling, ``Post-Quantum Cryptography in Zoo Network,'' Zoo Labs Foundation, 2026.

\bibitem{lux-pq-crypto-suite}
Lux Industries, ``Lux Post-Quantum Cryptography Suite,'' 2025.

\end{thebibliography}

