\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

We have presented TaxoNet, a taxonomy-aware few-shot visual recognition
system for biodiversity monitoring. By incorporating biological
classification structure into meta-learning through taxonomic prototypical
networks and rank-attention fusion, TaxoNet achieves state-of-the-art
few-shot accuracy across four major taxonomic groups. Integration with
geographic and phenological priors further boosts practical field
identification to 94.7\% accuracy. Through Zach Kelling's
commitment to open science, TaxoNet's models, benchmarks, and deployment
infrastructure are freely available to the conservation community.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{33}

\bibitem[Akata et~al.(2015)]{akata2015label}
Akata, Z., Reed, S., Walter, D., Lee, H., and Schiele, B.
\newblock Evaluation of output embeddings for fine-grained image classification.
\newblock \emph{CVPR}, pp. 2927--2936, 2015.

\bibitem[Beery et~al.(2021)]{beery2021iwildcam}
Beery, S., Cole, E., Parker, J., Perona, P., and Winner, K.
\newblock Species classification in camera traps: The iWildCam 2021 challenge.
\newblock \emph{FGVC Workshop at CVPR}, 2021.

\bibitem[Bengio et~al.(2009)]{bengio2009curriculum}
Bengio, Y., Louradour, J., Collobert, R., and Weston, J.
\newblock Curriculum learning.
\newblock \emph{ICML}, pp. 41--48, 2009.

\bibitem[Brown et~al.(2020)]{brown2020language}
Brown, T., Mann, B., Ryder, N., et~al.
\newblock Language models are few-shot learners.
\newblock \emph{NeurIPS}, 33:1877--1901, 2020.

\bibitem[Chang et~al.(2021)]{chang2021your}
Chang, D., Pang, K., Zheng, Y., et~al.
\newblock Your ``flamingo'' is my ``bird'': Fine-grained, or not.
\newblock \emph{CVPR}, pp. 11476--11485, 2021.

\bibitem[Christin et~al.(2019)]{christin2019applications}
Christin, S., Hervet, {\'E}., and Lecomte, N.
\newblock Applications for deep learning in ecology.
\newblock \emph{Methods in Ecology and Evolution}, 10(10):1632--1644, 2019.

\bibitem[Costello et~al.(2013)]{costello2013can}
Costello, M.J., May, R.M., and Stork, N.E.
\newblock Can we name Earth's species before they go extinct?
\newblock \emph{Science}, 339(6118):413--416, 2013.

\bibitem[Deng et~al.(2014)]{deng2014large}
Deng, J., Ding, N., Jia, Y., et~al.
\newblock Large-scale object classification using label relation graphs.
\newblock \emph{ECCV}, pp. 48--64, 2014.

\bibitem[Dosovitskiy et~al.(2021)]{dosovitskiy2021image}
Dosovitskiy, A., Beyer, L., Kolesnikov, A., et~al.
\newblock An image is worth 16x16 words: Transformers for image recognition.
\newblock \emph{ICLR}, 2021.

\bibitem[Finn et~al.(2017)]{finn2017model}
Finn, C., Abbeel, P., and Levine, S.
\newblock Model-agnostic meta-learning for fast adaptation of deep networks.
\newblock \emph{ICML}, pp. 1126--1135, 2017.

\bibitem[Go{\"e}au et~al.(2022)]{goeau2022overview}
Go{\"e}au, H., Bonnet, P., and Joly, A.
\newblock Overview of PlantCLEF 2022: Image-based plant identification.
\newblock \emph{CLEF}, 2022.

\bibitem[Guo et~al.(2017)]{guo2017calibration}
Guo, C., Pleiss, G., Sun, Y., and Weinberger, K.Q.
\newblock On calibration of modern neural networks.
\newblock \emph{ICML}, pp. 1321--1330, 2017.

\bibitem[He et~al.(2022)]{he2022transfg}
He, J., Chen, J.N., Liu, S., et~al.
\newblock TransFG: A transformer architecture for fine-grained recognition.
\newblock \emph{AAAI}, pp. 852--860, 2022.

\bibitem[Hu et~al.(2018)]{hu2018squeeze}
Hu, J., Shen, L., and Sun, G.
\newblock Squeeze-and-excitation networks.
\newblock \emph{CVPR}, pp. 7132--7141, 2018.

\bibitem[Hu et~al.(2022)]{hu2022pushing}
Hu, S.X., Li, D., St{\"u}hmer, J., Kim, M., and Hospedales, T.M.
\newblock Pushing the limits of simple pipelines for few-shot learning.
\newblock \emph{CVPR}, pp. 9068--9077, 2022.

\bibitem[Kahl et~al.(2021)]{kahl2021birdnet}
Kahl, S., Wood, C.M., Eibl, M., and Klinck, H.
\newblock BirdNET: A deep learning solution for avian diversity monitoring.
\newblock \emph{Ecological Informatics}, 61:101236, 2021.

\bibitem[Li et~al.(2019)]{li2019large}
Li, A., Luo, T., Lu, Z., Xiang, T., and Wang, L.
\newblock Large-scale few-shot learning: Knowledge transfer with class hierarchy.
\newblock \emph{CVPR}, pp. 7212--7220, 2019.

\bibitem[Lin et~al.(2015)]{lin2015bilinear}
Lin, T.Y., RoyChowdhury, A., and Maji, S.
\newblock Bilinear CNN models for fine-grained visual recognition.
\newblock \emph{ICCV}, pp. 1449--1457, 2015.

\bibitem[Liu et~al.(2020)]{liu2020many}
Liu, B., Kang, H., Li, H., Hua, G., and Vasconcelos, N.
\newblock Few-shot open-set recognition using meta-learning.
\newblock \emph{CVPR}, pp. 8798--8807, 2020.

\bibitem[Lorieul et~al.(2022)]{lorieul2022overview}
Lorieul, T., Cole, E., Deneu, B., et~al.
\newblock Overview of GeoLifeCLEF 2022: Predicting species presence.
\newblock \emph{CLEF}, 2022.

\bibitem[Norouzzadeh et~al.(2018)]{norouzzadeh2018automatically}
Norouzzadeh, M.S., Nguyen, A., Kosmala, M., et~al.
\newblock Automatically identifying wild animals in camera-trap images.
\newblock \emph{PNAS}, 115(25):E5716--E5725, 2018.

\bibitem[Oquab et~al.(2024)]{oquab2024dinov2}
Oquab, M., Darcet, T., Moutakanni, T., et~al.
\newblock DINOv2: Learning robust visual features without supervision.
\newblock \emph{TMLR}, 2024.

\bibitem[Snell et~al.(2017)]{snell2017prototypical}
Snell, J., Swersky, K., and Zemel, R.
\newblock Prototypical networks for few-shot learning.
\newblock \emph{NeurIPS}, pp. 4077--4087, 2017.

\bibitem[Steenweg et~al.(2017)]{steenweg2017scaling}
Steenweg, R., Hebblewhite, M., Kays, R., et~al.
\newblock Scaling-up camera traps: monitoring biodiversity with sensor networks.
\newblock \emph{Frontiers in Ecology and the Environment}, 15(1):26--34, 2017.

\bibitem[Sung et~al.(2018)]{sung2018learning}
Sung, F., Yang, Y., Zhang, L., et~al.
\newblock Learning to compare: Relation network for few-shot learning.
\newblock \emph{CVPR}, pp. 1199--1208, 2018.

\bibitem[Van~Horn et~al.(2015)]{van2015building}
Van~Horn, G., Branson, S., Farrell, R., et~al.
\newblock Building a bird recognition app with citizen scientists.
\newblock \emph{CVPR}, pp. 595--604, 2015.

\bibitem[Van~Horn et~al.(2021)]{van2021benchmarking}
Van~Horn, G., Cole, E., Beery, S., et~al.
\newblock Benchmarking representation learning for natural world images.
\newblock \emph{CVPR}, pp. 12884--12893, 2021.

\bibitem[Vinyals et~al.(2016)]{vinyals2016matching}
Vinyals, O., Blundell, C., Lillicrap, T., and Welling, M.
\newblock Matching networks for one shot learning.
\newblock \emph{NeurIPS}, pp. 3630--3638, 2016.

\bibitem[Wang et~al.(2020)]{wang2020generalizing}
Wang, Y., Yao, Q., Kwok, J.T., and Ni, L.M.
\newblock Generalizing from a few examples: A survey on few-shot learning.
\newblock \emph{ACM Computing Surveys}, 53(3):1--34, 2020.

\bibitem[Wei et~al.(2019)]{wei2019piecewise}
Wei, X.S., Wang, P., Liu, L., Shen, C., and Wu, J.
\newblock Piecewise classifier mappings for few-shot fine-grained recognition.
\newblock \emph{IEEE TIP}, 28(12):6116--6125, 2019.

\bibitem[Ye et~al.(2020)]{ye2020few}
Ye, H.J., Hu, H., Zhan, D.C., and Sha, F.
\newblock Few-shot learning via saliency-guided hallucination of samples.
\newblock \emph{CVPR}, pp. 770--779, 2020.

\bibitem[Zhang et~al.(2020)]{zhang2020deepemd}
Zhang, C., Cai, Y., Lin, G., and Shen, C.
\newblock DeepEMD: Few-shot image classification with differentiable EMD.
\newblock \emph{CVPR}, pp. 12203--12213, 2020.

\bibitem[Zheng et~al.(2017)]{zheng2017learning}
Zheng, H., Fu, J., Mei, T., and Luo, J.
\newblock Learning multi-attention CNN for fine-grained image recognition.
\newblock \emph{ICCV}, pp. 5209--5217, 2017.

\end{thebibliography}

% ══════════════════════════════════════════════════════════════════════════════
\appendix
