\section{Data Commons}

\label{sec:commons}
%----------------------------------------------------------------------

\subsection{DID-Authenticated Data Contributions}

The \Zoo{} Data Commons \cite{zoo-data-commons} is a decentralized repository where researchers contribute datasets with cryptographic provenance:

\begin{enumerate}
    \item The contributor authenticates via their \IChain{} \DID{}.
    \item The dataset is hashed (BLAKE3) and the hash is anchored on \IChain{} with the contributor's \DID{} as author.
    \item The dataset is stored on decentralized storage (IPFS/Filecoin) with the on-chain hash as the content address.
    \item Metadata (schema, license, domain, quality metrics) is stored on-chain.
\end{enumerate}

\subsection{Usage Receipts}

Every use of a dataset in a Research Capsule generates a \emph{usage receipt}---an on-chain record linking the consuming capsule to the contributing dataset:

\begin{verbatim}
struct UsageReceipt {
    dataset_hash:  bytes32,
    consumer_did:  bytes32,
    capsule_id:    bytes32,
    usage_type:    uint8,    // 0=training, 1=validation,
                             // 2=analysis, 3=reference
    timestamp:     uint64
}
\end{verbatim}

Usage receipts serve two purposes:
\begin{enumerate}
    \item \textbf{Attribution.} Data contributors receive verifiable credit for each use of their data.
    \item \textbf{Compensation.} Contributors earn $\KEEPER$ tokens proportional to usage. The default rate is 100 $\KEEPER$ per receipt, paid from the consuming capsule's budget.
\end{enumerate}

\subsection{Citation Graphs on G-Chain}

Research Capsules cite each other, creating a directed acyclic graph of scientific knowledge. This citation graph is stored on \GChain{} and provides:

\begin{itemize}
    \item \textbf{Impact Metrics.} Citation count, h-index, and PageRank-style importance scores computed on-chain.
    \item \textbf{Retroactive Funding.} High-impact work (measured by citations) can receive retroactive \DAO{} funding, rewarding quality after the fact rather than guessing impact before.
    \item \textbf{Dependency Tracking.} If a foundational dataset or method is retracted, all dependent capsules are flagged automatically.
\end{itemize}

\subsection{FAIR Compliance}

The Data Commons natively satisfies the FAIR principles \cite{wilkinson2016fair}:

\begin{itemize}
    \item \textbf{Findable.} Content-addressed with persistent identifiers on \IChain{}.
    \item \textbf{Accessible.} Open by default. Sensitive data is threshold-gated but discoverable (metadata is public).
    \item \textbf{Interoperable.} Schema metadata enforces standard formats. Cross-references use DIDs.
    \item \textbf{Reusable.} License and provenance are on-chain. Usage terms are machine-readable.
\end{itemize}

%----------------------------------------------------------------------
