\section{Conservation Science Applications}

\label{sec:conservation}
%----------------------------------------------------------------------

\Zoo{} Labs Foundation's mission centers on conservation. The DeSci platform provides immediate value through three existing research programs.

\subsection{Satellite Ecology}

The \Zoo{} satellite ecology program \cite{zoo-satellite-ecology} uses remote sensing data for habitat monitoring:

\begin{itemize}
    \item \textbf{Data.} Sentinel-2 and Landsat imagery contributed to the Data Commons by governmental and institutional partners.
    \item \textbf{Analysis.} \AChain{} runs land-use classification models on satellite imagery, detecting deforestation, habitat fragmentation, and urban encroachment.
    \item \textbf{Provenance.} Analysis results are anchored in Research Capsules with full reproducibility (container image + input data hash + model weights hash).
    \item \textbf{Funding.} Satellite ecology research is funded through Research ZIPs, with milestone deliverables including classified imagery and habitat change reports.
\end{itemize}

\subsection{Species Classification}

The species classification pipeline \cite{zoo-species-classification} uses AI for biodiversity assessment:

\begin{itemize}
    \item \textbf{Data.} Camera trap images, acoustic recordings, and citizen-science observations contributed to the Data Commons.
    \item \textbf{Models.} Classification models trained on \AChain{} with federated learning across institutions \cite{zoo-federated-wildlife}. Each institution contributes training data without sharing raw images (privacy preserved via federated aggregation).
    \item \textbf{Validation.} Expert validation of model predictions is incentivized: domain experts earn $\KEEPER$ for confirming or correcting classifications.
    \item \textbf{Output.} Species distribution maps, population estimates, and trend analyses published as Research Capsules.
\end{itemize}

\subsection{Habitat Modeling}

Predictive habitat modeling integrates satellite, species, and climate data:

\begin{itemize}
    \item \textbf{Multi-Source Fusion.} \AChain{} inference combines satellite-derived land cover, species occurrence records, and climate projections.
    \item \textbf{Scenario Analysis.} Models evaluate habitat viability under different climate scenarios (RCP 2.6, 4.5, 8.5).
    \item \textbf{Conservation Planning.} Output feeds directly into conservation action plans, identifying priority areas for protection.
    \item \textbf{Impact Bonds.} Conservation outcomes (measured via satellite monitoring) can trigger \Zoo{} impact bond payouts \cite{zoo-impact-bonds}, creating a financial feedback loop between science and action.
\end{itemize}

%----------------------------------------------------------------------
