\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{shapes,arrows,positioning,calc,decorations.pathreplacing}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{longtable}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}[section]

% Custom commands (match per-LLM chains style)
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Zen}{\textsc{Zen}}
\newcommand{\Quasar}{\textsc{Quasar}}
\newcommand{\QuasarGPU}{\textsc{Quasar}\textsc{gpu}}
\newcommand{\Liquidity}{\textsc{Liquidity}}
\newcommand{\PoAI}{\textsc{PoAI}}
\newcommand{\TEE}{\textsc{tee}}
\newcommand{\LLM}{\textsc{llm}}
\newcommand{\GPU}{\textsc{gpu}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\AMM}{\textsc{amm}}
\newcommand{\DEX}{\textsc{dex}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\FHE}{\textsc{fhe}}
\newcommand{\ZKP}{\textsc{zkp}}
\newcommand{\BLS}{\textsc{bls}}
\newcommand{\MLDSA}{\textsc{ml-dsa}}
\newcommand{\Corona}{\textsc{Corona}}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray},
    tabsize=2,
    captionpos=b,
}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\title{\textbf{Zoo 4.0 --- \GPU{}-Native Sovereign L1}\\
\large From \Lux{} L2 application to sovereign L1 with 100\%
\GPU{}-powered consensus, \textsc{evm}, AI mining, \DEX{}, and \FHE{}}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{February 14, 2026}

\begin{document}

\zoocoverpage

\begin{abstract}
\Zoo{} 4.0 is the \GPU{}-native graduation of the \Zoo{} stack. The
prior eras lead up to this release: \Zoo{} 1.0 (2021, BSC) coined the
``NFT Liquidity Protocol''; \Zoo{} 2.0 (October 2024) re-launched on
\Lux{} as an \textsc{evm}-compatible L2 with first-pass post-quantum
precompiles \cite{zoo-2-0}; \Zoo{} 3.0 (2025) brought the entire stack
to a fully post-quantum settlement model with triple-cert consensus
inheritance from \Lux{} \Quasar{} 3.0 \cite{zoo-3-0,lp-020}. \Zoo{} 4.0
takes the next step: every hot-path primitive --- consensus signature
aggregation, \BLS{}12-381 verification, \Corona{} threshold proofs,
\MLDSA{}-65 identity proofs, \textsc{evm} bytecode execution, \AMM{}
curve math, \FHE{} ciphertext arithmetic, and Groth16/Halo2 \ZKP{}
verification --- runs as deterministic \GPU{} kernels in pure C++ via
the \QuasarGPU{} adapter (\textsc{lp-132}) under the \GPU{}-Residency
Invariant (\textsc{lp-137}). With this release, \Zoo{} graduates from
a \Lux{} L2 application to a sovereign \Lux{} L1 with its own validator
set, \Quasar{}-certified consensus, and independent economic security,
while continuing to settle and bridge to the broader \Lux{} chain
topology (\textsc{lp-134}). This paper documents the 4.0 architecture:
the L1 graduation, the \GPU{}-native execution stack, the \Zoo{} \DEX{}
(native \AMM{} V2/V3, V4 as a precompile to the \Lux{} \DEX{}), the
\Zoo{} AI Desktop application hosting the \Zen{} model family for free
local private inference, the \texttt{zoo-bot} pure-Go re-write of the
\Hanzo{} OpenClaw bot, the \Zoo{} Bridge multi-ecosystem connector,
dual-path privacy via F-Chain \FHE{} (\textsc{lp-013}) and Z-Chain
\ZKP{} (\textsc{lp-063}), and the \NFT{} migration program. Spec
freeze: 2026-01-15. Activation alongside \Lux{} \Quasar{} 4.0:
2026-02-14.

\textbf{Keywords}: \Zoo{} 4.0, \GPU{}-native, L1 graduation, \Quasar{} 4.0, \Zoo{} \DEX{}, \Zoo{} AI Desktop, \texttt{zoo-bot}, \Zoo{} Bridge, \FHE{}, \ZKP{}, \NFT{} migration
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-thesis}
\input{sections/01a-triumvirate}
\input{sections/02-l1-graduation}
\input{sections/03-gpu-native}
\input{sections/04-zoo-dex}
\input{sections/05-ai-desktop}
\input{sections/06-zoo-bot}
\input{sections/07-zoo-bridge}
\input{sections/08-privacy}
\input{sections/09-nft-strategy}
\input{sections/10-wallet-desktop}
\input{sections/11-activation}

\begin{thebibliography}{99}

\bibitem{lp-009}
Lux Foundation.
\newblock LP-009: \GPU{}-Native \textsc{evm}.
\newblock Lux Improvement Proposal, December 2025.
\newblock \url{https://github.com/luxfi/lps/lp-009}.

\bibitem{lp-013}
Lux Foundation.
\newblock LP-013: F-Chain \FHE{}-Protected Inference.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-020}
Lux Foundation.
\newblock LP-020: \Quasar{} Consensus 3.0 (Triple Consensus: \BLS{} + \Corona{} + \MLDSA{}).
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-063}
Lux Foundation.
\newblock LP-063: Z-Chain Zero-Knowledge Rollup and \ZKP{} Registry.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-067}
Lux Foundation.
\newblock LP-067: Confidential ERC-20 on F-Chain.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-105}
Lux Foundation.
\newblock LP-105: \Quasar{} Consensus Family Specification (Photon, Wave, Focus, Nova, Nebula).
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-132}
Lux Foundation.
\newblock LP-132: \QuasarGPU{} Execution Adapter for Deterministic \GPU{} Scheduling.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-133}
Lux Foundation.
\newblock LP-133: \Quasar{}-Native App Stack for Sovereign Appchains.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-134}
Lux Foundation.
\newblock LP-134: \Lux{} Chain Topology (P/C/X/Q/Z/A/B/M/F-Chains).
\newblock Lux Improvement Proposal, December 2025.

\bibitem{lp-137}
Lux Foundation.
\newblock LP-137: \GPU{}-Residency Invariant.
\newblock Lux Improvement Proposal, December 2025.

\bibitem{hanzo-ai-chain}
Hanzo AI.
\newblock The \Hanzo{} AI Chain: Inference Marketplace and Commerce Layer.
\newblock Hanzo Whitepaper, December 2025.

\bibitem{zoo-per-llm-chains}
Zoo Labs Foundation.
\newblock Per-\LLM{} Chains: One Sovereign Chain per Model on \Quasar{} 3.0.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{zoo-2-0}
Zoo Labs Foundation.
\newblock Zoo \textsc{evm}: A Post-Quantum Safe L2 Blockchain with AI-Native Precompiles on the \Lux{} Network.
\newblock Zoo Whitepaper, October 2024.

\bibitem{zoo-3-0}
Zoo Labs Foundation.
\newblock Zoo 3.0 --- Full Post-Quantum Stack.
\newblock Zoo Whitepaper, October 31, 2025.

\bibitem{zoo-dex-launch}
Zoo Labs Foundation.
\newblock Zoo \DEX{} --- Native Securities Trading on \Lux{} \Quasar{} with \Liquidity{} Protocol Integration.
\newblock Zoo Whitepaper, April 20, 2026.

\bibitem{zoo-2025-securities-and-dao}
Zoo Labs Foundation.
\newblock Zoo 2025: Digital Securities, Quantum-Secure Settlement, and the Holographic Consensus DAO.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{zoo-dex}
Zoo Labs Foundation.
\newblock \Zoo{} \DEX{}: \AMM{} V2/V3 Native, V4 Precompile.
\newblock Zoo Whitepaper, December 2025.

\bibitem{lux-dex}
Lux Network.
\newblock \Lux{} \DEX{} Specification.
\newblock \url{https://github.com/luxfi/dex}, December 2025.

\bibitem{zen-models}
Zen LM.
\newblock The \Zen{} Model Family: Open-Weight Frontier Models.
\newblock \url{https://huggingface.co/zenlm}, December 2025.

\bibitem{uniswap-v3}
Adams, H., et al.
\newblock Uniswap v3 Core.
\newblock Uniswap Whitepaper, March 2021.

\bibitem{tee-attestation}
Costan, V., Devadas, S.
\newblock Intel SGX Explained.
\newblock IACR Cryptology ePrint Archive, 2016.

\end{thebibliography}

\end{document}
