\section{Introduction}

\label{sec:intro}
%----------------------------------------------------------------------

\subsection{The MEV Crisis}

Maximal Extractable Value (\MEV{}) is the profit available to block producers by reordering, inserting, or censoring transactions within a block. On Ethereum, \MEV{} extraction exceeded \$1.5 billion in 2024 alone \cite{flashbots2024mev}, borne entirely by ordinary users through:

\begin{enumerate}
    \item \textbf{Front-Running.} A searcher observes a pending swap in the mempool, submits an identical swap with higher gas to execute first, and profits from the price impact.
    \item \textbf{Sandwich Attacks.} A searcher brackets a victim's swap with a buy-before and sell-after, capturing the slippage as profit.
    \item \textbf{Just-in-Time Liquidity.} A searcher provides liquidity immediately before a large swap (capturing fees) and removes it immediately after (avoiding impermanent loss).
    \item \textbf{Time-Bandit Attacks.} Block producers reorg recent blocks to capture high-value \MEV{} from past transactions.
\end{enumerate}

These attacks share a common root cause: \textbf{transaction content is visible before execution.} The mempool is a public bulletin board. Anyone who can read it faster than others can extract value.

\subsection{Why Existing Solutions Fall Short}

\begin{table}[h]
\centering
\begin{tabular}{lccc}
\toprule
\textbf{Solution} & \textbf{Orderflow Privacy} & \textbf{On-Chain Matching} & \textbf{No Trusted Party} \\
\midrule
Flashbots Protect & Partial (relay sees orders) & No (off-chain relay) & No \\
MEV-Share & Partial (hints leaked) & No (off-chain) & No \\
Penumbra & Full (shielded pool) & No (MASP, not matching) & Yes \\
Renegade & Full (MPC matching) & No (off-chain MPC) & Partial \\
\Zoo{} Confidential DeFi & \textbf{Full (FHE)} & \textbf{Yes} & \textbf{Yes} \\
\bottomrule
\end{tabular}
\caption{Comparison of \MEV{} mitigation approaches}
\label{tab:mev_comparison}
\end{table}

Flashbots and MEV-Share reduce \MEV{} by hiding orderflow from searchers, but the relay operator still observes all transactions. Penumbra provides shielded transfers but does not support on-chain order matching. Renegade uses MPC for dark-pool matching but relies on off-chain computation with a limited participant set.

\subsection{Our Approach}

\Zoo{} Confidential DeFi encrypts the entire orderflow lifecycle:

\begin{enumerate}
    \item Orders are encrypted under a collective \FHE{} public key before entering the mempool.
    \item The \DEX{} VM evaluates encrypted orders homomorphically---price comparisons, volume matching, and settlement calculations are performed on ciphertexts.
    \item Only matched order pairs are decrypted via \TChain{} threshold protocol.
    \item Unmatched orders remain encrypted; no information is revealed about their content.
\end{enumerate}

This eliminates \MEV{} at the source: if no one can read the orders, no one can front-run them.

\subsection{Paper Organization}

Section~\ref{sec:primitives} introduces the cryptographic primitives. Section~\ref{sec:encrypted_orderflow} describes the encrypted orderflow protocol. Section~\ref{sec:matching} presents the FHE order matching engine. Section~\ref{sec:auctions} covers sealed-bid auctions. Section~\ref{sec:amm} describes the private AMM. Section~\ref{sec:settlement} details threshold settlement. Section~\ref{sec:comparison} compares against existing solutions. Section~\ref{sec:benchmarks} presents performance benchmarks. Section~\ref{sec:security} analyzes security properties. Section~\ref{sec:conclusion} concludes.

%----------------------------------------------------------------------
