\section{FHE Order Matching}

\label{sec:matching}
%----------------------------------------------------------------------

\subsection{Price Comparison Circuit}

The core matching operation is comparing an encrypted bid price $\hat{p}_b$ against an encrypted ask price $\hat{p}_a$. Using \TFHE{}, we evaluate a greater-than-or-equal circuit on the bit representations:

\begin{algorithm}
\caption{Encrypted Price Comparison (\TFHE{})}
\begin{algorithmic}[1]
\REQUIRE Encrypted bid price $\hat{p}_b = (\hat{b}_{63}, \ldots, \hat{b}_0)$, encrypted ask price $\hat{p}_a = (\hat{a}_{63}, \ldots, \hat{a}_0)$
\STATE $\hat{r} \leftarrow \text{TFHE.Encrypt}(pk, 0)$ \COMMENT{result bit}
\FOR{$i = 63$ \textbf{downto} $0$}
    \STATE $\hat{e}_i \leftarrow \text{TFHE.XNOR}(\hat{b}_i, \hat{a}_i)$ \COMMENT{bits equal?}
    \STATE $\hat{g}_i \leftarrow \text{TFHE.AND}(\hat{b}_i, \text{TFHE.NOT}(\hat{a}_i))$ \COMMENT{bid bit > ask bit?}
    \STATE $\hat{r} \leftarrow \text{TFHE.MUX}(\hat{e}_i, \hat{r}, \hat{g}_i)$ \COMMENT{propagate or update}
\ENDFOR
\RETURN $\hat{r}$ \COMMENT{encrypted bit: 1 if $p_b \geq p_a$, 0 otherwise}
\end{algorithmic}
\end{algorithm}

This circuit evaluates in $O(b)$ where $b = 64$ is the bit width of the price representation. With programmable bootstrapping after each gate, noise does not accumulate.

\subsection{Volume Matching}

When a price match is found ($\hat{r} = \hat{1}$), the fill quantity is computed via \CKKS{}:

\begin{equation}
\hat{q}_{\textit{fill}} = \text{CKKS.Min}(\hat{q}_b, \hat{q}_a)
\end{equation}

where $\text{CKKS.Min}$ is implemented using the sign function approximated by a Chebyshev polynomial of degree 16:

\begin{equation}
\text{CKKS.Min}(\hat{x}, \hat{y}) = \frac{\hat{x} + \hat{y} - |\hat{x} - \hat{y}|}{2}
\end{equation}

with $|\hat{z}| \approx \hat{z} \cdot \text{sign}(\hat{z})$ evaluated homomorphically.

\subsection{Matching Engine Architecture}

The matching engine processes orders in price-time priority:

\begin{algorithm}
\caption{FHE Order Matching Engine}
\begin{algorithmic}[1]
\REQUIRE Encrypted order books $\hat{\mathcal{B}}$ (bids), $\hat{\mathcal{A}}$ (asks), sorted by submission time
\STATE $\textit{matches} \leftarrow \emptyset$
\FOR{each bid $\hat{O}_b \in \hat{\mathcal{B}}$}
    \FOR{each ask $\hat{O}_a \in \hat{\mathcal{A}}$}
        \STATE $\hat{r} \leftarrow \text{TFHE.GEQ}(\hat{p}_b, \hat{p}_a)$ \COMMENT{price match?}
        \STATE $\hat{q}_f \leftarrow \text{CKKS.Min}(\hat{q}_b, \hat{q}_a)$ \COMMENT{fill quantity}
        \STATE $\hat{m} \leftarrow \text{TFHE.AND}(\hat{r}, \text{TFHE.GT}(\hat{q}_f, \hat{0}))$ \COMMENT{valid match?}
        \STATE Append $(\hat{O}_b, \hat{O}_a, \hat{q}_f, \hat{m})$ to $\textit{matches}$
        \IF{$\hat{m}$ decrypts to 1 (via threshold)}
            \STATE Update remaining quantities: $\hat{q}_b \leftarrow \hat{q}_b - \hat{q}_f$, $\hat{q}_a \leftarrow \hat{q}_a - \hat{q}_f$
        \ENDIF
    \ENDFOR
\ENDFOR
\RETURN $\textit{matches}$
\end{algorithmic}
\end{algorithm}

\begin{remark}
The naive $O(|\mathcal{B}| \cdot |\mathcal{A}|)$ matching is acceptable because: (a) each comparison is independent and parallelizable across GPU cores, and (b) active order book depth on \Zoo{} \DEX{} is bounded (typically $< 1{,}000$ orders per pair).
\end{remark}

\subsection{Batch Matching Optimization}

Rather than matching orders one-at-a-time, the engine batches all orders received within a block period (2 seconds) and matches them simultaneously. This provides:

\begin{itemize}
    \item \textbf{Fairness.} All orders in a batch are treated equally---no time advantage within a block.
    \item \textbf{Efficiency.} GPU parallelism is maximized when processing a batch.
    \item \textbf{MEV elimination.} Reordering within a batch does not affect matching outcomes when price-time priority considers only batch ordering.
\end{itemize}

%----------------------------------------------------------------------
