\section{Sealed-Bid Auctions}

\label{sec:auctions}
%----------------------------------------------------------------------

\subsection{Auction Structure}

A sealed-bid auction on \Zoo{} operates as follows:

\begin{enumerate}
    \item \textbf{Commitment Phase.} Bidders encrypt their bids under the collective \FHE{} key and submit them on-chain. The auction contract records encrypted bids but cannot read them.
    \item \textbf{Evaluation Phase.} After the bidding deadline, the auction contract evaluates encrypted bids homomorphically to determine the winner.
    \item \textbf{Reveal Phase.} Only the winning bid (and, in second-price auctions, the second-highest bid) is decrypted via \TChain{} threshold protocol. Losing bids are never decrypted.
\end{enumerate}

\subsection{TFHE Winner Determination}

For a first-price auction with $n$ encrypted bids $\hat{b}_1, \ldots, \hat{b}_n$:

\begin{algorithm}
\caption{Encrypted First-Price Auction}
\begin{algorithmic}[1]
\REQUIRE Encrypted bids $\hat{b}_1, \ldots, \hat{b}_n$
\STATE $\hat{w} \leftarrow \hat{b}_1$ \COMMENT{current winner}
\STATE $\hat{j} \leftarrow \text{TFHE.Encrypt}(pk, 1)$ \COMMENT{winner index}
\FOR{$i = 2$ to $n$}
    \STATE $\hat{r} \leftarrow \text{TFHE.GT}(\hat{b}_i, \hat{w})$
    \STATE $\hat{w} \leftarrow \text{TFHE.MUX}(\hat{r}, \hat{b}_i, \hat{w})$
    \STATE $\hat{j} \leftarrow \text{TFHE.MUX}(\hat{r}, \text{TFHE.Encrypt}(pk, i), \hat{j})$
\ENDFOR
\STATE Submit $(\hat{w}, \hat{j})$ to \TChain{} for threshold decryption
\RETURN Winner index $j$, winning bid $w$
\end{algorithmic}
\end{algorithm}

\subsection{Second-Price (Vickrey) Auction}

In a Vickrey auction, the winner pays the second-highest bid. The circuit extends the above to track both the highest and second-highest encrypted values:

\begin{equation}
(\hat{w}_1, \hat{w}_2) = \text{TFHE.TopTwo}(\hat{b}_1, \ldots, \hat{b}_n)
\end{equation}

The winner is determined by $\hat{w}_1$; the settlement price is $\hat{w}_2$. Only $\hat{w}_2$ is decrypted for settlement. The actual winning bid amount is never revealed---a property impossible in traditional auctions.

\subsection{Auction Applications}

\begin{itemize}
    \item \textbf{NFT Auctions.} Sealed-bid NFT sales without sniping or bid-shading.
    \item \textbf{Block Space Auctions.} Validators auction inclusion rights; builders submit encrypted bids for block space.
    \item \textbf{Token Launch Auctions.} Fair price discovery for new tokens without front-running.
    \item \textbf{Liquidation Auctions.} DeFi protocol liquidations without predatory bidding.
\end{itemize}

%----------------------------------------------------------------------
