\section{Comparison to Existing Solutions}

\label{sec:comparison}
%----------------------------------------------------------------------

\subsection{Flashbots Protect}

Flashbots Protect routes transactions through a private relay, hiding them from the public mempool. The relay operator (Flashbots) observes all transactions. Users must trust the relay not to extract \MEV{} itself. \Zoo{} Confidential DeFi requires no trusted relay; orders are encrypted under a collective key that no single party holds.

\subsection{MEV-Share}

MEV-Share allows users to share partial information (``hints'') about their transactions with searchers, receiving a portion of extracted \MEV{} in return. This reduces but does not eliminate \MEV{}---users still lose value, just less of it. \Zoo{} eliminates \MEV{} entirely: there is no information to share because there is no information to extract.

\subsection{Penumbra}

Penumbra implements a shielded pool (Multi-Asset Shielded Pool, MASP) for private transfers and swaps. Swaps use a batch auction mechanism where all swaps in a block execute at the same clearing price. However, Penumbra does not support limit orders, on-chain order books, or sealed-bid auctions. \Zoo{} provides full order book functionality with encrypted limit orders.

\subsection{Renegade}

Renegade implements a dark pool using multiparty computation (MPC). Two-party matching is performed via garbled circuits. Limitations: (a) matching is pairwise (two parties at a time), not batch; (b) both parties must be online simultaneously; (c) the MPC coordinator observes metadata. \Zoo{}'s \FHE{}-based approach handles arbitrary batch sizes, operates asynchronously, and reveals no metadata to validators.

\subsection{Summary}

\begin{table}[h]
\centering
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Property} & \textbf{Flashbots} & \textbf{MEV-Share} & \textbf{Penumbra} & \textbf{Renegade} & \textbf{\Zoo{}} \\
\midrule
Full orderflow privacy & No & No & Yes & Yes & \textbf{Yes} \\
On-chain matching & No & No & Batch only & No & \textbf{Yes} \\
Limit orders & N/A & N/A & No & Yes & \textbf{Yes} \\
Sealed auctions & No & No & No & No & \textbf{Yes} \\
No trusted party & No & No & Yes & Partial & \textbf{Yes} \\
Asynchronous & Yes & Yes & Yes & No & \textbf{Yes} \\
\bottomrule
\end{tabular}
\caption{Feature comparison across \MEV{} mitigation approaches}
\label{tab:feature_comparison}
\end{table}

%----------------------------------------------------------------------
