\section{Securities, Holographic DAO, and Quantum-Secure Settlement}
\label{sec:securities-and-dao}

This section extends the per-\LLM{} chain framework into four
dimensions that the activation roadmap (\S\ref{sec:activation}) does
not separately specify: digital-securities positioning, the holographic
consensus \DAO{}, the quantum-secure settlement floor, and the
democratized-access mission. The full-length treatment is in the
companion paper \emph{Zoo 2025: Digital Securities, Quantum-Secure
Settlement, and the Holographic Consensus DAO}~\cite{zoo-2025-securities-and-dao}.
What follows is the per-\LLM{}-chain-specific summary.

\subsection{\$113 trillion digital-securities market positioning}
\label{sec:113t}

The total addressable market for digital securities globally is
estimated at approximately \$113 trillion across listed equities,
fixed income, derivatives, and tokenized real-world assets. This is a
market-size estimate, not a \Zoo{}-specific TVL claim. \Zoo{} sits in
this stack as the open-AI / wildlife / metaverse layer; \Lux{}
provides the regulated chain infrastructure (per \textsc{lp-020}); and
a regulated securities partner (\Zoo{}'s exclusive
partner for retail-scale securities trading) carries the regulated ATS / broker-dealer / transfer-agent
perimeter. Per-\LLM{} chains do not custody securities, but inference
revenue and research-attribution flows from per-\LLM{} chains to the
regulated-securities-partner-operated trading stack are an in-protocol use of the
chain receipts.

\paragraph{Concrete capabilities.}
Through the \Zoo{} + \Lux{} + regulated-securities-partner stack, retail users get
zero-commission trading, fractional shares to the cent, instant
settlement on \Quasar{} 3.0, tokenized stocks/bonds/commodities/REITs
with the official transfer-agent record at the regulated securities partner, 24/7 markets,
and a public verifiable order book. Tagline: \emph{Robinhood-style
capabilities for everyone --- actually make money like the 1\%
finally.}

\subsection{Quantum-secure settlement}
\label{sec:pq-settlement}

Every per-\LLM{} chain block, every research-artifact commit, and
every receipts-tree update settles under \Quasar{} 3.0
(\Lux{} \textsc{lp-020}). \Quasar{} 3.0 composes three independent
hardness assumptions:

\newcommand{\BLS}{\textsc{bls}}
\newcommand{\MLDSA}{\textsc{ml-dsa}}

\begin{itemize}[nosep,leftmargin=1.2em]
\item $\BLS{}_{12\text{-}381}$ aggregate (classical), per
   \Lux{} \textsc{lp-075}.
\item Corona Ring-LWE threshold (\textsc{pq} lattice), per
   \Lux{} \textsc{lp-073}.
\item $\MLDSA{}\text{-}65$ identity proof via Z-Chain Groth16 rollup
   (\textsc{pq} lattice with succinct proof), per \Lux{} \textsc{lp-070}.
\end{itemize}

An adversary must break all three to forge a per-\LLM{} chain block.
A \BLS{} compromise alone (Q-day) does not break finality: the
Corona + $\MLDSA{}$ pair remains secure. Custody on M-Chain
(\Lux{} \textsc{lp-019}) and FHE substrate on F-Chain
(\Lux{} \textsc{lp-013}) inherit the same composition. The retail
user can therefore say: \emph{trade post-quantum-safe, custody
post-quantum-safe, audit-trail post-quantum-safe.}

\subsection{Homomorphic / holographic consensus DAO}
\label{sec:dao-perllm}

Each per-\LLM{} chain operates a sub-\DAO{} that inherits the
meta-\DAO{} governance topology: weighted voting under FHE tally on
F-Chain (\Lux{} \textsc{lp-013}), threshold key custody on M-Chain
(\Lux{} \textsc{lp-019}), and contributor attestation on A-Chain
(\Lux{} \textsc{lp-134}). The \DAO{} is described as
\emph{homomorphic} because votes are tallied while encrypted
(individual ballots stay private; only the aggregate decrypts), and
\emph{holographic} because every sub-\DAO{} is structurally a small
copy of the meta-\DAO{}. A per-\LLM{} chain can fork into its own
sub-chain and keep the same governance machinery without bespoke
integration.

\paragraph{Weighted voting.}
Voting weight per participant is

\[
\text{weight}(p) \;=\; \alpha \cdot a_p \;+\; \beta \cdot i_p \;+\; \gamma \cdot c_p \;+\; \delta \cdot s_p
\]

\noindent with $\alpha + \beta + \gamma + \delta = 1$. The four axes
are advocacy ($a_p$, attestation-rooted social signal), involvement
($i_p$, contribution-graph activity attested on A-Chain), contribution
($c_p$, landmark research artifacts with reproducibility attestation),
and token stake ($s_p$, locked $\$$ZOO). Default initial weights at
2025-12-25 activation:
$(\alpha, \beta, \gamma, \delta) = (0.20,\ 0.30,\ 0.40,\ 0.10)$, set
by meta-\DAO{} vote. The bound on $\delta$ prevents plutocracy; the
A-Chain attestation gate prevents Sybil inflation of $i_p$ and $c_p$.

\subsection{Democratized end-to-end access to information}
\label{sec:democratized}

The per-\LLM{} chain is the mechanism that makes ``democratized
end-to-end'' a protocol invariant rather than a marketing claim:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{User} --- selective disclosure via Z-Chain Groth16
   attestations: prove eligibility without revealing identity.
\item \textbf{Model} --- weights, architecture, and training-data
   lineage all on the per-\LLM{} chain.
\item \textbf{Data} --- every training shard's hash is committed to
   the chain's receipts tree.
\item \textbf{Output} --- every inference call produces a
   \textsc{tee}-attested receipt anchored on A-Chain via Hanzo AI
   Chain~\cite{hanzo-ai-chain}.
\item \textbf{Reward} --- royalties flow back along the on-chain
   contribution graph automatically; no central treasurer.
\end{itemize}

Closed AI labs (OpenAI, Anthropic) treat weights, data, and
architecture as competitive secrets. \Zoo{} per-\LLM{} chains treat
them as public goods whose provenance is the protocol. This is the
durable distinction.

\subsection{Cross-references to companion paper and ZIPs}

The full-length specification of these four planks --- with formal
definitions, receipts-tree layout, and the contribution-graph
attestation pipeline --- is in~\cite{zoo-2025-securities-and-dao}.
Relevant Zoo \textsc{zip}s: \textsc{zip-0017} (DAO Governance
Framework), \textsc{zip-0026} (Ecosystem Reputation System),
\textsc{zip-0407} (Decentralized AI Training Architecture),
\textsc{zip-0413} (Foundation Language Model Architecture),
\textsc{zip-0606} (Reproducibility Attestation), and \textsc{zip-0570}
(Zoo Labs Impact Thesis).
