\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{caption}
\usepackage{listings}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={Zoo 3.0 - Full Post-Quantum Stack},
    pdfauthor={Antje Worring, Zach Kelling},
    pdfsubject={Zoo Labs Foundation - Full Post-Quantum Stack 2025},
    pdfkeywords={Zoo 3.0, post-quantum, Ringtail, ML-DSA-65, BLS12-381, triple cert, Quasar 3.0}
}

% Macros (match Zoo 2.0 / 4.0 style)
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Zen}{\textsc{Zen}}
\newcommand{\Quasar}{\textsc{Quasar}}
\newcommand{\Liquidity}{\textsc{Liquidity}}
\newcommand{\PoAI}{\textsc{PoAI}}
\newcommand{\TEE}{\textsc{tee}}
\newcommand{\LLM}{\textsc{llm}}
\newcommand{\GPU}{\textsc{gpu}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\AMM}{\textsc{amm}}
\newcommand{\DEX}{\textsc{dex}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\FHE}{\textsc{fhe}}
\newcommand{\ZKP}{\textsc{zkp}}
\newcommand{\BLS}{\textsc{bls}}
\newcommand{\MLDSA}{\textsc{ml-dsa}}
\newcommand{\Ringtail}{\textsc{Ringtail}}
\newcommand{\PQ}{\textsc{pq}}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray},
    tabsize=2,
    captionpos=b,
}

\title{\textbf{Zoo 3.0 --- Full Post-Quantum Stack}\\
\large Every signature, every commitment, every proof in PQ-safe schemes}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{October 31, 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
\noindent \Zoo{} 3.0 is the full post-quantum era of the \Zoo{} stack.
\Zoo{} 2.0 (October 2024) shipped a post-quantum L2 \textsc{evm} on
\Lux{} with a precompile suite covering NIST PQC standards
\cite{zoo-2-0}; \Zoo{} 3.0 (2025) hardens \emph{every} hot-path
primitive across the \Zoo{} ecosystem to PQ-safe schemes.
Specifically: every consensus signature is a triple cert
($\BLS{}_{12\text{-}381}$ classical fast path, \Ringtail{} Ring-LWE
threshold proof, $\MLDSA{}\text{-}65$ identity proof) inherited from
the \Lux{} \Quasar{} 3.0 lane (\textsc{lp-020}); every governance
commitment is a $\MLDSA{}\text{-}65$-signed Merkle root; every loan
agreement under the Zoo \NFT{} Liquidity Protocol is \Ringtail{}-signed
and bridge-attested; every inference receipt and training-data
attribution carries PQ-safe witnesses. The \Zoo{} Bridge committee
re-keys to threshold \MLDSA{} + \Ringtail{} on M-Chain. The Zoo \DAO{}
migrates governance keys to threshold \MLDSA{}-65. zLLM commitments
for training artefacts, governance votes, and inference receipts
adopt PQ-safe Merkle accumulators. This paper documents the full
PQ stack, the migration roadmap, and the cross-references to \Lux{}
LP-020 (Quasar 3.0), LP-070 (\MLDSA{}), LP-073 (\Ringtail{}), and
LP-075 (\BLS{}). Spec freeze: 2025-10-31. Phased rollout through
2025; full activation aligned with \Lux{} \Quasar{} 3.0 cutover.

\textbf{Keywords}: \Zoo{} 3.0, post-quantum, \Ringtail{}, $\MLDSA{}\text{-}65$,
$\BLS{}_{12\text{-}381}$, triple cert, \Quasar{} 3.0, zLLM, NFT Liquidity
Protocol
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-context}
\input{sections/02-triple-cert}
\input{sections/03-zllm-pq}
\input{sections/04-bridge-pq}
\input{sections/05-nft-liquidity-pq}
\input{sections/06-dao-migration}
\input{sections/07-roadmap}

\begin{thebibliography}{99}

\bibitem{lp-009}
Lux Foundation.
\newblock LP-009: \GPU{}-Native \textsc{evm}.
\newblock Lux Improvement Proposal, 2025.
\newblock \url{https://github.com/luxfi/lps/lp-009}.

\bibitem{lp-016}
Lux Foundation.
\newblock LP-016: B-Chain Bridge VM and Cross-Chain Attestation.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-017}
Lux Foundation.
\newblock LP-017: Threshold Bridge Committee Key Schedule.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-020}
Lux Foundation.
\newblock LP-020: \Quasar{} Consensus 3.0 (Triple Consensus: \BLS{} + \Ringtail{} + \MLDSA{}).
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-070}
Lux Foundation.
\newblock LP-070: $\MLDSA{}\text{-}65$ Identity Proof Lane.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-073}
Lux Foundation.
\newblock LP-073: \Ringtail{} Ring-LWE Threshold Signature.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-075}
Lux Foundation.
\newblock LP-075: $\BLS{}_{12\text{-}381}$ Aggregate Signature Lane.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-105}
Lux Foundation.
\newblock LP-105: \Quasar{} Consensus Family Specification (Photon, Wave, Focus, Nova, Nebula).
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-134}
Lux Foundation.
\newblock LP-134: \Lux{} Chain Topology (P/C/X/Q/Z/A/B/M/F-Chains).
\newblock Lux Improvement Proposal, 2025.

\bibitem{zoo-2-0}
Zoo Labs Foundation.
\newblock Zoo \textsc{evm}: A Post-Quantum Safe L2 Blockchain with AI-Native Precompiles on the \Lux{} Network.
\newblock Zoo Whitepaper, October 2024.

\bibitem{zoo-4-0}
Zoo Labs Foundation.
\newblock Zoo 4.0 --- \GPU{}-Native Sovereign L1.
\newblock Zoo Whitepaper, February 14, 2026.

\bibitem{zoo-2021}
Antje Worring et al.
\newblock Zoo: A Decentralized Wildlife Conservation Platform.
\newblock Zoo Whitepaper (Original), October 31, 2021.

\bibitem{zoo-2025-securities-and-dao}
Zoo Labs Foundation.
\newblock Zoo 2025: Digital Securities, Quantum-Secure Settlement, and the Holographic Consensus \DAO{}.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{nist-fips-204}
NIST.
\newblock FIPS 204: Module-Lattice-Based Digital Signature Standard ($\MLDSA{}$).
\newblock NIST Federal Information Processing Standards, August 2024.

\bibitem{ringtail-paper}
Lux Foundation.
\newblock The \Ringtail{} Ring-LWE Threshold Signature Scheme: Construction and Analysis.
\newblock Lux Cryptography Note, 2025.

\end{thebibliography}

\end{document}
