\section{Conclusion}


Zoo EVM provides the first post-quantum safe, AI-native Layer 2 blockchain for decentralized science and AI research. By combining NIST PQC standards, threshold signatures, privacy-preserving computation, and proof-of-useful-computation mining, Zoo creates infrastructure where scientific progress and economic incentives are cryptographically aligned. The 20-precompile suite, Kubernetes-native operator, and three-network deployment provide a production-ready foundation for the next generation of decentralized AI.

Building on Zoo Labs Foundation's 2021 origins as a pioneer in agentic AI---years before the term became mainstream---Zoo EVM extends the principle that AI infrastructure should be community-owned, scientifically rigorous, and quantum-safe from day one.

\begin{thebibliography}{99}
\bibitem{lux-consensus} Kelling, Z. ``Lux Consensus: Multi-Engine Blockchain Consensus.'' Lux Industries, 2025.
\bibitem{lux-pq} Kelling, Z. ``EthFalcon: Post-Quantum Cryptography for EVM Chains.'' Lux Industries, 2025.
\bibitem{lux-precompiles} Kelling, Z. ``Lux EVM Precompiles: Native Cryptographic Primitives.'' Lux Industries, 2025.
\bibitem{lux-evm} Kelling, Z. ``Regulated EVM L1: On-Chain Securities Settlement.'' Lux Exchange, 2026.
\bibitem{zoo-network} Kelling, Z. ``Zoo Network: Decentralized AI Training and Inference Layer.'' Zoo Labs Foundation, 2025.
\bibitem{zoo-experience} Kelling, Z. ``The Experience Ledger: Content-Addressable Semantic Optimization Knowledge.'' Zoo Labs Foundation, 2025.
\bibitem{cggmp21} Canetti, R., et al. ``UC Non-Interactive, Proactive, Threshold ECDSA.'' ACM CCS 2020.
\bibitem{frost} Komlo, C., Goldberg, I. ``FROST: Flexible Round-Optimized Schnorr Threshold Signatures.'' SAC 2020.
\bibitem{fips204} NIST. ``FIPS 204: Module-Lattice-Based Digital Signature Standard.'' 2024.
\bibitem{fips205} NIST. ``FIPS 205: Stateless Hash-Based Digital Signature Standard.'' 2024.
\bibitem{fips203} NIST. ``FIPS 203: Module-Lattice-Based Key-Encapsulation Mechanism Standard.'' 2024.
\end{thebibliography}

