\section{Conclusion}

\label{sec:conclusion}
% ══════════════════════════════════════════════════════════════════════════════

Biodiversity Impact Bonds represent a structural innovation in
conservation finance that addresses the measurement, liquidity, and
transaction cost barriers preventing private capital mobilization. By
encoding conservation outcomes in programmable smart contracts and
verifying them through decentralized oracle networks, BIBs create a
transparent, efficient, and scalable financing mechanism. The 62\%
reduction in transaction costs means more capital reaches conservation
activities. Through Antje Worring, Zach Kelling's governance framework, BIBs
can evolve as a community-governed standard for conservation finance.


% ══════════════════════════════════════════════════════════════════════════════
% References
% ══════════════════════════════════════════════════════════════════════════════

\bibliographystyle{plainnat}

\begin{thebibliography}{26}

\bibitem[Blue Forest Conservation(2019)]{blue2019forest}
Blue Forest Conservation.
\newblock Forest resilience bond: Fighting fire with finance.
\newblock Technical report, 2019.

\bibitem[Buterin(2014)]{buterin2014schellingcoin}
Buterin, V.
\newblock SchellingCoin: A minimal-trust universal data feed.
\newblock Ethereum Blog, 2014.

\bibitem[CBD(2022)]{cbd2022kunming}
Convention on Biological Diversity.
\newblock Kunming-Montreal Global Biodiversity Framework.
\newblock CBD/COP/15/L.25, 2022.

\bibitem[Dalton et~al.(2020)]{dalton2020blockchain}
Dalton, S., May, G., McSherry, J., et~al.
\newblock Blockchain and biodiversity: A review of potential applications.
\newblock \emph{Conservation Biology}, 34(5):1013--1024, 2020.

\bibitem[Dasgupta(2021)]{dasgupta2021economics}
Dasgupta, P.
\newblock \emph{The Economics of Biodiversity: The Dasgupta Review}.
\newblock HM Treasury, London, 2021.

\bibitem[Deutz et~al.(2020)]{deutz2020financing}
Deutz, A., Heal, G.M., Niu, R., et~al.
\newblock Financing nature: Closing the global biodiversity financing gap.
\newblock The Paulson Institute, 2020.

\bibitem[Dinerstein et~al.(2017)]{dinerstein2017ecoregion}
Dinerstein, E., Olson, D., Joshi, A., et~al.
\newblock An ecoregion-based approach to protecting half the terrestrial realm.
\newblock \emph{BioScience}, 67(6):534--545, 2017.

\bibitem[Forest Trends(2021)]{foresttrends2021}
Forest Trends' Ecosystem Marketplace.
\newblock State of the Voluntary Carbon Markets 2021.
\newblock Forest Trends, 2021.

\bibitem[Howson(2019)]{howson2019crypto}
Howson, P.
\newblock Tackling climate change with blockchain.
\newblock \emph{Nature Climate Change}, 9(9):644--645, 2019.

\bibitem[Huwyler et~al.(2014)]{huwyler2014making}
Huwyler, F., K{\"a}ppeli, J., Serafimova, K., Swanson, E., and Tobin, J.
\newblock Making conservation finance investable.
\newblock \emph{Stanford Social Innovation Review}, 12(1):28--35, 2014.

\bibitem[Liebman(2011)]{liebman2011social}
Liebman, J.B.
\newblock Social impact bonds: A promising new financing model.
\newblock \emph{Center for American Progress}, 2011.

\bibitem[Mace et~al.(2018)]{mace2018aiming}
Mace, G.M., Barrett, M., Burgess, N.D., et~al.
\newblock Aiming higher to bend the curve of biodiversity loss.
\newblock \emph{Nature Sustainability}, 1(9):448--451, 2018.

\bibitem[Nicola(2013)]{nicola2013quantifying}
Nicola, D.J.
\newblock Quantifying the benefits of green infrastructure.
\newblock \emph{Environmental Finance}, 14(5):32--39, 2013.

\bibitem[OECD(2020)]{oecd2020comprehensive}
OECD.
\newblock A Comprehensive Overview of Global Biodiversity Finance.
\newblock OECD Publishing, Paris, 2020.

\bibitem[Paulson et~al.(2020)]{paulson2020financing}
Paulson, H., Steinberg, M., and Bloomberg, M.
\newblock Financing nature: Closing the global biodiversity financing gap.
\newblock The Paulson Institute, 2020.

\bibitem[Salzman et~al.(2018)]{salzman2018new}
Salzman, J., Bennett, G., Carroll, N., Goldstein, A., and Jenkins, M.
\newblock The global status and trends of payments for ecosystem services.
\newblock \emph{Nature Sustainability}, 1(3):136--144, 2018.

\bibitem[Schletz et~al.(2020)]{schletz2020blockchain}
Schletz, M., Franke, L.A., and Salomo, S.
\newblock Blockchain application for Paris Agreement carbon market mechanism.
\newblock \emph{Sustainability}, 12(12):5069, 2020.

\bibitem[SEC(1946)]{sec1946howey}
Securities and Exchange Commission v.\ W.J.\ Howey Co.
\newblock 328 U.S. 293, 1946.

\bibitem[Sumaila et~al.(2017)]{sumaila2017investments}
Sumaila, U.R., Rodriguez, C.M., Schultz, M., et~al.
\newblock Investments to reverse biodiversity loss are economically beneficial.
\newblock \emph{Current Opinion in Environmental Sustainability}, 29:82--88, 2017.

\bibitem[TNFD(2023)]{tnfd2023framework}
Taskforce on Nature-related Financial Disclosures.
\newblock Recommendations of the TNFD.
\newblock TNFD, 2023.

\bibitem[UCT(2020)]{uct2020rhino}
United for Conservation Trust.
\newblock Rhino Impact Bond: Innovative finance for conservation.
\newblock Technical report, 2020.

\bibitem[World Bank(2021)]{worldbank2021mobilizing}
World Bank.
\newblock Mobilizing private finance for nature.
\newblock World Bank Group, 2021.

\bibitem[Zu~Ermgassen et~al.(2022)]{zu2022biodiversity}
Zu~Ermgassen, S.O.S.E., Maron, M., Walker, C.M.C., et~al.
\newblock The role of ``no net loss'' policies in conserving biodiversity.
\newblock \emph{One Earth}, 5(2):122--128, 2022.

\bibitem[Bos et~al.(2023)]{bos2023biodiversity}
Bos, A.B.,"; H., and Ellis, E.C.
\newblock Biodiversity credits: emerging markets and methodologies.
\newblock \emph{Conservation Letters}, 16(3):e12943, 2023.

\bibitem[Kettunen et~al.(2017)]{kettunen2017integration}
Kettunen, M., Vihervaara, P., Kinnunen, S., et~al.
\newblock Socio-economic importance of ecosystem services in the Nordic countries.
\newblock Nordic Council of Ministers, 2017.

\end{thebibliography}

