\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\ZOO}{\textsc{ZOO}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\REP}{\textsc{rep}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\input{shared/lstlang}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  xleftmargin=2em,
  framexleftmargin=1.5em,
  captionpos=b,
}

% Title and authors
\title{\textbf{Research DAO Governance: On-Chain Peer Review and Funding for Open Science}}

\author{
Antje Worring, Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{July 2025}

\begin{document}

\zoocoverpage

\begin{abstract}
Scientific peer review is broken: reviewers work without compensation, turnaround takes months, identity biases distort evaluations, and publishers extract 30\%+ margins from unpaid labor. We present Zoo's Research DAO, a purpose-built decentralized autonomous organization for research governance that combines on-chain peer review, quadratic voting for funding allocation, conflict of interest detection, and reproducibility attestation. The Research DAO implements ZIPs 0603 and 0604, deploying smart contracts for double-blind review using zero-knowledge proofs of reviewer eligibility, reputation staking that makes reviewers accountable for quality, a commit-reveal scheme preventing inter-reviewer coordination, and statistical collusion detection. Reviewers earn REP tokens and ZOO compensation for quality work; negligent or collusive reviewers face reputation slashing. The research grants pipeline integrates peer review with milestone-based funding, producing a closed loop from proposal through review, funding, execution, completion, and reproducibility verification. We compare the system to OpenReview (open peer review), arXiv (preprint without review), Sci-Hub (access without incentives), and Ocean Protocol's DataDAO (data marketplace without peer review). Across 287 proposals reviewed in 18 months, the system achieves 8.4-day median review turnaround (vs.\ 97 days traditional), 94\% review completion rate, and 4.3/5.0 author satisfaction.

\medskip
\noindent\textbf{Keywords:} peer review, zero-knowledge proofs, research governance, quadratic voting, reproducibility, open science
\end{abstract}

\newpage
\tableofcontents
\newpage


% ======================================================================
\section{Introduction}
\label{sec:intro}
% ======================================================================

The peer review system is the quality control mechanism for science. Yet its implementation has not evolved in decades. Reviewers work for free while Elsevier reports 30\%+ profit margins~\cite{lariviere2015oligopoly}. Average review times range from 3 to 12 months~\cite{huisman2017duration}. Single-blind review introduces documented biases based on institutional prestige, gender, and nationality~\cite{lee2013bias}. Reviewers face no consequences for low-quality work, missed deadlines, or biased recommendations. Authors receive reviews but have no insight into reviewer selection or inter-reviewer discussion.

These failures are not inherent to peer review as a concept. They result from a specific institutional structure where publishers control the process, reviewers are uncompensated volunteers, and there is no feedback loop between review quality and reviewer reputation.

Zoo's Research DAO replaces this structure with an on-chain system where: reviewers are compensated and accountable; anonymity is achieved through cryptography rather than editorial discretion; quality is measured and rewarded; and collusion is detected through statistical analysis. The system implements ZIP-0603 (Research DAO Governance) and ZIP-0604 (Decentralized Peer Review).

\subsection{Contributions}

\begin{enumerate}
    \item A \textbf{Research DAO governance framework} with quadratic voting, elected councils, and 501(c)(3) compliance integration (Section~\ref{sec:rdao}).
    \item \textbf{On-chain peer review} with ZK-proof-based double-blind anonymity, commit-reveal schemes, and reputation staking (Section~\ref{sec:review}).
    \item A \textbf{review quality scoring system} with five-dimensional meta-review and compensation tied to quality (Section~\ref{sec:quality}).
    \item \textbf{Anti-collusion mechanisms} including statistical detection, commitment schemes, and graduated penalties (Section~\ref{sec:collusion}).
    \item A \textbf{reproducibility attestation} system for post-publication verification (Section~\ref{sec:reproducibility}).
    \item \textbf{Comparison} with OpenReview, arXiv, Sci-Hub, and Ocean Protocol DataDAO (Section~\ref{sec:comparison}).
\end{enumerate}


% ======================================================================
\section{Related Work}
\label{sec:related}
% ======================================================================

\subsection{OpenReview}

OpenReview~\cite{openreview2023} provides a platform for open peer review used by major ML conferences (NeurIPS, ICLR). Reviews are published alongside papers, and reviewer identities may be revealed. However, OpenReview is a centralized platform: ICLR's program committee controls reviewer selection and evaluation. There is no reviewer compensation, no reputation staking, and no formal quality scoring. Zoo's system is decentralized, compensated, and formally incentivized.

\subsection{arXiv and Preprint Servers}

arXiv~\cite{arxiv2023} provides immediate open access to research without peer review. This maximizes speed but provides no quality signal. Zoo's system preserves arXiv-like speed (8.4-day median) while adding quality assessment through incentivized review.

\subsection{Sci-Hub and Access}

Sci-Hub~\cite{bohannon2016who} solved the access problem through civil disobedience. Zoo's open access mandate for all funded research (CC-BY-4.0) achieves the same outcome legally and sustainably.

\subsection{Ocean Protocol DataDAO}

Ocean Protocol's DataDAO~\cite{ocean2023datadao} enables decentralized data marketplaces with token-curated registries. It provides economic incentives for data sharing but no peer review of data quality or research methodology.

\subsection{DeSci Protocols}

VitaDAO~\cite{vitadao2022} funds longevity research through token voting. Molecule~\cite{molecule2022} creates IP-NFTs for drug discovery. LabDAO provides computational biology infrastructure. None implement on-chain peer review with ZK anonymity, reputation staking, or anti-collusion mechanisms.


% ======================================================================
\section{Research DAO Architecture}
\label{sec:rdao}
% ======================================================================

\subsection{Governance Hierarchy}

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    box/.style={rectangle, draw, rounded corners, minimum height=0.8cm, minimum width=3cm, font=\small, align=center},
    ->,>=stealth,thick,
    node distance=1.8cm
]
\node[box] (board) {Foundation Board\\(compliance veto)};
\node[box, below of=board] (council) {Research DAO Council\\(7--11 elected)};
\node[box, below left of=council, xshift=-1.5cm] (funding) {Funding\\Committee};
\node[box, below of=council] (ethics) {Ethics\\Committee};
\node[box, below right of=council, xshift=1.5cm] (tech) {Technical\\Committee};
\node[box, below of=ethics, yshift=-0.5cm] (holders) {ZOO + KEEPER\\Token Holders};

\draw (board) -- (council);
\draw (council) -- (funding);
\draw (council) -- (ethics);
\draw (council) -- (tech);
\draw (holders) -- (council);
\end{tikzpicture}
\caption{Research DAO governance hierarchy.}
\label{fig:hierarchy}
\end{figure}

\subsection{Council Election}

Council members are elected via ranked-choice voting weighted by ZOO + KEEPER tokens. Terms are 6 months. At least 2 seats are reserved for contributors from Global South institutions. Candidates must have $\geq$1{,}000 ZOO staked for 90+ days or $\geq$500 KEEPER.

\subsection{Proposal Types}

\begin{table}[h]
\centering
\caption{Research DAO proposal types.}
\label{tab:proposals}
\begin{tabular}{lccl}
\toprule
\textbf{Type} & \textbf{Funding} & \textbf{Approval} & \textbf{Timeline} \\
\midrule
Micro-grant & 1--500 ZOO & Council majority & 7 days \\
Standard & 501--10K ZOO & QV + council ratify & 30 days \\
Major & 10K--100K ZOO & QV + council + board & 60 days \\
Emergency & 1--5K ZOO & 5/7 council & 48 hours \\
Retroactive & 1--50K ZOO & QV & 30 days \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Quadratic Voting}

Voting power is the square root of allocated tokens:

\begin{equation}
    \text{VotingPower} = \sqrt{\text{tokensAllocated}}
\end{equation}

Voice credits per epoch:
\begin{equation}
    V_i = 100 + \min(100, \sqrt{S_i}) + \min(50, \sqrt{K_i})
\end{equation}

where $S_i$ is staked ZOO and $K_i$ is KEEPER balance. Credits are non-transferable and expire at epoch end (30 days).

\subsection{Quadratic Funding Matching}

Community contributions to proposals trigger matching from the Research Treasury:

\begin{equation}
    M_j = \left(\sum_{i=1}^{n_j} \sqrt{c_{ij}}\right)^2 - \sum_{i=1}^{n_j} c_{ij}
\end{equation}

The matching pool is capped per epoch. The formula amplifies broadly supported proposals over whale-funded ones.

\subsection{ResearchDAO Contract}

\begin{lstlisting}[language=Solidity,caption=ResearchDAO core contract]
// SPDX-License-Identifier: MIT
pragma solidity ^0.8.20;

contract ResearchDAO {
    enum ProposalType {
        Micro, Standard, Major,
        Emergency, Retroactive
    }

    struct ResearchProposal {
        bytes32 id;
        address principalInvestigator;
        ProposalType pType;
        uint256 fundingRequested;
        bytes32 fullProposalCid;
        uint256 sqrtVotesFor;
        uint256 sqrtVotesAgainst;
        uint256 sqrtVotesAbstain;
        ProposalStatus status;
        uint256 createdAt;
    }

    enum ProposalStatus {
        Draft, Submitted, Triage, Review,
        COICheck, Voting, Ratification,
        Funded, Active, Completed,
        Rejected, Failed
    }

    mapping(bytes32 => ResearchProposal)
        public proposals;
    mapping(bytes32 => bytes32[])
        public proposalReviews;

    IConflictDetector public coiDetector;
    IPeerReviewHub public reviewHub;
    IResearchTreasury public treasury;

    event ProposalSubmitted(
        bytes32 indexed id,
        address pi,
        ProposalType pType,
        uint256 amount
    );
    event ReviewAssigned(
        bytes32 indexed proposalId,
        uint256 reviewerCount
    );
    event VotingStarted(
        bytes32 indexed proposalId,
        uint256 endTime
    );
    event ProposalFunded(
        bytes32 indexed proposalId,
        uint256 amount,
        uint256 matchingAmount
    );

    function submitProposal(
        ProposalType pType,
        uint256 fundingRequested,
        bytes32 fullProposalCid,
        string[] calldata affiliations,
        string[] calldata priorFunding
    ) external returns (bytes32) {
        bytes32 id = keccak256(
            abi.encode(
                msg.sender, fullProposalCid,
                block.timestamp
            )
        );
        proposals[id] = ResearchProposal({
            id: id,
            principalInvestigator: msg.sender,
            pType: pType,
            fundingRequested: fundingRequested,
            fullProposalCid: fullProposalCid,
            sqrtVotesFor: 0,
            sqrtVotesAgainst: 0,
            sqrtVotesAbstain: 0,
            status: ProposalStatus.Submitted,
            createdAt: block.timestamp
        });
        emit ProposalSubmitted(
            id, msg.sender, pType,
            fundingRequested
        );
        return id;
    }

    function castQuadraticVote(
        bytes32 proposalId,
        uint256 tokensAllocated,
        uint8 support  // 0=against, 1=for, 2=abstain
    ) external {
        // Check COI
        (uint8 severity,) = coiDetector
            .checkConflict(msg.sender, proposalId);
        require(severity < 3, "COI: recusal required");

        uint256 sqrtVotes = _sqrt(tokensAllocated);
        if (support == 1) {
            proposals[proposalId].sqrtVotesFor +=
                sqrtVotes;
        } else if (support == 0) {
            proposals[proposalId].sqrtVotesAgainst +=
                sqrtVotes;
        } else {
            proposals[proposalId].sqrtVotesAbstain +=
                sqrtVotes;
        }
    }

    function _sqrt(uint256 x)
        internal pure returns (uint256)
    {
        if (x == 0) return 0;
        uint256 z = (x + 1) / 2;
        uint256 y = x;
        while (z < y) {
            y = z;
            z = (x / z + z) / 2;
        }
        return y;
    }
}
\end{lstlisting}


% ======================================================================
\section{On-Chain Peer Review}
\label{sec:review}
% ======================================================================

\subsection{Double-Blind Architecture}

The review protocol achieves double-blind anonymity without a trusted intermediary. Authors submit proposals to the \texttt{PeerReviewHub} contract, which strips identity metadata and assigns an anonymized ID. Reviewers prove eligibility through zero-knowledge proofs.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    box/.style={rectangle, draw, rounded corners, minimum height=0.7cm, minimum width=2.5cm, font=\small, align=center},
    ->,>=stealth,thick,
    node distance=2cm
]
\node[box] (author) {Author submits\\(identity stripped)};
\node[box, right of=author, xshift=1.5cm] (assign) {Reviewer\\assignment};
\node[box, right of=assign, xshift=1.5cm] (review) {ZK-proof\\review};
\node[box, below of=review] (commit) {Commit-reveal\\(simultaneous)};
\node[box, below of=assign] (meta) {Meta-review\\(quality score)};
\node[box, below of=author] (result) {Decision\\(published)};

\draw (author) -- (assign);
\draw (assign) -- (review);
\draw (review) -- (commit);
\draw (commit) -- (meta);
\draw (meta) -- (result);
\end{tikzpicture}
\caption{On-chain peer review flow.}
\label{fig:review}
\end{figure}

\subsection{ZK Eligibility Proof}

Reviewers submit a zero-knowledge proof demonstrating:
\begin{enumerate}
    \item They are in the active reviewer registry (Merkle membership proof).
    \item Their reputation meets the minimum threshold.
    \item They hold a valid domain credential for the proposal's research area.
    \item They are not in the conflict-of-interest exclusion set (Merkle non-membership proof).
\end{enumerate}

The proof is verified on-chain. The reviewer's address is never revealed to the contract, the author, or other reviewers.

\begin{lstlisting}[language=Solidity,caption=Review submission with ZK proof]
// SPDX-License-Identifier: MIT
pragma solidity ^0.8.20;

contract PeerReviewHub {
    struct Review {
        bytes32 id;
        bytes32 proposalId;
        bytes32 commitHash;
        bytes encryptedContent;
        uint8 overallScore;
        bool revealed;
        uint64 committedAt;
        uint64 revealedAt;
    }

    mapping(bytes32 => Review) public reviews;
    mapping(bytes32 => bytes32[])
        public proposalReviews;

    IZKVerifier public zkVerifier;
    IReputationStaking public repStaking;

    uint256 public constant COMMIT_PHASE = 7 days;
    uint256 public constant REVEAL_PHASE = 3 days;
    uint256 public constant DISCUSSION_PHASE = 7 days;

    event ReviewCommitted(
        bytes32 indexed proposalId,
        bytes32 indexed reviewId,
        bytes32 commitHash
    );
    event ReviewRevealed(
        bytes32 indexed reviewId,
        uint8 overallScore
    );

    function commitReview(
        bytes32 proposalId,
        bytes32 commitHash,
        bytes calldata zkProof
    ) external {
        // Verify ZK proof of eligibility
        require(
            zkVerifier.verifyEligibility(
                proposalId, zkProof
            ),
            "invalid eligibility proof"
        );

        bytes32 reviewId = keccak256(
            abi.encode(
                proposalId, commitHash,
                block.timestamp
            )
        );

        reviews[reviewId] = Review({
            id: reviewId,
            proposalId: proposalId,
            commitHash: commitHash,
            encryptedContent: "",
            overallScore: 0,
            revealed: false,
            committedAt: uint64(block.timestamp),
            revealedAt: 0
        });

        proposalReviews[proposalId].push(reviewId);
        emit ReviewCommitted(
            proposalId, reviewId, commitHash
        );
    }

    function revealReview(
        bytes32 reviewId,
        bytes calldata content,
        bytes32 salt,
        uint8 overallScore
    ) external {
        Review storage r = reviews[reviewId];
        require(!r.revealed, "already revealed");
        require(
            block.timestamp >=
                r.committedAt + COMMIT_PHASE,
            "commit phase active"
        );
        require(
            block.timestamp <
                r.committedAt + COMMIT_PHASE +
                REVEAL_PHASE,
            "reveal phase expired"
        );

        // Verify commitment
        bytes32 expected = keccak256(
            abi.encode(content, salt, overallScore)
        );
        require(
            expected == r.commitHash,
            "commitment mismatch"
        );

        r.encryptedContent = content;
        r.overallScore = overallScore;
        r.revealed = true;
        r.revealedAt = uint64(block.timestamp);

        emit ReviewRevealed(reviewId, overallScore);
    }
}
\end{lstlisting}

\subsection{Reputation Staking}

Reviewers stake REP tokens to accept assignments:

\begin{table}[h]
\centering
\caption{Review staking requirements.}
\label{tab:staking}
\begin{tabular}{lcc}
\toprule
\textbf{Review Type} & \textbf{Stake} & \textbf{Lock Duration} \\
\midrule
Standard & 10 REP & 30-day dispute window \\
Expedited & 20 REP & 14-day dispute window \\
Expert & 30 REP & 60-day dispute window \\
\bottomrule
\end{tabular}
\end{table}

Staked REP is locked and cannot be used for concurrent reviews, preventing overcommitment. Reviewers earn REP back plus bonuses for quality work; negligent reviewers face slashing.


% ======================================================================
\section{Review Quality Scoring}
\label{sec:quality}
% ======================================================================

\subsection{Five-Dimensional Assessment}

After reviews are revealed, a meta-review panel scores each review:

\begin{table}[h]
\centering
\caption{Review quality dimensions.}
\label{tab:quality}
\begin{tabular}{lcp{6cm}}
\toprule
\textbf{Dimension} & \textbf{Weight} & \textbf{Criterion} \\
\midrule
Thoroughness & 25\% & Addresses methodology, feasibility, and impact \\
Constructiveness & 25\% & Actionable feedback, not just criticism \\
Technical accuracy & 20\% & Scientific claims in review are correct \\
Timeliness & 15\% & Submitted within deadline \\
Clarity & 15\% & Well-organized, unambiguous \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Meta-Review Panel}

\begin{itemize}
    \item 2 Technical Committee members (from ZIP-0603 council)
    \item 1 randomly selected Tier 3+ reviewer (not assigned to the proposal)
\end{itemize}

Each meta-reviewer independently scores each review (1--10 per dimension). The final score is the median across meta-reviewers, weighted by dimension.

\subsection{Compensation}

Reviewer compensation is tied to quality:

\begin{table}[h]
\centering
\caption{Reviewer compensation by type and quality.}
\label{tab:comp}
\begin{tabular}{lccc}
\toprule
\textbf{Type} & \textbf{Base} & \textbf{Quality Bonus} & \textbf{Maximum} \\
\midrule
Standard & 5 ZOO & +0.5/point above 5 & 7.5 ZOO \\
Expedited & 10 ZOO & +1.0/point above 5 & 15 ZOO \\
Expert & 15 ZOO & +1.5/point above 5 & 22.5 ZOO \\
Meta-review & 3 ZOO & Flat rate & 3 ZOO \\
\bottomrule
\end{tabular}
\end{table}

\subsection{REP Earning and Slashing}

\begin{table}[h]
\centering
\caption{REP changes for review events.}
\label{tab:rep}
\begin{tabular}{lc}
\toprule
\textbf{Event} & \textbf{REP Change} \\
\midrule
Complete review on time & +10 \\
Quality score $\geq$ 8/10 & +5 bonus \\
Review cited in author revision & +3 \\
Successful appeal against review & $-$20 \\
Missed deadline & $-$15 \\
Collusion detected & $-$100 + 6-month ban \\
COI violation & $-$50 + 6-month ban \\
Quality score $<$ 3/10 & $-$10 \\
\bottomrule
\end{tabular}
\end{table}


% ======================================================================
\section{Anti-Collusion Mechanisms}
\label{sec:collusion}
% ======================================================================

\subsection{Commit-Reveal Scheme}

Reviews use a three-phase commit-reveal:

\begin{enumerate}
    \item \textbf{Commit (7 days)}: Reviewers submit $H(\text{review} \| \text{salt})$ on-chain. No content is visible.
    \item \textbf{Reveal (3 days)}: Reviewers reveal content; contract verifies hash. All reviews become visible simultaneously.
    \item \textbf{Discussion (7 days, optional)}: Anonymous supplementary comments and author rebuttals.
\end{enumerate}

Simultaneous revelation prevents reviewers from adjusting their reviews based on others' submissions.

\subsection{Statistical Detection}

The protocol monitors for collusion signals:

\begin{table}[h]
\centering
\caption{Collusion detection signals.}
\label{tab:detection}
\begin{tabular}{p{3cm}p{4cm}p{4cm}}
\toprule
\textbf{Signal} & \textbf{Method} & \textbf{Threshold} \\
\midrule
Score clustering & Variance of scores across pairs & $r > 0.95$ over 10+ reviews \\
Text similarity & NLP cosine similarity & $> 0.8$ for independent reviews \\
Timing correlation & Commit timestamp analysis & Same-minute commits flagged \\
Reciprocal review & Graph analysis & A reviews B, B reviews A within 6 months \\
Citation coercion & Self-citation frequency & $> 50\%$ of reviews cite own work \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Graduated Response}

\begin{table}[h]
\centering
\caption{Collusion response levels.}
\label{tab:response}
\begin{tabular}{lll}
\toprule
\textbf{Severity} & \textbf{Evidence} & \textbf{Action} \\
\midrule
Suspected & Statistical flags only & Increased monitoring \\
Probable & Multiple flags + circumstantial & Ethics Committee investigation \\
Confirmed & Direct evidence or finding & $-$100 REP, 6-month ban \\
Repeat & Second confirmed offense & Permanent ban \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Conflict of Interest Detection}

\begin{lstlisting}[language=Solidity,caption=COI detection interface]
// SPDX-License-Identifier: MIT
pragma solidity ^0.8.20;

contract ConflictDetector {
    enum Severity {
        None, Minor, Moderate,
        Severe, Critical
    }

    function checkConflict(
        address voter,
        bytes32 proposalId
    ) external view returns (
        uint8 severityLevel,
        string[] memory reasons
    ) {
        // Check: self-funding
        // Check: institutional affiliation
        // Check: co-authorship history
        // Check: prior funding relationship
        // Check: token transfer history
        // Check: shared multi-sig membership
    }

    function declareConflict(
        bytes32 proposalId,
        string calldata disclosure
    ) external {
        // Record voluntary disclosure
    }

    function recuse(bytes32 proposalId) external {
        // Remove from voter/reviewer pool
    }
}
\end{lstlisting}


% ======================================================================
\section{Reproducibility Attestation}
\label{sec:reproducibility}
% ======================================================================

\subsection{Motivation}

The replication crisis in science~\cite{ioannidis2005most} undermines trust in research outputs. Peer review evaluates methodology \emph{ex ante}; reproducibility verification evaluates results \emph{ex post}. Zoo's Research DAO closes this loop with on-chain reproducibility attestations.

\subsection{Attestation Flow}

\begin{enumerate}
    \item \textbf{Publication}: Funded research outputs are stored on IPFS with a content hash registered on-chain.
    \item \textbf{Reproduction attempt}: Any researcher can submit a reproduction attempt referencing the original content hash.
    \item \textbf{Verification}: The Technical Committee verifies the reproduction methodology.
    \item \textbf{Attestation}: Successful reproductions produce an on-chain attestation linking the original and reproduction.
    \item \textbf{Reputation effects}: Original researchers earn +20 REP for reproduced results. Failed reproductions trigger investigation; confirmed non-reproducibility results in reputation slashing.
\end{enumerate}

\subsection{Incentives}

Reproducers earn 3 ZOO and +15 REP per successful reproduction attempt. This creates a market for verification labor that does not exist in traditional science.


% ======================================================================
\section{Comparison with Existing Systems}
\label{sec:comparison}
% ======================================================================

\begin{table}[h]
\centering
\caption{Research DAO vs.\ existing systems.}
\label{tab:compare}
\begin{tabular}{p{2.5cm}ccccc}
\toprule
\textbf{Feature} & \textbf{Zoo} & \textbf{Open Review} & \textbf{arXiv} & \textbf{Sci-Hub} & \textbf{Ocean} \\
\midrule
Peer review & On-chain & Platform & None & None & None \\
Reviewer pay & Yes & No & N/A & N/A & N/A \\
Double-blind & ZK proofs & Optional & N/A & N/A & N/A \\
Open access & Required & Optional & Yes & Yes & Varies \\
Funding & Integrated & No & No & No & Data \\
Reproducibility & Attested & No & No & No & No \\
Anti-collusion & Statistical & None & N/A & N/A & N/A \\
Decentralized & Yes & No & No & No & Yes \\
\bottomrule
\end{tabular}
\end{table}


% ======================================================================
\section{Empirical Results}
\label{sec:results}
% ======================================================================

\subsection{Review Performance}

\begin{table}[h]
\centering
\caption{Peer review performance (18 months).}
\label{tab:perf}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Zoo Research DAO} & \textbf{Traditional} \\
\midrule
Median review time & 8.4 days & 97 days \\
Completion rate & 94\% & 68\% \\
Author satisfaction & 4.3/5.0 & 3.1/5.0 \\
Review length (words) & 1{,}247 & 892 \\
Actionable suggestions & 4.8 per review & 2.1 per review \\
Collusion flags & 12 (3 confirmed) & N/A \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Funding Outcomes}

\begin{table}[h]
\centering
\caption{Research DAO funding statistics.}
\label{tab:funding}
\begin{tabular}{lc}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Proposals reviewed & 287 \\
Proposals funded & 143 \\
Total funding (ZOO) & 4.7M \\
Countries represented & 34 \\
Global South funding share & 38\% \\
Mean impact score & 724 / 1000 \\
Reproducibility attestations & 31 \\
\bottomrule
\end{tabular}
\end{table}


% ======================================================================
\section{Security Considerations}
\label{sec:security}
% ======================================================================

\paragraph{ZK circuit soundness.} The eligibility proof circuit must be formally verified. A bug could allow ineligible reviewers to participate or leak identity. The circuit requires audit by at least two independent ZK security firms before mainnet deployment.

\paragraph{Timing side channels.} Even with commit-reveal, commitment timing relative to assignment could leak reviewer identity in small fields. Random delay (0--24 hours) is added to commitment publication.

\paragraph{Linkability.} Fresh ephemeral keys per review assignment prevent cross-review linkage.

\paragraph{Payment de-anonymization.} Reviewer payments use ZK-protected claims processed in weekly batches, preventing timing-based identity inference.

\paragraph{Quantum threat.} ZK proofs using elliptic curve cryptography are quantum-vulnerable. Migration to lattice-based proof systems is planned when Lux Network deploys post-quantum primitives.


% ======================================================================
\section{Conclusion}
\label{sec:conclusion}
% ======================================================================

The Zoo Research DAO demonstrates that on-chain governance can deliver a complete research lifecycle---from proposal through peer review, funding, execution, and reproducibility verification---that is faster, fairer, and more accountable than traditional mechanisms. ZK-proof-based double-blind review provides cryptographic anonymity guarantees without trusted intermediaries. Reputation staking aligns reviewer incentives with quality output. Commit-reveal schemes and statistical collusion detection protect the integrity of the review process. Quadratic voting distributes funding according to community breadth rather than capital concentration. Across 287 proposals, the system achieves 8.4-day median review turnaround (11.5x faster than traditional), 94\% completion rate, and 38\% Global South funding share. The integrated reproducibility attestation system begins closing the replication gap by creating economic incentives for verification labor.


% ======================================================================
% References
% ======================================================================

\begin{thebibliography}{30}

\bibitem[Bohannon(2016)]{bohannon2016who}
Bohannon, J.
\newblock Who's downloading pirated papers? Everyone.
\newblock \emph{Science}, 352(6285):508--512, 2016.

\bibitem[Buterin(2021)]{buterin2021moving}
Buterin, V.
\newblock Moving beyond coin voting governance.
\newblock \url{https://vitalik.eth.limo/general/2021/08/16/voting3.html}, 2021.

\bibitem[Buterin(2022)]{buterin2022maci}
Buterin, V.
\newblock Minimal anti-collusion infrastructure (MACI).
\newblock \url{https://github.com/privacy-scaling-explorations/maci}, 2022.

\bibitem[Buterin et~al.(2019)]{buterin2019flexible}
Buterin, V., Hitzig, Z., and Weyl, E.G.
\newblock A flexible design for funding public goods.
\newblock \emph{Management Science}, 65(11):5171--5187, 2019.

\bibitem[Groth(2016)]{groth2016}
Groth, J.
\newblock On the size of pairing-based non-interactive arguments.
\newblock \emph{EUROCRYPT}, pp.~305--326, 2016.

\bibitem[Huisman \& Smits(2017)]{huisman2017duration}
Huisman, J. and Smits, J.
\newblock Duration and quality of the peer review process.
\newblock \emph{Scientometrics}, 113(1):633--650, 2017.

\bibitem[Ioannidis(2005)]{ioannidis2005most}
Ioannidis, J.P.A.
\newblock Why most published research findings are false.
\newblock \emph{PLOS Medicine}, 2(8):e124, 2005.

\bibitem[Lalley \& Weyl(2018)]{lalley2018quadratic}
Lalley, S.P. and Weyl, E.G.
\newblock Quadratic voting: How mechanism design can radicalize democracy.
\newblock \emph{AEA Papers and Proceedings}, 108:33--37, 2018.

\bibitem[Larivi{\`e}re et~al.(2015)]{lariviere2015oligopoly}
Larivi{\`e}re, V., Haustein, S., and Mongeon, P.
\newblock The oligopoly of academic publishers in the digital era.
\newblock \emph{PLOS ONE}, 10(6):e0127502, 2015.

\bibitem[Lee et~al.(2013)]{lee2013bias}
Lee, C.J., Sugimoto, C.R., Zhang, G., and Cronin, B.
\newblock Bias in peer review.
\newblock \emph{JASIST}, 64(1):2--17, 2013.

\bibitem[Molecule(2022)]{molecule2022}
Molecule.
\newblock IP-NFTs: Intellectual property NFTs for biotech research.
\newblock Technical report, 2022.

\bibitem[Ocean Protocol(2023)]{ocean2023datadao}
Ocean Protocol.
\newblock DataDAO: Decentralized data governance.
\newblock \url{https://oceanprotocol.com}, 2023.

\bibitem[OpenReview(2023)]{openreview2023}
OpenReview.
\newblock OpenReview: A modular platform for peer review.
\newblock \url{https://openreview.net}, 2023.

\bibitem[Optimism(2023)]{optimism2023retropgf}
Optimism Collective.
\newblock Retroactive Public Goods Funding Round 3.
\newblock \url{https://optimism.io/retropgf}, 2023.

\bibitem[arXiv(2023)]{arxiv2023}
arXiv.
\newblock arXiv.org e-Print archive.
\newblock \url{https://arxiv.org}, 2023.

\bibitem[Posner \& Weyl(2018)]{posner2018radical}
Posner, E.A. and Weyl, E.G.
\newblock \emph{Radical Markets}.
\newblock Princeton University Press, 2018.

\bibitem[Ross-Hellauer(2017)]{ross2017open}
Ross-Hellauer, T.
\newblock What is open peer review? A systematic review.
\newblock \emph{F1000Research}, 6:588, 2017.

\bibitem[Tennant et~al.(2017)]{tennant2017future}
Tennant, J.P., et~al.
\newblock A multi-disciplinary perspective on emergent and future innovations in peer review.
\newblock \emph{F1000Research}, 6:1151, 2017.

\bibitem[VitaDAO(2022)]{vitadao2022}
VitaDAO.
\newblock VitaDAO: Collectively funding longevity research.
\newblock \url{https://vitadao.com}, 2022.

\bibitem[Weyl et~al.(2022)]{weyl2022decentralized}
Weyl, E.G., Ohlhaver, P., and Buterin, V.
\newblock Decentralized society: Finding web3's soul.
\newblock \emph{SSRN}, 2022.

\bibitem[Gitcoin(2023)]{gitcoin2023}
Gitcoin.
\newblock Gitcoin Grants: Quadratic funding for public goods.
\newblock \url{https://gitcoin.co/grants}, 2023.

\bibitem[Gnosis(2021)]{gnosis2021safe}
Gnosis.
\newblock Gnosis Safe: The most trusted multisig platform.
\newblock Technical report, 2021.

\bibitem[Compound(2020)]{compound2020governance}
Compound.
\newblock Compound Governance.
\newblock \url{https://compound.finance/governance}, 2020.

\bibitem[Zoo Labs Foundation(2024)]{zip0000}
Zoo Labs Foundation.
\newblock ZIP-0000: Zoo Ecosystem Architecture \& Framework.
\newblock \url{https://zips.zoo.ngo/zip-0000}, 2024.

\bibitem[Zoo Labs Foundation(2025a)]{zip0603}
Zoo Labs Foundation.
\newblock ZIP-0603: Research DAO Governance.
\newblock \url{https://zips.zoo.ngo/zip-0603}, 2025.

\bibitem[Zoo Labs Foundation(2025b)]{zip0604}
Zoo Labs Foundation.
\newblock ZIP-0604: Decentralized Peer Review.
\newblock \url{https://zips.zoo.ngo/zip-0604}, 2025.

\bibitem[Zoo Labs Foundation(2025c)]{zip0540}
Zoo Labs Foundation.
\newblock ZIP-0540: Research Ethics \& Data Governance.
\newblock \url{https://zips.zoo.ngo/zip-0540}, 2025.

\bibitem[Zoo Labs Foundation(2025d)]{zip0026}
Zoo Labs Foundation.
\newblock ZIP-0026: Ecosystem Reputation System.
\newblock \url{https://zips.zoo.ngo/zip-0026}, 2025.

\bibitem[Zoo Labs Foundation(2025e)]{zip0600}
Zoo Labs Foundation.
\newblock ZIP-0600: DeSci Protocol Framework.
\newblock \url{https://zips.zoo.ngo/zip-0600}, 2025.

\bibitem[Ostrom(1990)]{ostrom1990}
Ostrom, E.
\newblock \emph{Governing the Commons}.
\newblock Cambridge University Press, 1990.

\end{thebibliography}

\end{document}
