\section{Team and Mission}
\label{sec:team}

\subsection{Founding}

\Zoo{} Labs Foundation Inc. is a 501(c)(3) non-profit founded by
\textbf{Antje Worring}, who authored the original Zoo whitepaper on
October~31, 2021. That document is the canonical founding artifact of
the foundation; it describes a Game-Fi multiverse for endangered species,
emotionally-intelligent NFT animals, and a \DAO{} that routes economic
value back to wildlife preservation. Worring also coined the term
\emph{NFT Liquidity Protocol} in the January~2022 deck, which is the
lineage that today underwrites the broader Liquidity Protocol family of
on-chain settlement primitives.

\subsection{Mission}

The \Zoo{} mission, as carried forward to the \Quasar{} 3.0 activation
(2025-12-25), is:

\begin{quote}\itshape
An open AI protocol for research aligned with environment-saving,
endangered AI, and the intersection of AI, blockchain, metaverse, and
gaming.
\end{quote}

This articulation is intentionally broader than ``conservation crypto.''
\Zoo{} now spans frontier-AI research (the \textsc{Zen} model family,
per-\textsc{llm} chains, Hamiltonian \textsc{llm}s, Z-JEPA, BitDelta),
metaverse and gaming infrastructure (Game-Fi, virtual habitats,
play-to-conserve mechanics), conservation and DeSci (federated wildlife
monitoring, satellite ecology, citizen science protocols), and
governance (the holographic \DAO{} described in \S\ref{sec:dao}).

\subsection{Core team}

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{Antje Worring} --- Founder, Zoo Labs Foundation.
   Author of the original Oct~31, 2021 whitepaper.
   \texttt{antje@zoo.ngo}.
\item \textbf{Zach Kelling} --- Chief Scientist / engineering lead.
   Co-author on the Foundation Mission paper, the per-\textsc{llm}
   chains paper, and on this document. Drives the \textsc{Zen} model
   family, the per-\textsc{llm} chain integration with \Lux{} \Quasar{}
   3.0, and the Hanzo / Lux / Zoo cross-stack. \texttt{zach@zoo.ngo}.
\item \textbf{Research staff} --- authors and maintainers of the
   80+ paper series and the 130+ \textsc{zip} corpus
   (\texttt{research@zoo.ngo}).
\item \textbf{Engineering staff} --- maintainers of \texttt{zoo/node}
   (Rust), \texttt{zoo/app} (Tauri + React desktop client),
   \texttt{zoo/zoo} (Next.js front-end), and the Zen LM training stack.
\end{itemize}

\subsection{Allied organizations}

\Zoo{} operates inside an explicitly layered ecosystem:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{\Lux{} Network} provides regulated chain infrastructure
   and the \Quasar{} 3.0 post-quantum consensus engine
   (\Lux{} \textsc{lp-020}, see \S\ref{sec:quantum}).
\item \textbf{\Hanzo{} AI Inc.} (Techstars '17) operates the Hanzo AI
   Chain inference and commerce marketplace and co-develops the
   \textsc{Zen} model family.
\item \textbf{A regulated securities partner} is \Zoo{}'s exclusive
   partner for retail-scale digital-securities trading; it brings the regulated
   ATS / broker-dealer / transfer-agent perimeter
   (see \S\ref{sec:securities}).
\end{itemize}

\subsection{Advisors and contributors}

The foundation maintains a rolling advisory roster across conservation
science, AI research, DeSci / open-research, and governance.
The advisor set is governed by the meta-\DAO{} and updated through the
contribution-graph attestation pipeline anchored on \Lux{} A-Chain
(\Lux{} \textsc{lp-134}). Detailed roster and join paths are in
\texttt{zips/team.md} (canonical team page).

\subsection{Provenance vs.\ today}

The Oct~31, 2021 founding vision named three pillars: open AI,
blockchain, and metaverse. Four years on, those pillars have hardened
into specific protocols: per-\textsc{llm} chains for open AI (one
sovereign \Quasar{} appchain per model, see~\cite{zoo-per-llm-chains}),
\Lux{} \Quasar{} 3.0 for blockchain, and a metaverse layer that
manifests across the Game-Fi, gaming, and AR/VR pillars of the original
whitepaper. The remainder of this paper specifies how those pillars now
extend into digital securities, quantum-secure settlement, holographic
governance, and democratized information access.
