\section{\$113 Trillion Digital Securities Market --- Powered by Lux + Liquidity}
\label{sec:securities}

\subsection{Market size}

The total addressable market for digital securities globally is
estimated at approximately \textbf{\$113 trillion}. This figure is the
standard market-size estimate aggregating global capitalization across
listed equities, fixed income, derivatives, and tokenized real-world
assets~\cite{bis-tokenization,wfe-market-cap}. We are explicit that
this is a \emph{market-size estimate}, not a \Zoo{}-specific
total-value-locked (\textsc{tvl}) claim. The \Zoo{} ecosystem is not
attempting to custody \$113 trillion; it is positioning itself as the
open AI / wildlife / metaverse layer of an ecosystem whose addressable
opportunity is that large.

\subsection{Positioning}

The stack of responsibilities at the \Quasar{} 3.0 activation date is:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{\Zoo{} Labs Foundation} provides the open AI / wildlife /
  metaverse layer. \Zoo{} is the application surface --- per-\textsc{llm}
  chains, NFT-native conservation, AR/VR companions, Game-Fi.
\item \textbf{\Lux{} Network} provides the regulated chain
  infrastructure: \Quasar{} 3.0 consensus (\Lux{} \textsc{lp-020}),
  identity proof on Z-Chain, custody on M-Chain MPC, and FHE substrate
  on F-Chain. Settlement happens here.
\item \textbf{\Liquidity{}} is \Zoo{}'s \emph{exclusive partner} for
  actual securities trading at retail scale. \Liquidity{} carries the
  regulated ATS / broker-dealer / transfer-agent
  perimeter~\cite{liquidity-perimeter} and is where the regulated
  market center, control plane, and official records live.
\end{itemize}

\subsection{Tagline and concrete capabilities}

\begin{quote}\itshape
Robinhood-style capabilities for everyone --- actually make money like
the 1\% finally.
\end{quote}

What that resolves to in concrete economic terms:

\begin{longtable}{p{4.6cm}p{10cm}}
\toprule
\textbf{Capability} & \textbf{Description} \\
\midrule
\endhead
Zero-commission trading &
   No per-trade commission for retail orders routed through
   \Liquidity{}. Revenue is earned on spread, market-data, and
   subscription tiers, not on commission per share. \\
Fractional shares &
   Position sizes denominated to the cent, not to the share. A
   \$5 position in a \$200 stock is one-fortieth of a share. \\
Instant settlement on \Quasar{} &
   Trades settle in sub-second finality on \Quasar{} 3.0 consensus
   (\S\ref{sec:quantum}). Cash and securities legs both atomic. \\
Tokenized real-world securities &
   Stocks, bonds, commodities, REITs, and private-equity positions
   issued as on-chain tokens with the official transfer-agent record
   maintained by \Liquidity{}. \\
24/7 markets &
   The chain is always on; market hours are an exchange-side policy,
   not a settlement-layer constraint. Off-hours trading routes via
   \Liquidity{} as the market center permits. \\
Public verifiable order book &
   Public Merkle commitments of every match; every retail trader can
   verify their fill against the canonical order book without trusting
   a private feed. \\
Compliant by construction &
   KYC, AML, suitability, and best-execution checks are enforced at
   the \Liquidity{} perimeter \emph{before} an order is routed to the
   chain. The chain is not a regulatory loophole; it is the
   audit-eligible settlement layer behind a regulated front. \\
\bottomrule
\end{longtable}

\subsection{Compliance}

The full regulated stack is operated by \Liquidity{}, not by \Zoo{}.
\Liquidity{} carries:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \textbf{SEC ATS} registration --- the alternative trading system
  is the regulated market center. \Zoo{} does not operate one.
\item \textbf{FINRA broker-dealer} registration --- order routing,
  best execution, suitability, KYC/AML.
\item \textbf{Registered transfer agent} --- maintenance of the
  official record of securities ownership and transfer.
\end{itemize}

This separation is deliberate: \Zoo{} Labs Foundation is a 501(c)(3)
non-profit and cannot itself operate a regulated market center.
\Liquidity{} carries that perimeter so \Zoo{} can ship the open-AI /
wildlife / metaverse layer that lives on top of it.

\subsection{Quantum-secure settlement}

Every securities trade routed through this stack settles with a
\Quasar{} 3.0 certificate (see \S\ref{sec:quantum}): \BLS{}-aggregate
classical signature, Corona Ring-LWE post-quantum threshold
signature, and an $\MLDSA{}\text{-}65$ identity proof committed via the
Z-Chain Groth16 rollup. The retail user can therefore truthfully say:

\begin{itemize}[nosep,leftmargin=1.2em]
\item \emph{Your trade is post-quantum-safe.}
\item \emph{Your custody is post-quantum-safe.}
\item \emph{Your audit trail is post-quantum-safe.}
\end{itemize}

This is the first time those three properties hold simultaneously for a
retail-accessible securities stack.

\subsection{Why Zoo (and not a generic chain)}

\Zoo{} is the alignment layer that gives the stack a reason to exist
beyond pure financial throughput. Conservation-aligned issuance, NFT
companions that fund real-world sanctuaries, and per-\textsc{llm} chains
that track research provenance are the durable economic flywheels that
keep retail traders engaged longer than a fee-rebate program would. The
\Zoo{} Foundation's 501(c)(3) status binds a portion of every economic
flow back into endangered-species preservation and open AI research,
which is the lineage of the founding 2021 vision.
