\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
% geometry loaded by zoocover
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{abstract}
\usepackage{longtable}

% Custom commands
\newcommand{\ZOO}{\textsc{Zoo}}
\newcommand{\DAO}{\textsc{DAO}}

% Title information
\title{\textbf{Zoo Labs: A Game-Fi Multiverse for Conservation of Endangered Species}\\
\large Whitepaper}
\author{
Antje Worring \quad Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{October 2021}

\begin{document}

\zoocoverpage

\begin{abstract}
\noindent
As we stand on the cusp of a digital revolution, an innovative project known as Zoo is propelling us into the uncharted territories of the metaverse, Web3, and Game-Fi. This initiative represents a radical rethinking of how we perceive and interact with the world, both physical and digital.

Zoo is a cornerstone of the emerging metaverse, an expansive and interconnected virtual space that is the next frontier in internet evolution. The project invites users to explore a captivating digital world teeming with virtual wildlife and real-world assets. Each activity in the Zoo ecosystem has a higher purpose tied to it, contributing to the overarching goal of conservation of endangered species.

Zoo leverages the power of Augmented and Virtual Reality (AR/MR/VR) to create a sense of presence, making the virtual world feel real and enhancing the emotional connection between players and the game. The project's integration of blockchain technology and Game-Fi---a fusion of gaming and decentralized finance (DeFi)---enables in-game assets to hold tangible, real-world value. Through this model, players can earn returns from in-game activities, effectively turning their gaming efforts into potential financial rewards.

Blockchain provides the foundation for Zoo's decentralized structure, enabling transparency, trust, and user empowerment. Each player's actions, accomplishments, and interactions are stored on the blockchain, creating a permanent, tamper-proof record. Zoo implements robust security measures, reduces barriers to entry associated with blockchain technology, and adopts a balanced and sustainable economic model to manage inherent market volatility.

Zoo Labs represents an exciting intersection of gaming, DeFi, and blockchain, holding the promise to reshape not just the gaming industry, but also our broader understanding of digital economies.
\end{abstract}

\tableofcontents
\newpage

% Part I: Introduction
\input{sections/01-mission}
\input{sections/02-zoo-animal-utility}
\input{sections/03-sustainability}
\input{sections/04-market-opportunity}
\input{sections/05-supporting-non-profits}
\input{sections/06-zoo-foundation}

% Part II: Game Play
\input{sections/07-gameplay-functions}
\input{sections/08-ai-assistant}
\input{sections/09-augmented-reality-app}
\input{sections/10-collateral-backed-nfts}
\input{sections/11-feeding-growing-breeding}
\input{sections/12-metaverse-companion}
\input{sections/13-native-token}
\input{sections/14-nft-marketplace}
\input{sections/15-gen-0-nft-drop}
\input{sections/16-asset-transfer}
\input{sections/17-zoo-animal-rewards}
\input{sections/18-roadmap}

% Part III: Differentiators
\input{sections/19-zoo-dao}
\input{sections/20-nft-liquidity-protocol}
\input{sections/21-nfts-that-make-you-smile}
\input{sections/22-bridging-blockchains}

% Part IV: Community
\input{sections/23-open-source}
\input{sections/24-partnerships}

\end{document}
