\section{Conclusion}


Zoo Network deploys a comprehensive post-quantum cryptography suite: \MLKEM{} for key encapsulation, \MLDSA{} for transaction signatures, \SLHDSA{} for high-assurance certification, and \Corona{} for threshold consensus. Hybrid composition ensures security during the classical-to-quantum transition. GPU NTT acceleration achieves 75\% of classical throughput with full quantum resistance. All primitives are formally verified in Lean~4.

\begin{thebibliography}{99}

\bibitem{shor1997polynomial}
P.~W. Shor, ``Polynomial-time algorithms for prime factorization and discrete logarithms on a quantum computer,'' \textit{SIAM J. Comput.}, vol.~26, no.~5, pp.~1484--1509, 1997.

\bibitem{nist2024pqc}
NIST, ``Post-Quantum Cryptography Standardization,'' FIPS 203, 204, 205, 2024.

\bibitem{lyubashevsky2010ideal}
V.~Lyubashevsky, C.~Peikert, and O.~Regev, ``On ideal lattices and learning with errors over rings,'' in \textit{EUROCRYPT}, 2010.

\bibitem{zoo-consensus}
Z.~Kelling, ``Quasar Hybrid Consensus on Zoo Network: GPU-Accelerated BLS and Post-Quantum Finality,'' Zoo Labs Foundation, 2026.

\bibitem{zoo-key-mgmt}
Z.~Kelling, ``Key Management for Post-Quantum Multi-Chain Architectures on Zoo Network,'' Zoo Labs Foundation, 2026.

\bibitem{zoo-threshold-sigs}
Z.~Kelling, ``Threshold Signatures for Decentralized Validator Sets on Zoo Network,'' Zoo Labs Foundation, 2026.

\bibitem{lux-pq-crypto}
Lux Industries, ``Post-Quantum Cryptography Suite for the Lux Blockchain,'' Lux Technical Report, 2025.

\bibitem{lux-corona-pq}
Lux Industries, ``Corona: A Ring-LWE Threshold Signature Scheme for Blockchain Consensus,'' Lux Technical Report, 2025.

\end{thebibliography}

