\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{caption}
\usepackage{listings}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={Zoo DEX - Native Securities Trading on Lux Quasar with Liquidity Protocol Integration},
    pdfauthor={Antje Worring, Zach Kelling},
    pdfsubject={Zoo Labs Foundation - Native Securities DEX Launch},
    pdfkeywords={Zoo DEX, native securities, Liquidity Protocol, Quasar 4.0, AMM V4, Robinhood-style, RWA, perpetuals}
}

% Macros
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\Lux}{\textsc{Lux}}
\newcommand{\Hanzo}{\textsc{Hanzo}}
\newcommand{\Zen}{\textsc{Zen}}
\newcommand{\Quasar}{\textsc{Quasar}}
\newcommand{\QuasarGPU}{\textsc{Quasar}\textsc{gpu}}
\newcommand{\Liquidity}{\textsc{Liquidity}}
\newcommand{\PoAI}{\textsc{PoAI}}
\newcommand{\TEE}{\textsc{tee}}
\newcommand{\LLM}{\textsc{llm}}
\newcommand{\GPU}{\textsc{gpu}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\AMM}{\textsc{amm}}
\newcommand{\DEX}{\textsc{dex}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\FHE}{\textsc{fhe}}
\newcommand{\ZKP}{\textsc{zkp}}
\newcommand{\BLS}{\textsc{bls}}
\newcommand{\MLDSA}{\textsc{ml-dsa}}
\newcommand{\Ringtail}{\textsc{Ringtail}}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray},
    tabsize=2,
    captionpos=b,
}

\title{\textbf{Zoo \DEX{} --- Native Securities Trading on \Lux{}
\Quasar{} with \Liquidity{} Protocol Integration}\\
\large April 20, 2026 launch edition}

\author{
Antje Worring, Zach Kelling\thanks{research@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{April 20, 2026}

\begin{document}

\zoocoverpage

\begin{abstract}
\noindent The \Zoo{} \DEX{} launches on April 20, 2026 with native
trading of equities, fixed income, real-world assets, and tokenised
derivatives, all settling on \Zoo{} D-Chain (a \Lux{} D-Chain
white-label per \textsc{lp-134}) with sub-second \Quasar{} 4.0
triple-cert finality. The \DEX{} integrates the \Liquidity{} Protocol
formally launched by \Liquidity{}.io on April 1, 2026 and adopted by
\Lux{}, \Hanzo{}, \Zoo{}, and Pars on April 20, 2026
\cite{zoo-adopts-liquidity-protocol}, providing compliance-by-construction
through a dedicated \Liquidity{} precompile that applies KYC/AML at
the perimeter (SEC ATS, FINRA broker-dealer, and registered transfer
agent obligations are met by the \Liquidity{}.io control plane). The
architecture is built on \Zoo{} 4.0's \GPU{}-native sovereign L1
\cite{zoo-4-0}, activated February 14, 2026 and inheriting the full
post-quantum stack from \Zoo{} 3.0 \cite{zoo-3-0}. The \DEX{} ships
\AMM{} V2 (constant product) and V3 (concentrated liquidity), with V4
implemented as a precompile to the \Lux{} \DEX{} cross-chain
liquidity router; \Zoo{} additionally exposes an optional in-chain
local \DEX{} precompile for \Zoo{}-specific assets (per-\LLM{}
tokens, fractional \NFT{}s). Order types include limit, market,
stop, IOC, FOK, post-only, hidden, and iceberg, with both batch-auction
and continuous-limit-book matching. A Robinhood-style retail UX
delivers zero commission, fractional shares, 24/7 markets, and
sub-1.1s settlement. Optional FHE-encrypted orderflow runs through
F-Chain (\textsc{lp-013}); ZK selective disclosure runs through Z-Chain
(\textsc{lp-063}). Quantum-secure trading is end-to-end: every trade
settles under a triple-cert ($\BLS{}_{12\text{-}381}$ + \Ringtail{} +
$\MLDSA{}$). The \DEX{} is the migration destination for \NFT{}
markets and the realisation of the original ``NFT Liquidity
Protocol'' coinage from the October 31, 2021 \Zoo{} whitepaper
\cite{zoo-2021}.

\textbf{Keywords}: \Zoo{} \DEX{}, native securities, \Liquidity{}
Protocol, \Quasar{} 4.0, real-world assets, tokenised derivatives,
post-quantum trading, Robinhood-style retail, \FHE{} orderflow
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-architecture}
\input{sections/02-native-securities}
\input{sections/03-liquidity-integration}
\input{sections/04-order-types}
\input{sections/05-retail-ux}
\input{sections/06-quantum-secure}
\input{sections/07-privacy}
\input{sections/08-nft-migration}
\input{sections/09-operator-economics}
\input{sections/10-roadmap}

\begin{thebibliography}{99}

\bibitem{lp-009}
Lux Foundation.
\newblock LP-009: \GPU{}-Native \textsc{evm}.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-013}
Lux Foundation.
\newblock LP-013: F-Chain \FHE{}-Protected Inference and Confidential Compute.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-020}
Lux Foundation.
\newblock LP-020: \Quasar{} Consensus 3.0 (Triple Consensus: \BLS{} + \Ringtail{} + \MLDSA{}).
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-063}
Lux Foundation.
\newblock LP-063: Z-Chain Zero-Knowledge Rollup and \ZKP{} Registry.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-105}
Lux Foundation.
\newblock LP-105: \Quasar{} Consensus Family Specification (Photon, Wave, Focus, Nova, Nebula).
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-132}
Lux Foundation.
\newblock LP-132: \QuasarGPU{} Execution Adapter for Deterministic \GPU{} Scheduling.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-133}
Lux Foundation.
\newblock LP-133: \Quasar{}-Native App Stack for Sovereign Appchains.
\newblock Lux Improvement Proposal, 2025.

\bibitem{lp-134}
Lux Foundation.
\newblock LP-134: \Lux{} Chain Topology and White-Label D-Chain Template (P/C/X/Q/Z/A/B/M/F/D-Chains).
\newblock Lux Improvement Proposal, 2026.

\bibitem{lp-137}
Lux Foundation.
\newblock LP-137: \GPU{}-Residency Invariant.
\newblock Lux Improvement Proposal, 2025.

\bibitem{liquidity-protocol-formal}
Liquidity.io.
\newblock The \Liquidity{} Protocol: A Compliance-by-Construction Settlement Standard for Tokenised Securities. Formal Proof.
\newblock Liquidity Whitepaper and Soundness Proof, April 1, 2026.

\bibitem{zoo-adopts-liquidity-protocol}
Zoo Labs Foundation.
\newblock Zoo Adopts the \Liquidity{} Protocol.
\newblock Zoo Adoption Note, April 20, 2026.

\bibitem{zoo-2021}
Antje Worring et al.
\newblock Zoo: A Decentralized Wildlife Conservation Platform (NFT Liquidity Protocol).
\newblock Zoo Whitepaper (Original), October 31, 2021.

\bibitem{zoo-2-0}
Zoo Labs Foundation.
\newblock Zoo \textsc{evm}: A Post-Quantum Safe L2 Blockchain with AI-Native Precompiles on the \Lux{} Network.
\newblock Zoo Whitepaper, October 2024.

\bibitem{zoo-3-0}
Zoo Labs Foundation.
\newblock Zoo 3.0 --- Full Post-Quantum Stack.
\newblock Zoo Whitepaper, October 31, 2025.

\bibitem{zoo-4-0}
Zoo Labs Foundation.
\newblock Zoo 4.0 --- \GPU{}-Native Sovereign L1.
\newblock Zoo Whitepaper, February 14, 2026.

\bibitem{zoo-per-llm-chains}
Zoo Labs Foundation.
\newblock Per-\LLM{} Chains: One Sovereign Chain per Model on \Quasar{}.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{zoo-2025-securities-and-dao}
Zoo Labs Foundation.
\newblock Zoo 2025: Digital Securities, Quantum-Secure Settlement, and the Holographic Consensus \DAO{}.
\newblock Zoo Whitepaper, December 15, 2025.

\bibitem{lux-dex}
Lux Network.
\newblock \Lux{} \DEX{} Specification.
\newblock \url{https://github.com/luxfi/dex}, 2026.

\bibitem{uniswap-v3}
Adams, H., et al.
\newblock Uniswap v3 Core.
\newblock Uniswap Whitepaper, March 2021.

\end{thebibliography}

\end{document}
