\section{\Zoo{} L1 Chain Set and Tenant Hierarchy}
\label{sec:chain-set}

\Zoo{} 4.0 is a sovereign L1 with a chain set strictly broader than
the financial-rail core. The financial-rail core is \DEX{} + \textsc{evm}
+ \FHE{} --- the same three subsystems \Hanzo{} 4.0 carries as its
canonical \emph{triumvirate}\cite{hanzo-4-0}. \Zoo{} carries that core
and more: it inherits the majority of the \Lux{}-family primary chains
(\textsc{lp-134}\cite{lp-134}) and natively hosts L2/L3/L4 tenants on
top.

\subsection{Inherited \Lux{} primary chains}

\Zoo{} 4.0 enables every \Lux{} primary chain except those reserved for
the parent network's cross-cutting infrastructure:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{P-Chain} (PVM) --- platform, staking, validators, epochs.
  \item \textbf{C-Chain} (\textsc{evm}, cevm) --- general smart
        contracts, GPU-native bytecode evaluation per \textsc{lp-009}.
  \item \textbf{X-Chain} (XVM) --- UTXO assets, swaps, native \textsc{tx}s.
  \item \textbf{Q-Chain} (QVM) --- \Quasar{} threshold-key ceremony
        (Ringtail \textsc{dkg}).
  \item \textbf{F-Chain} (FVM) --- \FHE{} compute, encrypted
        \textsc{evm}, confidential ERC-20 (\textsc{lp-013} v2,
        \textsc{lp-067}).
  \item \textbf{D-Chain} (white-labeled \DEX{} template) --- the \Zoo{}
        \DEX{} (\S\ref{sec:zoo-dex}), with native securities,
        \NFT{}-fractional pools, and Liquidity Protocol routing.
\end{itemize}

\Zoo{} L1 \emph{does not} run M-Chain (MPC ceremonies), B-Chain
(cross-ecosystem bridge), A-Chain (attestation/AI provenance), or
Z-Chain (\textsc{zk} rollups). Those remain \Lux{}-primary
infrastructure and \Zoo{} accesses them \emph{natively} via the
\Lux{} primary network --- same \Quasar{} cert lanes, same
subject-binding, same finality, no wrapped assets, no cross-chain
hop tax. Bridge withdrawals run on \Lux{} B-Chain natively; AI
provenance is recorded on \Lux{} A-Chain (AIVM) natively;
attestation, MPC ceremonies, and ZK rollups follow the same
pattern. The boundary between \Zoo{}'s chain set and \Lux{} primary
is cert-lane subject-binding, not a bridge.

\subsection{\Zoo{}-specific chains}

On top of the inherited \Lux{} primary chains, \Zoo{} L1 ships:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{Per-\LLM{} chains}~\cite{zoo-per-llm-chains} --- one
        chain per frontier model, where weights are commitments and
        inference is block production. Provenance, attribution, and
        royalty distribution are first-class.
  \item \textbf{Conservation chains} --- domain-specific subchains
        for wildlife \NFT{}s, carbon credits, and impact bonds, with
        the 1--3\% sustainability tax routed to the Foundation
        Treasury per ZIP-0017 and ZIP-0034.
  \item \textbf{Experience-Ledger DSO chain} --- the curated
        experience library backing training-free model adaptation.
\end{itemize}

\subsection{Native L2/L3/L4 tenancy}
\label{sec:zoo-tenancy}

A \Zoo{} L1 design constraint at the 4.0 boundary is that any
sovereign chain template inherited from \Lux{} primary (per
\textsc{lp-134} white-label) is \emph{also} a tenant host. Every
\Zoo{} chain in the set above can host:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{L2} --- tenant appchains, sovereign over their own
        validator set, inheriting \Quasar{} cert lanes from \Zoo{} L1.
  \item \textbf{L3} --- per-product subchains, typically deployed by
        L2 tenants for isolation of orderflow, compute, or
        compliance domains.
  \item \textbf{L4} --- per-user/org subchains, ephemeral or
        long-lived, useful for confidential lanes and
        organisation-private DAOs.
\end{itemize}

The L1$\rightarrow$L4 ladder is identical to the \Lux{} primary
hierarchy and uses the same cert-lane subject-binding to prevent
cross-chain replay.

\subsection{Financial-rail core}

The financial-rail core (\DEX{} + \textsc{evm} + \FHE{}) is what makes
\Zoo{} L1 a complete venue for digital securities, conservation
finance, and confidential \DAO{} grant execution. \DEX{} provides the
venue, \textsc{evm} provides the logic, \FHE{} provides the
confidentiality. The remainder of this paper details how the
financial-rail core, the inherited \Lux{} primary chains, the
\Zoo{}-specific chains, and the L1$\rightarrow$L4 tenancy interlock at
the 2026-02-14 activation boundary.
