\section{\Zoo{} Bridge}
\label{sec:zoo-bridge}

\Zoo{} Bridge (at \texttt{\textasciitilde/work/zoo/bridge}) is the
cross-ecosystem bridge for \Zoo{} L1. It lets ZRC-20 fungibles, ZRC-721
\NFT{}s, ZRC-1155 multi-tokens, and the native \$ZOO gas token move
between \Zoo{} L1 and external chains while preserving conservation
metadata, provenance chains, and token-bound account state.

\subsection{Architecture}

\Zoo{} Bridge is layered:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{Bridge \textsc{vm} layer.} \Zoo{} Bridge sits on top of
        the \Lux{} B-Chain (BVM, \textsc{lp-134}). The B-Chain is the
        \Lux{} ecosystem's bridge substrate; \Zoo{} Bridge is the
        \Zoo{}-specific message and asset model on top of it.
  \item \textbf{Threshold signing.} Lock and mint operations require a
        2-of-3 supermajority across the \Zoo{} bridge committee, which
        signs via M-Chain (MVM) MPC ceremonies (CGGMP21 / FROST), with
        \Corona{} threshold proofs anchoring the post-quantum side.
  \item \textbf{Asset adapters.} One adapter per supported foreign
        chain: \Lux{} C/X/B/Z-Chains, Ethereum mainnet, Polygon, BSC
        (legacy migration path for the original 2021 \Zoo{}
        deployment), and arbitrary EVM-compatible chains via the
        generic adapter.
\end{itemize}

\subsection{Supported chains}

At activation, \Zoo{} Bridge supports:

\begin{longtable}{p{0.20\textwidth}p{0.30\textwidth}p{0.40\textwidth}}
\toprule
\textbf{Chain} & \textbf{Direction} & \textbf{Notes} \\
\midrule
\Lux{} C-Chain & both & native EVM, fastest \\
\Lux{} X-Chain & both & for UTXO-based assets \\
\Lux{} B-Chain & both & native bridge VM \\
\Lux{} Z-Chain & both & for ZK-rollup-bound assets \\
Ethereum mainnet & both & wrapped \$ZOO, NFT migration \\
BSC & inbound & legacy 2021 \Zoo{} migration only \\
Polygon & both & retail wallets and tooling \\
Arbitrary EVM & both & generic adapter, slower path \\
\bottomrule
\end{longtable}

\subsection{Conservation metadata preservation}

A wildlife ZRC-721 carries fields that generic bridges lose: provenance
chain (pedigree), conservation contributions, token-bound account
balances, and on-chain achievements (gym, breeding, feeding). \Zoo{}
Bridge encodes these into the lock event and replays them on the
mint side. A bridged \NFT{} arrives identical, not a synthetic
wrapper.

\subsection{Yield-bearing bridge bookkeeping}

When \$ZOO sits on the foreign side as a wrapped token, the locked
backing on \Zoo{} L1 can stake into the validator set. Yield is
accrued on a per-deposit basis; on burn-and-unlock, the depositor
receives principal plus accumulated yield. This follows the
yield-bearing bridge pattern (\textsc{lp-018}).

\subsection{Security model}

The threshold signing ceremony has three properties:

\begin{enumerate}[leftmargin=1.4em]
  \item \textbf{Liveness.} Any 2-of-3 of the bridge committee can
        sign; one offline signer does not stall the bridge.
  \item \textbf{Safety.} Any forged release requires compromising 2-of-3
        committee members \emph{and} forging a \Corona{} threshold
        proof on M-Chain. The \Corona{} layer is post-quantum.
  \item \textbf{Auditability.} Every signed message is logged on
        B-Chain with the full triple-consensus certificate.
\end{enumerate}

\subsection{Migration of 2021 \Zoo{} BSC deployment}

The original 2021 \Zoo{} deployment lives on BSC. \Zoo{} Bridge
provides a one-way migration adapter that reads the BSC ZRC-20 and
ZRC-721 holdings, escrows them under the BSC bridge contract, and
mints native equivalents on \Zoo{} L1. Users opt in by signing a
migration intent with the BSC private key; the bridge committee
counter-signs and the L1 token is delivered. The BSC token contract
is locked post-migration so legacy holders cannot double-spend.
