\section{\Lux{} Wallet and Desktop}
\label{sec:wallet-desktop}

\Zoo{} 4.0 plugs into the \Lux{} Wallet and the \Lux{} Desktop app. The
\Lux{} Wallet is the canonical signing surface for \Lux{}-ecosystem
chains; the \Lux{} Desktop app is the canonical management UI. \Zoo{}
L1 appears as a chain target on both, alongside \Lux{} P/C/X/Q/Z/A/B/M/F-
Chains.

\subsection{Account / portfolio view}

The portfolio screen aggregates balances across \Lux{} ecosystem
assets and \Zoo{}-specific assets:

\begin{itemize}[leftmargin=1.2em]
  \item \$LUX (P-Chain stake, C-Chain balance, X-Chain UTXO).
  \item \$AI (A-Chain mining rewards, Hanzo network earnings).
  \item \$ZOO (Zoo L1 native, plus bridged wrappers).
  \item Per-\LLM{} tokens (one per active per-\LLM{} chain
        \cite{zoo-per-llm-chains}: \texttt{ZEN4-NANO}, \texttt{ZEN4-MINI},
        \dots, \texttt{ZEN4-ULTRA-MAX}).
  \item ZRC-721 \NFT{}s with thumbnails, provenance chain, and
        token-bound account holdings.
  \item Optional confidential ERC-20 balances (F-Chain), shown as
        ``encrypted'' until the user unwraps locally.
\end{itemize}

\subsection{Send / receive}

The send flow signs with the user's local key, submits to the chosen
chain, and surfaces \Quasar{}Cert finality progress: classical
fast-path acknowledgement first, then \Corona{} threshold
confirmation, then \MLDSA{} identity proof. The wallet shows the
triple-consensus state visually, so the user understands the
post-quantum side of the certificate is in place before treating
the transfer as final.

\subsection{\DEX{} swap}

The swap screen exposes a chain selector --- \Lux{} C-Chain or \Zoo{}
L1 --- because the same trading pair can have liquidity on both. The
selector remembers per-pair preferences. Behind the scenes:

\begin{itemize}[leftmargin=1.2em]
  \item C-Chain swap routes via the \Lux{} \DEX{} (\textsc{lp-009}
        + \cite{lux-dex}).
  \item Zoo L1 swap routes via \Zoo{} \DEX{} V2 / V3
        (\S\ref{sec:zoo-dex}).
  \item Cross-chain swap (Zoo asset $\rightarrow$ Lux asset) routes
        via Zoo \DEX{} V4 (the precompile to \Lux{} \DEX{}) plus
        a B-Chain bridge hop.
\end{itemize}

\subsection{Staking}

The staking screen splits between \Lux{} validator delegation
(P-Chain) and \Zoo{} validator delegation (Zoo L1 native staking
contract). Slashing risk is shown per-validator with on-chain
attestation history pulled from A-Chain.

\subsection{Bridge}

The bridge screen is a wrapper around the B-Chain bridge VM and
\Zoo{} Bridge. The user picks source chain, destination chain, and
asset; the wallet computes the bridge route, surfaces the threshold
ceremony progress, and shows the final mint event.

\subsection{AI inference}

The AI inference screen integrates \Zoo{} AI Desktop:

\begin{itemize}[leftmargin=1.2em]
  \item Local-only mode: queries to a local \Zen{} model. No network.
  \item Zoo Cloud mode: queries dispatched to federated \GPU{}
        capacity, paid in \$AI, attested via A-Chain.
  \item Hanzo Marketplace mode: queries dispatched to the Hanzo AI
        Chain inference marketplace \cite{hanzo-ai-chain}, paid in
        \$AI, with HMM (Hanzo Model Marketplace) routing.
  \item Confidential mode toggle: routes via F-Chain \FHE{}
        (\S\ref{sec:privacy}) so neither the worker nor the chain
        sees the plaintext prompt.
\end{itemize}

\subsection{Privacy mode}

A toggle at the top of the wallet enables privacy mode for the active
session:

\begin{itemize}[leftmargin=1.2em]
  \item Transfers route via F-Chain Confidential ERC-20 (\textsc{lp-067}).
  \item Identity proofs route via Z-Chain selective disclosure
        (\textsc{lp-063}).
  \item Wallet displays show ciphertext indicators rather than
        cleartext balances when in privacy mode.
\end{itemize}

\subsection{Mobile and desktop parity}

The \Lux{} Wallet ships in three form factors --- browser extension,
mobile (iOS/Android), and \Lux{} Desktop --- with the same screen
inventory. The desktop variant additionally surfaces validator
operator panes (block production stats, slashing exposure,
threshold-ceremony participation) for users running Zoo or Lux
validators.

\subsection{Screen inventory summary}

The full screen specifications --- ASCII mocks, flow diagrams, and
edge cases --- live in the wallet repo at
\texttt{\textasciitilde/work/lux/wallet/SCREENS.md}, which is the
canonical UX reference for engineering.
