\section{Security Analysis}

\label{sec:security}

\subsection{Threat Model}

We consider adversaries who may:
\begin{enumerate}
    \item Attempt to drain agent wallet
    \item Manipulate agent to make bad trades
    \item Steal agent key material
    \item Perform Sybil attacks via agent cloning
\end{enumerate}

\subsection{Security Properties}

\begin{theorem}[Wallet Safety]
Under autonomy bounds, an agent cannot transfer more than $\min(\text{maxTx}, \text{dailyLimit})$ within any 24-hour period.
\end{theorem}

\begin{proof}
Each transaction requires autonomy check (Algorithm~\ref{alg:tx}). Transactions exceeding maxTx require owner signature. Daily limit is enforced by smart contract. Rate limiting prevents rapid draining.
\end{proof}

\begin{theorem}[Owner Override]
The owner can always halt agent activity and withdraw assets.
\end{theorem}

\begin{proof}
Owner key $K_O$ can:
\begin{enumerate}
    \item Set autonomy parameters to zero (disable all actions)
    \item Execute direct wallet withdrawal
    \item Revoke agent key authorization
\end{enumerate}
\end{proof}

\begin{theorem}[Key Separation]
Agent cannot access owner's external assets; owner cannot impersonate agent.
\end{theorem}

\begin{proof}
Agent key $K_A$ is derived from model hash and stored in TEE. Owner never has access to $sk_A$. Agent key only controls agent wallet, not owner's personal wallets.
\end{proof}

\subsection{Attack Mitigation}

\begin{table}[t]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Attack} & \textbf{Mitigation} & \textbf{Residual Risk} \\
\midrule
Wallet drain & Autonomy bounds & Low \\
Bad trades & Whitelist + limits & Medium \\
Key theft & TEE protection & Low \\
Agent cloning & Model hash binding & Low \\
Flash loan attack & Cooldown period & Medium \\
\bottomrule
\end{tabular}
\caption{Attack mitigation strategies.}
\label{tab:attacks}
\end{table}

