\section{Introduction}


The Zoo ecosystem spans three chains with distinct purposes:

\begin{itemize}
    \item \textbf{Lux L0} (chain ID 96369): The base layer providing consensus, security, and native token economics. LUX is the gas token.

    \item \textbf{Zoo L2} (chain ID 200200): A conservation-focused chain for AI/ML workloads, decentralized science, and conservation finance. ZOO is the native token.

    \item \textbf{Beluga L3} (chain ID 420420): A specialized chain for AI model economics---model licensing, inference markets, and compute credit trading~\cite{beluga-l3-whitepaper}. BLG is the native token.
\end{itemize}

Cross-chain asset transfer is essential for this ecosystem: researchers stake LUX on Lux L0, train models on Zoo L2, and monetize them on Beluga L3. Each transfer involves not just fungible tokens but also AI-specific assets (model NFTs, data commitments, inference credits) that require semantic bridging beyond simple lock-and-mint.

Existing bridge designs (lock-and-mint, liquidity pools, optimistic verification) were designed for fungible token transfers. AI assets present unique challenges: model NFTs must preserve metadata integrity, data commitments must maintain cryptographic binding, and inference credits must be atomically converted between chain-specific denominations.

\subsection{Contributions}

\begin{enumerate}
    \item \textbf{Asset Taxonomy}: Four bridgeable asset classes with chain-specific representations (Section~\ref{sec:assets}).

    \item \textbf{Warp Bridge}: Verified cross-chain messaging with PQ-safe BLS signatures for model NFTs and data commitments (Section~\ref{sec:warp}).

    \item \textbf{Teleport Bridge}: Instant token transfers for ZOO $\leftrightarrow$ LUX with sub-second finality (Section~\ref{sec:teleport}).

    \item \textbf{Beluga Relay}: Chained Warp messages for L0 $\to$ L2 $\to$ L3 transfers (Section~\ref{sec:beluga}).

    \item \textbf{Security Analysis}: Reduction of bridge safety to Warp protocol safety (Section~\ref{sec:security}).
\end{enumerate}

