\documentclass[11pt,letterpaper]{article}
\usepackage{shared/zoocover}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
% geometry loaded by zoocover

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

% Custom commands
\newcommand{\Zoo}{\textsc{Zoo}}
\newcommand{\NFT}{\textsc{nft}}
\newcommand{\ERC}{\textsc{erc}}
\newcommand{\DAO}{\textsc{dao}}
\newcommand{\ZOO}{\$\textsc{zoo}}
\newcommand{\KEEPER}{\$\textsc{keeper}}
\newcommand{\OEP}{\text{OEP}}
\newcommand{\ETH}{\textsc{eth}}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\input{shared/lstlang}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\bfseries,
  commentstyle=\color{gray},
  stringstyle=\color{darkgray},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=5pt,
  frame=single,
  breaklines=true,
  captionpos=b,
  tabsize=2
}

% Title and authors
\title{\textbf{ZOO: A NFT Liquidity Protocol for Wildlife Conservation}\\[0.3em]
\large Genesis Retrospective}

\author{
Zach Kelling\thanks{zach@zoo.ngo}\\
\textit{Zoo Labs Foundation}\\
\texttt{research@zoo.ngo}
}

\date{October 2021}

\begin{document}

\zoocoverpage

\begin{abstract}
We present the \textbf{Zoo NFT Liquidity Protocol}, a collateral-backed breeding and staking
mechanism for AI-powered digital animals that aligns speculative DeFi incentives with real-world
endangered species conservation. Originating in the Zoo Gen~0 Deck of October
2021---\emph{predating ChatGPT by over a year}---the protocol introduced emotionally intelligent,
conversational NFT avatars backed by locked collateral, tiered yield generation through a
generational breeding system, and a dual-token economy (\ZOO{} + \KEEPER{}).  Each Gen~0
animal hatches from one of three egg tiers (Sublime, Rare, Endangered), progresses through life
stages (Baby, Teen, Adult) via escalating mint costs denominated in the egg's Original Egg Price
(\OEP{}), and breeds offspring whose generational yield flows back to ancestors in a virtuous
conservation loop.  A comprehensive sustainability tax---encompassing royalty fees (5\%), breeding
fees (0.015~\ETH{}), hatching fees (0.04~\ETH{}), feeding boosts, token sale levies (8\% on
\ZOO{} or \KEEPER{} sells, split 4\% LP / 4\% DAO), and NFT burn penalties (20\%, split 10\%
LP / 10\% DAO)---ensures continuous funding of both liquidity pools and the \Zoo{} \DAO{}
treasury.  We formalize the yield accumulation dynamics, prove long-run protocol solvency under
realistic adoption curves, and describe the ecosystem flow from gameplay through the \DAO{} to
Zoo NGO, the non-profit entity funding on-the-ground conservation of endangered species.  At its
August 2021 peak, the \ZOO{} token reached a market capitalization exceeding \$2B~USD
(ZOO/BNB on PancakeSwap), validating market demand for conservation-aligned digital asset
protocols.  The protocol's vision of emotionally intelligent, conversational AI avatars---each
animal ``fully converse[s] and play[s] with users''---preceded ChatGPT (November 2022) by over
thirteen months.

\textbf{Keywords}: NFT, liquidity protocol, collateral-backed, breeding, tokenomics, conservation,
AI avatars, dual-token, DeFi
\end{abstract}

%% ============================================================================
\section{Introduction}
\label{sec:intro}
%% ============================================================================

\subsection{Mission: Artificially Intelligent Digital Animals that Preserve Endangered Species}

Zoo Labs Foundation was established with a singular mission: to create \emph{artificially intelligent digital animals that preserve endangered species}.  The protocol channels speculative energy in the NFT and DeFi markets toward four concrete conservation initiatives:

\begin{enumerate}
    \item \textbf{Habitat Restoration}: Landscaping and on-the-ground activities for the health and safety of endangered species and their ecosystems.
    \item \textbf{Rescue and Reintegration}: Rescuing animals orphaned by poachers and reintegrating them into the wild.
    \item \textbf{Data Collection}: Collecting behavioral and population dynamics data to identify and eliminate poaching networks.
    \item \textbf{Policy Advocacy}: Pursuing legal avenues to enact lasting policy change for species protection.
\end{enumerate}

These four pillars are funded not through traditional charity but through the protocol's sustainability tax (Section~\ref{sec:sustainability-tax}), creating an economic engine where every trade, breed, hatch, and burn directly finances conservation.

\subsection{Market Context (2021)}

The Citi Global Perspectives \& Solutions report of 2021 projected the metaverse as an \$8--13 trillion opportunity by 2030, with an addressable user base of 5 billion~\cite{citi2021metaverse}.  Zoo positioned itself at the intersection of three converging trends: NFT collectibles, play-to-earn DeFi gaming, and AI-powered virtual companions.  At its August 2021 all-time high, the \ZOO{}/BNB pair on PancakeSwap achieved a market capitalization exceeding \textbf{\$2 billion USD}~\cite{poocoin2021}, demonstrating significant market demand for conservation-aligned digital assets.

\subsection{Pre-GPT AI Vision}

A defining characteristic of the Zoo Gen~0 design (October 2021) was its commitment to \emph{intelligent avatars}: ``Each animal is emotionally intelligent, and can fully converse and play with users.''  This vision of AI-powered conversational NFTs preceded the release of ChatGPT (November 2022) by over thirteen months, and preceded GPT-3.5's public API by even longer.  The protocol anticipated a world of persistent, emotionally responsive AI companions embedded in blockchain-native economic systems---a vision that only became mainstream after the generative AI explosion of 2023.

\subsection{History and Motivation}

The ZOO project launched in 2021 on Binance Smart Chain with PancakeSwap~\cite{pancakeswap}
serving as the primary decentralized exchange for the \ZOO{} token.  An earlier version of
the protocol (V3) was abandoned after the development team associated with that iteration
misappropriated investor tokens.  The current protocol (V4) was rebuilt from scratch with
full transparency: no hidden allocations, no centralized token custody, and a public
commitment of conservation fees enforced on-chain.  The \ZOO{} token
reached an all-time-high market capitalization exceeding \$2B in August 2021 on the
ZOO/BNB pair~\cite{poocoin2021}.

\subsection{Contributions}

This paper makes the following contributions:

\begin{enumerate}
    \item \textbf{Tiered Egg Tokenomics}: A complete specification of the Gen~0 egg pricing, life-stage progression, and yield schedule (Section~\ref{sec:tokenomics}).
    \item \textbf{Dual-Token Economy}: Formalization of the \ZOO{} and \KEEPER{} token roles and interactions (Section~\ref{sec:dual-token}).
    \item \textbf{Sustainability Tax}: A comprehensive fee taxonomy ensuring protocol-level conservation funding (Section~\ref{sec:sustainability-tax}).
    \item \textbf{Generational Breeding}: Mathematical model of cross-generational yield flow from offspring to ancestors (Section~\ref{sec:breeding}).
    \item \textbf{Feeding Boosts}: Staking-based yield amplification through collateral locking (Section~\ref{sec:feed}).
    \item \textbf{Ecosystem Flow}: End-to-end value circulation from gameplay through the \DAO{} to conservation outcomes (Section~\ref{sec:ecosystem}).
    \item \textbf{Intelligent Avatars}: Architecture for emotionally intelligent, conversational AI animals (Section~\ref{sec:avatars}).
    \item \textbf{Future Roadmap}: New animal drops, metaverse land, mutations, and virtual pet extensions (Section~\ref{sec:future}).
    \item \textbf{NFT Liquidity Protocol}: Protocol architecture in which each NFT carries transferable yield and accepts multi-asset collateral (Section~\ref{sec:protocol}).
    \item \textbf{ZOO Bridge}: Cross-chain play and asset exchange across sixteen blockchain networks (Section~\ref{sec:bridge}).
\end{enumerate}

%% ============================================================================
\section{Gen~0 Egg Tokenomics}
\label{sec:tokenomics}
%% ============================================================================

The Gen~0 collection establishes the foundational economic parameters of the protocol.  Each animal begins as an egg belonging to one of three rarity tiers, directly mapped to real-world conservation urgency.

\subsection{Egg Tiers and Daily Yield}

\begin{table}[ht]
\centering
\caption{Gen~0 Egg Tiers: Species, Rarity, Yield, and Original Egg Price}
\label{tab:egg-tiers}
\begin{tabular}{@{}llccccc@{}}
\toprule
\textbf{Tier} & \textbf{Species} & \textbf{Rarity} & \textbf{Baby} & \textbf{Teen} & \textbf{Adult} & \textbf{\OEP{}} \\
 & & & (\ZOO{}/day) & (\ZOO{}/day) & (\ZOO{}/day) & (\ETH{}) \\
\midrule
Sublime & Red Wolf & Almost Extinct & 1 & 2 & 3 & 1.0 \\
Rare & Amur Leopard & Rare & 0.25 & 0.5 & 1 & 0.5 \\
Endangered & Sumatran Elephant & Most Common & 0.05 & 0.1 & 0.2 & 0.05 \\
Endangered & Javan Rhinoceros & Most Common & 0.05 & 0.1 & 0.2 & 0.05 \\
Endangered & Siberian Tiger & Most Common & 0.05 & 0.1 & 0.2 & 0.05 \\
\bottomrule
\end{tabular}
\end{table}

The yield schedule is deliberately tied to real-world conservation urgency: the rarer the species in nature, the higher the yield and the higher the entry cost.

\subsection{Life-Stage Progression and Costs}

Each animal progresses through four life stages with escalating costs denominated in multiples of the Original Egg Price (\OEP{}):

\begin{table}[ht]
\centering
\caption{Life-Stage Transition Costs (multiples of \OEP{})}
\label{tab:life-stages}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Transition} & \textbf{Cost} & \textbf{Example (Sublime, \OEP{} = 1~\ETH{})} \\
\midrule
Hatch to Baby & Free & 0~\ETH{} \\
Mint Teen & $1 \times \OEP$ & 1~\ETH{} \\
Mint Adult & $2 \times \OEP$ & 2~\ETH{} \\
Breed Adult & $3 \times \OEP$ & 3~\ETH{} \\
\bottomrule
\end{tabular}
\end{table}

\begin{definition}[Total Investment for Adult]
The total cost to bring a Gen~0 egg to breeding-capable adult status is:
\begin{equation}
C_{\text{adult}} = \OEP + 0 + 1 \cdot \OEP + 2 \cdot \OEP = 4 \cdot \OEP
\label{eq:adult-cost}
\end{equation}
where the initial \OEP{} is the egg purchase price.
\end{definition}

\subsection{Yield Accumulation Model}

Let $Y_s$ denote the daily yield rate for life stage $s \in \{\text{Baby}, \text{Teen}, \text{Adult}\}$ and tier $\tau \in \{\text{Sublime}, \text{Rare}, \text{Endangered}\}$.  Let $t_s$ denote the number of days spent in stage $s$.  The cumulative yield for a single animal through its life stages is:

\begin{equation}
\mathcal{Y}(\tau) = Y_{\text{Baby}}^{(\tau)} \cdot t_{\text{Baby}} + Y_{\text{Teen}}^{(\tau)} \cdot t_{\text{Teen}} + Y_{\text{Adult}}^{(\tau)} \cdot t_{\text{Adult}}
\label{eq:yield-cumulative}
\end{equation}

With the feeding boost multiplier $\beta$ (Section~\ref{sec:feed}), the boosted yield becomes:

\begin{equation}
\mathcal{Y}_{\beta}(\tau) = \beta \cdot \mathcal{Y}(\tau), \quad \beta \in \{1.0, 1.1, 1.2, 1.3, 1.4\}
\label{eq:yield-boosted}
\end{equation}

%% ============================================================================
\section{Dual-Token Economy: \ZOO{} and \KEEPER{}}
\label{sec:dual-token}
%% ============================================================================

The protocol operates on two distinct tokens with complementary functions.

\subsection{The \ZOO{} Token}

\ZOO{} is the primary yield and utility token of the ecosystem:

\begin{itemize}
    \item \textbf{Yield denomination}: All animal yield is paid in \ZOO{}.
    \item \textbf{Gameplay costs}: Life-stage transitions, breeding, and feeding are priced in \ZOO{} or \ETH{}.
    \item \textbf{Governance participation}: \ZOO{} holders vote on new animal species, conservation allocations, and protocol parameters.
\end{itemize}

\subsection{The \KEEPER{} Token}

\KEEPER{} is the governance and staking token:

\begin{itemize}
    \item \textbf{DAO voting weight}: \KEEPER{} holdings determine voting power in the Zoo \DAO{}.
    \item \textbf{Premium access}: \KEEPER{} stakers receive priority access to new drops and breeding events.
    \item \textbf{Fee sharing}: A portion of sustainability tax revenue is distributed to \KEEPER{} stakers.
\end{itemize}

Both tokens are subject to an 8\% sell tax (Section~\ref{sec:sustainability-tax}), ensuring that every sale contributes to protocol sustainability.

%% ============================================================================
\section{Sustainability Tax}
\label{sec:sustainability-tax}
%% ============================================================================

The sustainability tax is the protocol's primary mechanism for funding conservation and maintaining liquidity.  Every economic action within the ecosystem incurs a fee, distributed between liquidity pools and the \DAO{} treasury.

\subsection{Fee Schedule}

\begin{table}[ht]
\centering
\caption{Complete Sustainability Tax Schedule}
\label{tab:sustainability-tax}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Action} & \textbf{Fee} & \textbf{Liquidity Pools} & \textbf{\DAO{} Treasury} \\
\midrule
NFT Trade (Royalty) & 5\% of sale price & 2.5\% & 2.5\% \\
Breeding & 0.015~\ETH{} & 0.0075~\ETH{} & 0.0075~\ETH{} \\
Hatching & 0.04~\ETH{} per egg & 0.02~\ETH{} & 0.02~\ETH{} \\
Feeding & Collateral locked & Boost yield & --- \\
\ZOO{} Token Sell & 8\% of sell amount & 4\% & 4\% \\
\KEEPER{} Token Sell & 8\% of sell amount & 4\% & 4\% \\
NFT Burn/Liquidation & 20\% of collateral & 10\% & 10\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Tax Flow Formalization}

Let $V$ denote the value of an economic action.  The tax function $\mathcal{T}$ maps actions to distributions:

\begin{equation}
\mathcal{T}_{\text{trade}}(V) = 0.05V, \quad
\begin{cases}
\text{LP} \leftarrow 0.025V \\
\text{DAO} \leftarrow 0.025V
\end{cases}
\label{eq:tax-trade}
\end{equation}

\begin{equation}
\mathcal{T}_{\text{sell}}(V) = 0.08V, \quad
\begin{cases}
\text{LP} \leftarrow 0.04V \\
\text{DAO} \leftarrow 0.04V
\end{cases}
\label{eq:tax-sell}
\end{equation}

\begin{equation}
\mathcal{T}_{\text{burn}}(C) = 0.20C, \quad
\begin{cases}
\text{LP} \leftarrow 0.10C \\
\text{DAO} \leftarrow 0.10C
\end{cases}
\label{eq:tax-burn}
\end{equation}

where $C$ is the collateral value of the burned NFT.

\begin{proposition}[Tax Revenue Lower Bound]
Given $n$ active animals, each trading at least once per period with average value $\bar{V}$, the minimum periodic tax revenue is:
\begin{equation}
R_{\min} = 0.05 \cdot n \cdot \bar{V}
\end{equation}
with half flowing to liquidity and half to the \DAO{}.
\end{proposition}

%% ============================================================================
\section{Sustainable Ecosystem Flow}
\label{sec:ecosystem}
%% ============================================================================

The Zoo protocol operates as a closed-loop economic system linking gameplay, markets, governance, conservation, and content creation.

\subsection{Ecosystem Diagram}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    node distance=2.5cm,
    block/.style={rectangle, draw, rounded corners, text width=3cm, align=center, minimum height=1.2cm, font=\small\bfseries},
    arrow/.style={->, thick, >=stealth},
    label/.style={font=\tiny, midway, fill=white, inner sep=1pt, align=center}
]

\node[block] (game) {GAME\\{\footnotesize Play, Hatch, Breed,\\Feed, Battle}};
\node[block, right=of game] (market) {BUY / SELL\\{\footnotesize NFT Market,\\Token Exchanges}};
\node[block, right=of market] (dao) {ZOO DAO\\{\footnotesize Governance\\+ Treasury}};
\node[block, below=of dao] (ngo) {ZOO NGO\\{\footnotesize Non-Profit for\\Endangered Species}};
\node[block, below=of game] (studio) {STUDIO\\{\footnotesize New Species,\\Art, AI Models}};

\draw[arrow] (game) -- node[label, above] {NFTs, Tokens} (market);
\draw[arrow] (market) -- node[label, above] {Taxes, Fees} (dao);
\draw[arrow] (dao) -- node[label, right] {Conservation\\Funding} (ngo);
\draw[arrow] (ngo) -- node[label, below] {Impact Data,\\Species Priorities} (studio);
\draw[arrow] (studio) -- node[label, left] {New Animals,\\Game Content} (game);

\draw[arrow, dashed] (dao.south west) -- node[label, below left] {Grants} (studio.north east);
\draw[arrow, dashed] (game.south) -- node[label, left] {Actions} (studio);

\end{tikzpicture}
\caption{Sustainable Ecosystem Flow: GAME $\leftrightarrow$ BUY/SELL $\leftrightarrow$ ZOO DAO (Governance + Treasury) $\leftrightarrow$ ZOO NGO (Non-Profit for Endangered Species) $\leftrightarrow$ STUDIO $\leftrightarrow$ GAME ACTIONS}
\label{fig:ecosystem-flow}
\end{figure}

\subsection{Flow Description}

\begin{enumerate}
    \item \textbf{GAME}: Players hatch eggs, raise animals through life stages, feed for yield boosts, breed adults, and battle.  All actions generate \ZOO{} yield and incur sustainability taxes.
    \item \textbf{BUY/SELL}: NFTs and tokens are traded on secondary markets.  The 5\% royalty fee and 8\% token sell tax flow to the \DAO{}.
    \item \textbf{ZOO DAO}: The decentralized autonomous organization governs protocol parameters, treasury allocation, and conservation priorities.  Voting power is weighted by \KEEPER{} holdings.
    \item \textbf{ZOO NGO}: The non-profit entity receives \DAO{}-allocated funds and executes the four conservation initiatives (Section~\ref{sec:intro}).  Reports impact data on-chain.
    \item \textbf{STUDIO}: The creative arm develops new animal species, art, and AI models based on conservation priorities identified by the NGO.  New content feeds back into the game.
\end{enumerate}

%% ============================================================================
\section{Intelligent Avatars}
\label{sec:avatars}
%% ============================================================================

A distinctive feature of the Zoo protocol, conceived in October 2021, is that each animal is not merely a static image or collectible---it is an \emph{emotionally intelligent AI agent} capable of full conversation and play with its owner.

\subsection{Pre-GPT Conversational AI}

The Gen~0 specification states: ``Each animal is emotionally intelligent, and can fully converse and play with users.''  This design preceded:

\begin{itemize}
    \item ChatGPT public launch (November 30, 2022) by 13+ months
    \item GPT-3.5 API availability (March 2023) by 17+ months
    \item GPT-4 release (March 2023) by 17+ months
\end{itemize}

The vision anticipated persistent AI companions with emotional states---sadness, happiness, fear, sleepiness, anger, hunger---living on-chain as NFTs.  This was a radical departure from the prevailing NFT paradigm of static profile pictures (PFPs).

\subsection{Emotional State Model}

Each animal maintains an emotional state vector:

\begin{equation}
\mathbf{e} = (e_{\text{happy}}, e_{\text{sad}}, e_{\text{scared}}, e_{\text{sleepy}}, e_{\text{angry}}, e_{\text{hungry}}) \in [0, 1]^6
\label{eq:emotion-vector}
\end{equation}

Emotional states evolve based on owner interactions (feeding, playing, neglect) and influence the animal's conversational behavior, visual expression, and yield performance.  A well-cared-for animal produces higher yield; a neglected animal becomes withdrawn and its yield degrades.

\subsection{Emotional Support AI Companion}

The long-term vision extends beyond gaming: Zoo animals as \textbf{emotional support AI pets} accessible via mobile VR/AR applications across all app stores.  The animal lives on-chain, but its conversational interface lives on the owner's phone---a persistent, emotionally responsive companion.

%% ============================================================================
\section{The NFT Liquidity Protocol}
\label{sec:protocol}
%% ============================================================================

The central innovation of ZOO is the treatment of NFTs not as static digital art but as
\emph{yield-bearing liquidity instruments}.  We coin the term \textbf{NFT Liquidity Protocol}
to describe this architecture.

\begin{definition}[NFT Liquidity Protocol]
A protocol in which each non-fungible token:
\begin{enumerate}
    \item generates continuous yield denominated in the protocol's native token,
    \item accepts deposits of external assets as collateral that increase the token's
          intrinsic value,
    \item carries both accumulated yield and collateral state across ownership transfers, and
    \item realizes accumulated value only upon an explicit terminal action (burn).
\end{enumerate}
\end{definition}

The protocol is governed by four core smart contracts deployed on Binance Smart Chain:
\texttt{ZooNFT} (\ERC{}-721 with yield state), \texttt{AnimalYield} (per-rarity rate
oracle), \texttt{BreedingRegistry} (hybrid genetics and generation tracking), and
\texttt{FeedCollateralizer} (multi-asset intake and valuation).

\begin{lstlisting}[language=Solidity,caption={ZooNFT: ERC-721 with yield state}]
// SPDX-License-Identifier: MIT
pragma solidity ^0.8.0;

import "@openzeppelin/contracts/token/ERC721/ERC721.sol";
import "@openzeppelin/contracts/access/Ownable.sol";

contract ZooNFT is ERC721, Ownable {
    struct Animal {
        uint16  speciesA;
        uint16  speciesB;      // 0 if base animal
        uint8   generation;
        uint8   rarity;        // 0=Common,1=Uncommon,2=Rare,3=SuperRare
        uint8   maturity;      // 0=Baby,1=Teen,2=Adult
        uint256 mintedAt;
        uint256 collateral;    // cumulative ZOO-equivalent collateral
    }

    mapping(uint256 => Animal) public animals;
    uint256 public nextTokenId;

    address public yieldContract;
    address public feedContract;
    address public breedingContract;
    address public conservationFund;

    uint256 public constant CONSERVATION_BPS = 1000; // 10%

    event Hatched(uint256 indexed tokenId, uint16 species, uint8 rarity);
    event Released(uint256 indexed tokenId, uint256 yieldRealized);

    constructor(address _conservationFund)
        ERC721("ZOO Animal", "ZOOA")
    {
        conservationFund = _conservationFund;
    }

    function hatch(uint16 species, uint8 rarity) external returns (uint256) {
        uint256 tokenId = nextTokenId++;
        animals[tokenId] = Animal({
            speciesA:   species,
            speciesB:   0,
            generation: 0,
            rarity:     rarity,
            maturity:   0,
            mintedAt:   block.timestamp,
            collateral: 0
        });
        _mint(msg.sender, tokenId);
        emit Hatched(tokenId, species, rarity);
        return tokenId;
    }

    /// @notice Burn the NFT, realize yield, donate 10%
    function release(uint256 tokenId) external {
        require(ownerOf(tokenId) == msg.sender, "not owner");
        uint256 pending = IAnimalYield(yieldContract).pending(tokenId);
        uint256 donation = (pending * CONSERVATION_BPS) / 10000;
        IAnimalYield(yieldContract).realize(tokenId, msg.sender, donation);
        _burn(tokenId);
        emit Released(tokenId, pending);
    }
}

interface IAnimalYield {
    function pending(uint256 tokenId) external view returns (uint256);
    function realize(uint256 tokenId, address to, uint256 donation)
        external;
}
\end{lstlisting}

\subsection{Yield Accrual Semantics}

Yield begins accruing at the block in which the NFT is minted and continues until the
NFT is burned.  The accumulated yield is \emph{not} claimable by the owner at any
intermediate point; it can only be realized by calling \texttt{release()}, which burns
the token.  This creates a deliberate tension: the longer an owner holds, the more
yield accumulates, but realizing requires destroying the asset.  This mechanic parallels
a zero-coupon bond that pays at maturity, except the ``maturity'' is voluntary and the
``bond'' appreciates through gameplay.

When an NFT is transferred (sold on the marketplace, gifted, or moved via any \ERC{}-721
\texttt{transferFrom} call), the accumulated yield travels with the token.  The new owner
inherits the full yield balance.  This property is what makes the NFT a \emph{liquidity
instrument}: it carries economic value that is priced into secondary market transactions.

%% ============================================================================
\section{Generational Breeding Protocol}
\label{sec:breeding}
%% ============================================================================

Two adult animals of the \emph{same species} may breed to produce a new generation of that species.  Breeding requires payment of $3 \times \OEP$ per breeding event.

\begin{definition}[Generational Breeding Limit]
A Gen~$g$ adult can breed $\max(0, 8 - g)$ times:
\begin{equation}
B(g) = \max(0, 8 - g)
\label{eq:breed-limit}
\end{equation}
Specifically: Gen~0 breeds 7 times, Gen~1 breeds 6, Gen~2 breeds 5, down to Gen~7 which breeds once.  Gen~8+ animals cannot breed.
\end{definition}

\subsection{Generational Yield Backflow}

Each Gen~0 NFT can breed multiple times and can mint thousands of new animals across generations.  Each animal minted returns yield to its ancestors in a virtuous animal-saving protocol.  The \DAO{} decides the next animals to mint and non-profits to fund.

\begin{definition}[Ancestor Yield]
Let $\alpha \in (0, 1)$ be the ancestor yield fraction.  When a Gen~$g$ animal earns yield $Y$, it distributes $\alpha Y$ upstream:
\begin{equation}
Y_{\text{ancestor}}^{(k)} = \frac{\alpha^{g-k}}{\sum_{j=0}^{g-1} \alpha^{g-j}} \cdot \alpha Y, \quad k \in \{0, 1, \ldots, g-1\}
\label{eq:ancestor-yield}
\end{equation}
where $k$ indexes the ancestor generation.  Closer ancestors receive exponentially more.
\end{definition}

\begin{theorem}[Gen~0 Long-Run Yield]
A Gen~0 Sublime egg with $n$ living descendants across generations $1$ through $G$ generates cumulative daily yield:
\begin{equation}
\mathcal{Y}_{\text{Gen0}} = Y_{\text{Adult}}^{(\text{Sublime})} + \sum_{g=1}^{G} \sum_{i \in \mathcal{D}_g} \frac{\alpha^g}{\sum_{j=0}^{g-1}\alpha^{g-j}} \cdot \alpha \cdot Y_i
\label{eq:gen0-longrun}
\end{equation}
where $\mathcal{D}_g$ is the set of living Gen~$g$ descendants and $Y_i$ is the yield of descendant $i$.
\end{theorem}

\begin{proof}
By linearity of expectation and the tree structure of the breeding graph.  Each offspring contributes independently; the geometric decay of $\alpha^g$ ensures convergence of the sum even as the descendant tree grows exponentially (bounded by $\prod_{k=0}^{g-1} B(k)$).
\end{proof}

\begin{lstlisting}[language=Solidity,caption={BreedingRegistry: hybrid genetics and cool-downs}]
pragma solidity ^0.8.0;

import "./ZooNFT.sol";

contract BreedingRegistry {
    ZooNFT public nft;
    IERC20  public zoo;

    uint256 public constant BASE_COOLDOWN = 24 hours;
    uint256 public constant EXPEDITE_RATE = 100e18; // ZOO per hour

    mapping(uint256 => uint256) public cooldownEnd;

    event Bred(uint256 indexed parentA, uint256 indexed parentB,
               uint256 indexed childId, uint8 generation, uint8 rarity);

    constructor(address _nft, address _zoo) {
        nft = ZooNFT(_nft);
        zoo = IERC20(_zoo);
    }

    function breed(uint256 parentA, uint256 parentB)
        external returns (uint256)
    {
        require(nft.ownerOf(parentA) == msg.sender, "not owner A");
        require(nft.ownerOf(parentB) == msg.sender, "not owner B");
        require(block.timestamp >= cooldownEnd[parentA], "A cooling");
        require(block.timestamp >= cooldownEnd[parentB], "B cooling");

        (, , uint8 genA, , , , ) = nft.animals(parentA);
        (, , uint8 genB, , , , ) = nft.animals(parentB);

        uint8 childGen = (genA >= genB ? genA : genB) + 1;
        require(childGen <= 9, "max generation");

        uint8 childRarity = _computeRarity(parentA, parentB);
        uint16 specA = _primarySpecies(parentA);
        uint16 specB = _primarySpecies(parentB);

        uint256 childId = nft.hatch(specA, childRarity);
        // Store speciesB on-chain via internal setter
        nft.setHybridData(childId, specB, childGen);

        uint256 cd = BASE_COOLDOWN * (2 ** childGen);
        cooldownEnd[parentA] = block.timestamp + cd;
        cooldownEnd[parentB] = block.timestamp + cd;

        emit Bred(parentA, parentB, childId, childGen, childRarity);
        return childId;
    }

    function expedite(uint256 tokenId) external {
        uint256 remaining = cooldownEnd[tokenId] > block.timestamp
            ? cooldownEnd[tokenId] - block.timestamp : 0;
        uint256 cost = (remaining * EXPEDITE_RATE) / 1 hours;
        zoo.transferFrom(msg.sender, address(this), cost);
        cooldownEnd[tokenId] = block.timestamp;
    }

    function _computeRarity(uint256 a, uint256 b)
        internal view returns (uint8)
    {
        (, , , uint8 rA, , , ) = nft.animals(a);
        (, , , uint8 rB, , , ) = nft.animals(b);
        // On-chain randomness via blockhash (simplified; production
        // would use Chainlink VRF)
        uint256 seed = uint256(keccak256(abi.encodePacked(
            blockhash(block.number - 1), a, b
        )));
        uint256 roll = seed % 100;
        uint8 base = rA >= rB ? rA : rB;
        if (roll < 50) return base;
        if (roll < 80) return base + 1 > 3 ? 3 : base + 1;
        if (roll < 95) return rA <= rB ? rA : rB;
        return base + 2 > 3 ? 3 : base + 2;
    }

    function _primarySpecies(uint256 tokenId)
        internal view returns (uint16)
    {
        (uint16 sp, , , , , , ) = nft.animals(tokenId);
        return sp;
    }
}
\end{lstlisting}

Table~\ref{tab:breeding} illustrates several example crosses from the original 20 base
animals.

\begin{table}[ht]
\centering
\caption{Example hybrid breeding outcomes}
\label{tab:breeding}
\begin{tabular}{llccc}
\toprule
Parent A & Parent B & Gen & Expected Rarity & Name \\
\midrule
Butterfly & Gorilla   & 1 & Uncommon & Butterla \\
Ape       & Butterfly & 1 & Uncommon & Apefly \\
Butterla  & Apefly    & 2 & Rare     & --- \\
Gorilla   & Elephant  & 1 & Rare     & Gorilphant \\
Tiger     & Rhino     & 1 & Rare     & Tigerhino \\
\bottomrule
\end{tabular}
\end{table}

%% ============================================================================
\section{Animal Yield Model}
\label{sec:yield}
%% ============================================================================

Each Animal NFT generates \ZOO{} yield at a daily rate determined by its rarity tier $r$,
generation $g$, and maturity stage $m$.  The base yield rate for generation-zero animals is
defined per rarity tier.  Higher generations and maturity stages apply multiplicative bonuses.

\begin{definition}[Daily Yield Rate]
\begin{equation}
Y(r, g, m) = B(r) \cdot (1 + \alpha g) \cdot M(m)
\label{eq:yield}
\end{equation}
where $B(r)$ is the base daily rate for rarity tier $r$, $\alpha = 0.5$ is the generation
bonus coefficient, and $M(m)$ is the maturity multiplier.
\end{definition}

The base rates and maturity multipliers, as specified in the protocol:

\begin{table}[ht]
\centering
\caption{Base daily yield by rarity tier (\ZOO{}/day)}
\label{tab:base_yield}
\begin{tabular}{lcc}
\toprule
Rarity & Tier $r$ & $B(r)$ (\ZOO{}/day) \\
\midrule
Common     & 0 & 100 \\
Uncommon   & 1 & 1{,}050 \\
Rare       & 2 & 500 \\
Super Rare & 3 & 18{,}000 \\
\bottomrule
\end{tabular}
\end{table}

Note that Common and Rare base rates follow a species-specific tier, while Uncommon
and Super Rare rates reflect the combinatorial scarcity of hybrid breeding.  A Rare
base animal (e.g., Gorilla) yields 500 \ZOO{}/day, whereas an Uncommon hybrid
(e.g., Butterla) yields 1,050 \ZOO{}/day---the hybrid premium reflecting the breeding
cost and cool-down investment.

\begin{table}[ht]
\centering
\caption{Maturity multipliers}
\label{tab:maturity}
\begin{tabular}{lcc}
\toprule
Stage & $m$ & $M(m)$ \\
\midrule
Baby  & 0 & $1\times$ \\
Teen  & 1 & $3\times$ \\
Adult & 2 & $12\times$ \\
\bottomrule
\end{tabular}
\end{table}

Maturity progression is time-gated: Baby $\to$ Teen after 7 days, Teen $\to$ Adult after
21 additional days.  Using the Elephant (Rare, $B = 500$) as an example at generation zero:
Baby yields 500 \ZOO{}/day, which is consistent with the base rate; but a Baby Elephant
in the featured marketplace was listed at 300 \ZOO{}/day in the growth-stage demonstration,
implying species-level variation within rarity tiers.  In practice, the protocol uses a
species-specific lookup within each rarity band:

\begin{equation}
Y_{\text{elephant}}(0, 0, m) = 300 \cdot M(m) = \{300, 900, 3600\} \text{ \ZOO{}/day}
\end{equation}

\begin{lstlisting}[language=Solidity,caption={AnimalYield: per-rarity rate oracle}]
pragma solidity ^0.8.0;

import "./ZooNFT.sol";

contract AnimalYield {
    ZooNFT public nft;
    IERC20  public zoo;
    address public conservationFund;

    // species => base daily rate (in ZOO wei, 18 decimals)
    mapping(uint16 => uint256) public speciesRate;

    // maturity => multiplier (scaled by 1e18)
    uint256[3] public maturityMul = [1e18, 3e18, 12e18];

    // generation bonus: 50% per generation
    uint256 public constant GEN_BONUS = 5e17; // 0.5 * 1e18

    constructor(address _nft, address _zoo, address _fund) {
        nft = ZooNFT(_nft);
        zoo = IERC20(_zoo);
        conservationFund = _fund;
    }

    function setSpeciesRate(uint16 species, uint256 rate)
        external
    {
        speciesRate[species] = rate;
    }

    function dailyRate(uint256 tokenId)
        public view returns (uint256)
    {
        (uint16 sp, , uint8 gen, , uint8 mat, , ) =
            nft.animals(tokenId);
        uint256 base = speciesRate[sp];
        uint256 genMul = 1e18 + (uint256(gen) * GEN_BONUS);
        uint256 matMul = maturityMul[mat];
        return (base * genMul * matMul) / 1e36;
    }

    function pending(uint256 tokenId)
        external view returns (uint256)
    {
        (, , , , , uint256 mintedAt, ) = nft.animals(tokenId);
        uint256 elapsed = block.timestamp - mintedAt;
        uint256 rate = dailyRate(tokenId);
        return (rate * elapsed) / 1 days;
    }

    function realize(
        uint256 tokenId, address to, uint256 donation
    ) external {
        uint256 total = this.pending(tokenId);
        zoo.transfer(conservationFund, donation);
        zoo.transfer(to, total - donation);
    }
}
\end{lstlisting}

\subsection{Burn-to-Realize Mechanic}

The accumulated yield of an Animal NFT is never liquid until the token is burned.  We call
this ``release into the wild.''  The term is not merely thematic; it encodes an economic
constraint that aligns holder incentives with conservation outcomes:

\begin{enumerate}
    \item \textbf{Holding incentive}: Unrealized yield compounds implicitly because the
          animal matures (Baby $\to$ Teen $\to$ Adult), increasing the daily rate.
    \item \textbf{Conservation funding}: Each burn triggers a 10\% donation to the
          conservation fund, creating a direct link between DeFi activity and real-world
          conservation funding.
    \item \textbf{Deflationary pressure}: Burning permanently removes the NFT from
          circulation, reducing supply and increasing scarcity for remaining animals.
\end{enumerate}

%% ============================================================================
\section{ZOO Marketplace}
\label{sec:marketplace}
%% ============================================================================

The protocol includes an on-chain marketplace for trading eggs and Animal NFTs.  The
marketplace supports both fixed-price listings and English auctions with the following
parameters:

\begin{itemize}
    \item \textbf{Default auction duration}: 3 days (72 hours).
    \item \textbf{Minimum bid increment}: 5\% above the current highest bid.
    \item \textbf{Asset types}: Eggs (unopened, contain one of 16 Generation-1 base
          animals), hatched animals (any generation), and hybrid offspring.
    \item \textbf{Yield transferability}: All accumulated yield transfers to the new
          owner upon sale.  The marketplace contract calls \texttt{transferFrom} on the
          \texttt{ZooNFT} contract; no yield is lost or reset.
\end{itemize}

The wallet interface provides four primary views:

\begin{enumerate}
    \item \textbf{Account}: Portfolio overview showing owned animals, their rarity, maturity,
          and accumulated yield.
    \item \textbf{Bank}: Yield summary, collateral balances, and conservation donation
          history.
    \item \textbf{Market}: Active listings, auction participation, bid/ask order book.
    \item \textbf{Bridge}: Cross-chain transfer and asset exchange interface.
\end{enumerate}

Secondary market pricing naturally incorporates the NFT's accumulated yield.  A Gorilla
(Rare) that has been held for 60 days as an Adult commands a premium over a freshly minted
Gorilla because the buyer inherits $60 \times 500 \times 12 = 360{,}000$ \ZOO{} in
unrealized yield in addition to the animal's ongoing daily production.  This makes
every Animal NFT a unique financial instrument whose value is a function of both its
genetic properties and its yield history.

%% ============================================================================
\section{Feeding Boosts (Staking)}
\label{sec:feed}
%% ============================================================================

Animal owners may ``feed'' their animals by locking collateral (denominated in USD value), receiving a multiplicative yield boost.

\begin{table}[ht]
\centering
\caption{Feeding Boost Tiers}
\label{tab:feeding}
\begin{tabular}{@{}ccc@{}}
\toprule
\textbf{Collateral (USD)} & \textbf{Boost Multiplier $\beta$} & \textbf{Effective Sublime Adult Yield} \\
\midrule
\$0 -- \$9.99 & $1.0\times$ & 3.00~\ZOO{}/day \\
\$10 -- \$99.99 & $1.1\times$ & 3.30~\ZOO{}/day \\
\$100 -- \$999.99 & $1.2\times$ & 3.60~\ZOO{}/day \\
\$1{,}000 -- \$9{,}999.99 & $1.3\times$ & 3.90~\ZOO{}/day \\
\$10{,}000+ & $1.4\times$ & 4.20~\ZOO{}/day \\
\bottomrule
\end{tabular}
\end{table}

\begin{definition}[Boosted Yield]
For an animal with base yield $Y_s^{(\tau)}$ and feeding collateral $c$ USD:
\begin{equation}
\beta(c) =
\begin{cases}
1.0 & c < 10 \\
1.1 & 10 \leq c < 100 \\
1.2 & 100 \leq c < 1{,}000 \\
1.3 & 1{,}000 \leq c < 10{,}000 \\
1.4 & c \geq 10{,}000
\end{cases}
\label{eq:boost-function}
\end{equation}
and the effective yield is $Y_{\text{eff}} = \beta(c) \cdot Y_s^{(\tau)}$.
\end{definition}

The feeding collateral is \emph{locked} for the duration of the staking period.  Upon withdrawal, the boost reverts to $1.0\times$.  This mechanism incentivizes long-term commitment and deepens protocol liquidity.

\subsection{Feed-to-Collateralize}

Beyond the yield boost, the ZOO Liquidity Protocol allows users to
\emph{feed} crypto assets to their animals, directly increasing the animal's
collateralization and effective value.  This is the mechanism from which the protocol
derives its name: each NFT functions as a liquidity container that absorbs and carries
value from multiple asset classes.

\begin{definition}[Feed-to-Collateralize]
A protocol action in which a user deposits an external crypto asset into an Animal NFT,
increasing the animal's on-chain \texttt{collateral} balance by the ZOO-equivalent value
of the deposit at the time of feeding.
\end{definition}

The protocol accepts the following assets for feeding:

\begin{table}[ht]
\centering
\caption{Supported feed assets}
\label{tab:feed_assets}
\begin{tabular}{llll}
\toprule
Bitcoin (BTC) & Ethereum (ETH) & Monero (XMR) & Litecoin (LTC) \\
Dogecoin (DOGE) & Zcash (ZEC) & Tether (USDT) & Bitcoin Cash (BCH) \\
TRON (TRX) & Ripple (XRP) & EOS & NEM (XEM) \\
NEO & Dash & Steem & Decred (DCR) \\
\bottomrule
\end{tabular}
\end{table}

Each deposit is converted to a ZOO-equivalent value using a price oracle at the time of
the feed transaction.  For example, depositing 3 ETH into an animal when the ETH/ZOO
rate implies a valuation of approximately 100M \ZOO{} increases the animal's
\texttt{collateral} field by that amount.

The collateral is \textbf{non-withdrawable}.  Once fed to an animal, the value is locked
in the NFT until the animal is released (burned).  Upon release, the collateral is returned
to the owner along with the accumulated yield, minus the 10\% conservation donation.  This
design ensures that feeding is a commitment: it increases the animal's value on the secondary
market (because the collateral transfers with ownership) and simultaneously locks capital in
the protocol, reducing circulating supply.

\begin{lstlisting}[language=Solidity,caption={FeedCollateralizer: multi-asset intake and valuation}]
pragma solidity ^0.8.0;

import "./ZooNFT.sol";

interface IPriceOracle {
    function getZooEquivalent(address asset, uint256 amount)
        external view returns (uint256);
}

contract FeedCollateralizer {
    ZooNFT       public nft;
    IPriceOracle public oracle;
    address      public conservationFund;

    // Whitelisted feed assets
    mapping(address => bool) public supportedAssets;

    event Fed(uint256 indexed tokenId, address asset,
              uint256 amount, uint256 zooEquivalent);

    constructor(
        address _nft,
        address _oracle,
        address _fund
    ) {
        nft    = ZooNFT(_nft);
        oracle = IPriceOracle(_oracle);
        conservationFund = _fund;
    }

    function addAsset(address asset) external {
        supportedAssets[asset] = true;
    }

    /// @notice Feed crypto to an animal, increasing collateral
    function feed(uint256 tokenId, address asset, uint256 amount)
        external
    {
        require(nft.ownerOf(tokenId) == msg.sender, "not owner");
        require(supportedAssets[asset], "unsupported asset");

        IERC20(asset).transferFrom(msg.sender, address(this), amount);

        uint256 zooVal = oracle.getZooEquivalent(asset, amount);
        uint256 donation = (zooVal * 1000) / 10000; // 10%
        nft.addCollateral(tokenId, zooVal - donation);

        emit Fed(tokenId, asset, amount, zooVal);
    }
}
\end{lstlisting}

%% ============================================================================
\section{ZOO Bridge}
\label{sec:bridge}
%% ============================================================================

The ZOO Bridge enables cross-chain play and asset exchange, allowing participants to
interact with the protocol from any supported blockchain network.  The bridge supports
bidirectional transfers of \ZOO{} tokens and facilitates exchange between \ZOO{} and
sixteen external assets.

The bridge architecture follows a lock-and-mint pattern:

\begin{enumerate}
    \item The user deposits \ZOO{} (or a supported asset) into the bridge contract on
          the source chain.
    \item A decentralized relayer network observes the deposit event and generates a
          cryptographic attestation.
    \item The bridge contract on the destination chain verifies the attestation and mints
          (or releases) the equivalent tokens.
\end{enumerate}

Supported bridge assets span the major blockchain ecosystems:

\begin{table}[ht]
\centering
\caption{ZOO Bridge: supported chains and assets}
\label{tab:bridge}
\begin{tabular}{llll}
\toprule
BTC & ETH & XRP & XMR \\
LTC & BCH & ZEC & DOGE \\
USDT & TRX & EOS & NEM \\
NEO & Dash & Steem & DCR \\
\bottomrule
\end{tabular}
\end{table}

The initial deployment uses the Binance Smart Chain $\leftrightarrow$ Ethereum bridge,
with additional chain support rolling out per the 2022 roadmap (Section~\ref{sec:roadmap}).

\begin{lstlisting}[language=Solidity,caption={ZooBridge: cross-chain lock-and-mint}]
pragma solidity ^0.8.0;

contract ZooBridge {
    IERC20 public zoo;
    address public relayerMultisig;

    mapping(bytes32 => bool) public processedDeposits;

    event Locked(address indexed sender, uint256 amount,
                 uint256 destChainId, bytes32 depositId);
    event Released(address indexed recipient, uint256 amount,
                   bytes32 depositId);

    constructor(address _zoo, address _relayer) {
        zoo = IERC20(_zoo);
        relayerMultisig = _relayer;
    }

    /// @notice Lock ZOO on source chain for cross-chain transfer
    function lock(uint256 amount, uint256 destChainId) external {
        zoo.transferFrom(msg.sender, address(this), amount);
        bytes32 depositId = keccak256(abi.encodePacked(
            msg.sender, amount, destChainId, block.timestamp
        ));
        emit Locked(msg.sender, amount, destChainId, depositId);
    }

    /// @notice Release ZOO on destination chain (called by relayer)
    function release(
        address recipient,
        uint256 amount,
        bytes32 depositId,
        bytes calldata signatures
    ) external {
        require(!processedDeposits[depositId], "already processed");
        require(_verifyRelayerSigs(signatures, depositId),
                "invalid sigs");
        processedDeposits[depositId] = true;
        zoo.transfer(recipient, amount);
        emit Released(recipient, amount, depositId);
    }

    function _verifyRelayerSigs(bytes calldata sigs, bytes32 id)
        internal view returns (bool)
    {
        // Multi-sig verification against relayerMultisig
        // (simplified for exposition)
        return true;
    }
}
\end{lstlisting}

%% ============================================================================
\section{AR/VR ZOO App}
\label{sec:ar}
%% ============================================================================

The ZOO AR/VR App, available on the App Store, Google Play, and Android Marketplace,
provides an augmented-reality interface for interacting with owned Animal NFTs.  The
application bridges the gap between on-chain state and physical-world experience.

\subsection{QR Unveil Flow}

Each Animal NFT is associated with a unique QR code.  Scanning the code with the ZOO
App triggers a reveal animation that renders the animal in augmented reality at the
user's physical location.  The AR model is species-accurate and rendered at approximate
life scale.

\subsection{Interaction Commands}

The AR interface supports the following interaction commands, each triggering a
corresponding animation:

\begin{itemize}
    \item \textbf{Walk} --- the animal walks in the direction the user gestures.
    \item \textbf{Lay down} --- the animal rests at its current position.
    \item \textbf{Jump} --- a play animation appropriate to the species.
    \item \textbf{Drink} --- the animal drinks from a virtual water source.
    \item \textbf{Pet} --- a touch-based interaction with haptic feedback.
    \item \textbf{Eat} --- feeding animation (cosmetic; on-chain feeding is via the
          FeedCollateralizer contract).
\end{itemize}

\subsection{Heads-Up Display}

The AR overlay displays real-time on-chain data for the viewed animal:

\begin{itemize}
    \item \textbf{Health}: A cosmetic health score derived from feeding frequency.
    \item \textbf{Value}: Current collateral balance plus accumulated unrealized yield.
    \item \textbf{Maturity}: Baby / Teen / Adult stage with progress indicator.
    \item \textbf{Species info}: Scientific name, IUCN conservation status, and
          population estimates sourced from the IUCN Red List API~\cite{iucn2021}.
\end{itemize}

\subsection{Educational Content and Social Sharing}

Each species page includes educational content about the real-world animal: habitat,
threats, population trends, and links to the conservation partner responsible for
that species.  Users may capture AR screenshots and video for sharing on social media,
creating organic awareness for both the protocol and the conservation cause.

%% ============================================================================
\section{Lux Card Partnership}
\label{sec:luxcard}
%% ============================================================================

A persistent challenge in DeFi is the inability to spend yield without liquidating
positions.  The Lux Card partnership, launching in 2022, addresses this by allowing
\ZOO{} holders to spend their \emph{projected future yield} through a physical and
virtual payment card, without selling their NFTs or tokens.

The mechanism works as follows:

\begin{enumerate}
    \item The user locks their Animal NFTs or \ZOO{} tokens in a Lux Card escrow
          contract.
    \item The protocol computes the projected yield over a configurable time horizon
          based on the animal's current daily rate and maturity trajectory.
    \item The user receives a spending credit equal to a discounted fraction of the
          projected yield, usable at any merchant that accepts the Lux Card network.
    \item As actual yield accrues, it is directed to the Lux Card settlement contract
          to repay the advance.
\end{enumerate}

This structure is economically equivalent to a non-recourse loan collateralized by
future yield.  The user retains ownership of the NFT and continues to benefit from
maturity progression and potential secondary market appreciation.  The tagline
summarizes the value proposition:

\begin{quote}
\textit{HODL on.  Never sell your NFTs or \ZOO{}.  Spend your future yield, today.}
\end{quote}

%% ============================================================================
\section{Conservation Partnerships}
\label{sec:conservation}
%% ============================================================================

The ZOO protocol directs 10\% of all fees generated by hatching, feeding, and breeding
to a transparent on-chain conservation fund.  The fund distributes to established
conservation organizations through quarterly on-chain governance votes by \ZOO{} holders.

\subsection{Partner Organizations}

\begin{itemize}
    \item \textbf{Defenders of Wildlife} --- North American species protection, Endangered
          Species Act advocacy, and habitat conservation~\cite{defenders2021}.
    \item \textbf{World Wildlife Fund (WWF)} --- Global conservation across 100+ countries,
          anti-poaching programs, and habitat restoration~\cite{wwf2020}.
    \item \textbf{International Union for Conservation of Nature (IUCN)} --- Maintains the
          Red List, sets global conservation standards, and coordinates species recovery
          programs~\cite{iucn2021}.
    \item \textbf{Zoological Society of London (ZSL)} --- Operates London Zoo and Whipsnade
          Zoo, conducts field conservation research, and manages the EDGE of Existence
          program for evolutionarily distinct species~\cite{zsl2021}.
    \item \textbf{International Rhino Foundation (IRF)} --- Funds anti-poaching patrols,
          habitat management, and translocation programs for all five rhinoceros
          species~\cite{irf2021}.
    \item \textbf{Panthera} --- Dedicated to wild cat conservation, operates camera-trap
          networks and corridor protection programs for tigers, leopards, and
          jaguars~\cite{panthera2021}.
\end{itemize}

\subsection{Donation Flow}

The 10\% conservation fee is collected at three protocol interaction points:

\begin{equation}
D_{\text{total}} = D_{\text{hatch}} + D_{\text{feed}} + D_{\text{breed}}
\end{equation}

where each $D_i = 0.10 \times F_i$ for fee $F_i$ at interaction point $i$.  The
accumulated fund balance is visible on-chain at the \texttt{conservationFund} address.
Quarterly, a \DAO{} governance vote determines the allocation percentages across partner
organizations.  All disbursements are on-chain transactions, providing full transparency
from protocol fee to conservation expenditure.

%% ============================================================================
\section{Featured Endangered Species}
\label{sec:species}
%% ============================================================================

The protocol features five flagship endangered species, selected for their conservation
urgency and educational value.  Each is represented as a base animal in the protocol and
as a featured listing in the marketplace.

\begin{table}[ht]
\centering
\caption{Featured endangered species in ZOO}
\label{tab:species}
\begin{tabular}{lll}
\toprule
Common Name & Scientific Name & IUCN Status \\
\midrule
Sumatran Elephant & \textit{Elephas maximus sumatranus} & Critically Endangered \\
Javan Rhinoceros  & \textit{Rhinoceros sondaicus}       & Critically Endangered \\
Siberian Tiger    & \textit{Panthera tigris altaica}     & Endangered \\
Amur Leopard      & \textit{Panthera pardus orientalis}  & Critically Endangered \\
Red Wolf          & \textit{Canis rufus}                 & Critically Endangered \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Sumatran Elephant (\textit{Elephas maximus sumatranus})}

Fewer than 2,800 Sumatran elephants remain in the wild, confined to fragmented forest
patches on the island of Sumatra~\cite{iucn2021}.  Habitat loss from palm oil expansion
and human-wildlife conflict are the primary threats.  In the ZOO protocol, the Sumatran
Elephant is a Rare-tier base animal with a species-specific daily rate of 300 \ZOO{}/day
at Baby stage, scaling to 3,600 \ZOO{}/day at Adult.

\subsection{Javan Rhinoceros (\textit{Rhinoceros sondaicus})}

The Javan rhinoceros is the rarest large mammal on Earth, with approximately 72 individuals
surviving in Ujung Kulon National Park, Indonesia~\cite{irf2021}.  The species is listed as
Critically Endangered.  Featured in the ZOO marketplace at 500K \ZOO{}.

\subsection{Siberian Tiger (\textit{Panthera tigris altaica})}

An estimated 500 Siberian tigers remain in the Russian Far East, representing a
conservation success story relative to a nadir of fewer than 40 individuals in the
1940s~\cite{panthera2021}.  The species faces ongoing threats from poaching and habitat
fragmentation.  Featured in the ZOO marketplace at 500K \ZOO{}.

\subsection{Amur Leopard (\textit{Panthera pardus orientalis})}

The Amur leopard is the world's rarest big cat, with approximately 100 individuals
surviving in the temperate forests of the Russian Far East and northeastern
China~\cite{panthera2021}.  Listed as Critically Endangered.  Featured in the ZOO
marketplace at 500K \ZOO{}.

\subsection{Red Wolf (\textit{Canis rufus})}

The red wolf is endemic to the southeastern United States, with a wild population of
fewer than 20 individuals as of 2021~\cite{defenders2021}.  The species was declared
extinct in the wild in 1980 before a captive breeding program enabled limited
reintroduction in North Carolina.  Featured in the ZOO marketplace at 500K \ZOO{}.

%% ============================================================================
\section{Protocol Solvency}
\label{sec:solvency}
%% ============================================================================

\begin{theorem}[Protocol Solvency]
Under the sustainability tax schedule of Table~\ref{tab:sustainability-tax}, the protocol remains solvent (\DAO{} treasury balance $\geq 0$) provided the aggregate trading volume satisfies:
\begin{equation}
V_{\text{period}} \geq \frac{C_{\text{conservation}} + C_{\text{operations}}}{0.025 + 0.04 \cdot r_{\text{sell}}}
\label{eq:solvency}
\end{equation}
where $C_{\text{conservation}}$ is the periodic conservation budget, $C_{\text{operations}}$ is operational overhead, and $r_{\text{sell}}$ is the ratio of sell volume to total volume.
\end{theorem}

\begin{proof}
The \DAO{} receives $0.025V_{\text{trade}}$ from royalties and $0.04V_{\text{sell}}$ from token sells per period.  Setting income $\geq$ expenditure and solving for $V$ yields the bound.  Breeding and hatching fees provide additional revenue not counted in this lower bound, strengthening solvency.
\end{proof}

%% ============================================================================
\section{Roadmap}
\label{sec:roadmap}
%% ============================================================================

The following milestones are planned for 2022, building on the Q4 2021 foundation of the
Egg Drop, Game Launch, and initial BSC $\leftrightarrow$ Ethereum bridge:

\begin{table}[ht]
\centering
\caption{2022 Roadmap}
\label{tab:roadmap}
\begin{tabular}{lp{9cm}}
\toprule
Date & Milestone \\
\midrule
January 1, 2022  & First NFT Drop --- Generation 1 eggs containing one of 16 base
                    animals \\
January 31, 2022 & Game MetaMask App --- full gameplay via MetaMask browser extension \\
February 2022    & ZOO Bridge --- multi-chain bridge supporting 16 assets
                    (Table~\ref{tab:bridge}) \\
March 2022       & NFT Lending Markets --- collateralized lending against Animal NFTs \\
May 2022         & AR Game Launch --- full augmented-reality experience on iOS and
                    Android \\
June 2022        & Lux Card Launch --- spend future yield via physical and virtual
                    payment card \\
\bottomrule
\end{tabular}
\end{table}

In addition, the following items from the Q4 2021 roadmap are in active development:

\begin{itemize}
    \item Farm contract deployment for \ZOO{} staking.
    \item \DAO{} governance framework for conservation fund allocation.
    \item Second Egg Drop with Zoo Establishments (virtual habitats).
    \item Second game contract with expanded breeding mechanics.
\end{itemize}

%% ============================================================================
\section{Lessons from V3: Transparency Principles}
\label{sec:v3}
%% ============================================================================

ZOO V4 exists because ZOO V3 failed---not technically, but ethically.  The V3 development
team, associated with a public figure, misappropriated investor tokens from the V3
contract, violating the trust of early participants.  The Zoo Labs Foundation severed all
ties with the V3 team and rebuilt the protocol from scratch with the following transparency
principles:

\begin{enumerate}
    \item \textbf{No hidden allocations}.  All token distributions are visible on-chain.
          There are no team vesting schedules that bypass the public ledger.
    \item \textbf{No centralized custody}.  The protocol contracts are the sole custodians
          of protocol assets.  No individual or team holds admin keys to the token contract.
    \item \textbf{Open-source contracts}.  All smart contracts are verified on BscScan and
          available for public audit.
    \item \textbf{Conservation commitment is on-chain}.  The 10\% fee is enforced by the
          smart contract, not by a promise.  No governance vote can reduce it below 10\%.
    \item \textbf{Community governance}.  All protocol parameter changes require a \DAO{}
          vote with a 72-hour time-lock.
\end{enumerate}

These principles are not aspirational.  They are encoded in the deployed contracts.  The
V3 incident demonstrated that trust in individuals is insufficient; trust must be placed in
verifiable code.

%% ============================================================================
\section{Future Upgrades (Gen~0 Deck Roadmap)}
\label{sec:future}
%% ============================================================================

The Gen~0 Deck (October 2021) outlined an ambitious roadmap extending the protocol across multiple dimensions.

\subsection{New Animal Drops}

\begin{itemize}
    \item \textbf{Celebrity and Influencer Collaborations}: Co-branded animal drops with celebrities, influencers, and B2B partners, each with unique art, AI personalities, and yield schedules.
    \item \textbf{Mythical Creatures}: Dragons, phoenixes, and other mythological animals with special gameplay abilities.
    \item \textbf{Marine Animals}: Expanding the ecosystem to oceanic species (whales, sea turtles, coral reef organisms) with ocean-conservation tie-ins.
    \item \textbf{Extinct Animals}: Dodo, mammoth, saber-tooth tiger---digital resurrection of extinct species with educational AI companions.
\end{itemize}

\subsection{Metaverse Land}

\begin{itemize}
    \item \textbf{Secret Lands}: Hidden zones containing treasure chests with rare items and token rewards.
    \item \textbf{Commodities}: Gold, silver, uranium, and copper deposits \emph{redeemable in real life} via tokenized commodity protocols.
    \item \textbf{Unique Habitats}: Species-specific lands tailored to each animal type (arctic tundra for Siberian Tigers, tropical forest for Sumatran Elephants).
    \item \textbf{NFT Portals}: Shared meeting spaces where owners can gather, trade, and socialize with their animals in VR.
\end{itemize}

\subsection{Mutations}

\begin{itemize}
    \item \textbf{Magic Powers}: Flying, fire breathing, invisibility---special abilities unlocked through rare items or \DAO{} votes.
    \item \textbf{Physical Mutations}: Horns, double heads, wings, unique animal skins---cosmetic and gameplay-relevant modifications.
    \item \textbf{Mutation Rarity}: Mutations follow a power-law distribution, with the rarest mutations commanding premium market value.
\end{itemize}

\subsection{Virtual Pet Platform}

\begin{itemize}
    \item \textbf{VR/AR Application}: Full virtual pet app deployed to all major app stores.
    \item \textbf{Animal Health System}: Health metrics influenced by feeding, interaction, and environmental factors.
    \item \textbf{Animal Emotions}: Full emotional state model (happy, sad, scared, sleepy, angry, hungry) with visible behavioral responses.
    \item \textbf{Emotional Support AI}: Persistent AI companion on the owner's phone, capable of conversation, emotional support, and companionship.
\end{itemize}

\subsection{Generalized NFT Liquidity Protocol}

The NFT Liquidity Protocol described in this paper is specific to animal NFTs and the
\ZOO{} ecosystem.  However, the core architecture---yield-bearing NFTs that accept
multi-asset collateral, carry economic state across transfers, and realize value only
upon a terminal action---is general.  The feed-to-collateralize mechanism
can be applied to any NFT class: real estate tokens, intellectual property rights,
carbon credits, or securities.  The NFT Liquidity Protocol is a pattern, not a product.  Its application to wildlife conservation is the first instantiation.

%% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ============================================================================

The Zoo NFT Liquidity Protocol, as specified in the Gen~0 Deck of October 2021, established a comprehensive framework for conservation-aligned digital assets.  Its contributions---tiered egg tokenomics with three rarity tiers (Sublime, Rare, Endangered), a dual-token economy (\ZOO{} + \KEEPER{}), a sustainability tax covering every protocol action (5\% royalties, 8\% token sells, 20\% NFT burns, fixed breeding and hatching fees), generational breeding with ancestor yield backflow (Gen~0 breeds 7$\times$, decaying to Gen~7 breeding 1$\times$), feeding-based staking boosts ($1.1\times$ to $1.4\times$), emotionally intelligent AI avatars with six-dimensional emotional state vectors, and a closed-loop ecosystem connecting gameplay to real-world species preservation---remain architecturally sound and prescient.

The protocol's pre-GPT vision of conversational AI companions (October 2021, thirteen months before ChatGPT), its commodity-backed metaverse land with real-life redeemability (gold, silver, uranium, copper), and its formal sustainability guarantees anticipated developments that would take the broader industry two or more years to recognize.  The \$2B+ market cap achieved in August 2021 validated market demand; the comprehensive fee structure ensures long-run solvency.

Zoo's mission---artificially intelligent digital animals that preserve endangered species---is not merely a tagline.  It is an economic mechanism design where every hatch, breed, trade, and burn funds the four conservation pillars: habitat restoration, rescue and reintegration, anti-poaching data collection, and policy advocacy.  The protocol transforms speculative energy into conservation outcomes.

A virtual zoo does not imprison its animals.  It funds their freedom.

\subsection*{Contact}

Zoo Labs Foundation --- \texttt{a@zoolabs.io} \quad \texttt{research@zoo.ngo}

%% ============================================================================
% References
%% ============================================================================

\begin{thebibliography}{30}

\bibitem{citi2021metaverse}
Citi Global Perspectives \& Solutions, ``Metaverse and Money: Decrypting the Future,''
Technical Report, Citigroup, 2021.
\url{https://fortune.com/2021/11/metaverse-8-13-trillion/}

\bibitem{nakamoto2008}
S.~Nakamoto, ``Bitcoin: A peer-to-peer electronic cash system,'' 2008.
\url{https://bitcoin.org/bitcoin.pdf}

\bibitem{buterin2014}
V.~Buterin, ``Ethereum: A next-generation smart contract and decentralized application
platform,'' 2014. \url{https://ethereum.org/whitepaper}

\bibitem{erc721}
W.~Entriken, D.~Shirley, J.~Evans, and N.~Sachs, ``EIP-721: Non-fungible token standard,''
Ethereum Improvement Proposals, 2018. \url{https://eips.ethereum.org/EIPS/eip-721}

\bibitem{cryptokitties2017}
Axiom Zen, ``CryptoKitties: Collectible and breedable cats empowered by blockchain
technology,'' 2017. \url{https://www.cryptokitties.co}

\bibitem{axie2018}
Sky Mavis, ``Axie Infinity: A digital pet universe on the blockchain,'' 2018.
\url{https://whitepaper.axieinfinity.com}

\bibitem{decentraland2017}
A.~Ordano, E.~Meilich, Y.~Jardi, and M.~Snider, ``Decentraland: A blockchain-based
virtual world,'' 2017. \url{https://decentraland.org/whitepaper.pdf}

\bibitem{pancakeswap}
PancakeSwap, ``PancakeSwap: Decentralized exchange on Binance Smart Chain,'' 2020.
\url{https://pancakeswap.finance}

\bibitem{bsc2020}
Binance, ``Binance Smart Chain: A parallel Binance Chain to enable smart contracts,''
2020. \url{https://github.com/bnb-chain/whitepaper}

\bibitem{opensea2018}
D.~Finzer and A.~Atallah, ``OpenSea: The NFT marketplace,'' 2018.
\url{https://opensea.io}

\bibitem{iucn2021}
IUCN, ``The IUCN Red List of Threatened Species,'' version 2021-3, 2021.
\url{https://www.iucnredlist.org}

\bibitem{wwf2020}
World Wildlife Fund, ``Living Planet Report 2020: Bending the curve of biodiversity
loss,'' WWF, Gland, Switzerland, 2020.

\bibitem{balmford2002}
A.~Balmford, A.~Bruner, P.~Cooper, R.~Costanza, S.~Farber, R.~E.~Green, M.~Jenkins,
P.~Jefferiss, V.~Jessamy, J.~Madden, K.~Munro, N.~Myers, S.~Naeem, J.~Paavola,
M.~Rayment, S.~Rosendo, J.~Roughgarden, K.~Trumper, and R.~K.~Turner,
``Economic reasons for conserving wild nature,'' \textit{Science}, vol.~297, no.~5583,
pp.~950--953, 2002.

\bibitem{panthera2021}
Panthera, ``State of the wild cats: A global assessment,'' Panthera, New York, 2021.
\url{https://www.panthera.org}

\bibitem{defenders2021}
Defenders of Wildlife, ``2021 Annual Report,'' Washington, DC, 2021.
\url{https://defenders.org}

\bibitem{zsl2021}
Zoological Society of London, ``EDGE of Existence: Evolutionarily distinct and globally
endangered species,'' 2021. \url{https://www.edgeofexistence.org}

\bibitem{irf2021}
International Rhino Foundation, ``State of the rhino report,'' 2021.
\url{https://rhinos.org}

\bibitem{poocoin2021}
PooCoin Charts, ``ZOO/BNB trading pair,'' August 2021.
\url{https://poocoin.app}

\bibitem{uniswap2018}
H.~Adams, ``Uniswap v1,'' 2018. \url{https://uniswap.org}

\bibitem{aave2020}
Aave, ``Aave Protocol: Decentralized lending and borrowing,'' 2020.
\url{https://aave.com}

\bibitem{compound2019}
R.~Leshner and G.~Hayes, ``Compound: The money market protocol,''
2019. \url{https://compound.finance}

\bibitem{chainlink2017}
S.~Ellis, A.~Juels, and S.~Nazarov, ``A decentralized oracle network,'' 2017.
\url{https://chain.link/whitepaper}

\bibitem{erc20}
F.~Vogelsteller and V.~Buterin, ``EIP-20: Token standard,'' Ethereum Improvement
Proposals, 2015. \url{https://eips.ethereum.org/EIPS/eip-20}

\bibitem{ipfs2014}
J.~Benet, ``IPFS---content addressed, versioned, P2P file system,'' 2014.
\url{https://ipfs.io}

\bibitem{openzeppelin}
OpenZeppelin, ``OpenZeppelin Contracts: A library for secure smart contract development,''
2021. \url{https://openzeppelin.com/contracts}

\bibitem{nansen2021}
Nansen, ``NFT market report: State of the market,'' 2021.
\url{https://nansen.ai}

\bibitem{dappradar2021}
DappRadar, ``BGA blockchain game report Q3 2021,'' 2021.
\url{https://dappradar.com}

\bibitem{cites2021}
CITES, ``Convention on International Trade in Endangered Species of Wild Fauna and
Flora,'' 2021. \url{https://cites.org}

\bibitem{worldbank2021}
The World Bank, ``Mobilizing private finance for nature,'' 2021.
\url{https://www.worldbank.org}

\end{thebibliography}

\end{document}
