\section{The PQZ Algorithm}


Post-Quantum Zero-Knowledge (PQZ) combines three primitives into a unified framework:

\subsection{Poseidon2 Commitments}

We replace Pedersen commitments (which rely on discrete log hardness) with Poseidon2 \cite{grassi2023poseidon2}, a ZK-friendly algebraic hash function secure under collision resistance assumptions that do not require elliptic curve hardness.

\begin{definition}[Poseidon2 Commitment]
For message $m \in \mathbb{F}_p^n$ and randomness $r \in \mathbb{F}_p$:
$$\text{Commit}(m, r) = \text{Poseidon2}(m \| r)$$
\end{definition}

\textbf{Performance.} Benchmarks on Zoo Network validators (NVIDIA H100):

\begin{table}[h]
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Operation} & \textbf{Poseidon2} & \textbf{Pedersen} & \textbf{Speedup} \\
\midrule
Commitment & 29 ns & 92 $\mu$s & 2000$\times$ \\
Note commit & 50 ns & 99 $\mu$s & 1980$\times$ \\
Merkle path (depth 32) & 928 ns & 2.9 ms & 3125$\times$ \\
Gas cost & 800 & 6,000 & 7.5$\times$ \\
\bottomrule
\end{tabular}
\caption{Poseidon2 vs Pedersen on GPU-accelerated validators.}
\end{table}

\subsection{STARK Proofs over Goldilocks}

For the core proving system, PQZ uses STARKs over the Goldilocks field ($p = 2^{64} - 2^{32} + 1$), which provides:

\begin{itemize}
\item \textbf{Transparency}: No trusted setup. Verification relies only on hash function collision resistance.
\item \textbf{Post-quantum security}: Security reduces to hash collision resistance (Poseidon2), not discrete log or factoring.
\item \textbf{Efficient FRI}: The Goldilocks field admits $2^{32}$-th roots of unity, enabling efficient Fast Reed-Solomon Interactive Oracle Proofs.
\end{itemize}

\subsection{ML-DSA Receipt Signing}

After STARK verification, the Z-Chain validator set signs the resulting receipt using ML-DSA (FIPS 204) threshold signatures coordinated through K-Chain:

\begin{enumerate}
\item Prover submits STARK proof to Z-Chain.
\item Z-Chain validators independently verify the proof.
\item Upon $\geq 67\%$ agreement, validators produce a threshold ML-DSA signature on the receipt.
\item The signed receipt is stored on-chain and indexed by receipt hash.
\end{enumerate}

\subsection{Groth16 Export for External EVMs}

For interoperability with Ethereum and other EVM chains that only support BN254 pairings, PQZ provides a Groth16 export:

\begin{enumerate}
\item Take a verified STARK receipt from Z-Chain.
\item Generate a Groth16 proof that ``I know a STARK proof that was verified and signed by the Zoo validator set.''
\item Export the Groth16 proof to the target EVM chain.
\end{enumerate}

This wraps the post-quantum STARK in a classical Groth16 for compatibility, while the original STARK receipt remains the source of truth on Z-Chain.

