\section{Conclusion}


Z-Chain provides the first post-quantum zero-knowledge infrastructure integrated natively into a blockchain consensus layer. The PQZ algorithm unifies STARK transparency with ML-DSA quantum resistance and Groth16 EVM compatibility. Universal receipts enable composable proofs across chains. Formal verification in Lean 4 ensures soundness, completeness, and zero-knowledge. GPU acceleration achieves practical throughput (12,000 verifications/sec) for production deployment.

\begin{thebibliography}{20}

\bibitem{grassi2023poseidon2}
Grassi, L., Khovratovich, D., Rechberger, C., et al., ``Poseidon2: A New Hash Function for Zero-Knowledge Proof Systems,'' IACR ePrint 2023/323, 2023.

\bibitem{fips203}
NIST, ``Module-Lattice-Based Key-Encapsulation Mechanism Standard (ML-KEM),'' FIPS 203, 2024.

\bibitem{fips204}
NIST, ``Module-Lattice-Based Digital Signature Standard (ML-DSA),'' FIPS 204, 2024.

\bibitem{fips205}
NIST, ``Stateless Hash-Based Digital Signature Standard (SLH-DSA),'' FIPS 205, 2024.

\bibitem{stark}
Ben-Sasson, E., Bentov, I., Horesh, Y., Riabzev, M., ``Scalable, Transparent, and Post-Quantum Secure Computational Integrity,'' IACR ePrint 2018/046, 2018.

\bibitem{groth16}
Groth, J., ``On the Size of Pairing-Based Non-Interactive Arguments,'' EUROCRYPT 2016.

\bibitem{plonk}
Gabizon, A., Williamson, Z., Ciobotaru, O., ``PLONK: Permutations over Lagrange-Bases for Oecumenical Noninteractive Arguments of Knowledge,'' IACR ePrint 2019/953, 2019.

\bibitem{zoo-consensus}
Kelling, Z., ``Quasar Hybrid Consensus on Zoo Network,'' Zoo Labs Foundation, 2024.

\bibitem{zoo-pq}
Kelling, Z., ``Post-Quantum Cryptography Suite for Zoo Network,'' Zoo Labs Foundation, 2024.

\bibitem{zoo-fhe}
Kelling, Z., ``Threshold Fully Homomorphic Encryption on Zoo Network,'' Zoo Labs Foundation, 2025.

\end{thebibliography}

